### 2019-07-16 Article.tex
Update on Overleaf.

\documentclass[times, twoside, watermark]{zHenriquesLab-StyleBioRxiv}
\usepackage{blindtext}

% Please give the surname of the lead author for the running footer
\leadauthor{Denovellis}

\begin{document}

\title{Categorizing the trajectory dynamics of replay}
\shorttitle{My Template}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[1,\Letter]{Eric L. Denovellis}
\author[2, 3]{Anna K. Gillespie}
\author[2, 3]{Michael E. Coulter}
\author[2, 3, 4]{Loren M. Frank}
\author[1]{Uri T. Eden}

\affil[1]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
\affil[2]{Department of Physiology, University of California, San Francisco, San Francisco, California}
\affil[3]{Kavli Institute for Fundamental Neuroscience, University of California, San Francisco, San Francisco, California}
\affil[4]{Howard Hughes Medical Institute, University of California, San Francisco, San Francisco, California}

\maketitle

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
During sleep and immobility, hippocampal place cells fire in sequences consistent with temporally compressed versions of trajectories previously run by the animal. These replayed sequences are hypothesized to be an important mechanism for the retrieval of spatial memory in service of consolidation and decision-making. Replay events are typically evaluated based on whether they activate sequences of place cells that represent spatially continuous trajectories through the environment, but recent work has shown that these events can have more complex dynamics. For example, sequences can alternate between hovering on a particular spatial location and continuous movement or can represent continuous trajectories in other spatial environments, which may appear spatially incoherent in the context of the current environment.

To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics. Each discrete latent “category” is associated with a type of continuous latent dynamic—such as hovering, spatially fragmented or spatially continuous. This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories. In addition, the model can function in 2D, avoiding linearization errors on more complicated environments. We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.

\end {abstract}
%TC:break main
%the command above serves to have a word count for the abstract

\begin{keywords}
Hippocampus | Replay | State Space | Point Process | Switching
\end{keywords}

\begin{corrauthor}
%\texttt{edeno{@}bu.edu}
edeno\at bu.edu
\end{corrauthor}

\section*{Introduction}
\Blindtext

\section*{Results}

Just for kicks here's a citation \cite{Gustafsson2016}. And a reference to a supplement \cref{note:Note1}. And \nameref{note:Note1}.
\Blindtext

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{Figures/Figure_1}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{fig:computerNo}
\end{figure}

\Blindtext

Figure \ref{fig:computerNo} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\section*{Conclusions}

blablaba \ref{fig:computerNo} 
\blindtext

\subsection*{Blabla} 
\blindtext

\section*{Conclusions}

blablaba \ref{fig:computerNo}
\blindtext

\begin{acknowledgements}
\blindtext
\end{acknowledgements}

\section*{Bibliography}
\bibliography{zHenriquesLab-Mendeley}

%% You can use these special %TC: tags to ignore certain parts of the text.
%TC:ignore
%the command above ignores this section for word count
\onecolumn
\newpage

\section*{Word Counts}
This section is \textit{not} included in the word count. 
\subsection*{Notes on Nature Methods Brief Communication}
\begin{itemize}
\item Abstract: 3 sentences, 70 words.
\item Main text: 3 pages, 2 figures, 1000-1500 words, more figures possible if under 3 pages
\end{itemize}

\subsection*{Statistics on word count}
\detailtexcount
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Supplementary Information %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup*{format=largeformat}
\section{Something about something} \label{note:Note1} 
\Blindtext

%TC:endignore
%the command above ignores this section for word count

\end{document}

### 2019-07-16 scripts/run_by_epoch.py
delete folder

@@ -17,9 +17,11 @@ from loren_frank_data_processing import save_xarray
 from src.analysis import (get_linear_position_order, get_place_field_max,
                           get_replay_info, reshape_to_segments)
 from src.load_data import load_data
-from src.parameters import (FIGURE_DIR, PROCESSED_DATA_DIR, SAMPLING_FREQUENCY,
-                            discrete_diag, model, model_kwargs, movement_var,
-                            place_bin_size, replay_speed)
+from src.parameters import (FIGURE_DIR, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR, SAMPLING_FREQUENCY,
+                            TRANSITION_TO_CATEGORY, discrete_diag, model,
+                            model_kwargs, movement_var, place_bin_size,
+                            replay_speed)
 from src.visualization import (plot_category_counts, plot_category_duration,
                                plot_neuron_place_field_2D_1D_position,
                                plot_ripple_decode)
@@ -28,15 +30,6 @@ FORMAT = '%(asctime)s %(message)s'
 
 logging.basicConfig(level='INFO', format=FORMAT, datefmt='%d-%b-%y %H:%M:%S')
 plt.switch_backend('agg')
-
-
-TRANSITION_TO_CATEGORY = {
-    'identity': 'hover',
-    'uniform': 'fragmented',
-    'random_walk': 'continuous',
-}
-
-PROBABILITY_THRESHOLD = 0.8
 
 
 def run_analysis(epoch_key, make_movies=False, data_type='sorted_spikes'):
@@ -225,12 +218,13 @@ def main():
                    stdout=PIPE, universal_newlines=True).stdout
     logging.info('Git Hash: {git_hash}'.format(git_hash=git_hash.rstrip()))
 
-    client = Client(n_workers=14, threads_per_worker=4,
-                    processes=True, memory_limit='25GB')
-    logging.info(client)
-    # Analysis Code
-    run_analysis(epoch_key, data_type='clusterless')
-    run_analysis(epoch_key, data_type='sorted_spikes')
+    client_params = dict(n_workers=14, threads_per_worker=4, processes=True,
+                         memory_limit='25GB')
+    with Client(**client_params) as client:
+        logging.info(client)
+        # Analysis Code
+        run_analysis(epoch_key, data_type='clusterless')
+        run_analysis(epoch_key, data_type='sorted_spikes')
 
 
 if __name__ == '__main__':

### 2019-07-16 src/parameters.py

@@ -39,8 +39,16 @@ replay_speed = 1
 model = NumbaKDE
 model_kwargs = {
     'bandwidth': np.array([24.0, 24.0, 24.0, 24.0, 6.0, 6.0])
-    }
+}
 discrete_diag = 1 - 1E-3
+
+TRANSITION_TO_CATEGORY = {
+    'identity': 'hover',
+    'uniform': 'fragmented',
+    'random_walk': 'continuous',
+}
+
+PROBABILITY_THRESHOLD = 0.8
 
 # Plotting Colors
 STATE_COLORS = {

### 2019-07-16 Article.tex
Update on Overleaf.

@@
 To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics.
-Each discrete latent “category” is associated with a type of continuous latent dynamic—such as hovering, spatially fragmented or spatially continuous.
+Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous.
 This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories.
@@
 \begin{keywords}
-Hippocampus | Replay | State Space | Point Process | Switching
+Hippocampus | Replay | State Space
 \end{keywords}
@@
 \centering
-\includegraphics[width=.8\linewidth]{Figures/Figure_1}
+\includegraphics[width=.8\linewidth]{figures/Figure_1}
 \caption{Placeholder image of Iris with a long example caption to show justification setting.}
@@
 
-\section*{Conclusions}
+\section*{Discussion}
 
@@
 \blindtext
-
-\section*{Conclusions}
 
@@
 
-%% You can use these special %TC: tags to ignore certain parts of the text.
-%TC:ignore
-%the command above ignores this section for word count
 \onecolumn
@@
 
-\section*{Word Counts}
-This section is \textit{not} included in the word count. 
-\subsection*{Notes on Nature Methods Brief Communication}
-\begin{itemize}
-\item Abstract: 3 sentences, 70 words.
-\item Main text: 3 pages, 2 figures, 1000-1500 words, more figures possible if under 3 pages
-\end{itemize}
+\section*{Materials and Methods}
 
-\subsection*{Statistics on word count}
-\detailtexcount
+\subsection*{Filter Derivation}
+\begin{enumerate}
+\item Use Bayes theorem to define the posterior in terms of a data-derived likelihood and a prior ($ posterior \propto likelihood * prior $)
+$$
+p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k-1})
+$$
+
+\item Use Chapman-Kolmogorov to define prior in terms of previous time step's posterior and state transitions
+$$
+\begin{align*}
+p(x_{k}, I_{k} \mid y_{1:k-1}) &= \sum_{I_{k-1}} \int p(x_{k}, x_{k-1}, I_{k}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}, y_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+\\ &= \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+\end{align*}$$
+
+where:
+$p(x_{k-1}, I_{k-1} \mid y_{1:k-1})$ is the previous posterior,
+$Pr(I_{k} \mid I_{k-1}, x_{k-1})$ is the discrete state transition, and $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ is the continuous state transition
+
+\item Final Filter
+$$
+p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+$$
+\end{enumerate}
+
+\subsection*{Smoother Derivation}
+
+\begin{enumerate}
+\item
+$$
+\begin{align*}
+p(x_{k}, I_{k} \mid y_{1:T}) &= \sum_{I_{k+1}} \int p(x_{k}, x_{k+1}, I_{k}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:T}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:k}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k}, I_{k}, x_{k+1}, I_{k+1} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
+\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
+\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
+\end{align*}$$
+
+\item
+$$
+\begin{align*}
+p(x_{k+1}, I_{k+1} \mid y_{1:k}) &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1}, x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
+\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
+\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
+\\ &= \sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
+\end{align*}$$
+
+\end{enumerate}
+
 \newpage

### 2019-07-16 Article.tex
Update on Overleaf.

@@
 \usepackage{blindtext}
+\usepackage{amsmath}
 
@@
 \\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-\\ &= \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+\\ &= \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
 \end{align*}$$
@@
 $p(x_{k-1}, I_{k-1} \mid y_{1:k-1})$ is the previous posterior,
-$Pr(I_{k} \mid I_{k-1}, x_{k-1})$ is the discrete state transition, and $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ is the continuous state transition
+$Pr(I_{k} \mid I_{k-1})$ is the discrete state transition, and $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ is the continuous state transition
 
@@
 $$
-p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}, x_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
+p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
 $$
@@
 \\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
-\\ &= \sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
+\\ &= \sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
 \end{align*}$$

### 2019-07-17 src/parameters.py
Fix colors

@@ -52,9 +52,9 @@ PROBABILITY_THRESHOLD = 0.8
 
 # Plotting Colors
 STATE_COLORS = {
-    'hover': '#ff6944',
-    'fragmented': '#ffe027',
-    'continuous': '#9f043a',
+    'hover': '#9f043a',
+    'fragmented': '#ff6944',
+    'continuous': '#521b65',
     '': '#c7c7c7',
 }
 

### 2019-07-17 scripts/generate_Figure2.py
Add script for generating figure

@@ -1,0 +1,54 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+
+from src.figure_utilities import (GOLDEN_RATIO, TWO_COLUMN, save_figure,
+                                  set_figure_defaults)
+from src.sorted_spikes_simulation import make_simulated_run_data
+
+
+def generate_figure():
+    set_figure_defaults()
+
+    (time, linear_distance, sampling_frequency,
+     spikes, place_fields) = make_simulated_run_data()
+    spike_ind, neuron_ind = np.nonzero(spikes)
+    cmap = plt.get_cmap('tab20')
+
+    fig, axes = plt.subplots(3, 1, figsize=(
+        TWO_COLUMN, TWO_COLUMN * GOLDEN_RATIO), constrained_layout=True)
+    for place_field, color in zip(place_fields.T, cmap.colors):
+        axes[0].plot(linear_distance, place_field, linewidth=3, color=color)
+    axes[0].set_xlabel('Position [cm]')
+    axes[0].set_ylabel('Firing Rate\n[spikes / s]')
+    axes[0].set_title('Simulated Neuron Place Fields')
+    axes[0].set_xlim((linear_distance.min(), linear_distance.max()))
+    axes[0].set_yticks([0, np.round(place_fields.max())])
+    axes[0].text(-0.16, 1.1, 'a', transform=axes[0].transAxes,
+                 size=20, weight='bold')
+
+    axes[1].plot(time, linear_distance, linewidth=3)
+    axes[1].set_ylabel('Position [cm]')
+    axes[1].set_title('Simulated Position and Spikes')
+    axes[1].set_yticks([0, np.round(linear_distance.max())])
+    axes[1].set_xticks([])
+    axes[1].set_xlim((0.0, 90.0))
+    axes[1].text(-0.16, 1.1, 'b', transform=axes[1].transAxes,
+                 size=20, weight='bold')
+
+    c = [cmap.colors[ind] for ind in neuron_ind]
+    axes[2].scatter(time[spike_ind], neuron_ind + 1, c=c, s=0.5)
+    axes[2].set_yticks((0, spikes.shape[1]))
+    axes[2].set_ylabel('Neuron Index')
+
+    axes[2].set_xlabel('Time [s]')
+    axes[2].set_xlim((0.0, 90.0))
+
+    sns.despine()
+
+    save_figure('Figure2', figure_format='pdf')
[…]

### 2019-07-17 src/figure_utilities.py

@@ -1,0 +1,34 @@
+import os
+
+import matplotlib.pyplot as plt
+import seaborn as sns
+from numpy import sqrt
+
+from src.parameters import FIGURE_DIR
+
+# Figure Parameters
+MM_TO_INCHES = 1.0 / 25.4
+ONE_COLUMN = 89.0 * MM_TO_INCHES
+TWO_COLUMN = 178.0 * MM_TO_INCHES
+GOLDEN_RATIO = (sqrt(5) - 1.0) / 2.0
+
+
+def set_figure_defaults():
+    # Set background and fontsize
+    rc_params = {
+        'pdf.fonttype': 42,  # Make fonts editable in Adobe Illustrator
+        'ps.fonttype': 42,  # Make fonts editable in Adobe Illustrator
+        'axes.labelcolor': '#222222',
+        'text.color': '#222222',
+        'font.sans-serif': 'Helvetica',
+        'text.usetex': False,
+        'figure.figsize': (TWO_COLUMN, TWO_COLUMN * GOLDEN_RATIO),
+    }
+    sns.set(style='white', context='paper', rc=rc_params,
+            font_scale=1.3)
+
+
+def save_figure(figure_name, figure_format='pdf'):
+    figure_name = os.path.join(FIGURE_DIR, figure_name)
+    plt.savefig(f'{figure_name}.{figure_format}', transparent=True,
+                dpi=300, bbox_inches='tight')

### 2019-07-17 scripts/generate_Figure2.py
No offset of letter

@@ -24,7 +24,7 @@ def generate_figure():
     axes[0].set_title('Simulated Neuron Place Fields')
     axes[0].set_xlim((linear_distance.min(), linear_distance.max()))
     axes[0].set_yticks([0, np.round(place_fields.max())])
-    axes[0].text(-0.16, 1.1, 'a', transform=axes[0].transAxes,
+    axes[0].text(-0.16, 1.0, 'a', transform=axes[0].transAxes,
                  size=20, weight='bold')
 
     axes[1].plot(time, linear_distance, linewidth=3)
@@ -33,7 +33,7 @@ def generate_figure():
     axes[1].set_yticks([0, np.round(linear_distance.max())])
     axes[1].set_xticks([])
     axes[1].set_xlim((0.0, 90.0))
-    axes[1].text(-0.16, 1.1, 'b', transform=axes[1].transAxes,
+    axes[1].text(-0.16, 1.0, 'b', transform=axes[1].transAxes,
                  size=20, weight='bold')
 
     c = [cmap.colors[ind] for ind in neuron_ind]

### 2019-07-17 Article.tex
Update on Overleaf.

@@
 \usepackage{amsmath}
+\pdfoptionalwaysusepdfpagebox=5
 
@@
 \title{Categorizing the trajectory dynamics of replay}
-\shorttitle{My Template}
+\shorttitle{Replay Dynamics}
 
@@
 Replay events are typically evaluated based on whether they activate sequences of place cells that represent spatially continuous trajectories through the environment, but recent work has shown that these events can have more complex dynamics.
-For example, sequences can alternate between hovering on a particular spatial location and continuous movement or can represent continuous trajectories in other spatial environments, which may appear spatially incoherent in the context of the current environment.
+For example, sequences can alternate between hovering on a particular spatial location and continuous movement or can represent continuous trajectories in other spatial environments, which may appear spatially incoherent in the context of the current environment.\newline
 
-To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics.
+To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent state to decompose place cell sequences into categories based on their latent dynamics.
 Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous.
@@
 \centering
-\includegraphics[width=.8\linewidth]{figures/Figure_1}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\includegraphics[width=1.0\linewidth]{figures/Figure2.pdf}
+\caption{Simulated data used to fit the encoding model.
+\textbf{a.} Place fields for each simulated neuron are evenly spaced Gaussians with a maximum firing rate of 15 spikes / s.
+\textbf{b.} Linear position is simulated as a sinusoid to mimic an animal moving back and forth on a linear track.
+Each neuron spikes according to its place field rate and Poisson distribution.}
 \label{fig:computerNo}

### 2019-07-17 scripts/generate_Figure3.py
Add figure3 and generation code

@@ -1,0 +1,120 @@
+import string
+import logging
+
+import matplotlib.gridspec as gridspec
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+from replay_trajectory_classification import SortedSpikesClassifier
+from replay_trajectory_classification.state_transition import \
+    estimate_movement_var
+
+from src.figure_utilities import TWO_COLUMN, save_figure, set_figure_defaults
+from src.parameters import STATE_COLORS, TRANSITION_TO_CATEGORY
+from src.sorted_spikes_simulation import (make_continuous_replay,
+                                          make_fragmented_continuous_fragmented_replay,
+                                          make_fragmented_hover_fragmented_replay,
+                                          make_fragmented_replay,
+                                          make_hover_continuous_hover_replay,
+                                          make_hover_replay,
+                                          make_simulated_run_data)
+
+replay_types = [
+    ('continuous', make_continuous_replay),
+    ('hover', make_hover_replay),
+    ('fragmented', make_fragmented_replay),
+    ('hover-cont.-hover', make_hover_continuous_hover_replay),
+    ('frag.-cont.-frag.', make_fragmented_continuous_fragmented_replay),
+    ('frag.-hover-frag.', make_fragmented_hover_fragmented_replay)
+]
+
+
+def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
+                        letter):
+    results = results.assign_coords(
+        state=lambda ds: ds.state.to_index()
+        .map(TRANSITION_TO_CATEGORY),
+        time=lambda ds: ds.time * 1000.0,
+    )
+    spike_time_ind, neuron_ind = np.nonzero(test_spikes)
+    replay_time = results.time.values
+
+    inner_grid = gridspec.GridSpecFromSubplotSpec(
+        nrows=3, ncols=1, subplot_spec=subplot_spec)
+
+    # Spikes
+    ax = plt.Subplot(fig, inner_grid[0])
+    ax.scatter(replay_time[spike_time_ind], neuron_ind, color='black',
+               zorder=1, marker='|', s=10, linewidth=1)
+    ax.set_yticks((0, test_spikes.shape[1]))
+    ax.set_xticks([])
+    ax.set_xlim((replay_time.min(), replay_time.max()))
+    ax.set_ylabel('Neuron Index')
[…]

### 2019-07-18 scripts/generate_Figure3.py
Add replay probabilities legend

@@ -1,5 +1,5 @@
-import string
 import logging
+import string
 
 import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
@@ -67,6 +67,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     ax.set_xlim((replay_time.min(), replay_time.max()))
     ax.set_ylim((-0.01, 1.05))
     fig.add_subplot(ax)
+    legend_handle, legend_labels = ax.get_legend_handles_labels()
 
     # Posterior
     ax = plt.Subplot(fig, inner_grid[2])
@@ -81,6 +82,8 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     fig.add_subplot(ax)
 
     sns.despine()
+
+    return legend_handle, legend_labels
 
 
 def generate_figure():
@@ -101,15 +104,20 @@ def generate_figure():
 
     # Make Figure
     logging.info('Making figure...')
-    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.1),
+    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.4),
                      constrained_layout=True)
-    outer_grid = fig.add_gridspec(nrows=2, ncols=3)
+    outer_grid = fig.add_gridspec(nrows=3, ncols=3, height_ratios=[10, 10, 1])
     for replay_ind, (replay_name, make_replay) in enumerate(replay_types):
         replay_time, test_spikes = make_replay()
         results = classifier.predict(test_spikes, time=replay_time)
         letter = string.ascii_lowercase[replay_ind]
-        plot_classification(test_spikes, results,
-                            outer_grid[replay_ind], fig, replay_name, letter)
+        legend_handle, legend_labels = plot_classification(
+            test_spikes, results, outer_grid[replay_ind], fig, replay_name,
+            letter)
+    legend_ax = fig.add_subplot(outer_grid[-1, :])
+    legend_ax.axis('off')
+    legend_ax.legend(legend_handle, legend_labels, loc='upper center',
+                     fancybox=False, shadow=False, ncol=3, frameon=False)
     sns.despine()
 
     save_figure('Figure3', figure_format='pdf')

### 2019-07-18 scripts/generate_Figure3.py
Add posterior colorbar

@@ -5,6 +5,7 @@ import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from replay_trajectory_classification import SortedSpikesClassifier
 from replay_trajectory_classification.state_transition import \
     estimate_movement_var
@@ -71,7 +72,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
 
     # Posterior
     ax = plt.Subplot(fig, inner_grid[2])
-    results.acausal_posterior.sum('state').plot(
+    quad_mesh = results.acausal_posterior.sum('state').plot(
         x='time', y='position', robust=True, vmin=0.0, ax=ax,
         add_colorbar=False)
     ax.set_ylabel('Position [cm]')
@@ -83,7 +84,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
 
     sns.despine()
 
-    return legend_handle, legend_labels
+    return legend_handle, legend_labels, quad_mesh
 
 
 def generate_figure():
@@ -111,13 +112,23 @@ def generate_figure():
         replay_time, test_spikes = make_replay()
         results = classifier.predict(test_spikes, time=replay_time)
         letter = string.ascii_lowercase[replay_ind]
-        legend_handle, legend_labels = plot_classification(
+        legend_handle, legend_labels, quad_mesh = plot_classification(
             test_spikes, results, outer_grid[replay_ind], fig, replay_name,
             letter)
     legend_ax = fig.add_subplot(outer_grid[-1, :])
     legend_ax.axis('off')
     legend_ax.legend(legend_handle, legend_labels, loc='upper center',
                      fancybox=False, shadow=False, ncol=3, frameon=False)
+
+    colorbar_ax = fig.add_subplot(outer_grid[-1, -1])
+    colorbar_ax.axis('off')
+    axins = inset_axes(colorbar_ax, width='50%', height='50%',
+                       bbox_transform=colorbar_ax.transAxes, borderpad=0,
+                       loc='lower center')
+    cbar = plt.colorbar(quad_mesh, cax=axins, orientation='horizontal',
+                        extend='max', label='posterior', ticks=[0.0, 0.035])
+    cbar.ax.set_xticklabels(['0', 'max'])
+    cbar.ax.tick_params(labelsize=10, width=0, length=0)
     sns.despine()
 
     save_figure('Figure3', figure_format='pdf')

### 2019-07-18 scripts/generate_Figure3.py
Fix positioning of colorbar

@@ -115,7 +115,7 @@ def generate_figure():
         legend_handle, legend_labels, quad_mesh = plot_classification(
             test_spikes, results, outer_grid[replay_ind], fig, replay_name,
             letter)
-    legend_ax = fig.add_subplot(outer_grid[-1, :])
+    legend_ax = fig.add_subplot(outer_grid[-1, :2])
     legend_ax.axis('off')
     legend_ax.legend(legend_handle, legend_labels, loc='upper center',
                      fancybox=False, shadow=False, ncol=3, frameon=False)

### 2019-07-18 scripts/generate_Figure3.py
Rasterize pcolormesh

@@ -74,7 +74,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     ax = plt.Subplot(fig, inner_grid[2])
     quad_mesh = results.acausal_posterior.sum('state').plot(
         x='time', y='position', robust=True, vmin=0.0, ax=ax,
-        add_colorbar=False)
+        add_colorbar=False, rasterized=True)
     ax.set_ylabel('Position [cm]')
     ax.set_xlim((replay_time.min(), replay_time.max()))
     ax.set_xticks((replay_time.min(), replay_time.max()))

### 2019-07-18 scripts/generate_Figure3.py
Exit after running script

@@ -1,3 +1,4 @@
+import sys
 import logging
 import string
 
@@ -136,4 +137,4 @@ def generate_figure():
 
 
 if __name__ == '__main__':
-    generate_figure()
+    sys.exit(generate_figure())

### 2019-07-18 scripts/generate_Figure3.py
Close distributed client to cleanup

@@ -6,6 +6,7 @@ import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
+from dask.distributed import Client
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from replay_trajectory_classification import SortedSpikesClassifier
 from replay_trajectory_classification.state_transition import \
@@ -90,6 +91,8 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
 
 def generate_figure():
     logging.basicConfig(level=logging.INFO)
+    client = Client()
+    logging.info(client)
     set_figure_defaults()
 
     logging.info('Simulating data...')
@@ -134,6 +137,7 @@ def generate_figure():
 
     save_figure('Figure3', figure_format='pdf')
     save_figure('Figure3', figure_format='png')
+    client.close()
 
 
 if __name__ == '__main__':

### 2019-07-18 scripts/generate_Figure2.py
Update on Overleaf.

@@ -1,3 +1,5 @@
+import sys
+
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
@@ -51,4 +53,4 @@ def generate_figure():
 
 
 if __name__ == '__main__':
-    generate_figure()
+    sys.exit(generate_figure())

### 2019-07-18 scripts/generate_Figure3.py
Try to center colorbar and legend

@@ -1,6 +1,6 @@
-import sys
 import logging
 import string
+import sys
 
 import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
@@ -109,9 +109,9 @@ def generate_figure():
 
     # Make Figure
     logging.info('Making figure...')
-    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.4),
+    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.1),
                      constrained_layout=True)
-    outer_grid = fig.add_gridspec(nrows=3, ncols=3, height_ratios=[10, 10, 1])
+    outer_grid = fig.add_gridspec(nrows=3, ncols=3, height_ratios=[1, 1, 0.13])
     for replay_ind, (replay_name, make_replay) in enumerate(replay_types):
         replay_time, test_spikes = make_replay()
         results = classifier.predict(test_spikes, time=replay_time)
@@ -122,13 +122,14 @@ def generate_figure():
     legend_ax = fig.add_subplot(outer_grid[-1, :2])
     legend_ax.axis('off')
     legend_ax.legend(legend_handle, legend_labels, loc='upper center',
-                     fancybox=False, shadow=False, ncol=3, frameon=False)
+                     fancybox=False, shadow=False, ncol=3, frameon=False,
+                     bbox_to_anchor=(0.1, 0, 1, 1))
 
     colorbar_ax = fig.add_subplot(outer_grid[-1, -1])
     colorbar_ax.axis('off')
     axins = inset_axes(colorbar_ax, width='50%', height='50%',
                        bbox_transform=colorbar_ax.transAxes, borderpad=0,
-                       loc='lower center')
+                       loc='center', bbox_to_anchor=(0.1, 0, 1, 1))
     cbar = plt.colorbar(quad_mesh, cax=axins, orientation='horizontal',
                         extend='max', label='posterior', ticks=[0.0, 0.035])
     cbar.ax.set_xticklabels(['0', 'max'])

### 2019-07-18 scripts/generate_Figure3.py
Revert "Try to center colorbar and legend"
This reverts commit 3bf124ad0a49d9503e3d3591a7a5a88a954a2d6d.

@@ -1,6 +1,6 @@
+import sys
 import logging
 import string
-import sys
 
 import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
@@ -109,9 +109,9 @@ def generate_figure():
 
     # Make Figure
     logging.info('Making figure...')
-    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.1),
+    fig = plt.figure(figsize=(TWO_COLUMN, TWO_COLUMN * 1.4),
                      constrained_layout=True)
-    outer_grid = fig.add_gridspec(nrows=3, ncols=3, height_ratios=[1, 1, 0.13])
+    outer_grid = fig.add_gridspec(nrows=3, ncols=3, height_ratios=[10, 10, 1])
     for replay_ind, (replay_name, make_replay) in enumerate(replay_types):
         replay_time, test_spikes = make_replay()
         results = classifier.predict(test_spikes, time=replay_time)
@@ -122,14 +122,13 @@ def generate_figure():
     legend_ax = fig.add_subplot(outer_grid[-1, :2])
     legend_ax.axis('off')
     legend_ax.legend(legend_handle, legend_labels, loc='upper center',
-                     fancybox=False, shadow=False, ncol=3, frameon=False,
-                     bbox_to_anchor=(0.1, 0, 1, 1))
+                     fancybox=False, shadow=False, ncol=3, frameon=False)
 
     colorbar_ax = fig.add_subplot(outer_grid[-1, -1])
     colorbar_ax.axis('off')
     axins = inset_axes(colorbar_ax, width='50%', height='50%',
                        bbox_transform=colorbar_ax.transAxes, borderpad=0,
-                       loc='center', bbox_to_anchor=(0.1, 0, 1, 1))
+                       loc='lower center')
     cbar = plt.colorbar(quad_mesh, cax=axins, orientation='horizontal',
                         extend='max', label='posterior', ticks=[0.0, 0.035])
     cbar.ax.set_xticklabels(['0', 'max'])

### 2019-07-18 src/figure_utilities.py
Make labels and tick sizes smaller

@@ -9,7 +9,9 @@ from src.parameters import FIGURE_DIR
 # Figure Parameters
 MM_TO_INCHES = 1.0 / 25.4
 ONE_COLUMN = 89.0 * MM_TO_INCHES
+ONE_AND_HALF_COLUMN = 140.0 * MM_TO_INCHES
 TWO_COLUMN = 178.0 * MM_TO_INCHES
+PAGE_HEIGHT = 247.0 * MM_TO_INCHES
 GOLDEN_RATIO = (sqrt(5) - 1.0) / 2.0
 
 
@@ -19,10 +21,14 @@ def set_figure_defaults():
         'pdf.fonttype': 42,  # Make fonts editable in Adobe Illustrator
         'ps.fonttype': 42,  # Make fonts editable in Adobe Illustrator
         'axes.labelcolor': '#222222',
+        'axes.labelsize': 9,
         'text.color': '#222222',
         'font.sans-serif': 'Helvetica',
         'text.usetex': False,
-        'figure.figsize': (TWO_COLUMN, TWO_COLUMN * GOLDEN_RATIO),
+        'figure.figsize': (7.2, 4.45),
+        'xtick.major.size': 0.05,
+        'ytick.major.size': 0.05,
+        'axes.labelpad': 0.1,
     }
     sns.set(style='white', context='paper', rc=rc_params,
             font_scale=1.3)

### 2019-07-18 scripts/generate_Figure2.py
Tighten up figure

@@ -26,8 +26,8 @@ def generate_figure():
     axes[0].set_title('Simulated Neuron Place Fields')
     axes[0].set_xlim((linear_distance.min(), linear_distance.max()))
     axes[0].set_yticks([0, np.round(place_fields.max())])
-    axes[0].text(-0.16, 1.0, 'a', transform=axes[0].transAxes,
-                 size=20, weight='bold')
+    axes[0].text(-0.1, 1.0, 'a', transform=axes[0].transAxes,
+                 size=15, weight='bold')
 
     axes[1].plot(time, linear_distance, linewidth=3)
     axes[1].set_ylabel('Position [cm]')
@@ -35,13 +35,13 @@ def generate_figure():
     axes[1].set_yticks([0, np.round(linear_distance.max())])
     axes[1].set_xticks([])
     axes[1].set_xlim((0.0, 90.0))
-    axes[1].text(-0.16, 1.0, 'b', transform=axes[1].transAxes,
-                 size=20, weight='bold')
+    axes[1].text(-0.1, 1.0, 'b', transform=axes[1].transAxes,
+                 size=15, weight='bold')
 
     c = [cmap.colors[ind] for ind in neuron_ind]
     axes[2].scatter(time[spike_ind], neuron_ind + 1, c=c, s=0.5)
-    axes[2].set_yticks((0, spikes.shape[1]))
-    axes[2].set_ylabel('Neuron Index')
+    axes[2].set_yticks((1, spikes.shape[1]))
+    axes[2].set_ylabel('Cells')
 
     axes[2].set_xlabel('Time [s]')
     axes[2].set_xlim((0.0, 90.0))

### 2019-07-18 scripts/generate_Figure3.py
Adjust spacing and make figure bigger

@@ -1,6 +1,6 @@
-import sys
 import logging
 import string
+import sys
 
 import matplotlib.gridspec as gridspec
 import matplotlib.pyplot as plt
@@ -12,7 +12,8 @@ from replay_trajectory_classification import SortedSpikesClassifier
 from replay_trajectory_classification.state_transition import \
     estimate_movement_var
 
-from src.figure_utilities import TWO_COLUMN, save_figure, set_figure_defaults
+from src.figure_utilities import (PAGE_HEIGHT, TWO_COLUMN, save_figure,
+                                  set_figure_defaults)
 from src.parameters import STATE_COLORS, TRANSITION_TO_CATEGORY
 from src.sorted_spikes_simulation import (make_continuous_replay,
                                           make_fragmented_continuous_fragmented_replay,
@@ -43,18 +44,18 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     replay_time = results.time.values
 
     inner_grid = gridspec.GridSpecFromSubplotSpec(
-        nrows=3, ncols=1, subplot_spec=subplot_spec)
+        nrows=3, ncols=1, subplot_spec=subplot_spec, hspace=0.3)
 
     # Spikes
     ax = plt.Subplot(fig, inner_grid[0])
     ax.scatter(replay_time[spike_time_ind], neuron_ind, color='black',
                zorder=1, marker='|', s=10, linewidth=1)
-    ax.set_yticks((0, test_spikes.shape[1]))
+    ax.set_yticks((1, test_spikes.shape[1]))
     ax.set_xticks([])
     ax.set_xlim((replay_time.min(), replay_time.max()))
-    ax.set_ylabel('Neuron Index')
-    ax.text(-0.6, 1.0, letter, transform=ax.transAxes,
-            size=20, weight='bold')
+    ax.set_ylabel('Cells')
+    ax.text(-0.3, 1.0, letter, transform=ax.transAxes,
+            size=15, weight='bold')
     ax.set_title(replay_name)
     fig.add_subplot(ax)
 
@@ -62,8 +63,8 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     ax = plt.Subplot(fig, inner_grid[1])
     replay_probability = results.acausal_posterior.sum('position')
     for state, prob in replay_probability.groupby('state'):
-        ax.plot(prob.time, prob.values, linewidth=2,
[…]

### 2019-07-19 scripts/generate_Figure2.py
Update place fields name

@@ -23,7 +23,7 @@ def generate_figure():
         axes[0].plot(linear_distance, place_field, linewidth=3, color=color)
     axes[0].set_xlabel('Position [cm]')
     axes[0].set_ylabel('Firing Rate\n[spikes / s]')
-    axes[0].set_title('Simulated Neuron Place Fields')
+    axes[0].set_title('Simulated Place Fields')
     axes[0].set_xlim((linear_distance.min(), linear_distance.max()))
     axes[0].set_yticks([0, np.round(place_fields.max())])
     axes[0].text(-0.1, 1.0, 'a', transform=axes[0].transAxes,

### 2019-07-19 scripts/generate_Figure3.py
Use threads to avoid conflicts with other analyses

@@ -92,7 +92,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
 
 def generate_figure():
     logging.basicConfig(level=logging.INFO)
-    client = Client()
+    client = Client(processes=False)
     logging.info(client)
     set_figure_defaults()
 

### 2019-07-19 scripts/generate_Figure2.py
Make panel letters extra bold

@@ -27,7 +27,7 @@ def generate_figure():
     axes[0].set_xlim((linear_distance.min(), linear_distance.max()))
     axes[0].set_yticks([0, np.round(place_fields.max())])
     axes[0].text(-0.1, 1.0, 'a', transform=axes[0].transAxes,
-                 size=15, weight='bold')
+                 size=15, weight='extra bold')
 
     axes[1].plot(time, linear_distance, linewidth=3)
     axes[1].set_ylabel('Position [cm]')
@@ -36,7 +36,7 @@ def generate_figure():
     axes[1].set_xticks([])
     axes[1].set_xlim((0.0, 90.0))
     axes[1].text(-0.1, 1.0, 'b', transform=axes[1].transAxes,
-                 size=15, weight='bold')
+                 size=15, weight='extra bold')
 
     c = [cmap.colors[ind] for ind in neuron_ind]
     axes[2].scatter(time[spike_ind], neuron_ind + 1, c=c, s=0.5)

### 2019-07-19 scripts/generate_Figure3.py

@@ -55,7 +55,7 @@ def plot_classification(test_spikes, results, subplot_spec, fig, replay_name,
     ax.set_xlim((replay_time.min(), replay_time.max()))
     ax.set_ylabel('Cells')
     ax.text(-0.3, 1.0, letter, transform=ax.transAxes,
-            size=15, weight='bold')
+            size=15, weight='extra bold')
     ax.set_title(replay_name)
     fig.add_subplot(ax)
 

### 2019-07-19 src/figure_utilities.py
Use default font

@@ -23,7 +23,6 @@ def set_figure_defaults():
         'axes.labelcolor': '#222222',
         'axes.labelsize': 9,
         'text.color': '#222222',
-        'font.sans-serif': 'Helvetica',
         'text.usetex': False,
         'figure.figsize': (7.2, 4.45),
         'xtick.major.size': 0.05,

### 2019-07-19 src/figure_utilities.py
Make tick labels smaller

@@ -22,6 +22,8 @@ def set_figure_defaults():
         'ps.fonttype': 42,  # Make fonts editable in Adobe Illustrator
         'axes.labelcolor': '#222222',
         'axes.labelsize': 9,
+        'xtick.labelsize': 9,
+        'ytick.labelsize': 9,
         'text.color': '#222222',
         'text.usetex': False,
         'figure.figsize': (7.2, 4.45),

### 2019-07-19 Article.tex

--- 2019-07-16 Article.tex
@@
 \usepackage{blindtext}
-\usepackage{amsmath}
+\pdfoptionalwaysusepdfpagebox=5
+\DeclareMathOperator*{\E}{\mathbb{E}}
+\newcommand\norm[1]{\left\lVert#1\right\rVert}
 
@@
 
-To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics.
+To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent state to decompose place cell sequences into categories based on their latent dynamics.
 Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous.
@@
 \centering
-\includegraphics[width=.8\linewidth]{figures/Figure_1}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\includegraphics[width=1.0\linewidth]{figures/Figure2.pdf}
+\caption{Simulated data used to fit the encoding model.
+\textbf{a.} Place fields for each cell are simulated as a Gaussian with a maximum firing rate of 15 spikes / s.
+\textbf{b.} Position is simulated as a sinusoid to mimic an animal moving back and forth on a linear track.
+Each cell is simulated as a inhomogeneous Poisson process with a rate that depends on position according to its place field in a.}
 \label{fig:computerNo}
@@
 Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=1.0\linewidth]{figures/Figure3.pdf}
+\caption{Model can classify each category of trajectory dynamic (and combinations of dynamics) on simulated data.
+Each panel shows (from top to bottom) the simulated cell population activity, the probability of the three trajectory dynamics at each time, and the decoded posterior probability of position at each time.
+\textbf{a.} Continuous.
+\textbf{b.} Hover.
+\textbf{c.} Fragmented.
+\textbf{d.} Hover-Continuous-Hover.
+\textbf{e.} Fragmented-Continuous-Fragmented.
+\textbf{f.} Fragmented-Hover-Fragmented.}
+\label{fig:computerNo}
+\end{figure}
 
@@
 \section*{Materials and Methods}
+\subsection*{The Model}
+The 
+\subsection*{Encoding}
+In order to encode each cell's place preference, a generalized linear model was fit to each neuron's spiking activity during movement (defined as time periods when the running speed is greater than 4 cm/s).
+The model used log link function.
+
+$$\lambda(t) = \exp{(\sum_{i} f_{i}(position)\beta_{i})}$$
+
+A small L2 penalization term $-\lambda\norm{\beta_{i}}_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low.
+A third degree B-splines basis over position (or tensor products of the B-splines when position is two dimensional) is used to allow firing activity to smoothly vary over position and to reduce the total number of model parameters.
+\subsection*{Decoding}
+
+\subsection*{Simulated Data}
+
+\subsection*{Data}
+
+\subsection*{Software and Code availability}
+Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
+Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
+All code is open-source and licensed under the MIT Software License.
 

### 2020-03-03 src/analysis.py
Fix docstring.

@@ -373,7 +373,6 @@ def calculate_replay_distance(track_graph, map_estimate, actual_positions,
     -------
     replay_distance_from_actual_position : ndarray, shape (n_time,)
     replay_distance_from_center_well : ndarray, shape (n_time,)
-    replay_linear_position : ndarray, shape (n_time,)
 
     '''
 

### 2020-03-04 Article.tex
Update on Overleaf.

@@
 \DeclareMathOperator*{\E}{\mathbb{E}}
-\newcommand\norm[1]{\left\lVert#1\right\rVert}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
 
@@
 
-\title{Categorizing the trajectory dynamics of replay}
+\title{A state space model for discovering the trajectory dynamics of replay}
 \shorttitle{Replay Dynamics}
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[1,\Letter]{Eric L. Denovellis}
+\author[4]{Eric L. Denovellis}
 \author[2, 3]{Anna K. Gillespie}
 \author[2, 3]{Michael E. Coulter}
-\author[2, 3, 4]{Loren M. Frank}
 \author[1]{Uri T. Eden}
+\author[2, 3, 4, \Letter]{Loren M. Frank}
 
@@
 \begin{corrauthor}
-%\texttt{edeno{@}bu.edu}
-edeno\at bu.edu
+%\texttt{loren{@}phy.ucsf.edu}
+loren\at phy.ucsf.edu
 \end{corrauthor}
@@
 \section*{Introduction}
-\Blindtext
+The hippocampus is a brain area that is important for short term learning and memory.
+Neurons in the hippocampus preferentially fire in response to specific locations in an environment, which results in sequences of cells firing as an animal moves through an environment that reflect the trajectory of the animal.
+When the animal is asleep or immobile, cells also fire in sequence, forming trajectory-like representations that can recapitulate sequences previously observed when the animal was running.
+These sequences are thought to be important for memory consolidation or prospective decision making.
+(cite some examples).
+Therefore, characterizing the dynamics of hippocampal sequences is important for understanding memory consolidation and prospective decision making.
+
+Current methods analyze hippocampal sequences by assuming a constant velocity trajectory (Davidson et al.
+2009), corresponding to a single dynamic.
+This is done by computing the probability of position at each time using the place field information from the population of neurons and then fitting a line to maximize the probability of position over time.
+If the probabilities near the best fit line are high relative to shuffled distribution, the replay is analyzed.
+Often, trajectories that travel less than a certain distance (Pfeiffer and Foster 2013, Pfeiffer and Foster 2015) are excluded.
+
+However, there is a growing evidence that trajectories that have more complex dynamics.
+Farooq and Dragoi (2019) found that juvenile rats .
+For example, Pfeiffer and Foster (2015) found that trajectories alternated between stationary hovering trajectories and continuous movement trajectories on a gamma timescale.
+
+Our mathematical model precisely specifies the dyna
 
@@
 \caption{Model can classify each category of trajectory dynamic (and combinations of dynamics) on simulated data.
-Each panel shows (from top to bottom) the simulated cell population activity, the probability of the three trajectory dynamics at each time, and the decoded posterior probability of position at each time.
+Each panel has three plots that show (from top to bottom) the simulated cell population activity, the probability of the three trajectory dynamics at each time, and the decoded posterior probability of position at each time.
 \textbf{a.} Continuous.
@@
 
-\section*{Bibliography}
+\section*{References}
 \bibliography{zHenriquesLab-Mendeley}
@@
 
-\section*{Materials and Methods}
+\section*{Methods}
 \subsection*{The Model}
 The
-\subsection*{Encoding}
-In order to encode each cell's place preference, a generalized linear model was fit to each neuron's spiking activity during movement (defined as time periods when the running speed is greater than 4 cm/s).
-The model used log link function.
 
-$$\lambda(t) = \exp{(\sum_{i} f_{i}(position)\beta_{i})}$$
+\subsection*{Encoding - Clusterless}
+In order to encode how each tetrode's unsorted spiking activity relates to position, we use a marked point process framework.
+For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features (the marks) and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
+In our case, the waveform features correspond to the max amplitude observed on each tetrode wire at the time of the multiunit spike.
+This array is used as the training samples in a kernel density estimator during the decoding step.
+
+\subsection*{Encoding - Sorted Spikes}
+In order to encode how each cell's spiking activity relates to position (the place field), we fit a generalized linear model (GLM) with a Poisson response distribution to each cell's spiking activity during movement (defined as time periods when the running speed is greater than 4 cm/s).
+We estimate the parameters $\beta$, which consist of $\beta_{0}$, the average firing rate over time, and $\beta_{i}$, weights for third degree B-splines basis functions $f_{i}$ over position (or tensor products of the B-splines when position is two dimensional).
+B-spline basis functions are used because place field firing activity is assumed to vary smoothly over position and this prior knowledge can be exploited to reduce the total number of model parameters needed.
+Each basis function is spaced every 5 cm over the range of the position and zero constrained so that the change encoded by the parameters is relative to the baseline firing rate.
+We use a log link function to convert the linear combination of parameters to an instantaneous firing rate over time $\lambda(t)$ to ensure the rate is always positive.
+
+$$log(\lambda(t)) = \beta_{0} + \sum_{i} f_{i}(position)\beta_{i}$$
 
 A small L2 penalization term $-\lambda\norm{\beta_{i}}_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low.
-A third degree B-splines basis over position (or tensor products of the B-splines when position is two dimensional) is used to allow firing activity to smoothly vary over position and to reduce the total number of model parameters.
-\subsection*{Decoding}
+We set this to 0.5 for all cells.
+Fitting is done by maximizing the penalized likelihood using a Newton-Raphson algorithm.
+
+\subsection*{Decoding - Clusterless}
 
@@
 All code is open-source and licensed under the MIT Software License.
+Classifier code can be easily installed as a python package with all requisite dependencies using pip or conda.
+See software repositories for specific details.
 
@@
 \\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:k}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k}, I_{k}, x_{k+1}, I_{k+1} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
 \\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
@@
 
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=1.0\linewidth]{figures/Supplemental_Figure3.pdf}
+\caption{Increasing the probability of transitioning within category.}
+\label{fig:Supplemental Figure 3}
+\end{figure*}
+
 %TC:endignore

### 2020-03-04 Article.tex

@@
 
-\subsection*{Filter Derivation}
-\begin{enumerate}
-\item Use Bayes theorem to define the posterior in terms of a data-derived likelihood and a prior ($ posterior \propto likelihood * prior $)
-$$
-p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k-1})
-$$
-
-\item Use Chapman-Kolmogorov to define prior in terms of previous time step's posterior and state transitions
-$$
-\begin{align*}
-p(x_{k}, I_{k} \mid y_{1:k-1}) &= \sum_{I_{k-1}} \int p(x_{k}, x_{k-1}, I_{k}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}, y_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-\\ &= \sum_{I_{k-1}} \int p(x_{k}, I_{k} \mid x_{k-1}, I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-\\ &= \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-\end{align*}$$
-
-where:
-$p(x_{k-1}, I_{k-1} \mid y_{1:k-1})$ is the previous posterior,
-$Pr(I_{k} \mid I_{k-1})$ is the discrete state transition, and $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ is the continuous state transition
-
-\item Final Filter
-$$
-p(x_{k}, I_{k} \mid y_{1:k}) \propto p(y_{k} \mid x_{k}, I_{k}) * \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) * Pr(I_{k} \mid I_{k-1}) * p(x_{k-1}, I_{k-1} \mid y_{1:k-1}) * dx_{k-1}
-$$
-\end{enumerate}
-
-\subsection*{Smoother Derivation}
-
-\begin{enumerate}
-\item
-$$
-\begin{align*}
-p(x_{k}, I_{k} \mid y_{1:T}) &= \sum_{I_{k+1}} \int p(x_{k}, x_{k+1}, I_{k}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:T}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\\ &= \sum_{I_{k+1}} \int p(x_{k}, I_{k} \mid x_{k+1}, I_{k+1}, y_{1:k}) * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
-\\ &= \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1} 
-\\ &= p(x_{k}, I_{k} \mid y_{1:k}) * \sum_{I_{k+1}} \int \frac{p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}, x_{k})}{p(x_{k+1}, I_{k+1} \mid y_{1:k})} * p(x_{k+1}, I_{k+1} \mid y_{1:T}) * dx_{k+1}
-\end{align*}$$
-
-\item
-$$
-\begin{align*}
-p(x_{k+1}, I_{k+1} \mid y_{1:k}) &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1}, x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
-\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}, y_{1:k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
-\\ &= \sum_{I_{k}} \int p(x_{k+1}, I_{k+1} \mid x_{k}, I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
-\\ &= \sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) * Pr(I_{k+1} \mid I_{k}) * p(x_{k}, I_{k} \mid y_{1:k}) * dx_{k}
-\end{align*}$$
-
-\end{enumerate}
-
 \newpage

### 2020-03-04 Article.tex
Update on Overleaf.

@@
 
-Just for kicks here's a citation \cite{Gustafsson2016}. And a reference to a supplement \cref{note:Note1}. And \nameref{note:Note1}.
+Just for kicks here's a citation \cite{DavidsonHippocampalReplayExtended2009}. And a reference to a supplement \cref{note:Note1}. And \nameref{note:Note1}.
 \Blindtext
 
-\begin{figure}%[tbhp]
+\begin{figure*}%[tbhp]
 \centering
-\includegraphics[width=1.0\linewidth]{figures/Figure2.pdf}
-\caption{Simulated data used to fit the encoding model.
-\textbf{a.} Place fields for each cell are simulated as a Gaussian with a maximum firing rate of 15 spikes / s.
-\textbf{b.} Position is simulated as a sinusoid to mimic an animal moving back and forth on a linear track.
-Each cell is simulated as a inhomogeneous Poisson process with a rate that depends on position according to its place field in a.}
+\includegraphics[width=.8\linewidth]{figures/Figure1/Figure1_v3.pdf}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
 \label{fig:computerNo}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}%[tbhp]
+\begin{figure*}%[tbhp]
 \centering
-\includegraphics[width=1.0\linewidth]{figures/Figure3.pdf}
+\includegraphics[width=1.0\linewidth]{figures/Figure2/Figure2_v3.pdf}
 \caption{Model can classify each category of trajectory dynamic (and combinations of dynamics) on simulated data.
@@
 \label{fig:computerNo}
-\end{figure}
+\end{figure*}
 
@@
 \section*{References}
-\bibliography{zHenriquesLab-Mendeley}
+\bibliography{refs}
 
@@
 \centering
-\includegraphics[width=1.0\linewidth]{figures/Supplemental_Figure3.pdf}
+\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v1_supplemental1.pdf}
 \caption{Increasing the probability of transitioning within category.}

### 2020-03-05 Article.tex
Update on Overleaf.

@@
-\documentclass[times, twoside, watermark]{zHenriquesLab-StyleBioRxiv}
+\documentclass[times, twoside]{zHenriquesLab-StyleBioRxiv}
 \usepackage{blindtext}
-\pdfoptionalwaysusepdfpagebox=5
+\usepackage[utf8]{inputenc}
+\usepackage{graphicx}
+
 \DeclareMathOperator*{\E}{\mathbb{E}}
@@
 
-Current methods analyze hippocampal sequences by assuming a constant velocity trajectory (Davidson et al.
-2009), corresponding to a single dynamic.
+Current methods analyze hippocampal sequences by assuming a constant velocity trajectory \cite{DavidsonHippocampalReplayExtended2009}, corresponding to a single dynamic.
 This is done by computing the probability of position at each time using the place field information from the population of neurons and then fitting a line to maximize the probability of position over time.
 If the probabilities near the best fit line are high relative to shuffled distribution, the replay is analyzed.
-Often, trajectories that travel less than a certain distance (Pfeiffer and Foster 2013, Pfeiffer and Foster 2015) are excluded.
+Often, trajectories that travel less than a certain distance \cite{PfeifferAutoassociativedynamicsgeneration2015} are excluded.
 
 However, there is a growing evidence that trajectories that have more complex dynamics.
-Farooq and Dragoi (2019) found that juvenile rats .
-For example, Pfeiffer and Foster (2015) found that trajectories alternated between stationary hovering trajectories and continuous movement trajectories on a gamma timescale.
+Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} found that juvenile rats .
+For example, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that trajectories alternated between stationary hovering trajectories and continuous movement trajectories on a gamma timescale.
 
@@
 
-Just for kicks here's a citation \cite{DavidsonHippocampalReplayExtended2009}. And a reference to a supplement \cref{note:Note1}. And \nameref{note:Note1}.
-\Blindtext
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3.pdf}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{Figure1}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=.8\linewidth]{figures/Figure1/Figure1_v3.pdf}
+\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v3}
 \caption{Placeholder image of Iris with a long example caption to show justification setting.}
-\label{fig:computerNo}
+\label{Figure2}
 \end{figure*}
-
-\Blindtext
-
-Figure \ref{fig:computerNo} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
 
@@
 \centering
-\includegraphics[width=1.0\linewidth]{figures/Figure2/Figure2_v3.pdf}
-\caption{Model can classify each category of trajectory dynamic (and combinations of dynamics) on simulated data.
-Each panel has three plots that show (from top to bottom) the simulated cell population activity, the probability of the three trajectory dynamics at each time, and the decoded posterior probability of position at each time.
-\textbf{a.} Continuous.
-\textbf{b.} Hover.
-\textbf{c.} Fragmented.
-\textbf{d.} Hover-Continuous-Hover.
-\textbf{e.} Fragmented-Continuous-Fragmented.
-\textbf{f.} Fragmented-Hover-Fragmented.}
-\label{fig:computerNo}
+\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v3}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{Figure3}
+\end{figure*}
+
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v2}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{Figure4}
 \end{figure*}
@@
 
-blablaba \ref{fig:computerNo} 
-\blindtext
-
 \subsection*{Blabla}
-\blindtext
-
-blablaba \ref{fig:computerNo}
-\blindtext
 
 \begin{acknowledgements}
-\blindtext
 \end{acknowledgements}
@@
 \centering
-\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v1_supplemental1.pdf}
-\caption{Increasing the probability of transitioning within category.}
-\label{fig:Supplemental Figure 3}
+\includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v1_supplemental1}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{fig:Figure1-Figure supplement 1}
+\end{figure*}
+
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v2-supplemental1}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{fig:Figure2-Figure supplement 1}
+\end{figure*}
+
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v2-supplemental2}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{fig:Figure2-Figure supplement 2}
+\end{figure*}
+
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v2_supplemental1}
+\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\label{fig:Figure3-Figure supplement 1}
 \end{figure*}

### 2020-03-05 src/load_data.py
Fix merge headers

@@ -86,53 +86,47 @@ def load_data(epoch_key, brain_areas=None):
     position_info = (
         get_interpolated_position_dataframe(
             epoch_key, ANIMALS, _time_function)
-<< << << < HEAD
-        .dropna(subset=['linear_distance', 'linear_speed']))
-
-
-== == == =
         .dropna(subset=['linear_position', 'speed']))
->> >>>> > analyze - all - immobility
 
-    time=position_info.index
+    time = position_info.index
 
-    tetrode_info=make_tetrode_dataframe(ANIMALS, epoch_key = epoch_key)
-    is_brain_areas=(
+    tetrode_info = make_tetrode_dataframe(ANIMALS, epoch_key=epoch_key)
+    is_brain_areas = (
         tetrode_info.area.astype(str).str.upper().isin(brain_areas))
-    tetrode_keys=tetrode_info.loc[is_brain_areas].index
-    lfps=get_LFPs(tetrode_keys, ANIMALS)
-    lfps=lfps.resample('2ms').mean().fillna(method = 'pad').reindex(time)
+    tetrode_keys = tetrode_info.loc[is_brain_areas].index
+    lfps = get_LFPs(tetrode_keys, ANIMALS)
+    lfps = lfps.resample('2ms').mean().fillna(method='pad').reindex(time)
 
     try:
-        neuron_info=make_neuron_dataframe(ANIMALS).xs(
-            epoch_key, drop_level = False)
-        neuron_info=neuron_info.loc[
+        neuron_info = make_neuron_dataframe(ANIMALS).xs(
+            epoch_key, drop_level=False)
+        neuron_info = neuron_info.loc[
             (neuron_info.numspikes > 100) &
             neuron_info.area.isin(brain_areas) &
             (neuron_info.type == 'principal')]
         spikes = get_all_spike_indicators(
             neuron_info.index, ANIMALS, _time_function).reindex(time)
     except KeyError:
-        spikes=None
+        spikes = None
 
-    tetrode_info=tetrode_info.loc[is_brain_areas]
-    multiunit=(get_all_multiunit_indicators(
+    tetrode_info = tetrode_info.loc[is_brain_areas]
+    multiunit = (get_all_multiunit_indicators(
         tetrode_info.index, ANIMALS, _time_function)
         .sel(features=_MARKS)
         .reindex({'time': time}))
-    multiunit_spikes=(np.any(~np.isnan(multiunit.values), axis=1)
[…]

### 2020-03-05 src/load_data.py
Fix so valid ripple appears

@@ -46,7 +46,8 @@ def get_ripple_times(epoch_key, sampling_frequency=1500,
         .dropna(subset=['linear_position', 'speed']))
     speed = position_info['speed']
     time = position_info.index
-    tetrode_info = make_tetrode_dataframe(ANIMALS, epoch_key=epoch_key)
+    tetrode_info = make_tetrode_dataframe(ANIMALS).xs(
+        epoch_key, drop_level=False)
     if ~np.all(np.isnan(tetrode_info.validripple.astype(float))):
         tetrode_keys = tetrode_info.loc[
             (tetrode_info.validripple == 1)].index

### 2020-03-05 src/figure2.py
Figure updates from lab meeting

@@ -58,12 +58,12 @@ def plot_1D_results(spike_times, data, results_1D,
          (ripple_end - ripple_start) / np.timedelta64(1, "s"))
     )
 
-    axes[0].set_yticks((1, n_tetrodes))
+    axes[0].set_yticks((0, n_tetrodes))
 
     if data_type == "sorted_spikes":
         axes[0].set_ylabel("Cells")
     else:
-        axes[0].set_ylabel("Tetrodes")
+        axes[0].set_ylabel("Tet.")
 
     # axis 1
     probability = results_1D.acausal_posterior.sum(["position"])
@@ -74,11 +74,11 @@ def plot_1D_results(spike_times, data, results_1D,
             MILLISECONDS_TO_SECONDS *
             probability.time / np.timedelta64(1, "s"),
             prob,
-            linewidth=3,
+            linewidth=2,
             color=STATE_COLORS[state],
         )
 
-    axes[1].set_ylim((0, 1.01))
+    axes[1].set_ylim((0, 1.05))
     axes[1].set_yticks((0, 1))
     axes[1].set_ylabel("Prob.")
 
@@ -103,7 +103,7 @@ def plot_1D_results(spike_times, data, results_1D,
     cmap = copy.copy(plt.cm.get_cmap(cmap))
     cmap.set_bad(color="lightgrey", alpha=1.0)
 
-    posterior_mesh = (
+    (
         results_1D.assign_coords(
             time=lambda ds: MILLISECONDS_TO_SECONDS *
             ds.time / np.timedelta64(1, "s")
@@ -145,8 +145,7 @@ def plot_1D_results(spike_times, data, results_1D,
 
     plot_1D_wtrack_landmarks(data, max_time, ax=axes[2])
     axes[2].set_ylabel("Position [cm]")
-    
-    plt.suptitle(data_type.replace("_", " ").title(), y=1.05)
+
     sns.despine()
 
     fig_name = (
@@ -155,8 +154,6 @@ def plot_1D_results(spike_times, data, results_1D,
         f"{data_type}_1D_acasual_classification"
     )
     save_figure(fig_name)
-    
-    return posterior_mesh
 
 
 def plot_1D_projected_to_2D(epoch_key, ripple_number, data,

### 2020-03-05 src/figure2_supplemental1.py

@@ -1,16 +1,12 @@
-import copy
-
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 
 from src.analysis import (get_is_classified, get_probability,
                           maximum_a_posteriori_estimate)
-from src.figure_utilities import (ONE_COLUMN, PAGE_HEIGHT, TWO_COLUMN,
-                                  save_figure)
+from src.figure_utilities import PAGE_HEIGHT, TWO_COLUMN, save_figure
 from src.parameters import PROBABILITY_THRESHOLD, STATE_COLORS
-from src.visualization import (plot_1D_wtrack_landmarks,
-                               plot_2D_position_with_color_time)
+from src.visualization import plot_2D_position_with_color_time
 
 MILLISECONDS_TO_SECONDS = 1000
 
@@ -70,7 +66,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
             MILLISECONDS_TO_SECONDS *
             probability.time / np.timedelta64(1, "s"),
             prob,
-            linewidth=3,
+            linewidth=2,
             color=STATE_COLORS[state],
         )
 
@@ -84,7 +80,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
     )
     axes[1].set_xlabel("Time [ms]")
 
-    axes[1].set_ylim((0, 1.01))
+    axes[1].set_ylim((0, 1.05))
     axes[1].set_yticks((0, 1))
     axes[1].set_ylabel("Probability")
 
@@ -190,5 +186,5 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
         f"{data_type}_2D_acasual_classification"
     )
     save_figure(fig_name)
-    
+
     return cbar

### 2020-03-05 src/figure2_supplemental2.py

@@ -61,11 +61,11 @@ def plot_clusterless_1D_results(multiunit_times, data, results,
             MILLISECONDS_TO_SECONDS *
             probability.time / np.timedelta64(1, "s"),
             prob,
-            linewidth=3,
+            linewidth=2,
             color=STATE_COLORS[state],
         )
 
-    axes[1].set_ylim((0, 1.01))
+    axes[1].set_ylim((0, 1.05))
     axes[1].set_yticks((0, 1))
     axes[1].set_ylabel("Prob.")
     probability2 = get_probability(results)

### 2020-03-05 src/figure3.py

@@ -61,11 +61,11 @@ def plot_clusterless_1D_results(multiunit_times, data, results,
             MILLISECONDS_TO_SECONDS *
             probability.time / np.timedelta64(1, "s"),
             prob,
-            linewidth=3,
+            linewidth=2,
             color=STATE_COLORS[state],
         )
 
-    axes[1].set_ylim((0, 1.01))
+    axes[1].set_ylim((0, 1.05))
     axes[1].set_yticks((0, 1))
     axes[1].set_ylabel("Probability")
     probability2 = get_probability(results)

### 2020-03-05 src/parameters.py

@@ -1,7 +1,6 @@
 from os.path import abspath, dirname, join, pardir
 
 import numpy as np
-
 from loren_frank_data_processing import Animal
 from replay_trajectory_classification.misc import NumbaKDE
 
@@ -62,12 +61,19 @@ PROBABILITY_THRESHOLD = 0.8
 STATE_ORDER = ['Hover', 'Hover-Continuous-Mix', 'Continuous',
                'Fragmented-Continuous-Mix', 'Fragmented']
 
+SHORT_STATE_ORDER = ['Hover', 'Hover-Cont.-Mix', 'Cont.',
+                     'Frag.-Cont.-Mix', 'Frag.']
+
 # Plotting Colors
 STATE_COLORS = {
     'Hover': '#9f043a',
     'Fragmented': '#ff6944',
+    'Frag.': '#ff6944',
     'Continuous': '#521b65',
+    'Cont.': '#521b65',
     'Hover-Continuous-Mix': '#61c5e6',
+    'Hover-Cont.-Mix': '#61c5e6',
     'Fragmented-Continuous-Mix': '#2a586a',
+    'Frag.-Cont.-Mix': '#2a586a',
     '': '#c7c7c7',
 }

### 2020-03-05 src/visualization.py

@@ -1,23 +1,36 @@
+
+import os
+
 import matplotlib.animation as animation
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import seaborn as sns
-from matplotlib.collections import LineCollection
-from matplotlib.colorbar import ColorbarBase, make_axes
-
 from loren_frank_data_processing.position import (get_position_dataframe,
                                                   make_track_graph)
 from loren_frank_data_processing.track_segment_classification import (
     get_track_segments_from_graph, plot_track, project_points_to_segment)
+from matplotlib.collections import LineCollection
+from matplotlib.colorbar import ColorbarBase, make_axes
+
 from src.analysis import maximum_a_posteriori_estimate
-from src.parameters import STATE_COLORS, STATE_ORDER
+from src.parameters import (FIGURE_DIR, SAMPLING_FREQUENCY, SHORT_STATE_ORDER,
+                            STATE_COLORS, STATE_ORDER)
 
 try:
     from upsetplot import UpSet
 except ImportError:
     class Upset:
         pass
+
+
+SHORT_STATE_NAMES = {
+    "Hover": "Hover",
+    "Hover-Continuous-Mix": "Hover-Cont.-Mix",
+    "Continuous": "Cont.",
+    "Fragmented-Continuous-Mix": "Frag.-Cont.-Mix",
+    "Fragmented": "Frag."
+}
 
 
 def plot_2D_position_with_color_time(time, position, ax=None, cmap='plasma',
@@ -287,11 +300,21 @@ def plot_neuron_place_field_2D_1D_position(
 
 
 def plot_category_counts(replay_info):
-    upset = UpSet(replay_info.set_index(STATE_ORDER[::-1]),
+    df = (replay_info
+          .rename(columns=SHORT_STATE_NAMES)
+          .set_index(SHORT_STATE_ORDER[::-1]))
+    upset = UpSet(df,
                   sum_over=False, sort_sets_by=None, show_counts=False,
                   sort_by='cardinality', intersection_plot_elements=6,
                   element_size=32)
-    return upset.plot()
+    axes = upset.plot()
+    axes["intersections"].set_ylabel(
+        "Number of\nRipples per\nCombination",
+        rotation="horizontal",
+        ha="right",
+        va="center",
+    )
+    return axes
 
 
 def _plot_category(replay_info, category, kind='strip', ax=None,
[…]

### 2020-03-06 src/analysis.py
Add sleep analysis code

@@ -2,6 +2,7 @@ import networkx as nx
 import numpy as np
 import pandas as pd
 import xarray as xr
+
 from loren_frank_data_processing.track_segment_classification import (
     get_track_segments_from_graph, project_points_to_segment)
 
@@ -86,6 +87,59 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
     center_well_id = 0
     replay_info['max_linear_distance'] = list(
         classifier.distance_between_nodes_[center_well_id].values())[-1]
+
+    return replay_info
+
+
+def get_sleep_replay_info(results, ripple_spikes, ripple_times, position_info,
+                          sampling_frequency, probability_threshold,
+                          epoch_key):
+    '''
+
+    Parameters
+    ----------
+    results : xarray.Dataset, shape (n_ripples, n_position_bins, n_states,
+                                     n_ripple_time)
+    ripple_spikes : pandas.DataFrame (n_ripples * n_ripple_time, n_neurons)
+    ripple_times : pandas.DataFrame (n_ripples, 2)
+    position_info : pandas.DataFrame (n_time, n_covariates)
+    sampling_frequency : float
+    probability_threshold : float
+    epoch_key : tuple
+
+    Returns
+    -------
+    replay_info : pandas.DataFrame, shape (n_ripples, n_covariates)
+
+    '''
+    probability = get_probability(results)
+    is_classified = get_is_classified(probability, probability_threshold)
+
+    duration = (is_classified.sum('time') / sampling_frequency)
+    duration = duration.to_dataframe().unstack(level=1)
+    duration.columns = list(duration.columns.get_level_values('state'))
+    duration = duration.rename(
+        columns=lambda column_name: column_name + '_duration')
+    is_category = (duration > 0.0).rename(columns=lambda c: c.split('_')[0])
+    duration = pd.concat((duration, is_category), axis=1)
+    duration['is_classified'] = np.any(duration > 0.0, axis=1)
+    duration['n_unique_spiking'] = get_n_unique_spiking(ripple_spikes)
+    duration['n_total_spikes'] = get_n_total_spikes(ripple_spikes)
+
+    ripple_position_info = reshape_to_segments(position_info, ripple_times)
+    duration['actual_x_position'] = ripple_position_info.groupby(
+        'ripple_number').x_position.mean()
+    duration['actual_y_position'] = ripple_position_info.groupby(
+        'ripple_number').y_position.mean()
[…]

### 2020-03-06 src/load_data.py

@@ -1,3 +1,4 @@
+import itertools
 from logging import getLogger
 
 import numpy as np
@@ -7,7 +8,8 @@ from scipy.signal import filtfilt
 from loren_frank_data_processing import (get_all_multiunit_indicators,
                                          get_all_spike_indicators,
                                          get_interpolated_position_dataframe,
-                                         get_LFPs, get_trial_time,
+                                         get_LFPs, get_position_dataframe,
+                                         get_trial_time, make_epochs_dataframe,
                                          make_neuron_dataframe,
                                          make_tetrode_dataframe)
 from ripple_detection import (Kay_ripple_detector,
@@ -141,3 +143,168 @@ def load_data(epoch_key, brain_areas=None):
         'multiunit_firing_rate': multiunit_firing_rate,
         'sampling_frequency': SAMPLING_FREQUENCY,
     }
+
+
+def get_sleep_interpolated_position_info(
+    epoch_key, animals, time_function=get_trial_time
+):
+    time = time_function(epoch_key, ANIMALS)
+    position_df = get_position_dataframe(
+        epoch_key, animals, skip_linearization=True)
+    new_index = pd.Index(
+        np.unique(np.concatenate((position_df.index, time))), name="time"
+    )
+    position_df = (
+        position_df.reindex(index=new_index)
+        .interpolate(method="linear")
+        .reindex(index=time)
+    )
+
+    position_df.loc[position_df.speed < 0, "speed"] = 0.0
+
+    return position_df
+
+
+def get_sleep_ripple_times(epoch_key, brain_areas=["CA1", "CA2", "CA3"]):
+    position_info = (get_sleep_interpolated_position_info(epoch_key, ANIMALS)
+                     .dropna(subset=["speed"]))
+    speed = position_info["speed"]
+    time = position_info.index
+    tetrode_info = make_tetrode_dataframe(
+        ANIMALS).xs(epoch_key, drop_level=False)
+    if ~np.all(np.isnan(tetrode_info.validripple.astype(float))):
+        tetrode_keys = tetrode_info.loc[(tetrode_info.validripple == 1)].index
+    else:
+        is_brain_areas = tetrode_info.area.astype(
+            str).str.upper().isin(brain_areas)
[…]

### 2020-03-06 scripts/run_by_sleep_epoch.py
Add sleep analysis run script

@@ -1,0 +1,348 @@
+import logging
+import os
+import sys
+from argparse import ArgumentParser
+from signal import SIGUSR1, SIGUSR2, signal
+from subprocess import PIPE, run
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import xarray as xr
+from scipy.ndimage import label
+from tqdm.auto import tqdm
+
+from loren_frank_data_processing import save_xarray
+from replay_trajectory_classification import (ClusterlessClassifier,
+                                              SortedSpikesClassifier)
+from src.analysis import (get_place_field_max, get_sleep_replay_info,
+                          reshape_to_segments)
+from src.load_sleep_data import get_sleep_and_prev_run_epochs, load_sleep_data
+from src.parameters import (FIGURE_DIR, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR, SAMPLING_FREQUENCY,
+                            TRANSITION_TO_CATEGORY)
+from src.visualization import (plot_category_counts, plot_category_duration,
+                               plot_ripple_decode_1D)
+
+FORMAT = '%(asctime)s %(message)s'
+
+logging.basicConfig(level='INFO', format=FORMAT, datefmt='%d-%b-%y %H:%M:%S')
+plt.switch_backend('agg')
+
+
+def sorted_spikes_analysis_1D(sleep_epoch_key, prev_run_epoch_key,
+                              plot_ripple_figures=False):
+    data_type, dim = 'sorted_spikes', '1D'
+
+    logging.info('Loading data...')
+    data = load_sleep_data(sleep_epoch_key)
+
+    model_name = os.path.join(
+        PROCESSED_DATA_DIR,
+        (f'{prev_run_epoch_key[0]}_{prev_run_epoch_key[1]:02}_'
+         f'{prev_run_epoch_key[2]:02}_{data_type}_{dim}_model.pkl'))
+    try:
+        results = xr.open_dataset(
+            os.path.join(
+                PROCESSED_DATA_DIR,
+                (f'{sleep_epoch_key[0]}_{sleep_epoch_key[1]:02}'
+                 f'_{sleep_epoch_key[2]:02}.nc')),
+            group=f'/{data_type}/{dim}/classifier/ripples/')
+        logging.info('Found existing results. Loading...')
+        ripple_times = data['ripple_times'].loc[:, ['start_time', 'end_time']]
[…]

### 2020-03-06 scripts/run_by_sleep_epoch.py
Fix name of import

@@ -9,15 +9,15 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import xarray as xr
-from scipy.ndimage import label
-from tqdm.auto import tqdm
-
 from loren_frank_data_processing import save_xarray
 from replay_trajectory_classification import (ClusterlessClassifier,
                                               SortedSpikesClassifier)
+from scipy.ndimage import label
+from tqdm.auto import tqdm
+
 from src.analysis import (get_place_field_max, get_sleep_replay_info,
                           reshape_to_segments)
-from src.load_sleep_data import get_sleep_and_prev_run_epochs, load_sleep_data
+from src.load_data import get_sleep_and_prev_run_epochs, load_sleep_data
 from src.parameters import (FIGURE_DIR, PROBABILITY_THRESHOLD,
                             PROCESSED_DATA_DIR, SAMPLING_FREQUENCY,
                             TRANSITION_TO_CATEGORY)

### 2020-03-06 src/visualization.py
Don't ignore NaNs

@@ -242,7 +242,7 @@ def plot_ripple_decode_1D(posterior, ripple_position, ripple_spikes,
     axes[1].legend(loc='center left', bbox_to_anchor=(1, 0.5),
                    fancybox=False, shadow=False, ncol=1, frameon=False)
 
-    posterior.sum('state').plot(
+    posterior.sum('state', skipna=False).plot(
         x='time', y='position', robust=True, vmin=0.0, ax=axes[2])
     axes[2].set_ylabel('Position [cm]')
     axes[2].set_xlim((min_time, max_time))

### 2020-03-06 src/visualization.py
Ignore if arm names aren't available in the position info.

@@ -249,19 +249,22 @@ def plot_ripple_decode_1D(posterior, ripple_position, ripple_spikes,
     axes[2].set_xticks((min_time, max_time))
     axes[-1].set_xlabel('Time [ms]')
 
-    max_df = position_info.groupby('arm_name').linear_position.max()
-    min_df = position_info.groupby('arm_name').linear_position.min()
-    axes[2].set_ylim((0, position_info.linear_position.max()))
-    axes[2].set_yticks((0, position_info.linear_position.max()))
+    try:
+        max_df = position_info.groupby('arm_name').linear_position.max()
+        min_df = position_info.groupby('arm_name').linear_position.min()
+        axes[2].set_ylim((0, position_info.linear_position.max()))
+        axes[2].set_yticks((0, position_info.linear_position.max()))
 
-    for arm_name, max_position in max_df.iteritems():
-        axes[2].axhline(max_position, color='lightgrey',
-                        linestyle='-', linewidth=1)
-        axes[2].text(min_time, max_position - 5, arm_name, color='white',
-                     fontsize=8, verticalalignment='top')
-    for arm_name, min_position in min_df.iteritems():
-        axes[2].axhline(min_position, color='lightgrey',
-                        linestyle='-', linewidth=1)
+        for arm_name, max_position in max_df.iteritems():
+            axes[2].axhline(max_position, color='lightgrey',
+                            linestyle='-', linewidth=1)
+            axes[2].text(min_time, max_position - 5, arm_name, color='white',
+                         fontsize=8, verticalalignment='top')
+        for arm_name, min_position in min_df.iteritems():
+            axes[2].axhline(min_position, color='lightgrey',
+                            linestyle='-', linewidth=1)
+    except KeyError:
+        pass
     axes[2].plot(ripple_time, ripple_position, color='white', linestyle='--',
                  linewidth=2, alpha=0.7)
 

### 2020-03-06 src/analysis.py
Need arm min max for plots.

@@ -2,7 +2,6 @@ import networkx as nx
 import numpy as np
 import pandas as pd
 import xarray as xr
-
 from loren_frank_data_processing.track_segment_classification import (
     get_track_segments_from_graph, project_points_to_segment)
 
@@ -93,7 +92,7 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
 
 def get_sleep_replay_info(results, ripple_spikes, ripple_times, position_info,
                           sampling_frequency, probability_threshold,
-                          epoch_key):
+                          epoch_key, classifier):
     '''
 
     Parameters
@@ -140,6 +139,24 @@ def get_sleep_replay_info(results, ripple_spikes, ripple_times, position_info,
     replay_info['animal'] = animal
     replay_info['day'] = int(day)
     replay_info['epoch'] = int(epoch)
+
+    min_max = (
+        classifier
+        ._nodes_df[classifier._nodes_df.is_bin_edge]
+        .groupby('edge_id')
+        .aggregate(['min', 'max']))
+
+    replay_info['center_well_position'] = min_max.loc[0].linear_position.min()
+    replay_info['choice_position'] = min_max.loc[0].linear_position.max()
+
+    replay_info['left_arm_start'] = min_max.loc[1].linear_position.min()
+    replay_info['left_well_position'] = min_max.loc[3].linear_position.max()
+
+    replay_info['right_arm_start'] = min_max.loc[2].linear_position.min()
+    replay_info['right_well_position'] = min_max.loc[4].linear_position.max()
+    center_well_id = 0
+    replay_info['max_linear_distance'] = list(
+        classifier.distance_between_nodes_[center_well_id].values())[-1]
 
     return replay_info
 

### 2020-03-06 scripts/run_by_epoch.py
Use classifier to label arms.

@@ -9,14 +9,14 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import xarray as xr
-from scipy.ndimage import label
-from tqdm.auto import tqdm
-
 from loren_frank_data_processing import save_xarray
 from loren_frank_data_processing.position import (EDGE_ORDER, EDGE_SPACING,
                                                   make_track_graph)
 from replay_trajectory_classification import (ClusterlessClassifier,
                                               SortedSpikesClassifier)
+from scipy.ndimage import label
+from tqdm.auto import tqdm
+
 from src.analysis import (get_linear_position_order, get_place_field_max,
                           get_replay_info, reshape_to_segments)
 from src.load_data import load_data
@@ -176,7 +176,7 @@ def sorted_spikes_analysis_1D(epoch_key, plot_ripple_figures=False):
                 plot_ripple_decode_1D(
                     posterior, ripple_position.loc[ripple_number],
                     ripple_spikes.loc[ripple_number], linear_position_order,
-                    data['position_info'])
+                    data['position_info'], classifier)
                 plt.suptitle(
                     f'ripple number = {animal}_{day:02d}_{epoch:02d}_'
                     f'{ripple_number:04d}')
@@ -453,7 +453,7 @@ def clusterless_analysis_1D(epoch_key, plot_ripple_figures=False):
                 plot_ripple_decode_1D(
                     posterior, ripple_position.loc[ripple_number],
                     ripple_spikes.loc[ripple_number], linear_position_order,
-                    data['position_info'], spike_label='Tetrodes')
+                    data['position_info'], classifier, spike_label='Tetrodes')
                 plt.suptitle(
                     f'ripple number = {animal}_{day:02d}_{epoch:02d}_'
                     f'{ripple_number:04d}')

### 2020-03-06 scripts/run_by_sleep_epoch.py

@@ -134,7 +134,7 @@ def sorted_spikes_analysis_1D(sleep_epoch_key, prev_run_epoch_key,
                 plot_ripple_decode_1D(
                     posterior, ripple_position,
                     ripple_spikes.loc[ripple_number], linear_position_order,
-                    data['position_info'])
+                    data['position_info'], classifier)
                 plt.suptitle(
                     f'ripple number = {sleep_epoch_key[0]}_'
                     f'{sleep_epoch_key[1]:02d}_{sleep_epoch_key[2]:02d}'
@@ -265,7 +265,7 @@ def clusterless_analysis_1D(sleep_epoch_key, prev_run_epoch_key,
                 plot_ripple_decode_1D(
                     posterior, ripple_position,
                     ripple_spikes.loc[ripple_number], linear_position_order,
-                    data['position_info'], spike_label='Tetrodes')
+                    data['position_info'], classifier, spike_label='Tetrodes')
                 plt.suptitle(
                     f'ripple number = {sleep_epoch_key[0]}_'
                     f'{sleep_epoch_key[1]:02d}_{sleep_epoch_key[2]:02d}'

### 2020-03-06 src/visualization.py

@@ -211,7 +211,7 @@ def plot_ripple_decode_2D(posterior, ripple_position,
 
 
 def plot_ripple_decode_1D(posterior, ripple_position, ripple_spikes,
-                          linear_position_order, position_info,
+                          linear_position_order, position_info, classifier,
                           spike_label='Cells', figsize=(10, 7)):
     ripple_spikes = np.asarray(ripple_spikes)
     spike_time_ind, neuron_ind = np.nonzero(
@@ -249,24 +249,28 @@ def plot_ripple_decode_1D(posterior, ripple_position, ripple_spikes,
     axes[2].set_xticks((min_time, max_time))
     axes[-1].set_xlabel('Time [ms]')
 
-    try:
-        max_df = position_info.groupby('arm_name').linear_position.max()
-        min_df = position_info.groupby('arm_name').linear_position.min()
-        axes[2].set_ylim((0, position_info.linear_position.max()))
-        axes[2].set_yticks((0, position_info.linear_position.max()))
+    min_max = (
+        classifier
+        ._nodes_df[classifier._nodes_df.is_bin_edge]
+        .groupby('edge_id')
+        .aggregate(['min', 'max']))
 
-        for arm_name, max_position in max_df.iteritems():
-            axes[2].axhline(max_position, color='lightgrey',
-                            linestyle='-', linewidth=1)
-            axes[2].text(min_time, max_position - 5, arm_name, color='white',
-                         fontsize=8, verticalalignment='top')
-        for arm_name, min_position in min_df.iteritems():
-            axes[2].axhline(min_position, color='lightgrey',
-                            linestyle='-', linewidth=1)
-    except KeyError:
-        pass
+    center_arm_position = min_max.loc[0].linear_position.max()
+    left_well_position = min_max.loc[3].linear_position.max()
+    right_well_position = min_max.loc[4].linear_position.max()
+
+    axes[2].set_ylim((0, np.floor(left_well_position)))
+    axes[2].set_yticks((0, np.floor(left_well_position)))
+
+    axes[2].text(min_time, left_well_position, "Left Arm", color='white',
+                 fontsize=8, verticalalignment='top')
+    axes[2].text(min_time, right_well_position, "Right Arm", color='white',
+                 fontsize=8, verticalalignment='top')
+    axes[2].text(min_time, center_arm_position, "Center Arm", color='white',
+                 fontsize=8, verticalalignment='top')
[…]

### 2020-03-06 src/visualization.py
Alter ratio of subplots

@@ -221,7 +221,8 @@ def plot_ripple_decode_1D(posterior, ripple_position, ripple_spikes,
 
     fig, axes = plt.subplots(
         3, 1, figsize=figsize,
-        constrained_layout=True, sharex=True)
+        constrained_layout=True, sharex=True,
+        gridspec_kw={"height_ratios": [1, 1, 3]},)
 
     axes[0].scatter(ripple_time[spike_time_ind], neuron_ind, color='black',
                     zorder=1, marker='|', s=20, linewidth=1)

### 2020-03-07 Article.tex
Update on Overleaf.

@@
 \section*{Results}
+\subsection*{Test}
 
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3.pdf}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3}
+\caption{Components of the state space model and demonstration on simulated data.
+\textbf{(A)} 19 simulated place cells with peaks evenly spaced on a 180 cm track that exhibit three different sequence patterns.
+For the first third of time, one cell fires repeatedly, representing one stationary location.
+For the second third of time, the cells fire in sequence, representing a smooth continuous trajectory through space.
+For the last third cells fire in an incoherent pattern, representing a non-smooth fragmented trajectory through space.
+\textbf{(B)}} Like the Bayesian Decoder, the state space encoding model uses estimates of the cell's place fields to compute a likelihood of position.
+\textbf{(C)}
 \label{Figure1}
@@
 \section*{Discussion}
-
-\subsection*{Blabla}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\beginsupplement
 \captionsetup*{format=largeformat}
-\section{Something about something} \label{note:Note1} 
-\Blindtext
 

### 2020-03-11 Article.tex

@@
 \caption{Components of the state space model and demonstration on simulated data.
-\textbf{(A)} 19 simulated place cells with peaks evenly spaced on a 180 cm track that exhibit three different sequence patterns.
+\textbf{(A)} 19 simulated place cells that exhibit three different sequence patterns.
 For the first third of time, one cell fires repeatedly, representing one stationary location.
@@
 For the last third cells fire in an incoherent pattern, representing a non-smooth fragmented trajectory through space.
-\textbf{(B)}} Like the Bayesian Decoder, the state space encoding model uses estimates of the cell's place fields to compute a likelihood of position.
-\textbf{(C)}
+\textbf{(B)} Like the Bayesian Decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
+\textbf{(C)} This is combined with an explicit model of each dynamic state--which determines how latent position can change based on the posterior from the previous time step.
+In the hover state, this is modeled as the identity matrix because the position remains the same.
+In the continuous state, the dynamics are modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close.
+In the fragmented state, the dynamics are modeled as uniform, meaning the next position is equally likely to be anywhere.
+Note that the states displayed only reflect the within-state dynamics.
+\textbf{(D)} The probability of remaining in a particular state versus switching to another state.
+This is modeled as having a high probability of remaining in a particular state with a small probability of switching to one of the other states at each time step.
+\textbf{(F)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
+The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic state over time (lines).
+This can also be used to get an estimate of latent position over time (right panel).
+\textbf{(G)} The probability of each state given a particular constant speed trajectory.
+Each speed corresponds to a simulated spiking sequence at that speed.
+The lines correspond to the average probability of that state over the sequence.
+Dotted line represents the 0.8 classification threshold we use to classify each range of speeds and shaded regions correspond to the classification we have given the range of speeds.
+}
 \label{Figure1}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v3}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{Demonstration of the state space model on real data using both sorted and clusterless spikes on the same SWR.
+\textbf{(A)} Decoding using sorted spikes (arranged top to bottom).
+In the top panel, 31 cells on a W-track ordered according to linearized position by their place field peaks.
+The second panel shows the probability of each dynamic over time as in Figure 1F.
+Shaded regions correspond to the speed classifications in Figure 1G.
+The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
+L, R, C correspond to the position of the left, right and center wells respectively.
+The animal's position is indicated by the red triangle.
+\textbf{(B)} Decoding using clusterless spikes.
+The top panel shows multiunit spiking activity from each tetrode.
+Other panels have the same convention as(A).
+\textbf{(C)} 1D MAP estimate of the latent position in (A) projected back into 2D.
+Color indicates time.
+The animal's position is denoted by the red circle.
+Grey lines show the animal's 2D position over the entire epoch.
+L, R, and C correspond to the left, right and center well as in (A).
+\textbf{(D)} 1D MAP estimate of the latent position in (B) projected back into 2D.
+Figure conventions are the same as in (C).
+}
 \label{Figure2}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v3}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{A-D.
+Four examples of replays that would not be well-described using the Bayesian decoder.
+Figure conventions are the same as in Figure 2.
+\textbf{(A)} A replay that starts down the center arm away from the animal's position at the center well, slows down, and returns back.
+\textbf{(B)} A replay that stays at the left well for the entirety of the SWR while the animal is in the center arm.
+\textbf{(C)} A replay that begins stationary at the left well and then jumps to the middle of the right arm and proceeds up the right arm to the right well.
+\textbf{(D)} A replay that begins stationary at the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
+\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories (Hover, Hover-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
+\textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Hover, Hover-Continuous-Mix, or Continuous content.
+\textbf{(G)} Of the classified content, the percentage of SWR in each epoch that has some Continuous content.
+}
 \label{Figure3}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v2}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{\textbf{(A)} UpSet plot of the most common combinations of dynamics within each ripple.
+Each row represents a possible category present in the SWR.
+Each column represents a set of dynamics categories, where filled-in black dots with an edge between the dots indicates that multiple categories are present in the SWR (at temporally distinct times).
+The sets are ordered by how often they occurred as indicated by the bar plot above each category.
+The total number of each category is indicated by the rotated bar plot on the left.
+\textbf{(B)} Box plot of the duration of each dynamic category within a ripple.
+\textbf{(C)} The distance of the latent position from the animal's position for each dynamic category within each ripple.
+\textbf{(D)} Probability density estimate
+}
 \label{Figure4}

### 2020-03-11 Article.tex
Update on Overleaf.

@@
 \section*{Results}
-\subsection*{Test}
+\subsection*{The model can successfully detect short-time scale changes in dynamics}
 
@@
 
+\subsection*{The model can find canonical continuous replays}
+
 \begin{figure*}%[tbhp]
@@
 
+\subsection*{The model also finds replays that would not normally be classified by the Bayesian Decoder}
+\subsection*{Most replays are spatially coherent, but often not at the the speed posited by canonical replay}
+
 \begin{figure*}%[tbhp]
@@
 \caption{A-D.
-Four examples of replays that would not be well-described using the Bayesian decoder.
+Four examples of replays that would not be well-characterized by using the Bayesian decoder.
 Figure conventions are the same as in Figure 2.
@@
 
+\subsection*{Most replays only have one dynamic but some do not}
+
 \begin{figure*}%[tbhp]
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v2}
-\caption{\textbf{(A)} UpSet plot of the most common combinations of dynamics within each ripple.
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v3}
+\caption{
+Analysis of SWRs from all animals.
+\textbf{(A)} UpSet plot of the most common combinations of dynamics within each ripple.
 Each row represents a possible category present in the SWR.
@@
 The total number of each category is indicated by the rotated bar plot on the left.
-\textbf{(B)} Box plot of the duration of each dynamic category within a ripple.
-\textbf{(C)} The distance of the latent position from the animal's position for each dynamic category within each ripple.
-\textbf{(D)} Probability density estimate
+\textbf{(B)} Distribution of the duration of each dynamic category within a SWR.
+\textbf{(C)} The distance of the average latent position from the animal's position for each dynamic category within each SWR.
+\textbf{(D)} Position of hovers on the W-track at least 30 cm away from the animal's position.
+\textbf{(E)} Distribution of multiunit spike rates for each dynamic category within each SWR.
 }
@@
 \end{figure*}
+
+\subsection*{Many replays have time periods where the latent position remains at just one position and that position can be both local and non-local to the animal's position}
+\subsection*{Non-local hovers, though rare, occur all over the track, but mostly at reward wells and choice points.}
+
+\subsection*{Removing spatial information from the model leads to fewer classified ripples and shorter duration dynamic categories}
 
@@
 \section*{Methods}
+\subsection*{Experiment and Electrophysiology}
+\subsection*{Linearization}
+\subsection*{Ripple Detection}
+\subsection*{Spike Sorting}
+\subsection*{Clusterless Data}
+\subsection*{Simulated Data}
+
 \subsection*{The Model}
-The
 
@@
 
-\subsection*{Decoding - Clusterless}
+\subsection*{Decoding}
 
-\subsection*{Simulated Data}
-
-\subsection*{Data}
+\subsection*{Classification of Dynamic States}
+\subsection*{Replay distance from animal}
 
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v1_supplemental1}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{The model is robust to changes in the discrete transition matrix.
+Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value--which governs the probability of remaining in that state.
+The off-diagonal values--the probability of switching to one of the other states--are set to be equally likely with the remainder of the probability, as in Figure 1D.
+The first plot (upper left) shows the case when all states are equally likely.
+The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero--i.e.
+the case where there is no probability of switching to another state.}
 \label{fig:Figure1-Figure supplement 1}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v2-supplemental1}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{Decoding the same SWR in Figure 2 using 2D clusterless and sorted spikes.
+\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure 2.
+The middle plot shows the probability of each dynamic over time from the 2D decode.
+Shaded regions indicate classification category as in Figure 1G and 2.
+The rightmost plot shows the MAP estimate of the latent position with color indicating time.
+The latent position posterior summed over time is shown in purple underneath.
+The light grey lines represent the position of the animal over the entire epoch and the red dot represents the animal's position.
+\textbf{(B)} Same as in A, but with clusterless decoding.}
 \label{fig:Figure2-Figure supplement 1}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v2-supplemental2}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{A-C.
+Examples of continuous replays.
+Left panel uses the same conventions as Figure 2A and 2B.
+Right panel shows the 1D MAP estimate projected back to 2D as in Figure 2C.
+Color indicates time.
+Light grey lines indicate the animal's position over the entire epoch.}
 \label{fig:Figure2-Figure supplement 2}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v2_supplemental1}
-\caption{Placeholder image of Iris with a long example caption to show justification setting.}
+\caption{More examples of replays that would not be well-characterized by using the Bayesian decoder.
+Conventions are the same as in Figure 3.
+\textbf{(A)} A continuous replay that starts in the left arm back toward the center well, jumps to the right arm and continuous back toward the center well.
+\textbf{(B)} A continuous replay that travels down the right arm toward the center well, proceeds past the choice point toward the left well and then returns back to the choice point.
+\textbf{(C)} A replay that starts down the left arm toward the center arm and turns into a hover near the choice point.
+\textbf{(D)} The first half is a fragmented replay and the second half is unclassified.
+\textbf{(E)} A hover in the first half of the SWR on the right arm and then a hover in the SWR at the left well.
+\textbf{(F)} A hover that starts at the animal's position at the center well, then jumps to the left well and comes back toward the center well.
+}
 \label{fig:Figure3-Figure supplement 1}

### 2020-03-12 Article.tex

@@
 The second panel shows the probability of each dynamic over time as in Figure 1F.
-Shaded regions correspond to the speed classifications in Figure 1G.
+Shaded regions correspond to the speed classifications as in Figure 1G.
 The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
@@
 The animal's position is denoted by the red circle.
-Grey lines show the animal's 2D position over the entire epoch.
+Light grey lines show the animal's 2D position over the entire epoch.
 L, R, and C correspond to the left, right and center well as in (A).
@@
 \subsection*{Ripple Detection}
+Ripple detection performed as in
 \subsection*{Spike Sorting}

### 2020-03-25 src/figure2.py
Save to appropriate figure directories

@@ -1,13 +1,12 @@
 import copy
+import os
 
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 
-from src.analysis import (get_is_classified, get_probability,
-                          maximum_a_posteriori_estimate)
-from src.figure_utilities import (ONE_COLUMN, PAGE_HEIGHT, TWO_COLUMN,
-                                  save_figure)
+from src.analysis import get_is_classified, get_probability
+from src.figure_utilities import ONE_COLUMN, PAGE_HEIGHT, save_figure
 from src.parameters import PROBABILITY_THRESHOLD, STATE_COLORS
 from src.visualization import (plot_1D_wtrack_landmarks,
                                plot_2D_position_with_color_time)
@@ -153,7 +152,7 @@ def plot_1D_results(spike_times, data, results_1D,
         f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
         f"{data_type}_1D_acasual_classification"
     )
-    save_figure(fig_name)
+    save_figure(os.path.join("Figure2", fig_name))
 
 
 def plot_1D_projected_to_2D(epoch_key, ripple_number, data,
@@ -233,4 +232,4 @@ def plot_1D_projected_to_2D(epoch_key, ripple_number, data,
         f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
         f"{data_type}_1D_acasual_classification"
     )
-    save_figure(fig_name)
+    save_figure(os.path.join("Figure2", fig_name))

### 2020-03-25 src/figure2_supplemental1.py

@@ -1,3 +1,5 @@
+import os
+
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
@@ -185,6 +187,6 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
         f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
         f"{data_type}_2D_acasual_classification"
     )
-    save_figure(fig_name)
+    save_figure(os.path.join("Figure2-supplemental1", fig_name))
 
     return cbar

### 2020-03-25 src/figure2_supplemental2.py

@@ -1,4 +1,5 @@
 import copy
+import os
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -140,7 +141,7 @@ def plot_clusterless_1D_results(multiunit_times, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"clusterless_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure2-supplemental2", fig_name))
 
 
 def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
@@ -221,4 +222,4 @@ def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"{data_type}_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure2-supplemental2", fig_name))

### 2020-03-25 src/figure3.py

@@ -1,4 +1,5 @@
 import copy
+import os
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -140,7 +141,7 @@ def plot_clusterless_1D_results(multiunit_times, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"clusterless_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure3", fig_name))
 
 
 def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
@@ -220,4 +221,4 @@ def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"{data_type}_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure3", fig_name))

### 2020-03-25 src/figure3_supplemental1.py

@@ -1,4 +1,5 @@
 import copy
+import os
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -140,7 +141,7 @@ def plot_clusterless_1D_results(multiunit_times, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"clusterless_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure3-supplemental1", fig_name))
 
 
 def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
@@ -220,4 +221,4 @@ def plot_1D_projected_to_2D(epoch_key, ripple_number, data, results,
             f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
             f"{data_type}_1D_acasual_classification"
         )
-        save_figure(fig_name)
+        save_figure(os.path.join("Figure3-supplemental1", fig_name))

### 2020-03-25 Article.tex
Update on Overleaf.

@@
 \section*{Results}
-\subsection*{The model can successfully detect short-time scale changes in dynamics}
+\subsection*{Overview of the model}
+
+In order to validate the model's effectiveness in detecting short time scale changes in spiking dynamics, we first simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
+Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
+We then applied our decoding algorithm to the spiking sequence in Figure \ref{1}A.
+For the first third of this sequence, a single place cell fires repeatedly.
+For the middle third of the sequence, the cells fire in sequential spatial order.
+For the last third, the cells fire in a incoherent spatial pattern.
+
+Decoding the spiking sequence requires specifying two elements: the data model and the dynamics model.
+For the data model, our decoder is the same as the Bayesian decoder.
+We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
+This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
+In contrast to the typical Bayesian decoder, we use small time bins (2 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
+This allows us to detect changes on smaller time scales than would be possible with the 20 ms bin size used with the Bayesian decoder.
+
+Next, we specify a dynamics model for how latent position can evolve at a given time (Figure \ref{1}C).
+We are interested in identifying three types of different dynamics---hover, continuous, and fragmented---which are modeled with
 
@@
 }
-\label{Figure1}
+\label{1}
 \end{figure*}
@@
 \subsection*{The model also finds replays that would not normally be classified by the Bayesian Decoder}
-\subsection*{Most replays are spatially coherent, but often not at the the speed posited by canonical replay}
+
+\subsection*{Most replays are spatially coherent, but often not at the speed posited by canonical replay}
 
@@
 \subsection*{Many replays have time periods where the latent position remains at just one position and that position can be both local and non-local to the animal's position}
-\subsection*{Non-local hovers, though rare, occur all over the track, but mostly at reward wells and choice points.}
+\subsection*{Non-local hovers occur all over the track, but mostly at reward wells and choice points.}
 
-\subsection*{Removing spatial information from the model leads to fewer classified ripples and shorter duration dynamic categories}
+\subsection*{Removing spatial information from the model leads to fewer classified ripples}
 
@@
 \section*{Methods}
-\subsection*{Experiment and Electrophysiology}
+\subsection*{Recording Locations and Techniques}
+Ten male Long Evans rats (500-700 g, 4-9 months old) were trained on a W-track spatial alternation task.
+XX rats contributed to a previous study \cite{KarlssonAwakereplayremote2009}.
+Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
+We only used hippocampal tetrodes (CA1, CA2, CA3) in this study.
+
+\subsection*{Behavioral Task}
+The behavioral task is a W-track spatial alternation task as presented in \cite{KarlssonAwakereplayremote2009}.
+For each day, animals alternated between 15 minute sessions in a rest box and 15 minutes in the W-track.
+The behavioral task is a W-track spatial alternation task as presented in \cite{KarlssonAwakereplayremote2009}.
+For each day, animals alternated between 15 minute sessions in a rest box and 15 minutes in the W-track.
+Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
+
+Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
+
 \subsection*{Linearization}
-\subsection*{Ripple Detection}
-Ripple detection performed as in
+To decrease the time it takes to run the model, the 2D position of the animal is converted into a 1D position.
+This is done by first defining a graph representation of the track, where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
+Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to classify which track segment the animal is on.
+This algorithm was developed to track cars on roads with GPS signals.
+Using the HMM prevents sudden jumps from one track segment to another, which is particularly important near intersections when we are tracking the head position.
+Briefly, the observation model of the HMM is Gaussian and it models the likelihood of being on a track segment as the Gaussian distance to that segment with a 5 cm standard deviation.
+The state transition model is an empirically estimated state transition that changes with each time point that tries to ensure that the euclidean distance between successive position estimates is similar to the shortest path distance along the graph between successive position estimates.
+A slight bias of 0.1 is given to the diagonal to encourage staying on the same track segment.
+The most likely track segment the animal is on is computed using the Viterbi algorithm.
+After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the track segment.
+This allows us to define a distance from the center well in terms of shortest path length on the track, where 0 cm represents the center well position.
+The linear distance can then be converted into a linear position by assigning each track segment a position in 1D space.
+The code used for linearization can be found at \url{https://github.com/Eden-Kramer-Lab/loren_frank_data_processing}.
+
+\subsection*{SWR Detection}
+Sharp wave ripples were detected using the same method as in \cite{Kayhippocampalnetworkspatial2016}.
+Briefly, each LFP was filtered between 150-250 Hz, squared and then summed across tetrodes--forming a single population trace over time.
+This trace was smoothed with a Gaussian with a 4 ms standard deviation and the square root of this trace was taken to get an estimate of the population ripple band power.
+Candidate SWR times were found by z-scoring the population power trace and finding times when the z-score exceeded 2 standard deviations for a minimum of 15 ms and the speed of the animal was less than 4 cm/s.
+The SWR times were then defined as times when the z-score was greater than the mean and contained the candidate SWR times.
+The code used for ripple detection can be found at \url{https://github.com/Eden-Kramer-Lab/ripple_detection}.
+
 \subsection*{Spike Sorting}
-\subsection*{Clusterless Data}
+
 \subsection*{Simulated Data}
@@
 
-\subsection*{Decoding}
+\subsection*{Decoding - Clusterless}
 
 \subsection*{Classification of Dynamic States}
+We used a threshold of 0.8 to classify the probability of each state into 5 categories.
+Time periods during sharp wave ripples are labeled as Hover, Continuous, or Fragmented when the probability of each state is above 0.8.
+Time periods are labeled as Hover-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Hover and Continuous or Fragmented and Continuous are above 0.8, respectively.
+Time periods where none of these criterion are met are considered unclassified because there is not strong enough evidence for one or more of the states.
+
 \subsection*{Replay distance from animal}
+Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's projected position on the track graph (see Linearization) and the MAP estimate of the joint posterior marginalized over state, which is defined as the center of the position bin with the greatest posterior value.
 

### 2020-03-25 src/parameters.py
Add documentation of why the diagonal is 0.98

@@ -1,6 +1,7 @@
 from os.path import abspath, dirname, join, pardir
 
 import numpy as np
+
 from loren_frank_data_processing import Animal
 from replay_trajectory_classification.misc import NumbaKDE
 
@@ -42,7 +43,16 @@ model_kwargs = {
 }
 knot_spacing = 5
 spike_model_penalty = 0.5
+
+'''
+1. Geometric mean of duration is 1 / (1 - p)
+2. So p = 1 - (1 / n_time_steps).
+3. Want `n_time_steps` to equal 100 ms.
+4. If our timestep is 2 ms, then n_time_steps = 50
+5. So p = 0.98
+'''
 discrete_diag = 0.98
+
 continuous_transition_types = (
     [['random_walk', 'uniform', 'identity'],  # noqa
      ['uniform',     'uniform', 'uniform'],   # noqa

### 2020-03-25 src/parameters.py
Add clarification

@@ -45,7 +45,7 @@ knot_spacing = 5
 spike_model_penalty = 0.5
 
 '''
-1. Geometric mean of duration is 1 / (1 - p)
+1. Geometric mean of duration is: n_time_steps = 1 / (1 - p)
 2. So p = 1 - (1 / n_time_steps).
 3. Want `n_time_steps` to equal 100 ms.
 4. If our timestep is 2 ms, then n_time_steps = 50

### 2020-03-25 src/parameters.py
More notes on parameters

@@ -35,6 +35,12 @@ _BRAIN_AREAS = ['CA1', 'CA2', 'CA3']
 
 # Classifier Parameters
 place_bin_size = 3.0
+'''
+In one time step (2 ms),
+68% likely to be within 2.45 cm (12.25 m / s)
+95% likely to be within 4.90 cm (24.5 m / s)
+99.7% likely to be within 7.35 cm (36.75 m / s)
+'''
 movement_var = 6.0
 replay_speed = 1
 model = NumbaKDE

### 2020-04-21 Article.tex
Intro Draft 1

@@
 
-\title{A state space model for discovering the trajectory dynamics of replay}
+\title{A state space model for characterizing the trajectory dynamics of replay}
 \shorttitle{Replay Dynamics}
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[4]{Eric L. Denovellis}
+\author[1]{Eric L. Denovellis}
 \author[2, 3]{Anna K. Gillespie}
 \author[2, 3]{Michael E. Coulter}
-\author[1]{Uri T. Eden}
-\author[2, 3, 4, \Letter]{Loren M. Frank}
+\author[4]{Uri T. Eden}
+\author[1, 2, 3, \Letter]{Loren M. Frank}
 
-\affil[1]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
+\affil[1]{Howard Hughes Medical Institute, University of California, San Francisco, San Francisco, California}
 \affil[2]{Department of Physiology, University of California, San Francisco, San Francisco, California}
 \affil[3]{Kavli Institute for Fundamental Neuroscience, University of California, San Francisco, San Francisco, California}
-\affil[4]{Howard Hughes Medical Institute, University of California, San Francisco, San Francisco, California}
+\affil[4]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
 
@@
 To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent state to decompose place cell sequences into categories based on their latent dynamics.
-Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous.
+Each discrete latent “category” is associated with a type of continuous latent dynamic—spatially stationary, spatially fragmented or spatially continuous.
 This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories.
@@
 \section*{Introduction}
-The hippocampus is a brain area that is important for short term learning and memory.
-Neurons in the hippocampus preferentially fire in response to specific locations in an environment, which results in sequences of cells firing as an animal moves through an environment that reflect the trajectory of the animal.
-When the animal is asleep or immobile, cells also fire in sequence, forming trajectory-like representations that can recapitulate sequences previously observed when the animal was running.
+Neurons in the hippocampus preferentially fire in response to specific locations in an environment, firing in sequence as the animal moves through space.
+When the animal is asleep or immobile, cells also fire in sequence, forming trajectory-like representations that can recapitulate sequences previously observed when the animal was running (cite).
 These sequences are thought to be important for memory consolidation or prospective decision making.
-(cite some examples).
-Therefore, characterizing the dynamics of hippocampal sequences is important for understanding memory consolidation and prospective decision making.
+Therefore, characterizing the trajectory of hippocampal sequences is important for understanding memory consolidation and prospective decision making.
 
-Current methods analyze hippocampal sequences by assuming a constant velocity trajectory \cite{DavidsonHippocampalReplayExtended2009}, corresponding to a single dynamic.
-This is done by computing the probability of position at each time using the place field information from the population of neurons and then fitting a line to maximize the probability of position over time.
+An often used method for characterizing the trajectories of hippocampal sequences is the Bayesian decoder (cite).
+The Bayesian decoder consists of two steps: first, the probability of position is computed at each time bin using the previously estimated place fields from the population of neurons.
+Second---assuming the trajectory of the replay is of constant velocity---a line is fit that maximizes the probability of position over time.
 If the probabilities near the best fit line are high relative to shuffled distribution, the replay is analyzed.
-Often, trajectories that travel less than a certain distance \cite{PfeifferAutoassociativedynamicsgeneration2015} are excluded.
+Trajectories that travel less than a certain distance are often excluded (cite).
 
-However, there is a growing evidence that trajectories that have more complex dynamics.
-Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} found that juvenile rats .
-For example, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that trajectories alternated between stationary hovering trajectories and continuous movement trajectories on a gamma timescale.
+Although the Bayesian decoder has been useful in characterizing replay---particularly in the case of linear tracks---the assumption of constant velocity is a strong assumption about the dynamics of replay.
+Because replays are typically judged based on how well their trajectories conform to the constant velocity assumption, only a subset of sharp wave ripples or multiunit high synchrony events are typically analyzed (37\% of events\cite{KarlssonAwakereplayremote2009}, 16\%  of events\cite{DavidsonHippocampalReplayExtended2009} 5-20\% Tingley and Peyrache 2020).
+This potentially excludes replays with more complicated, non-constant velocity dynamics.
 
-Our mathematical model precisely specifies the dyna
+Recognizing this, several studies have moved away from the constant velocity assumption, using the weighted mean or maximum of each posterior time bin and connecting each time bin with a line (cite).
+For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that replays can alternate between representing a single location and sequential spatial trajectories.
+On the other hand, Stella et al.
+found that replays are more spatially continuous, following Brownian diffusion dynamics.
+One possibility for the difference between the studies is that this method is dependent on the size of the time bin, does not take into account the uncertainty of the posterior, and can lead to highly variable trajectories.
+
+State space models are a well-understood, well-known statistical solution to these problems.
+By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
+State space models have successfully been applied to decoding movement from the spiking of cells in the hippocampus (Brown et al 1998) and in classifying replay trajectories in real-time settings (Deng et al.
+2016).
+However, these state space models have generally assumed one dynamic persists through the replay, which would not capture more complicated, non-constant velocity trajectories.
+
+Here, we define a state space model of replay that characterizes trajectories as a mixture of three speed dynamics: stationary trajectories, continuous trajectories many times the typical speed of the animal, and fragmented trajectories.
+These dynamics represent the categories neuroscientists studying replay typically focus on when investigating replay.
+We show how this model can take advantage of clusterless decoding---which relates multiunit spike wave form features to position without spike sorting--giving us more information about the population spiking activity.
+Using this, we find that this model can identify replays with spatially coherent trajectories as well as replays with more complicated representations that can have a mixture of stationary, continuous, and fragmented dynamics.
+We also show that, by using this model, we are able to interpret a larger fraction of the replays than previous studies.
+Surprisingly, we find that many replay trajectories, although spatially coherent, are slower in speed than reported in previous studies including stationary events, which represent a single location, and slowly moving trajectories that are more consistent with the animal's movement speed.
+This challenges the notion that replay trajectories only consist of trajectories that are many times faster than the animal's behavior.
 
 \section*{Results}
-\subsection*{Overview of the model}
-
-In order to validate the model's effectiveness in detecting short time scale changes in spiking dynamics, we first simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
-Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then applied our decoding algorithm to the spiking sequence in Figure \ref{1}A.
-For the first third of this sequence, a single place cell fires repeatedly.
-For the middle third of the sequence, the cells fire in sequential spatial order.
-For the last third, the cells fire in a incoherent spatial pattern.
-
-Decoding the spiking sequence requires specifying two elements: the data model and the dynamics model.
-For the data model, our decoder is the same as the Bayesian decoder.
-We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
-This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the typical Bayesian decoder, we use small time bins (2 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the 20 ms bin size used with the Bayesian decoder.
-
-Next, we specify a dynamics model for how latent position can evolve at a given time (Figure \ref{1}C).
-We are interested in identifying three types of different dynamics---hover, continuous, and fragmented---which are modeled with
-
 \begin{figure*}%[tbhp]
@@
 \end{figure*}
+\subsection*{Overview of the model}
+In order to introduce and validate the model, we demonstrate how the model works on a simulated dataset (Figure \ref{1}).
+We simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
+Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
+We then apply our decoding algorithm to the spiking sequence in Figure \ref{1}A.
+For the first third of this sequence, a single place cell fires repeatedly.
+For the middle third of the sequence, the cells fire in sequential spatial order.
+For the last third, the cells fire in a incoherent spatial order.
+These firing patterns represent three different types of sequence dynamics, which we call stationary, continuous, and fragmented, respectively.
+The goal of our model is to characterize replay events in terms of these three dynamical states at every time point.
 
-\subsection*{The model can find canonical continuous replays}
+Decoding the spiking sequence dynamics requires specifying two elements: the data model and the dynamics model.
+For the data model, our decoder is the same as the Bayesian decoder.
+We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
+This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
+In contrast to the Bayesian decoder, we use small time bins (2 ms vs.
+20 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
+This allows us to detect changes on smaller time scales than would be possible with the Bayesian decoder.
+
+Next, we specify a dynamics model for how latent position--the "mental" position of the animal represented by the cells--can evolve in each of these sequences (Figure \ref{1}C).
+We do this by defining a state transition matrix, which defines how the latent position can change from the previous time step..
+In the stationary state, we expect latent position to not change between time steps, so we use an identity state transition matrix.
+In the continuous state, we expect the latent position to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
+In our case, since we are interested in identifying replays that move at speeds much faster than the animal's speed, we set this to 6.0 cm.
+This means that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or 24.5 m/s) which is consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
+In the fragmented state, we expect that the latent position can move to any available position.
+We model this using a uniform state transition matrix, which makes all positions equally likely.
+
+Finally, we specify how likely we are to stay in a dynamic state versus changing to another state (Figure \ref{1}D).
+In order to be conservative, we assume that each dynamic is likely to dominate for the entirety of the replay, with a small probability of switching to one of the other states.
+Therefore, we set the probability of staying in a state to 0.98 for each 2 ms time step.
+Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular state.
+
+Once we have specified the data and dynamics model, we have fully specified the state space model.
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state.
+We summarize the resulting posterior with two quantities: the probability of each state over time and the probability of position over time (Figure \ref{1}F, left and right plot respectively).
+From these two summaries, we can see that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+The probability of the stationary state has high probability at the beginning of the simulated replay, reflecting the consistent spiking from one cell.
+This gives way to the continuous state, reflecting the trajectory-like spiking from the population of cells.
+Then the fragmented state dominates for the last third when the population spiking is not spatially coherent.
+We can also see that these three states are reflected in the estimate of latent position of the animal.
+Importantly, we get a much less variable estimate due to the correct specification of the dynamics.
+
+One way to interpret the probabilities in our model is as a classifier of speeds of replay.
+To demonstrate this, we applied the model to simulated replays of increasing speed(Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
+From this, we can see that not only are there regions of speed that correspond to our three states being high probable (where highly probable means over 0.8), there are in-between speeds where two of the states dominate with high probability; that is, the sum of two of the states' probabilities is over 0.8.
+In this manuscript, we will refer to these as mixture states.
+For example, when the stationary state has a probability of 0.5 and the continuous has a probability of 0.5, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
+By using this classification scheme, we can characterize the speed or set of speeds in a replay.
+
+In order to verify that the model was robust to the choice of probability of staying in a state, we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
+We found that for a large range of plausible probabilities of staying in one of the states (0.9 - 0.9999), the model still correctly identified the dynamic states with high probability.
 
@@
 }
-\label{Figure2}
+\label{2}
 \end{figure*}
+\subsection*{The model finds known continuous replays with sorted and clusterless spikes}
+Next, we wanted to validate our model on real hippocampal data.
+We first applied the decoding algorithm to a sharp wave ripple (SWR) with clear sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on the W-track.
+We see that the probability of being in the continuous state is high throughout the SWR, with the probability of being in a stationary state being slightly higher at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+Using our speed classification, this means that the speed of the replay starts slower---as a mixture of continuous and stationary states---and then speeds up and slows down again.
+We can see this in the posterior probability of linear position over time as the replay travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel) as well as the projection of the maximum of this trajectory to 2D (Figure \ref{2}C).
+Moreover, when we apply the same model using 2D position, we get a similar result (Figure 2-Supplemental 1A).
 
-\subsection*{The model also finds replays that would not normally be classified by the Bayesian Decoder}
+One problem with using spikes from cells identified using spike sorting is that there is potentially a lot of information lost by determining the identity of each cell---many spikes are discarded with information about position because it could not be assigned to one cell.
+Therefore, we wanted to use a method which directly relates multiunit spike waveform features to position without spike sorting, which we call clusterless decoding.
+Clusterless decoding has been successfully used to identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016}, Kay et al.
+2020).
+Here, if we apply it to the same SWR as the sorted spikes, we get similar decoding and classification of the sequence (Figure \ref{2}B, D), both in using 1D linearized and 2D position (Figure 2-Supplemental 1B).
+Note that, using clusterless decoding, the estimate of the posterior probability of position is much more certain than the corresponding posterior probability of position using sorted spikes (Figure \ref{2}D vs C).
 
-\subsection*{Most replays are spatially coherent, but often not at the speed posited by canonical replay}
+\subsection*{The model finds replays that would not normally be identified by the Bayesian Decoder}
+After validating our model on a single SWR, we then applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009}).
+We observed many replays that were classified as continuous throughout the entirety of SWR and would have been identified by the Bayesian Decoder (Figure 2-Supplemental 2A-C) .
+However, we also observed replays that did not have this structure, but could be interpreted in terms of our model.
+For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A), trajectories that stayed in one position (Figure \ref{3}A), trajectories that started in a single location trajectories, jumped to a
+
+\subsection*{Most replays are spatially coherent, but slower than reported}
 
@@
 \subsection*{Removing spatial information from the model leads to fewer classified ripples}
-
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v1}
+\caption{Comparison of real vs.
+position shuffled data for two epochs from different animals.
+Line represents the percent of SWR classified in that epoch for real data.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
+\label{Figure5}
+\end{figure}
 \section*{Discussion}

### 2020-04-24 Article.tex
Update on Overleaf.

--- 2020-03-25 Article.tex
@@
 \begin{abstract}
-During sleep and immobility, hippocampal place cells fire in sequences consistent with temporally compressed versions of trajectories previously run by the animal.
-These replayed sequences are hypothesized to be an important mechanism for the retrieval of spatial memory in service of consolidation and decision-making.
-Replay events are typically evaluated based on whether they activate sequences of place cells that represent spatially continuous trajectories through the environment, but recent work has shown that these events can have more complex dynamics.
-For example, sequences can alternate between hovering on a particular spatial location and continuous movement or can represent continuous trajectories in other spatial environments, which may appear spatially incoherent in the context of the current environment.
-
-To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent state to decompose place cell sequences into categories based on their latent dynamics.
-Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous.
-This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories.
-In addition, the model can function in 2D, avoiding linearization errors on more complicated environments.
-We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.
+Representations of past or possible future experiences play a critical role in decision-making processes.
+The hippocampus is a brain region known to express these representations across multiple states, including during sharp-wave ripple (SWR) events where sets of hippocampal place cells can be activated in time-compressed sequences consistent with a rapid mental traversal of a path.
+Efforts to understand these representations typically make arbitrary assumptions about the character and dynamics of the generative representations.
+As a result, we lack a clear picture of the nature and diversity of SWR representations.
+Here we develop a flexible state space model that uses a combination of discrete and latent states to describe the content and time evolution of SWR events.
+Application of this model reveals that a much larger proportion of SWR events contain coherent spatial content that previously reported, and that the dynamics of events span a surprisingly wide range.
 
@@
 \section*{Introduction}
-The hippocampus is a brain area that is important for short term learning and memory.
-Neurons in the hippocampus preferentially fire in response to specific locations in an environment, which results in sequences of cells firing as an animal moves through an environment that reflect the trajectory of the animal.
-When the animal is asleep or immobile, cells also fire in sequence, forming trajectory-like representations that can recapitulate sequences previously observed when the animal was running.
-These sequences are thought to be important for memory consolidation or prospective decision making.
-(cite some examples).
-Therefore, characterizing the dynamics of hippocampal sequences is important for understanding memory consolidation and prospective decision making.
+The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
+These internal "generative" representations can be identified because the underlying neural activity patterns resemble those seen during actual events.
+In particular, an encoding model that relates neural firing to external covariates can be constructed during an experience, and this model can inverted to decode an internal representation of the covariates from the neural activity at a different time.
 
-Current methods analyze hippocampal sequences by assuming a constant velocity trajectory \cite{DavidsonHippocampalReplayExtended2009}, corresponding to a single dynamic.
-This is done by computing the probability of position at each time using the place field information from the population of neurons and then fitting a line to maximize the probability of position over time.
-If the probabilities near the best fit line are high relative to shuffled distribution, the replay is analyzed.
-Often, trajectories that travel less than a certain distance \cite{PfeifferAutoassociativedynamicsgeneration2015} are excluded.
+Mental events do not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to those actions.
+Consistent with this, generative representations can unfold rapidly.
+As a result, decoding generative representations requires making assumptions about their timescale and temporal structure, and these assumptions can limit the interpretation of the data.
 
-However, there is a growing evidence that trajectories that have more complex dynamics.
-Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} found that juvenile rats .
-For example, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that trajectories alternated between stationary hovering trajectories and continuous movement trajectories on a gamma timescale.
+Hippocampal replay events are a prototypical example of a generative representation.
+As animals move through space, neurons in the hippocampus preferentially fire in at specific locations in an environment, and thus sets of cells can fire in sequence as the animal moves through a series of locations.
+When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
+A subset of SWRs contain sequential firing similar to that seen during a previous or potential future experience, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed.
 
-Our mathematical model precisely specifies the dyna
+Identifying these events typically involves multiple steps and assumptions about the nature of the event.
+First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
+Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter.
+The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
+Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed.
+A statistical test is then used to determining whether the fit is better than the fit to shuffled versions of the data.
+
+While this approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bin acts as a boxcar smoother that limits the potential movement speeds of the representation.
+The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
+Moreover, there is now substantial evidence that a subset of these events express stationary representations, which are often excluded.
+In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
+
+ Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the weighted mean or maximum of each posterior time bin and connecting each time bin with a line (cite).
+For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that replays can alternate between representing a single location and sequential spatial trajectories.
+On the other hand, Stella et al.
+found that replays are more spatially continuous, following Brownian diffusion dynamics.
+Both methods still used fixed time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
+
+An ideal approach to identifying and quantifying the dynamics of generative representations would circumvent these problems.
+It would use all of the available spiking data to yield the most accurate decoded positions.
+It would be applicable to either thresholded events or to all of the data to permit an unbiased assessment of the prevalence and nature of generative activity.
+It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
+It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for the dynamic.
+Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
+
+We therefore developed a state space model (cite Brown, Deng) of generative activity that achieves all of those goals.
+State space models are a well-understood, well-known statistical solution to the problems described above.
+By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
+Our model goes beyond previous approaches (refs) and characterizes represented trajectories as a mixture of three underlying patterns of dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
+We show how this model can take advantage of clusterless decoding---which relates multiunit spike wave form features to position without spike sorting---giving us more information about the population spiking activity.
+We apply this model to data from 10 animals, focusing on SWR events to permit a direct comparison to previous work.
+We find that the vast majority of SWRs contain spatially coherent content.
+Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at speeds consistent with real animal movement, suggesting that high speed replay is not the most common manifestation of generative activity.
+Our method provides a useful and powerful tool for understanding the dynamics of mental representations, and could be adapted for use in many brain areas.
 
 \section*{Results}
-\subsection*{Overview of the model}
-
-In order to validate the model's effectiveness in detecting short time scale changes in spiking dynamics, we first simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
-Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then applied our decoding algorithm to the spiking sequence in Figure \ref{1}A.
-For the first third of this sequence, a single place cell fires repeatedly.
-For the middle third of the sequence, the cells fire in sequential spatial order.
-For the last third, the cells fire in a incoherent spatial pattern.
-
-Decoding the spiking sequence requires specifying two elements: the data model and the dynamics model.
-For the data model, our decoder is the same as the Bayesian decoder.
-We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
-This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the typical Bayesian decoder, we use small time bins (2 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the 20 ms bin size used with the Bayesian decoder.
-
-Next, we specify a dynamics model for how latent position can evolve at a given time (Figure \ref{1}C).
-We are interested in identifying three types of different dynamics---hover, continuous, and fragmented---which are modeled with
-
 \begin{figure*}%[tbhp]
@@
 In the fragmented state, the dynamics are modeled as uniform, meaning the next position is equally likely to be anywhere.
-Note that the states displayed only reflect the within-state dynamics.
+Note that the dynamics displayed only reflect the within-state dynamics.
 \textbf{(D)} The probability of remaining in a particular state versus switching to another state.
-This is modeled as having a high probability of remaining in a particular state with a small probability of switching to one of the other states at each time step.
+This is modeled as having a high probability of remaining in a particular state with a small probability of switching to one of the other dynamics at each time step.
 \textbf{(F)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
@@
 \end{figure*}
+\subsection*{Overview of the model}
+In order to introduce and validate the model, we demonstrate how the model works on a simulated dataset (Figure \ref{1}).
+We simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
+Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
+We then apply our decoding algorithm to the spiking sequence in Figure \ref{1}A.
+For the first third of this sequence, a single place cell fires repeatedly.
+For the middle third of the sequence, the cells fire in sequential spatial order.
+For the last third, the cells fire in a incoherent spatial order.
+These firing patterns represent three different types of sequence dynamics, which we call stationary, continuous, and fragmented, respectively.
+The goal of our model is to characterize replay events in terms of a mixture of these three dynamics at every time point.
 
-\subsection*{The model can find canonical continuous replays}
+Decoding the spiking sequence dynamics requires specifying two elements: the data model and the dynamics model.
+For the data model, our decoder is the same as the Bayesian decoder.
+We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
+This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
+In contrast to the Bayesian decoder, we use small time bins (2 ms vs.
+20 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
+This allows us to detect changes on smaller time scales than would be possible with the Bayesian decoder.
+
+Next, we specify a dynamics model for how latent position---the "mental" position of the animal represented by the cells---can evolve in each of these sequences (Figure \ref{1}C).
+We do this by defining a state transition matrix, which defines how the latent position can change from the previous time step.
+In the stationary state, we expect latent position to not change between time steps, so we use an identity state transition matrix.
+In the continuous state, we expect the latent position to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
+In our case, since we are interested in identifying replays that move at speeds much faster than the animal's speed, we set this to 6.0 cm.
+This means that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or 24.5 m/s), consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
+In the fragmented state, we expect that the latent position can move to any available position.
+We model this using a uniform state transition matrix, which makes all positions equally likely.
+
+Finally, we specify how likely we are to stay in a dynamic state versus changing to another state (Figure \ref{1}D).
+In order to be conservative, we assume that each dynamic is likely to dominate for the entirety of the replay, with a small probability of switching to one of the other dynamics.
+Therefore, we set the probability of staying in a state to 0.98 for each 2 ms time step.
+Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular state.
+
+Once we have specified the data and dynamics model, we have fully specified the state space model.
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state.
+We summarize the resulting posterior with two quantities: the mixture probability of each state over time and the probability of position over time (Figure \ref{1}F, left and right plot respectively).
+From these two summaries, we can see that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+The probability of the stationary state has high probability at the beginning of the simulated replay, reflecting the consistent spiking from one cell.
+This gives way to the continuous state, reflecting the trajectory-like spiking from the population of cells.
+Then the fragmented state dominates for the last third when the population spiking is not spatially coherent.
+We can also see that these three dynamics are reflected in the estimate of latent position of the animal.
+Importantly, we get a much less variable estimate due to the correct specification of the dynamics.
+
+One way to interpret the probabilities in our model is as a classifier of speeds of replay.
+To demonstrate this, we applied the model to simulated replays of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
+From this, we can see that not only are there regions of speed that correspond to our three dynamics being high probable (where we define highly probable to be over 0.8 probability), there are in-between speeds where two of the dynamics dominate with high probability; that is, the sum of two of the dynamics' probabilities is over 0.8.
+In this manuscript, we will refer to these as mixture dynamics.
+For example, when the stationary state has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
+By using this classification scheme, we can characterize the speed or set of speeds in a replay by its dynamics.
+
+In order to verify that the model was robust to the choice of probability of staying in a state, we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
 
@@
 }
-\label{Figure2}
+\label{2}
 \end{figure*}
-
-\subsection*{The model also finds replays that would not normally be classified by the Bayesian Decoder}
-
-\subsection*{Most replays are spatially coherent, but often not at the speed posited by canonical replay}
+\subsection*{The model finds known continuous replays with sorted and clusterless spikes}
 
@@
 }
-\label{Figure3}
+\label{3}
 \end{figure*}
 
-\subsection*{Most replays only have one dynamic but some do not}
+Next, we wanted to validate our model on real hippocampal data.
+We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with clear sequential population activity (Figure \ref{2}A, top panel).
+We observed that the probability of being in the continuous state is high throughout for this SWR, with the probability of being in a stationary state being slightly higher at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+Using our speed classification, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+This is also evident in the posterior probability of linear position over time.
+This shows that the replay travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
+We can also see this when we project the maximum of the posterior of this trajectory to 2D (Figure \ref{2}C) to better see the spatial trajectory.
+Moreover, when we apply the same model using the 2D position of the animal, we get a similar result as when we used the 1D linearized position (Figure 2-Supplemental 1A).
 
+One problem with using spikes from cells identified using spike sorting is that there is potentially a lot of information lost by determining the identity of each cell---many spikes with position information are discarded because the spikes could not be assigned to one cell.
+Therefore, we wanted to use a method which directly relates multiunit spike waveform features to position without spike sorting, which we call clusterless decoding.
+Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016}, Kay et al.
+2020).
+Here, if we apply it to the same SWR as the sorted spikes, we get similar decoding and classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure 2-Supplemental 1B).
+Note that, using clusterless decoding, the estimate of the posterior probability of position is much more narrow, indicating that there is more certainty in the true latent position at that time compared the corresponding posterior probability of position using sorted spikes (Figure \ref{2}D vs C).
+
+\subsection*{The model finds replays with non-constant speed dynamics}
+After validating our model on a single SWR, we then applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009}).
+We observed many replays that were classified as continuous throughout the entirety of SWR and likely would have been identified by the most commonly used Bayesian Decoder (Figure 2-Supplemental 2A-C).
+However, we also observed replays that did not have this structure, but could be interpreted in terms of our model.
+For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that stayed in one position (Figure \ref{3}B), trajectories that jumped between arms (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that remained spatially incoherent throughout the SWR (Figure3-Supplemental1D).
+Overall, we were able to classify XX\% of ripples as containing at least one of the five dynamics.
+To verify that this depended on the place field coding of hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
+There were far fewer SWR classified when spatial information was removed (p=0.XX for epoch 1, p=0.XX for epoch 2, Figure 3-Supplemental 2).
+
+To better understand the distribution of dynamics during SWRs, we then asked what percentage of classified replays contained spatially coherent structure---which we define as a SWR containing any of stationary, stationary-continuous mixture, or continuous categories.
+We found that XX\% of SWR had some spatially coherent structure while XX\% had some spatially incoherent structure.
+This was consistent across animals (Figure \ref{3}F).
+We wondered what percentage of SWR contained continuous content, like would typically be analyzed when using by the Bayesian decoder.
+We found that, consistent with previous reports, only XX\% had continuous content (Figure \ref{3}G).
+This highlights that much of the SWR replay content is missed when analyzing SWRs with the Bayesian decoder.
+
+\subsection*{Many replay trajectories are closer to the animal's movement speed}
 \begin{figure*}%[tbhp]
@@
 Analysis of SWRs from all animals.
-\textbf{(A)} UpSet plot of the most common combinations of dynamics within each ripple.
+\textbf{(A)} UpSet plot---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each ripple.
 Each row represents a possible category present in the SWR.
@@
 }
-\label{Figure4}
+\label{4}
 \end{figure*}
 
-\subsection*{Many replays have time periods where the latent position remains at just one position and that position can be both local and non-local to the animal's position}
-\subsection*{Non-local hovers occur all over the track, but mostly at reward wells and choice points.}
+We then looked at which sets of dynamics are most common.
+Many, but not all, of the SWR ripples were well described by one dynamic (XX/YY, XX\%).
+Surprisingly, the most common SWRs were ones that were categorized exclusively as stationary-continuous mixtures (XX\%, Figure \ref{4}A).
+These trajectories had similar duration to the continuous events (median XX ms, Figure \ref{4}B) and occurred away from the animal (median XX cm, Figure \ref{4}C).
+This indicates that the trajectories of these SWRs do not represent the animal's position and move at a speed that is closer to the animal's actual movement speed when running, rather than many times the animal's movement speed.
 
-\subsection*{Removing spatial information from the model leads to fewer classified ripples}
+The next most common category is those SWRs that were exclusively stationary (XX\%).
+Unlike the Stationary-Continuous Mixtures, most of these trajectories represented the animal's position (Figure \ref{4}C), but there were many that represented positions as far as XX cm from the animal.
+We call the ones that are more than 30 cm away from the animal "non-local stationary trajectories".
+There were XX/YY or XX\% non-local stationary trajectories.
+These non-local stationary trajectories represented positions all over the track, but mostly at reward wells and choice points (Figure \ref{4}D).
 
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v1}
+\caption{Comparison of real vs.
+position shuffled data for two epochs from different animals.
+Line represents the percent of SWR classified in that epoch for real data.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
+\label{Figure5}
+\end{figure}
 \section*{Discussion}
@@
 
-\subsection*{Classification of Dynamic States}
+\subsection*{Classification of Dynamics}
 We used a threshold of 0.8 to classify the probability of each state into 5 categories.
@@
 Time periods are labeled as Hover-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Hover and Continuous or Fragmented and Continuous are above 0.8, respectively.
-Time periods where none of these criterion are met are considered unclassified because there is not strong enough evidence for one or more of the states.
+Time periods where none of these criterion are met are considered unclassified because there is not an interpretation for all three dynamics being equally likely.
 
@@
 Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value--which governs the probability of remaining in that state.
-The off-diagonal values--the probability of switching to one of the other states--are set to be equally likely with the remainder of the probability, as in Figure 1D.
-The first plot (upper left) shows the case when all states are equally likely.
+The off-diagonal values--the probability of switching to one of the other dynamics--are set to be equally likely with the remainder of the probability, as in Figure 1D.
+The first plot (upper left) shows the case when all dynamics are equally likely.
 The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero--i.e.

### 2020-05-20 Article.tex
Update on Overleaf.

@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v3}
+\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v4}
 \caption{Components of the state space model and demonstration on simulated data.
@@
 \textbf{(C)} This is combined with an explicit model of each dynamic state--which determines how latent position can change based on the posterior from the previous time step.
-In the hover state, this is modeled as the identity matrix because the position remains the same.
+In the stationary state, this is modeled as the identity matrix because the position remains the same.
 In the continuous state, the dynamics are modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v3}
+\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v4}
 \caption{Demonstration of the state space model on real data using both sorted and clusterless spikes on the same SWR.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v3}
+\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v4}
 \caption{A-D.
@@
 \textbf{(D)} A replay that begins stationary at the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
-\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories (Hover, Hover-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
-\textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Hover, Hover-Continuous-Mix, or Continuous content.
+\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
+\textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Stationary, Stationary-Continuous-Mix, or Continuous content.
 \textbf{(G)} Of the classified content, the percentage of SWR in each epoch that has some Continuous content.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v3}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v4}
 \caption{
@@
 \textbf{(C)} The distance of the average latent position from the animal's position for each dynamic category within each SWR.
-\textbf{(D)} Position of hovers on the W-track at least 30 cm away from the animal's position.
+\textbf{(D)} Position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
 \textbf{(E)} Distribution of multiunit spike rates for each dynamic category within each SWR.
@@
 
-\begin{figure}%[tbhp]
+\begin{figure*}%[tbhp]
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v1}
-\caption{Comparison of real vs.
-position shuffled data for two epochs from different animals.
-Line represents the percent of SWR classified in that epoch for real data.
-The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v2}
+\caption{}
 \label{Figure5}
-\end{figure}
+\end{figure*}
 \section*{Discussion}
@@
 We used a threshold of 0.8 to classify the probability of each state into 5 categories.
-Time periods during sharp wave ripples are labeled as Hover, Continuous, or Fragmented when the probability of each state is above 0.8.
-Time periods are labeled as Hover-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Hover and Continuous or Fragmented and Continuous are above 0.8, respectively.
+Time periods during sharp wave ripples are labeled as Stationary, Continuous, or Fragmented when the probability of each state is above 0.8.
+Time periods are labeled as Stationary-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Stationary and Continuous or Fragmented and Continuous are above 0.8, respectively.
 Time periods where none of these criterion are met are considered unclassified because there is not an interpretation for all three dynamics being equally likely.
@@
 % Supplementary Information %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%x%%%%%%%%%
 \beginsupplement
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v1_supplemental1}
+\includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v2_supplemental1}
 \caption{The model is robust to changes in the discrete transition matrix.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v2-supplemental1}
+\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}
 \caption{Decoding the same SWR in Figure 2 using 2D clusterless and sorted spikes.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v2-supplemental2}
+\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v3-supplemental2}
 \caption{A-C.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v2_supplemental1}
+\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v3_supplemental1}
 \caption{More examples of replays that would not be well-characterized by using the Bayesian decoder.
@@
 \textbf{(B)} A continuous replay that travels down the right arm toward the center well, proceeds past the choice point toward the left well and then returns back to the choice point.
-\textbf{(C)} A replay that starts down the left arm toward the center arm and turns into a hover near the choice point.
+\textbf{(C)} A replay that starts down the left arm toward the center arm and turns into a stationary trajectory near the choice point.
 \textbf{(D)} The first half is a fragmented replay and the second half is unclassified.
-\textbf{(E)} A hover in the first half of the SWR on the right arm and then a hover in the SWR at the left well.
-\textbf{(F)} A hover that starts at the animal's position at the center well, then jumps to the left well and comes back toward the center well.
+\textbf{(E)} A stationary trajectory in the first half of the SWR on the right arm and then a stationary trajectory in the SWR at the left well.
+\textbf{(F)} A stationary trajectory that starts at the animal's position at the center well, then jumps to the left well and comes back toward the center well.
 }
@@
 \end{figure*}
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v1}
+\caption{Comparison of real vs.
+position shuffled data for two epochs from different animals.
+Line represents the percent of SWR classified in that epoch for real data.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
+\label{Figure3-Figure supplement 2}
+\end{figure}
 

### 2020-05-20 src/figure4.py
Add figure4 plots

@@ -1,0 +1,191 @@
+import os
+from glob import glob
+
+import numpy as np
+import pandas as pd
+import seaborn as sns
+
+import matplotlib.pyplot as plt
+from loren_frank_data_processing import make_tetrode_dataframe
+from src.figure_utilities import PAGE_HEIGHT, TWO_COLUMN, save_figure
+from src.parameters import (_BRAIN_AREAS, ANIMALS, PROCESSED_DATA_DIR,
+                            SHORT_STATE_ORDER, STATE_COLORS, STATE_ORDER)
+from src.visualization import (SHORT_STATE_NAMES, _plot_category,
+                               plot_category_duration,
+                               plot_linear_position_markers,
+                               plot_replay_distance_from_actual_position)
+from upsetplot import UpSet
+
+
+def load_replay_info():
+    tetrode_info = make_tetrode_dataframe(ANIMALS)
+
+    data_type, dim = "clusterless", "1D"
+    n_unique_spiking = 2
+
+    file_paths = glob(
+        os.path.join(PROCESSED_DATA_DIR,
+                     f"*_{data_type}_{dim}_replay_info.csv")
+    )
+    replay_info = pd.concat(
+        [pd.read_csv(file_path) for file_path in file_paths], axis=0,
+    ).set_index(["animal", "day", "epoch", "ripple_number"])
+    replay_info = replay_info.loc[
+        replay_info.n_unique_spiking >= n_unique_spiking
+    ].sort_index()
+
+    is_brain_areas = tetrode_info.area.astype(
+        str).str.upper().isin(_BRAIN_AREAS)
+    n_tetrodes = (
+        tetrode_info.loc[is_brain_areas]
+        .groupby(["animal", "day", "epoch"])
+        .tetrode_id.count()
+        .rename("n_tetrodes")
+    )
+    replay_info = pd.merge(
+        replay_info.reset_index(), pd.DataFrame(n_tetrodes).reset_index()
+    ).set_index(["animal", "day", "epoch", "ripple_number"])
+
+    for state in STATE_ORDER:
+        replay_info[f"{state}_pct_unique_spiking"] = (
+            replay_info[f"{state}_n_unique_spiking"] /
+            replay_info["n_tetrodes"]
+        )
+    replay_info = replay_info.rename(index={"Cor": "cor"}).rename_axis(
+        index={"animal": "Animal ID"}
+    )
+
[…]

### 2020-05-31 Article.tex
Update on Overleaf.

@@
 The hippocampus is a brain region known to express these representations across multiple states, including during sharp-wave ripple (SWR) events where sets of hippocampal place cells can be activated in time-compressed sequences consistent with a rapid mental traversal of a path.
-Efforts to understand these representations typically make arbitrary assumptions about the character and dynamics of the generative representations.
+Efforts to understand these representations typically make assumptions about the character and dynamics of the generative representations.
 As a result, we lack a clear picture of the nature and diversity of SWR representations.
@@
 
-Mental events do not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to those actions.
+Mental events do not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
 Consistent with this, generative representations can unfold rapidly.
@@
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous or potential future experience, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed.
+A subset of SWRs contain sequential firing similar to that seen during a previous or potential future experience, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed \cite{DavidsonHippocampalReplayExtended2009}.
 
@@
 
- Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the weighted mean or maximum of each posterior time bin and connecting each time bin with a line (cite).
+ Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the weighted mean or maximum of each posterior time bin and connecting each time bin with a line.
 For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that replays can alternate between representing a single location and sequential spatial trajectories.
 On the other hand, Stella et al.
-found that replays are more spatially continuous, following Brownian diffusion dynamics.
+\cite{StellaHippocampalReactivationRandom2019} found that replays are more spatially continuous, following Brownian diffusion dynamics.
 Both methods still used fixed time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
@@
 
-We therefore developed a state space model (cite Brown, Deng) of generative activity that achieves all of those goals.
+We therefore developed a state space model of generative activity that achieves all of those goals.
 State space models are a well-understood, well-known statistical solution to the problems described above.
@@
 We then apply our decoding algorithm to the spiking sequence in Figure \ref{1}A.
-For the first third of this sequence, a single place cell fires repeatedly.
-For the middle third of the sequence, the cells fire in sequential spatial order.
+For the first third of this sequence, a single place cell fires repeatedly, representing a single location.
+For the middle third of the sequence, the cells fire in sequential spatial order, representing a fast moving trajectory.
 For the last third, the cells fire in a incoherent spatial order.
 These firing patterns represent three different types of sequence dynamics, which we call stationary, continuous, and fragmented, respectively.
-The goal of our model is to characterize replay events in terms of a mixture of these three dynamics at every time point.
+The goal of our model is to characterize SWRs in terms of a mixture of these three dynamics at every time point.
 
-Decoding the spiking sequence dynamics requires specifying two elements: the data model and the dynamics model.
+Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the dynamics model--how the position can change over time.
 For the data model, our decoder is the same as the Bayesian decoder.
@@
 We do this by defining a state transition matrix, which defines how the latent position can change from the previous time step.
-In the stationary state, we expect latent position to not change between time steps, so we use an identity state transition matrix.
-In the continuous state, we expect the latent position to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+In the stationary dynamic, we expect latent position to not change between time steps, so we use an identity state transition matrix, which predicts the next position will be the same as the last position.
+In the continuous dynamic, we expect the latent position to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
 This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
 In our case, since we are interested in identifying replays that move at speeds much faster than the animal's speed, we set this to 6.0 cm.
-This means that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or 24.5 m/s), consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
-In the fragmented state, we expect that the latent position can move to any available position.
+This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or 24.5 m/s), consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
+In the fragmented dynamic, we expect that the latent position can move to any available position instantaneously.
 We model this using a uniform state transition matrix, which makes all positions equally likely.
 
-Finally, we specify how likely we are to stay in a dynamic state versus changing to another state (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for the entirety of the replay, with a small probability of switching to one of the other dynamics.
+Finally, we specify how likely the dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
+In order to be conservative, we assume that each dynamic is likely to dominate for the duration of the SWR, with a small probability of switching to one of the other dynamics.
 Therefore, we set the probability of staying in a state to 0.98 for each 2 ms time step.
@@
 We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state.
-We summarize the resulting posterior with two quantities: the mixture probability of each state over time and the probability of position over time (Figure \ref{1}F, left and right plot respectively).
+With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time and the probability of position over time, irrespective of dynamic (Figure \ref{1}F, left and right plot respectively).
 From these two summaries, we can see that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
-The probability of the stationary state has high probability at the beginning of the simulated replay, reflecting the consistent spiking from one cell.
-This gives way to the continuous state, reflecting the trajectory-like spiking from the population of cells.
-Then the fragmented state dominates for the last third when the population spiking is not spatially coherent.
+The probability of the stationary dynamic has high probability at the beginning of the simulated replay, reflecting the consistent spiking from one cell.
+This gives way to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
+Then the fragmented dynamic dominates for the last third when the population spiking is not spatially coherent.
 We can also see that these three dynamics are reflected in the estimate of latent position of the animal.
@@
 One way to interpret the probabilities in our model is as a classifier of speeds of replay.
-To demonstrate this, we applied the model to simulated replays of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
+To demonstrate this, we applied the model to simulated trajectories of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
 From this, we can see that not only are there regions of speed that correspond to our three dynamics being high probable (where we define highly probable to be over 0.8 probability), there are in-between speeds where two of the dynamics dominate with high probability; that is, the sum of two of the dynamics' probabilities is over 0.8.
@@
 
-Next, we wanted to validate our model on real hippocampal data.
+Next, we wanted to validate the model on real hippocampal data.
 We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with clear sequential population activity (Figure \ref{2}A, top panel).
-We observed that the probability of being in the continuous state is high throughout for this SWR, with the probability of being in a stationary state being slightly higher at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
-Using our speed classification, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+We observed that the probability of being in the continuous dynamic is high throughout for this SWR, with the probability of being in a stationary state being slightly higher at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
 This is also evident in the posterior probability of linear position over time.
@@
 
-One problem with using spikes from cells identified using spike sorting is that there is potentially a lot of information lost by determining the identity of each cell---many spikes with position information are discarded because the spikes could not be assigned to one cell.
-Therefore, we wanted to use a method which directly relates multiunit spike waveform features to position without spike sorting, which we call clusterless decoding.
-Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016}, Kay et al.
-2020).
+One problem with using spikes from cells identified using spike sorting is that there is potentially a lot of information lost by determining the identity of each cell---many spikes are discarded because the spikes could not be uniquely assigned to one cell.
+Therefore, we wanted to use a method which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding.
+Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}).
 Here, if we apply it to the same SWR as the sorted spikes, we get similar decoding and classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure 2-Supplemental 1B).
@@
 After validating our model on a single SWR, we then applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009}).
-We observed many replays that were classified as continuous throughout the entirety of SWR and likely would have been identified by the most commonly used Bayesian Decoder (Figure 2-Supplemental 2A-C).
+We observed many replays that were classified as continuous throughout the entirety of SWR and likely would have been identified by the typical Bayesian Decoder (Figure 2-Supplemental 2A-C).
 However, we also observed replays that did not have this structure, but could be interpreted in terms of our model.
 For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that stayed in one position (Figure \ref{3}B), trajectories that jumped between arms (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that remained spatially incoherent throughout the SWR (Figure3-Supplemental1D).
-Overall, we were able to classify XX\% of ripples as containing at least one of the five dynamics.
-To verify that this depended on the place field coding of hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
-There were far fewer SWR classified when spatial information was removed (p=0.XX for epoch 1, p=0.XX for epoch 2, Figure 3-Supplemental 2).
+Overall, we were able to classify 89\% ( 23382/26160) of SWRs as containing at least one of the five dynamics.
+To verify that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
+There were far fewer SWR classified when spatial information was removed (XX fewer SWRs, CI [], p=0.XX for epoch 1, p=0.XX for epoch 2, Figure 3-Supplemental 2).
 
-To better understand the distribution of dynamics during SWRs, we then asked what percentage of classified replays contained spatially coherent structure---which we define as a SWR containing any of stationary, stationary-continuous mixture, or continuous categories.
-We found that XX\% of SWR had some spatially coherent structure while XX\% had some spatially incoherent structure.
+To better understand the distribution of dynamics during SWRs, we then asked what percentage of classified replays contained spatially coherent content---which we define as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics.
+We found that 22478 of 23382 or 96\% of SWR had some spatially coherent structure while 3353 of 23382 or 14\% had some spatially incoherent structure.
 This was consistent across animals (Figure \ref{3}F).
-We wondered what percentage of SWR contained continuous content, like would typically be analyzed when using by the Bayesian decoder.
-We found that, consistent with previous reports, only XX\% had continuous content (Figure \ref{3}G).
+We then asked what percentage of SWR contained continuous content, like would typically be analyzed when using the Bayesian decoder.
+We found that, consistent with previous reports, only 4511 of  23382 or 19\% of SWRs had some continuous content (Figure \ref{3}G).
 This highlights that much of the SWR replay content is missed when analyzing SWRs with the Bayesian decoder.
 
-\subsection*{Many replay trajectories are closer to the animal's movement speed}
+To verify that partitioning the data into spatially coherent and incoherent time segments by dynamic was meaningful, we quantified the cumulative spatial extent of the 95\% highest posterior density (HPD) values of the posterior linear position.
+Larger values of the cumulative HPD indicate the model is less certain about position, because the HPD values are spread out over more of the track at a given time point, whereas lower values indicate that the model is more certain about the estimate of position, because the HPD values are more concentrated and cover less of the track.
+For example, Figure \ref{4}A and B show two SWRs that were classified as having stationary and continuous dynamics, respectively.
+The bulk of the HPD values at each time step in these SWRs is concentrated on a small portion of the track and so the cumulative HPD is low throughout the SWRs.
+In contrast, the Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD is much higher.
+This can change over the time course of a SWR, as in Figure \ref{4}D, where the cumulative HPD is small for most of the SWR until the end, where the uncertainty of position becomes much higher.
+When we looked at the average cumulative HPD for each SWR by dynamic, we found a clear bi-modal split between spatially congruent dynamics and spatially incongruent dynamics (Figure \ref{4}E).
+For the spatially congruent dynamics, the cumulative HPD was much lower than the spatially incongruent dynamics (median XX cm, spatially congruent vs.
+median XX cm, spatially incongruent).
+
+## Reminder: need colorbar and legend
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v2}
+\caption{95 \% Highest Posterior Density of the Dynamics.
+\textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density.
+Top panel.
+Probability of dynamic over time.
+Shading and labels indicates dynamic categories.
+2nd panel.
+Posterior linear position over time.
+Red triangle indicates animal's position.
+3rd panel.
+Black indicates the position bin is in the 95\% highest posterior density (HPD) values.
+4th panel.
+Cumulative HPD.
+\textbf{(E)} Average Cumulative 95\% HPD for each dynamic category.}
+\label{4}
+\end{figure*}
+
+\subsection*{Many replay trajectories have slower movement speeds}
 \begin{figure*}%[tbhp]
@@
 }
-\label{4}
+\label{5}
 \end{figure*}
 
-We then looked at which sets of dynamics are most common.
-Many, but not all, of the SWR ripples were well described by one dynamic (XX/YY, XX\%).
-Surprisingly, the most common SWRs were ones that were categorized exclusively as stationary-continuous mixtures (XX\%, Figure \ref{4}A).
-These trajectories had similar duration to the continuous events (median XX ms, Figure \ref{4}B) and occurred away from the animal (median XX cm, Figure \ref{4}C).
-This indicates that the trajectories of these SWRs do not represent the animal's position and move at a speed that is closer to the animal's actual movement speed when running, rather than many times the animal's movement speed.
+We next looked at which sets of dynamics are most common.
+Most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by one dynamic.
+Surprisingly, the most common SWRs were ones that were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
+These trajectories had a slightly shorter duration to events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+continuous 95 ms, Figure \ref{5}B) and occurred at a similar average distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
+continuous 42 cm, Figure \ref{5}C).
+This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's position and move at a speed that is closer to the animal's actual movement speed when running, rather than many times the animal's movement speed.
 
-The next most common category is those SWRs that were exclusively stationary (XX\%).
-Unlike the Stationary-Continuous Mixtures, most of these trajectories represented the animal's position (Figure \ref{4}C), but there were many that represented positions as far as XX cm from the animal.
-We call the ones that are more than 30 cm away from the animal "non-local stationary trajectories".
-There were XX/YY or XX\% non-local stationary trajectories.
-These non-local stationary trajectories represented positions all over the track, but mostly at reward wells and choice points (Figure \ref{4}D).
+The next most common category is those SWRs that were exclusively stationary (5149 / 23382 or 22\%).
+Unlike the stationary-continuous mixtures, most of these trajectories represented the animal's position (Figure \ref{5}C), but there were many that represented positions as far as 226 cm from the animal.
+Duration.
+We define the ones that are more than 30 cm away from the animal as "non-local stationary trajectories".
+There were 1585 of 23382 classified SWRs or 7\% non-local stationary trajectories and 731 of these non-local stationary trajectories were stationary throughout the duration of the SWR.
+These non-local stationary trajectories represented positions all over the track, but were most common at reward wells and choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}.
+The average multiunit firing rate during stationary dynamics was not lower than the other dynamics (Figure \ref{5}E.
+## Need to fix multiunit panel
 
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v2}
-\caption{}
-\label{Figure5}
-\end{figure*}
 \section*{Discussion}
+
+Summary
+
+Why our results are different from the Bayesian decoder
+seek to describe all SWR events, clusterless decoding, more permissive, descriptive model
+
+Challenges the notion that replays consist only of high speed events.
+
+model represents middle ground between HMM and typical Bayesian decoder.
+allows SWR to quickly be identified because of the discrete states.
+Can extend to incorporate other previously experienced environments.
+Can work in 2D.
+Can precisely identify timing of dynamics because of acausal decoding
+
+adjudicates/clarifies results of Pfiefer and Foster and Stella et al.
 

### 2020-06-08 Article.tex

@@
 For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that stayed in one position (Figure \ref{3}B), trajectories that jumped between arms (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that remained spatially incoherent throughout the SWR (Figure3-Supplemental1D).
-Overall, we were able to classify 89\% ( 23382/26160) of SWRs as containing at least one of the five dynamics.
-To verify that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
-There were far fewer SWR classified when spatial information was removed (XX fewer SWRs, CI [], p=0.XX for epoch 1, p=0.XX for epoch 2, Figure 3-Supplemental 2).
+Overall, we were able to classify 89\% ( 23382 of 26160) of SWRs as containing at least one of the five dynamics.
+To check that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
+There were far fewer SWR classified when spatial information was removed (99\% true classified, 9\% median resampled distribution, p=0.02 for epoch 1, p=0.02 for epoch 2, Figure 3-Supplemental 2).
 
@@
 When we looked at the average cumulative HPD for each SWR by dynamic, we found a clear bi-modal split between spatially congruent dynamics and spatially incongruent dynamics (Figure \ref{4}E).
-For the spatially congruent dynamics, the cumulative HPD was much lower than the spatially incongruent dynamics (median XX cm, spatially congruent vs.
-median XX cm, spatially incongruent).
+For the spatially congruent dynamics, the cumulative HPD was much lower than the spatially incongruent dynamics (median 24 cm, spatially congruent vs.
+median 238 cm, spatially incongruent).
 
-## Reminder: need colorbar and legend
-\begin{figure*}%[tbhp]
+\begin{figure*}%[tb2]
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v2}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v5}
 \caption{95 \% Highest Posterior Density of the Dynamics.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v4}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v3}
 \caption{
@@
 We next looked at which sets of dynamics are most common.
-Most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by one dynamic.
+Most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic.
 Surprisingly, the most common SWRs were ones that were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
@@
 
-The next most common category is those SWRs that were exclusively stationary (5149 / 23382 or 22\%).
+The next most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\%).
 Unlike the stationary-continuous mixtures, most of these trajectories represented the animal's position (Figure \ref{5}C), but there were many that represented positions as far as 226 cm from the animal.
-Duration.
 We define the ones that are more than 30 cm away from the animal as "non-local stationary trajectories".
@@
 The average multiunit firing rate during stationary dynamics was not lower than the other dynamics (Figure \ref{5}E.
-## Need to fix multiunit panel
 
@@
 The behavioral task is a W-track spatial alternation task as presented in \cite{KarlssonAwakereplayremote2009}.
-For each day, animals alternated between 15 minute sessions in a rest box and 15 minutes in the W-track.
-The behavioral task is a W-track spatial alternation task as presented in \cite{KarlssonAwakereplayremote2009}.
-For each day, animals alternated between 15 minute sessions in a rest box and 15 minutes in the W-track.
-Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
-
-Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
+In brief, each day, animals alternated between 15 minute rest sessions in a rest box and 15 minutes run session in the W-track.
+On the W-track, animals had to run: inbound and outbound.
+On outbound runs, the Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
 
@@
 Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's projected position on the track graph (see Linearization) and the MAP estimate of the joint posterior marginalized over state, which is defined as the center of the position bin with the greatest posterior value.
+
+\subsection{Highest Posterior Density}
 

### 2020-07-01 Article.tex
Update on Overleaf.

@@
 \usepackage{graphicx}
-
+\usepackage{mathtools}
+\usepackage{blkarray, bigstrut}
+\usepackage{bbold}
 \DeclareMathOperator*{\E}{\mathbb{E}}
@@
 
-\title{A state space model for characterizing the trajectory dynamics of replay}
+\title{A state space model reveals the rich and varied dynamics of replay}
 \shorttitle{Replay Dynamics}
@@
 \begin{abstract}
-Representations of past or possible future experiences play a critical role in decision-making processes.
+Representations of past or possible future experiences play a critical role in memory and decision-making processes.
 The hippocampus is a brain region known to express these representations across multiple states, including during sharp-wave ripple (SWR) events where sets of hippocampal place cells can be activated in time-compressed sequences consistent with a rapid mental traversal of a path.
-Efforts to understand these representations typically make assumptions about the character and dynamics of the generative representations.
-As a result, we lack a clear picture of the nature and diversity of SWR representations.
-Here we develop a flexible state space model that uses a combination of discrete and latent states to describe the content and time evolution of SWR events.
-Application of this model reveals that a much larger proportion of SWR events contain coherent spatial content that previously reported, and that the dynamics of events span a surprisingly wide range.
+Efforts to understand these "replay" events typically analyze a small subset of events and assume rapid mental movement, and as a result we lack a clear picture of both the prevalence and the range of dynamics associated with replay.
+Here we develop a flexible state space model that uses a combination of discrete and latent states to capture the content and time evolution of SWR events.
+Application of this model reveals the large majority of SWR events contain coherent spatial content that evolves in accordance with a surprisingly wide range of mental movement speeds.
 
@@
 The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
-These internal "generative" representations can be identified because the underlying neural activity patterns resemble those seen during actual events.
+These internal "generative" representations \cite{KayConstantSubsecondCycling2020} can be identified because the underlying neural activity patterns resemble those seen during actual events.
 In particular, an encoding model that relates neural firing to external covariates can be constructed during an experience, and this model can inverted to decode an internal representation of the covariates from the neural activity at a different time.
 
-Mental events do not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
+Mental events may not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
 Consistent with this, generative representations can unfold rapidly.
-As a result, decoding generative representations requires making assumptions about their timescale and temporal structure, and these assumptions can limit the interpretation of the data.
+As a result, decoding generative representations often involves  making assumptions about their timescale and temporal structure, and these assumptions can limit the interpretation of the data.
 
@@
 \subsection*{Overview of the model}
-In order to introduce and validate the model, we demonstrate how the model works on a simulated dataset (Figure \ref{1}).
+We begin with an application of the model to simulated data, both to validate the model and to provide intuition.
+(Figure \ref{1}).
 We simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
@@
 For the last third, the cells fire in a incoherent spatial order.
-These firing patterns represent three different types of sequence dynamics, which we call stationary, continuous, and fragmented, respectively.
+These firing patterns represent three different types of sequence dynamics that could be expressed during replay events, which we call stationary, continuous, and fragmented, respectively.
 The goal of our model is to characterize SWRs in terms of a mixture of these three dynamics at every time point.
@@
 This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the Bayesian decoder, we use small time bins (2 ms vs.
-20 ms) because we are able to take advantage of the prior placed on the dynamics by the state space model.
+In contrast to the Bayesian decoder, we can use arbitrarily small time bins (in our case 2 ms vs.
+20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
 This allows us to detect changes on smaller time scales than would be possible with the Bayesian decoder.
 
-Next, we specify a dynamics model for how latent position---the "mental" position of the animal represented by the cells---can evolve in each of these sequences (Figure \ref{1}C).
+Next, we specify a dynamics model for how latent position---the "mental" position of the animal represented by the cells--- evolves over time (Figure \ref{1}C).
 We do this by defining a state transition matrix, which defines how the latent position can change from the previous time step.
-In the stationary dynamic, we expect latent position to not change between time steps, so we use an identity state transition matrix, which predicts the next position will be the same as the last position.
-In the continuous dynamic, we expect the latent position to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+Previous findings suggest three distinct dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, smoothly moving \cite{DavidsonHippocampalReplayExtended2009}, and fragmented / jumpy \cite{PfeifferAutoassociativedynamicsgeneration2015}, which could correspond to both spatially incoherent representations and representations that jump from one position to another.
+We therefore developed state transition models to capture each of these dynamics.
+
+In the stationary dynamic, the latent position would not change between time steps, so we use an identity state transition matrix, which predicts the next position will be the same as the last position.
+In the continuous dynamic, the latent position would be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
 This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
 In our case, since we are interested in identifying replays that move at speeds much faster than the animal's speed, we set this to 6.0 cm.
-This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or 24.5 m/s), consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
-In the fragmented dynamic, we expect that the latent position can move to any available position instantaneously.
+This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (corresponding to a latent speed of up to ~25 m/s).
+This is consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
+Finally, in the fragmented dynamic, the latent position could move to any available position instantaneously.
 We model this using a uniform state transition matrix, which makes all positions equally likely.
@@
 Finally, we specify how likely the dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for the duration of the SWR, with a small probability of switching to one of the other dynamics.
-Therefore, we set the probability of staying in a state to 0.98 for each 2 ms time step.
+In order to be conservative, we assume that each dynamic is likely to dominate for the duration of the SWR, with a small probability of switching to one of the other dynamics (note that  below we show that our results are relatively insensitive to this parameter).
+Accordingly, we set the probability of staying in a state to 0.98 for each 2 ms time step.
 Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular state.
@@
 Once we have specified the data and dynamics model, we have fully specified the state space model.
-We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state.
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state (see methods).
 With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time and the probability of position over time, irrespective of dynamic (Figure \ref{1}F, left and right plot respectively).
-From these two summaries, we can see that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
-The probability of the stationary dynamic has high probability at the beginning of the simulated replay, reflecting the consistent spiking from one cell.
-This gives way to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
-Then the fragmented dynamic dominates for the last third when the population spiking is not spatially coherent.
-We can also see that these three dynamics are reflected in the estimate of latent position of the animal.
-Importantly, we get a much less variable estimate due to the correct specification of the dynamics.
 
-One way to interpret the probabilities in our model is as a classifier of speeds of replay.
+An examination of these two summaries reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+The stable firing of the one active neuron indicates a stationary representation, and the probability of the stationary dynamic is high at the beginning of the simulated replay.
+The change in the character of the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
+Subsequently, as the population activity becomes spatially incoherent, and as a result, the fragmented dynamic dominates for the last third.
+Importantly, these dynamics also inform the estimate of latent position of the animal, and the correct assignment of dynamic reduces the variance of the latent position estimate.
+
+The specification of these dynamics allows the model to capture a wide range of movement speeds for the latent position.
 To demonstrate this, we applied the model to simulated trajectories of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
-From this, we can see that not only are there regions of speed that correspond to our three dynamics being high probable (where we define highly probable to be over 0.8 probability), there are in-between speeds where two of the dynamics dominate with high probability; that is, the sum of two of the dynamics' probabilities is over 0.8.
+From this, we can see that not only are there regions of speed that correspond to our three dynamics being high probable (where we define highly probable to be over 0.8 probability), there are intermediate speeds where two of the dynamics dominate with high probability; that is, the sum of two of the dynamics' probabilities is over 0.8.
 In this manuscript, we will refer to these as mixture dynamics.
@@
 
-In order to verify that the model was robust to the choice of probability of staying in a state, we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
+We note here that the choice of any particular threshold is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
+Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95 because the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are likely to be meaningful to the brain.
+Nonetheless, we also verified that our central results hold with a higher threshold of 0.95.
+
+As mentioned above, the model is robust to the choice of probability of staying in a state.
+We decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
 We found that for a large range of plausible probabilities of staying in one of the dynamics (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
@@
 \caption{Demonstration of the state space model on real data using both sorted and clusterless spikes on the same SWR.
-\textbf{(A)} Decoding using sorted spikes (arranged top to bottom).
-In the top panel, 31 cells on a W-track ordered according to linearized position by their place field peaks.
+\textbf{(A)} Decoding using sorted spikes.
+The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
 The second panel shows the probability of each dynamic over time as in Figure 1F.
@@
 The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
-L, R, C correspond to the position of the left, right and center wells respectively.
-The animal's position is indicated by the red triangle.
+L, R, C correspond to the position of the left, right and center reward wells respectively.
+The animal's actual position is indicated by the red triangle.
 \textbf{(B)} Decoding using clusterless spikes.
@@
 Color indicates time.
-The animal's position is denoted by the red circle.
+The animal's actual position is denoted by the red circle.
 Light grey lines show the animal's 2D position over the entire epoch.
-L, R, and C correspond to the left, right and center well as in (A).
+L, R, and C correspond to the left, right and center reward well as in (A).
 \textbf{(D)} 1D MAP estimate of the latent position in (B) projected back into 2D.
@@
 \end{figure*}
-\subsection*{The model finds known continuous replays with sorted and clusterless spikes}
+\subsection*{Identification of continuous replays with sorted and clusterless spikes}
 
@@
 
-Next, we wanted to validate the model on real hippocampal data.
-We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with clear sequential population activity (Figure \ref{2}A, top panel).
-We observed that the probability of being in the continuous dynamic is high throughout for this SWR, with the probability of being in a stationary state being slightly higher at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+Our success with simulated data inspired a further validation of the model on real hippocampal data.
+We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with sequential population activity (Figure \ref{2}A, top panel).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this SWR, but the probability of being in a stationary state was noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
 Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
@@
 
-One problem with using spikes from cells identified using spike sorting is that there is potentially a lot of information lost by determining the identity of each cell---many spikes are discarded because the spikes could not be uniquely assigned to one cell.
-Therefore, we wanted to use a method which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding.
+One of our criteria for a more optimal method is that it use all of the available spiking data.
+Using only clustered spikes discards all of the spike events that cannot be uniquely assigned to a putative single neuron, reducing very substantially the quality of the resulting decoding.
+We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding (see Methods).
 Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}).
-Here, if we apply it to the same SWR as the sorted spikes, we get similar decoding and classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure 2-Supplemental 1B).
-Note that, using clusterless decoding, the estimate of the posterior probability of position is much more narrow, indicating that there is more certainty in the true latent position at that time compared the corresponding posterior probability of position using sorted spikes (Figure \ref{2}D vs C).
+Applying a clusterless decoder to the same SWR, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure 2-Supplemental 1B).
+The spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model, however.
+This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event, and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).
 
-\subsection*{The model finds replays with non-constant speed dynamics}
-After validating our model on a single SWR, we then applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009}).
-We observed many replays that were classified as continuous throughout the entirety of SWR and likely would have been identified by the typical Bayesian Decoder (Figure 2-Supplemental 2A-C).
-However, we also observed replays that did not have this structure, but could be interpreted in terms of our model.
-For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that stayed in one position (Figure \ref{3}B), trajectories that jumped between arms (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that remained spatially incoherent throughout the SWR (Figure3-Supplemental1D).
-Overall, we were able to classify 89\% ( 23382 of 26160) of SWRs as containing at least one of the five dynamics.
-To check that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, breaking the relationship between spiking and spatial information.
-There were far fewer SWR classified when spatial information was removed (99\% true classified, 9\% median resampled distribution, p=0.02 for epoch 1, p=0.02 for epoch 2, Figure 3-Supplemental 2).
+\subsection*{Replays with non-constant speed}
 
-To better understand the distribution of dynamics during SWRs, we then asked what percentage of classified replays contained spatially coherent content---which we define as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics.
-We found that 22478 of 23382 or 96\% of SWR had some spatially coherent structure while 3353 of 23382 or 14\% had some spatially incoherent structure.
+After testing our model on a single SWR, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020}).
+One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see methods) to ensure that we were included both the rare large events as well as the much more common small events.
+
+As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of SWR (Figure 2-Supplemental 2A-C).
+However, we also observed many events with spatially coherent content that did not have this structure.
+For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure3-Supplemental1D).
+
+ Using our 0.8 threshold, we were able to classify 89\% ( 23382 of 26160) of SWRs as containing at least one of the five dynamics.
+To ensure that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, shuffling the relationship between spiking and spatial information.
+We then decoded the same SWR events containing the same spikes.
+Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for epoch 1, p=0.02 for epoch 2, Figure 3-Supplemental 2).
+
+Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content [REFS].
+We therefore asked what fraction of events contained spatially coherent content in any form, defined as containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see methods).
+We found that 22478 of 23382 or 96\% of SWRs had some spatially coherent structure while 3353 of 23382 or 14\% had some spatially incoherent structure.
 This was consistent across animals (Figure \ref{3}F).
 We then asked what percentage of SWR contained continuous content, like would typically be analyzed when using the Bayesian decoder.
-We found that, consistent with previous reports, only 4511 of  23382 or 19\% of SWRs had some continuous content (Figure \ref{3}G).
-This highlights that much of the SWR replay content is missed when analyzing SWRs with the Bayesian decoder.
+Here our findings were consistent with previous reports: in our datasets 4511 of  23382 or 19\% of SWRs had some continuous content (Figure \ref{3}G).
+Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent but not continuous content.
 
-To verify that partitioning the data into spatially coherent and incoherent time segments by dynamic was meaningful, we quantified the cumulative spatial extent of the 95\% highest posterior density (HPD) values of the posterior linear position.
+Our model is specified in the context of a marked point process associated with different dynamics.
+This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time.
+To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) values of the posterior linear position.
 Larger values of the cumulative HPD indicate the model is less certain about position, because the HPD values are spread out over more of the track at a given time point, whereas lower values indicate that the model is more certain about the estimate of position, because the HPD values are more concentrated and cover less of the track.
-For example, Figure \ref{4}A and B show two SWRs that were classified as having stationary and continuous dynamics, respectively.
-The bulk of the HPD values at each time step in these SWRs is concentrated on a small portion of the track and so the cumulative HPD is low throughout the SWRs.
-In contrast, the Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD is much higher.
-This can change over the time course of a SWR, as in Figure \ref{4}D, where the cumulative HPD is small for most of the SWR until the end, where the uncertainty of position becomes much higher.
-When we looked at the average cumulative HPD for each SWR by dynamic, we found a clear bi-modal split between spatially congruent dynamics and spatially incongruent dynamics (Figure \ref{4}E).
+Thus, the HPD provides a complementary measures of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
+
+We found that spatially coherent events did indeed have lower cumulative HPDs.
+Figure \ref{4}A and B show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
+The bulk of the HPD values at each time step in these SWRs is concentrated in a small portion of the track and  the cumulative HPD is low throughout the SWRs.
+In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD is much higher.
+The HPD also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
+An example of this change is shown in Figure \ref{4}D, where the cumulative HPD is small for most of the SWR until the end, where the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examined the average cumulative HPD for each SWR, grouped by dynamic, we found a clear bi-modal split between spatially congruent dynamics and spatially incongruent dynamics (Figure \ref{4}E).
 For the spatially congruent dynamics, the cumulative HPD was much lower than the spatially incongruent dynamics (median 24 cm, spatially congruent vs.
@@
 
-We next looked at which sets of dynamics are most common.
-Most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic.
-Surprisingly, the most common SWRs were ones that were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These trajectories had a slightly shorter duration to events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
-continuous 95 ms, Figure \ref{5}B) and occurred at a similar average distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
+Our results indicate that most events with coherent spatial content are not continuous.
+We therefore examined these events in more details.
+We first noted that most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic.
+Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
+These events contain representations that move at rates consistent with real-world movement (Fig.
+1G), but were slightly shorter, on average,
+than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+continuous 95 ms, Figure \ref{5}B).
+Nonetheless, both these slow-moving events and continuous events  represented locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
 continuous 42 cm, Figure \ref{5}C).
-This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's position and move at a speed that is closer to the animal's actual movement speed when running, rather than many times the animal's movement speed.
+This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position.
 
 The next most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\%).
-Unlike the stationary-continuous mixtures, most of these trajectories represented the animal's position (Figure \ref{5}C), but there were many that represented positions as far as 226 cm from the animal.
-We define the ones that are more than 30 cm away from the animal as "non-local stationary trajectories".
-There were 1585 of 23382 classified SWRs or 7\% non-local stationary trajectories and 731 of these non-local stationary trajectories were stationary throughout the duration of the SWR.
-These non-local stationary trajectories represented positions all over the track, but were most common at reward wells and choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}.
-The average multiunit firing rate during stationary dynamics was not lower than the other dynamics (Figure \ref{5}E.
+Unlike the stationary-continuous mixtures, most of these events represented a location close to the the animal's actual position (Figure \ref{5}C).
+There were, however, a number of stationary events  that represented positions far from the animal.
+We set a threshold of 30 cm distance from the animal to identify "non-local stationary trajectories" and found these trajectories in 1585 of 23382 (~7\%) of classified SWRs.
+About half of these (731)  were stationary throughout the duration of the SWR.
+These non-local stationary trajectories most commonly represented  reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}.
+Finally, we examined the overall activity levels across dynamics and found similar total spiking rates, indicating ...
+
+ (Figure \ref{5}E.
 
@@
 
+relationship to Johnson and Redish 2008
+
+extension to previously experienced tracks
+
 \begin{acknowledgements}
@@
 \section*{Methods}
+
+\subsection*{Simulated Data}
+Encoding data for Figure \ref{1} and Figure \ref{fig:Figure1-Figure supplement 1} was generated by simulating 15 traversals of a 180 cm linear track.
+The spiking of 19 neurons was simulated using an inhomogeneous Poisson process where the instantaneous firing rate of each neuron changes according to a Gaussian place field.
+The Gaussian place fields had a variance of 36 cm and were spaced at 10 cm intervals over the 180 cm track.
+The decoding data for Figure \ref{1}G was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
+Each neuron spiked when the traversal passed through the peak location of its place field.
+
 \subsection*{Recording Locations and Techniques}
 Ten male Long Evans rats (500-700 g, 4-9 months old) were trained on a W-track spatial alternation task.
-XX rats contributed to a previous study \cite{KarlssonAwakereplayremote2009}.
+9 rats contributed to previous studies \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020}.
 Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
@@
 \subsection*{Behavioral Task}
-The behavioral task is a W-track spatial alternation task as presented in \cite{KarlssonAwakereplayremote2009}.
-In brief, each day, animals alternated between 15 minute rest sessions in a rest box and 15 minutes run session in the W-track.
-On the W-track, animals had to run: inbound and outbound.
-On outbound runs, the Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
+All animals performed a W-track spatial alternation task, which is described in detail in \cite{KarlssonAwakereplayremote2009}.
+In brief, each day, animals alternated between 15 minute rest epochs in a rest box and 15 minutes run epochs in the W-track.
+On the W-track, animals alternated between two types of runs: inbound and outbound.
+On inbound runs, the animal had to run from an outer arm to the center arm.
+On outbound runs, the animals had to run from the center arm to the least recently visited outer arm.
+Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
 
-\subsection*{Linearization}
-To decrease the time it takes to run the model, the 2D position of the animal is converted into a 1D position.
-This is done by first defining a graph representation of the track, where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
-Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to classify which track segment the animal is on.
-This algorithm was developed to track cars on roads with GPS signals.
-Using the HMM prevents sudden jumps from one track segment to another, which is particularly important near intersections when we are tracking the head position.
-Briefly, the observation model of the HMM is Gaussian and it models the likelihood of being on a track segment as the Gaussian distance to that segment with a 5 cm standard deviation.
-The state transition model is an empirically estimated state transition that changes with each time point that tries to ensure that the euclidean distance between successive position estimates is similar to the shortest path distance along the graph between successive position estimates.
+\subsection*{Position of the Animal and Linearization}
+The animal's 2D position was estimated from video of 2 LEDs placed on the head stage.
+In order to decrease the time it takes to run the model, the 2D position of the animal was converted into a 1D position.
+This is done by first defining a 2D graph representation of the track, where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
+Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to classify the most likely track segment the animal is on.
+Using the HMM takes into account the time dependent nature of the data and helps prevents sudden jumps from one track segment to another, which is particularly important near intersections.
+The observation model of the HMM is Gaussian and it models the likelihood of being on a track segment as the Gaussian distance to that segment with a 5 cm standard deviation.
+The state transition model is an empirically estimated state transition that changes with each time point that tries to ensure that the Euclidean distance between successive position estimates is similar to the shortest path distance along the graph between successive position estimates.
 A slight bias of 0.1 is given to the diagonal to encourage staying on the same track segment.
@@
 Sharp wave ripples were detected using the same method as in \cite{Kayhippocampalnetworkspatial2016}.
-Briefly, each LFP was filtered between 150-250 Hz, squared and then summed across tetrodes--forming a single population trace over time.
+Briefly, each CA1 LFP was filtered between 150-250 Hz, squared and then summed across tetrodes---forming a single population trace over time.
 This trace was smoothed with a Gaussian with a 4 ms standard deviation and the square root of this trace was taken to get an estimate of the population ripple band power.
@@
 
-\subsection*{Spike Sorting}
+\subsection*{The Model}
+Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and $I_{k}$ be a discrete latent variable that is an indicator for the dynamics we wish to characterize: stationary, continuous, and fragmented.
+The goal of the model is to simultaneously estimate the posterior probability of both variables $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed data for for all times from 1 to $T$, the final time.
+The observed data can be either spike trains $\Delta N_{1:T}^{(1:C)}$ from $C$ putative cells when decoding with sorted spikes or multiunit spikes $\Delta N_{1:T}^{(1:E)}$ and their associated wave form features $\Vec{m}$ from each tetrode $E$ when decoding with clusterless spikes.
 
-\subsection*{Simulated Data}
+We have previously shown [Asilomar REF] that the posterior $p(x_k, I_k \mid O_{1:T})$ can be directly estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$:
 
-\subsection*{The Model}
+$$
+p(x_{k}, I_{k} \mid O_{1:k}) \propto
+p(O_{k}  \mid x_{k}, I_{k}) \sum_{I_{k-1}} \int p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) 
+Pr(I_{k} \mid I_{k-1}) p(x_{k-1}, I_{k-1} \mid O_{1:k-1}) dx_{k-1}
+$$
+and then applying the acausal smoother equation, starting from the last estimate of the casual filter $p(x_{T}, I_{T} \mid O_{1:k})$ and iterating backwards to time 1:
+$$
+p(x_{k}, I_{k} \mid O_{1:T}) =
+p(x_{k}, I_{k} \mid O_{1:k})
+\sum_{I_{k+1}} \int \frac{p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k})
+Pr(I_{k+1} \mid I_{k})}{p(x_{k+1}, I_{k+1} \mid O_{1:k})}
+p(x_{k+1}, I_{k+1} \mid O_{1:T}) dx_{k+1}
+$$
+where:
+$$
+p(x_{k+1}, I_{k+1} \mid O_{1:k}) =
+\sum_{I_{k}} \int p(x_{k+1} \mid x_{k}, I_{k+1}, I_{k}) Pr(I_{k+1} \mid I_{k})
+p(x_{k}, I_{k} \mid O_{1:k}) dx_{k}
+$$
 
-\subsection*{Encoding - Clusterless}
-In order to encode how each tetrode's unsorted spiking activity relates to position, we use a marked point process framework.
-For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features (the marks) and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
-In our case, the waveform features correspond to the max amplitude observed on each tetrode wire at the time of the multiunit spike.
-This array is used as the training samples in a kernel density estimator during the decoding step.
+Therefore, to specify the model, we have to define or estimate the following quantities:
+\begin{enumerate}
+  \item $p(x_{0}, I_{0})$ - the initial conditions
+  \item $Pr(I_{k} \mid I_{k-1})$ - the dynamics transition matrix
+  \item $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$ - the dynamics movement model
+  \item $p(O_{k}  \mid x_{k}, I_{k})$ - the likelihood of the observations
+\end{enumerate}
+
+For the initial conditions $p(x_{0}, I_{0})$, we set each dynamic $I_0$ to have uniform probability over all possible positions $\mathcal{U}(\min x, \max x)$, reflecting the fact that we do not have any prior knowledge about which dynamic or position is more likely:
+
+\begin{equation*}
+  p(x_{0}, I_{0}) = 
+  \begin{blockarray}{*{1}{c} l}
+    \begin{block}{[*{1}{c}]>{$\footnotesize}l<{$}}
+      \mathcal{U}(\min x, \max x) \bigstrut[t]& $I_{0} = Stationary$ \\
+      \mathcal{U}(\min x, \max x) & $I_{0} = Continuous$ \\
+      \mathcal{U}(\min x, \max x) & $I_{0} = Fragmented$ \\
+    \end{block}
+  \end{blockarray}
+\end{equation*}
+
+We set the dynamics transition matrix $Pr(I_{k} \mid I_{k-1})$, which defines how likely the dynamic is to change to another dynamic versus persist in the same dynamic, to be:
+
+\begin{equation*}
+  Pr(I_{k} \mid I_{k-1}) = 
+  \begin{blockarray}{*{3}{c} l}
+    \begin{block}{*{3}{>{$\footnotesize}c<{$}} l}
+      $I_{k} = Stationary$ & $I_{k} = Continuous$ & $I_{k} = Fragmented$ \\
+    \end{block}
+    \begin{block}{[*{3}{c}]>{$\footnotesize}l<{$}}
+      0.98 & 0.01 & 0.01 \bigstrut[t]& $I_{k-1} = Stationary$ \\
+      0.01 & 0.98 & 0.01 & $I_{k-1} = Continuous$ \\
+      0.01 & 0.01 & 0.98 & $I_{k-1} = Fragmented$ \\
+    \end{block}
+  \end{blockarray}
+\end{equation*}
+
+to encode the prior expectation that each of the dynamics will last the average duration of 100 ms, with a small probability of changing to one of the other dynamics.
+
+We set the dynamics movement model $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$, which defines how likely the position $x_{k}$ is to change given the previous position $x_{k-1}$ and current $I_{k}$ and past dynamics $I_{k-1}$, to be:
+
+\begin{equation*}
+  p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1}) = 
+  \begin{blockarray}{*{3}{c} l}
+    \begin{block}{*{3}{>{$\footnotesize}c<{$}} l}
+      $I_{k} = Stationary$ & $I_{k} = Continuous$ & $I_{k} = Fragmented$ \\
+    \end{block}
+    \begin{block}{[*{3}{c}]>{$\footnotesize}l<{$}}
+      \delta(x_{k-1}) & \mathcal{N}(x_{k-1}, 6.0) & \mathcal{U}(\min x, \max x) \bigstrut[t]& $I_{k-1} = Stationary$ \\
+      \delta(x_{k-1}) & \mathcal{N}(x_{k-1}, 6.0)  & \mathcal{U}(\min x, \max x) & $I_{k-1} = Continuous$ \\
+      \mathcal{U}(\min x, \max x) & \mathcal{U}(\min x, \max x) & \mathcal{U}(\min x, \max x) & $I_{k-1} = Fragmented$ \\
+    \end{block}
+  \end{blockarray}
+\end{equation*}
+where $\delta(x_{k-1})$ is an identity transition matrix where position cannot change from the previous time step, $\mathcal{N}(x_{k-1}, 6.0)$ is a random walk from the previous position with variance 6.0, and $\mathcal{U}(\min x, \max x)$ is a uniform transition that allows transitions to any possible position.
+
+As discussed in the Results, this says that when persisting in the same dynamic, the stationary, continuous, and fragmented dynamics are defined by the identity transition, the random walk, and the uniform transition, respectively.
+When transitioning to or from the fragmented dynamic, we assume we do not have any information about the position, so the transition is uniform.
+Finally, when the transition is from the stationary to continuous, we assume the position is spatially close where it was previously, so we use a random walk.
+When the transition is from continuous to stationary, we assume that the position is no longer changing, so we use the identity transition.
+
+Lastly, we estimate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$. It has been shown [REFS, Zhang, Brown] that the likelihood with sorted spikes can be computed as:
+$$
+p(O_{k} \mid x_{k}, I_{k}) = p(\Delta N_{k}^{(1:C)} \mid x_{k}, I_{k}) \propto
+\prod^{C}_{i=1} [\lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
+    \exp[-\lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]
+$$
+where $N_{t_{k}}^{i}$ represents a spike at time $t_k$ from cell $i$, $\Delta_k$ is the time bin size, and $\lambda_{i}(t_{k} \mid x_{k}, I_{k})$ is the instantaneous firing rate of cell $i$ given position $x_{k}$ and dynamic $I_{k}$---this is the "place field" of the cell, which we estimate during the encoding step.
+
+Likewise, it has been shown \cite{ChenTransductiveneuraldecoding2012, DengRapidclassificationhippocampal2016, KloostermanBayesiandecodingusing2014} that the clusterless likelihood can be computed as: 
+$$
+p(O_{k} \mid x_{k}, I_{k}) = p(\Delta N_{1:T}^{(1:E)}, \vec{m} \mid x_{k}, I_{k}) \propto
+\prod^{E}_{i=1} [\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}, I_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
+    \exp[-\Lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]
+$$
+where $\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}, I_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k}, I_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
+Both of these quantities can be estimated by:
 
@@
 Fitting is done by maximizing the penalized likelihood using a Newton-Raphson algorithm.
+The code used to fit the GLMs is available at \url{https://github.com/Eden-Kramer-Lab/regularized_glm}.
+
+\subsection*{Encoding - Clusterless}
+
+In our case, $\Vec{m}$ corresponds to the peak amplitude of each of the four tetrode wires at the time of the multiunit spike.
+
+In order to encode how each tetrode's unsorted spiking activity relates to position, we use a marked point process framework [REFS].
+For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features (the marks) and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
+In our case, the waveform features correspond to the max amplitude observed on each tetrode wire at the time of the multiunit spike.
+This array is used as the training samples in a kernel density estimator during the decoding step.
+
+\subsection*{Decoding}
+
+position bin size 3 cm
+
+\subsection*{Decoding - Sorted Spikes}
 
 \subsection*{Decoding - Clusterless}
-
+24 mV smoothing for peak amplitude
+6.0 cm smoothing for position (1D and 2D)
+\subsection*{Mixture probability of the dynamics}
+The mixture probability of the dynamics is a quantity that indicates how much that dynamic is contributing to the posterior at a given time.
+We estimate the mixture probability of the dynamics by integrating out position the joint posterior:
+$$Pr(I_{k} \mid O_{1:T}) = \int p(x_{k}, I_{k} \mid O_{1:T}) dx_{k}$$
+As in other calculations, the integral is approximated with a Reimann sum.
+\subsection*{Posterior probability of position}
+The posterior probability of position is . We estimate it by summing over the dynamics.
+$$p(x_{k} \mid O_{1:T}) = \sum_{I_{k}} p(x_{k}, I_{k} \mid O_{1:T})$$
 \subsection*{Classification of Dynamics}
-We used a threshold of 0.8 to classify the probability of each state into 5 categories.
+We used a threshold of 0.8 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
 Time periods during sharp wave ripples are labeled as Stationary, Continuous, or Fragmented when the probability of each state is above 0.8.
 Time periods are labeled as Stationary-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Stationary and Continuous or Fragmented and Continuous are above 0.8, respectively.
-Time periods where none of these criterion are met are considered unclassified because there is not an interpretation for all three dynamics being equally likely.
+Time periods where none of these criterion are met are considered unclassified.
 
+\subsection*{Spatially Coherent Content}
+\subsection*{Spatially Incoherent Content}
+
+\subsection*{Highest Posterior Density}
+The 95\% highest posterior density (HPD) is a measurement of the uncertainty of the posterior and is defined as the region of the posterior that contains the top 95\% values of the posterior probability.
+It can be calculated by determining the maximum threshold value $h$ that fulfills the follow equality:
+$$
+\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} p(x_{k} \mid O_{1:T})dx = 0.95
+$$
+The 95\% HPD is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
+The cumulative spatial extent of the HPD is calculated by taking the integral of the members of this set:
+$$
+\int_{\{x: p(x_{k} \mid O_{1:T}) > h\}} \mathbb{1}dx
+$$
+which we approximate with a Reimann sum.
+\subsection*{Shuffle}
+\subsection*{Duration of dynamic}
 \subsection*{Replay distance from animal}
@@
 
-\subsection{Highest Posterior Density}
+\subsection*{Non-Local Stationary Position}
+\subsection*{Multiunit Rate of Each Dynamic}
 
-\subsection*{Software and Code availability}
+\subsection*{Software and Code Availability}
 Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v1}
+\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental2/Figure3_v1-supplemental2}
 \caption{Comparison of real vs.

### 2020-07-01 Article.tex
Update on Overleaf.

@@
 
-We have previously shown [Asilomar REF] that the posterior $p(x_k, I_k \mid O_{1:T})$ can be directly estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$:
+We have previously shown \cite{DenovellisCharacterizinghippocampalreplay2019} that the posterior $p(x_k, I_k \mid O_{1:T})$ can be directly estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$:
 
@@
 
-Lastly, we estimate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$. It has been shown [REFS, Zhang, Brown] that the likelihood with sorted spikes can be computed as:
+Lastly, we estimate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$.
+It has been shown \cite{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the likelihood with sorted spikes can be computed as:
 $$

### 2020-07-10 Article.tex

--- 2020-07-01 Article.tex
@@
 \usepackage{bbold}
+\usepackage{amsmath}
 \DeclareMathOperator*{\E}{\mathbb{E}}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
@@
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Moreover, there is now substantial evidence that a subset of these events express stationary representations, which are often excluded.
+Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which are often excluded.
 In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
@@
 \cite{StellaHippocampalReactivationRandom2019} found that replays are more spatially continuous, following Brownian diffusion dynamics.
-Both methods still used fixed time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
+Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
 
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches (refs) and characterizes represented trajectories as a mixture of three underlying patterns of dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} and characterizes represented trajectories as a mixture of three underlying patterns of dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike wave form features to position without spike sorting---giving us more information about the population spiking activity.
@@
 Once we have specified the data and dynamics model, we have fully specified the state space model.
-We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state (see methods).
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state (see Methods).
 With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time and the probability of position over time, irrespective of dynamic (Figure \ref{1}F, left and right plot respectively).
@@
 The stable firing of the one active neuron indicates a stationary representation, and the probability of the stationary dynamic is high at the beginning of the simulated replay.
-The change in the character of the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
-Subsequently, as the population activity becomes spatially incoherent, and as a result, the fragmented dynamic dominates for the last third.
+The change in the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
+Subsequently, as the population activity becomes spatially incoherent, the fragmented dynamic dominates for the last third.
 Importantly, these dynamics also inform the estimate of latent position of the animal, and the correct assignment of dynamic reduces the variance of the latent position estimate.
@@
 After testing our model on a single SWR, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020}).
-One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see methods) to ensure that we were included both the rare large events as well as the much more common small events.
+One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see Methods) to ensure that we were included both the rare large events as well as the much more common small events.
 
@@
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content [REFS].
-We therefore asked what fraction of events contained spatially coherent content in any form, defined as containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see methods).
-We found that 22478 of 23382 or 96\% of SWRs had some spatially coherent structure while 3353 of 23382 or 14\% had some spatially incoherent structure.
+We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see methods).
+We found that 22478 of 23382 or 96\% of classified SWRs had some spatially coherent structure.
+We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented.
+We found that 3353 of 23382 or 14\% of classified SWRs had some spatially incoherent structure.
 This was consistent across animals (Figure \ref{3}F).
-We then asked what percentage of SWR contained continuous content, like would typically be analyzed when using the Bayesian decoder.
-Here our findings were consistent with previous reports: in our datasets 4511 of  23382 or 19\% of SWRs had some continuous content (Figure \ref{3}G).
+We then asked what percentage of classified SWR contained continuous content, like would typically be analyzed when using the Bayesian decoder.
+Here our findings were consistent with previous reports: in our datasets 4511 of  23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G).
 Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent but not continuous content.
 
+We repeated this analysis with a higher threshold of 0.95 to ensure that our result was not dependent on this threshold.
+We found that, while this slightly reduced the total fraction of classified ripples (19180 / 25845 or 74\% SWRs), an even higher fraction of the classified ripples (19020 / 19180 or 99\% classified) had spatially continuous content.
+Classified SWRs containing spatially incoherent content similarly were a small fraction of the classified SWRs (spatially incoherent: 484 / 19180 or 3\%).
+
 Our model is specified in the context of a marked point process associated with different dynamics.
-This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time.
+This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
 To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) values of the posterior linear position.
@@
 
-Our results indicate that most events with coherent spatial content are not continuous.
+Our results indicate that most events with coherent spatial content are not continuous; that is, trajectories that move many times the running speed of the animal.
 We therefore examined these events in more details.
@@
 These events contain representations that move at rates consistent with real-world movement (Fig.
-1G), but were slightly shorter, on average,
-than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+1G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 We set a threshold of 30 cm distance from the animal to identify "non-local stationary trajectories" and found these trajectories in 1585 of 23382 (~7\%) of classified SWRs.
-About half of these (731)  were stationary throughout the duration of the SWR.
+About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
 These non-local stationary trajectories most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}.
-Finally, we examined the overall activity levels across dynamics and found similar total spiking rates, indicating ...
 
- (Figure \ref{5}E.
+Finally, we examined the overall activity levels across dynamics and found that all the dynamics had similar average multiunit spiking rates by tetrode (Figure \ref{5}E).
+This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures and importantly not just the continuous dynamics, were driven by sustained spiking information.
 
@@
 
-Why our results are different from the Bayesian decoder
-seek to describe all SWR events, clusterless decoding, more permissive, descriptive model
-
 Challenges the notion that replays consist only of high speed events.
 
-model represents middle ground between HMM and typical Bayesian decoder.
+adjudicates/clarifies results of Pfiefer and Foster and Stella et al. Brownian motion but couldn't tell the speeds. Pfiefer and Foster didn't know the percentage of events that acted this way
+
+Why our results are different from the Bayesian decoder
+seek to describe as many SWR events, clusterless decoding, more permissive model with different dynamics, explicit descriptive model instead of implicit noisy model.
+
+relationship to Johnson and Redish 2008, Maboudi, model represents middle ground between HMM and typical Bayesian decoder.
 allows SWR to quickly be identified because of the discrete states.
-Can extend to incorporate other previously experienced environments.
 Can work in 2D.
@@
 
-adjudicates/clarifies results of Pfiefer and Foster and Stella et al.
-
-relationship to Johnson and Redish 2008
-
-extension to previously experienced tracks
+Can extend to incorporate other previously experienced environments. Add prior on dynamics of speed
 
@@
 The decoding data for Figure \ref{1}G was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
-Each neuron spiked when the traversal passed through the peak location of its place field.
+Each simulated neuron "spiked" when the traversal passed through the peak location of its place field.
 
@@
 Ten male Long Evans rats (500-700 g, 4-9 months old) were trained on a W-track spatial alternation task.
-9 rats contributed to previous studies \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020}.
+9 rats contributed to previous studies \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, Kayhippocampalnetworkspatial2016, CarrTransientSlowGamma2012}.
 Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
@@
 All animals performed a W-track spatial alternation task, which is described in detail in \cite{KarlssonAwakereplayremote2009}.
-In brief, each day, animals alternated between 15 minute rest epochs in a rest box and 15 minutes run epochs in the W-track.
+In brief, each day, animals alternated between 20 minute rest epochs in a rest box and 15 minutes run epochs in the W-track.
 On the W-track, animals alternated between two types of runs: inbound and outbound.
@@
 On outbound runs, the animals had to run from the center arm to the least recently visited outer arm.
+Animals were rewarded for performing the task correctly.
 Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
@@
 \subsection*{Position of the Animal and Linearization}
-The animal's 2D position was estimated from video of 2 LEDs placed on the head stage.
+The animal's 2D position was estimated from digital video (30 Hz) of two infrared diodes placed on the head stage preamplifiers using a semi-automated analysis.
 In order to decrease the time it takes to run the model, the 2D position of the animal was converted into a 1D position.
@@
 The linear distance can then be converted into a linear position by assigning each track segment a position in 1D space.
+15 cm gaps were placed between the center arm, left arm, and right arms in 1D space to prevent any smoothing done in the model from bleeding over inappropriately.
 The code used for linearization can be found at \url{https://github.com/Eden-Kramer-Lab/loren_frank_data_processing}.
+
+\subsection*{Sorted Spikes, Multiunit Spikes and Waveform Features}
+To obtain the neural data used for decoding, hippocampus electrical potentials were recorded at 30 kHz, referenced to a corpus callosum tetrode, and then digitally filtered between 600 Hz and 6 kHz.
+Events were considered threshold crossing events if any one of the four tetrode wires exceeded a 60 mV threshold.
+These multiunit threshold crossing events and the associated peak amplitudes of the tetrodes wires---defined as the electrical potential value at the time of the maximum amplitude at the threshold crossing---were used as data for the clusterless version of the model.
+
+For decoding using sorted spikes, the multiunit events were processed further to assign events to putative single cells.
+Putative single cells were manually identified based on the clustering of three waveform features within a day: peak amplitude, principal components, and spike width.
+Only putative hippocampal pyramidal cells---identified based on spike width and average firing rate---were included in the analysis.
 
@@
 Sharp wave ripples were detected using the same method as in \cite{Kayhippocampalnetworkspatial2016}.
-Briefly, each CA1 LFP was filtered between 150-250 Hz, squared and then summed across tetrodes---forming a single population trace over time.
+Each CA1 LFP was obtained by downsampling the original 30 kHz electrical potential to 1.5 kHz, bandpass filtering between 0.5 Hz and 400 Hz.
+This was further bandpass filtered between 150-250 Hz for ripple band frequencies, squared and then summed across tetrodes---forming a single population trace over time.
 This trace was smoothed with a Gaussian with a 4 ms standard deviation and the square root of this trace was taken to get an estimate of the population ripple band power.
@@
 Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and $I_{k}$ be a discrete latent variable that is an indicator for the dynamics we wish to characterize: stationary, continuous, and fragmented.
-The goal of the model is to simultaneously estimate the posterior probability of both variables $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed data for for all times from 1 to $T$, the final time.
+The goal of the model is to simultaneously estimate the posterior probability of position and dynamics $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed spiking data from time 1 to time $T$.
 The observed data can be either spike trains $\Delta N_{1:T}^{(1:C)}$ from $C$ putative cells when decoding with sorted spikes or multiunit spikes $\Delta N_{1:T}^{(1:E)}$ and their associated wave form features $\Vec{m}$ from each tetrode $E$ when decoding with clusterless spikes.
 
-We have previously shown [Asilomar REF] that the posterior $p(x_k, I_k \mid O_{1:T})$ can be directly estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$:
+We have previously shown \cite{DenovellisCharacterizinghippocampalreplay2019} that the posterior probability $p(x_k, I_k \mid O_{1:T})$ can be estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$ and iterating to time T:
 
@@
 $$
-and then applying the acausal smoother equation, starting from the last estimate of the casual filter $p(x_{T}, I_{T} \mid O_{1:k})$ and iterating backwards to time 1:
+and then applying the acausal smoother equation, starting from the last estimate of the casual filter $p(x_{T}, I_{T} \mid O_{1:k})$ and recursively iterating backwards to time 1:
 $$
@@
 
-We set the dynamics transition matrix $Pr(I_{k} \mid I_{k-1})$, which defines how likely the dynamic is to change to another dynamic versus persist in the same dynamic, to be:
+For the dynamics transition matrix $Pr(I_{k} \mid I_{k-1})$, which defines how likely the dynamic is to change to another dynamic versus persist in the same dynamic, we set it to be:
 
@@
 
-We set the dynamics movement model $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$, which defines how likely the position $x_{k}$ is to change given the previous position $x_{k-1}$ and current $I_{k}$ and past dynamics $I_{k-1}$, to be:
+For the dynamics movement model $p(x_{k} \mid x_{k-1}, I_{k}, I_{k-1})$, which defines how likely the position $x_{k}$ is to change given the previous position $x_{k-1}$ and current $I_{k}$ and past dynamics $I_{k-1}$, we set it to be:
 
@@
 where $\delta(x_{k-1})$ is an identity transition matrix where position cannot change from the previous time step, $\mathcal{N}(x_{k-1}, 6.0)$ is a random walk from the previous position with variance 6.0, and $\mathcal{U}(\min x, \max x)$ is a uniform transition that allows transitions to any possible position.
-
 As discussed in the Results, this says that when persisting in the same dynamic, the stationary, continuous, and fragmented dynamics are defined by the identity transition, the random walk, and the uniform transition, respectively.
@@
 
-Lastly, we estimate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$. It has been shown [REFS, Zhang, Brown] that the likelihood with sorted spikes can be computed as:
+Lastly, we evaluate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$ based on an encoding model fit during the encoding period.
+We assume the likelihood is the same for each dynamic $I_k$, so we only need to estimate $p(O_{k} \mid x_{k})$.
+It has been shown \cite{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the likelihood with sorted spikes can be computed as:
 $$
-p(O_{k} \mid x_{k}, I_{k}) = p(\Delta N_{k}^{(1:C)} \mid x_{k}, I_{k}) \propto
-\prod^{C}_{i=1} [\lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
-    \exp[-\lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]
+p(O_{k} \mid x_{k}) = p(\Delta N_{k}^{(1:C)} \mid x_{k}) \propto
+\prod^{C}_{i=1} [\lambda_{i}(t_{k} \mid x_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
+    \exp[-\lambda_{i}(t_{k} \mid x_{k})\Delta_{k}]
 $$
-where $N_{t_{k}}^{i}$ represents a spike at time $t_k$ from cell $i$, $\Delta_k$ is the time bin size, and $\lambda_{i}(t_{k} \mid x_{k}, I_{k})$ is the instantaneous firing rate of cell $i$ given position $x_{k}$ and dynamic $I_{k}$---this is the "place field" of the cell, which we estimate during the encoding step.
+where $N_{t_{k}}^{i}$ represents a spike at time $t_k$ from cell $i$, $\Delta_k$ is the time bin size, and $\lambda_{i}(t_{k} \mid x_{k})$ is the instantaneous firing rate of cell $i$ given position $x_{k}$.
+$\lambda_{i}(t_{k} \mid x_{k})$ is the "place field" of the cell, which can be estimated via generalized linear model during the encoding period (which we explain further in the Model Implementation section).
 
@@
 $$
-p(O_{k} \mid x_{k}, I_{k}) = p(\Delta N_{1:T}^{(1:E)}, \vec{m} \mid x_{k}, I_{k}) \propto
-\prod^{E}_{i=1} [\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}, I_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
-    \exp[-\Lambda_{i}(t_{k} \mid x_{k}, I_{k})\Delta_{k}]
+p(O_{k} \mid x_{k}) = p(\Delta N_{k}^{(1:E)}, \vec{m_k} \mid x_{k}) \propto
+\prod^{E}_{i=1} [\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
+    \exp[-\Lambda_{i}(t_{k} \mid x_{k})\Delta_{k}]
 $$
-where $\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}, I_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k}, I_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
-Both of these quantities can be estimated by:
+where $\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
+Both of these quantities can be estimated via kernel density estimation based on multiunit spikes and their associated waveform features observed in the encoding period.
+
+\subsection*{Model Implementation}
+In order to implement the model, we used a grid based approximation of the latent position $x_k$ that respected the geometry of the track.
+For 1D linearized positions, we discretized the position space based on the same track graph we used for linearization by finding bins less than or equal to the chosen position bin size for each edge of the graph.
+For 2D positions, we discretized the position space by binning 2D positions occupied by the animal with equal sized bins of the chosen position bin size, followed by morphological opening to get rid of any holes smaller than the bin size.
+This was done using the scipy image processing tools \cite{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
+This grid based approximation allows us to use Reimann sums to approximate the integrals in the causal filter and acausal smoother equations.
+We also used this grid in combination with the track graph in order to construct the appropriate 1D random walk transition matrices that respected the track geometry.
+To do this, we inserted the position bin centers as nodes in the track graph, and then computed the shortest path distance between all pairs of position bin centers.
+We then evaluated a zero mean Gaussian with a variance of 6 cm on these distances to get the appropriate transition matrix.
 
@@
 \subsection*{Encoding - Clusterless}
-
-In our case, $\Vec{m}$ corresponds to the peak amplitude of each of the four tetrode wires at the time of the multiunit spike.
-
-In order to encode how each tetrode's unsorted spiking activity relates to position, we use a marked point process framework [REFS].
-For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features (the marks) and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
+In order to encode how each tetrode's unsorted spiking activity and waveform features relate to position, For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
 In our case, the waveform features correspond to the max amplitude observed on each tetrode wire at the time of the multiunit spike.
@@
 
+Both of these rates can be defined as:
+
+$$
+\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}) = \mu_i \frac{p_{i}(x_k, \vec{m}_k)}{\pi(x)}
+$$
+and
+$$
+\Lambda_{i}(t_{k} \mid x_{k}) = \mu_i \frac{p_{i}(x_k)}{\pi(x)}
+$$
+where $\mu_i$ is the mean firing rate for tetrode $i$, $\pi(x)$ is the spatial occupancy of the animal on the track, and $p_{i}(x_k, \vec{m}_k)$ and $p_{i}(x_k)$ are the joint probability of observing waveform features $\Vec{m}$ and position $x$ and the probability of observing position $x$, respectively.
+$\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$ can all be estimated using the kernel density estimator:
+$$
+kde(y) = \frac{1}{N h_1...h_D} \sum^{N}_{i=1} \prod^{D}_{d=1} K_{d}\left(\frac{y_d - y_{i,d}}{h_d}\right)
+$$
+where $y$ is the data with $D$ dimensions, $K$ is a 1 dimensional Gaussian kernel with bandwidth $h_d$, $y_{i, d}$ is the data observed during the encoding period, $N$ is the number of observations in the encoding period.
 \subsection*{Decoding}
@@
 
-\subsection*{Spatially Coherent Content}
-\subsection*{Spatially Incoherent Content}
-
 \subsection*{Highest Posterior Density}
@@
 \subsection*{Shuffle}
-\subsection*{Duration of dynamic}
+
 \subsection*{Replay distance from animal}
-Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's projected position on the track graph (see Linearization) and the MAP estimate of the joint posterior marginalized over state, which is defined as the center of the position bin with the greatest posterior value.
+Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+The shortest path is found using Dijkstra's algorithm.
 
 \subsection*{Non-Local Stationary Position}
-\subsection*{Multiunit Rate of Each Dynamic}
+The non-local stationary position is defined as replay distances at least 30 cm from the animal's position during a time period classified as stationary.
 

### 2020-07-14 Article.tex
Update on Overleaf.

@@
 \usepackage{amsmath}
-\DeclareMathOperator*{\E}{\mathbb{E}}
 \DeclareMathOperator*{\argmax}{arg\,max}
-\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
 
@@
 Next, we specify a dynamics model for how latent position---the "mental" position of the animal represented by the cells--- evolves over time (Figure \ref{1}C).
-We do this by defining a state transition matrix, which defines how the latent position can change from the previous time step.
+We do this by defining a movement state transition matrix, which defines how the latent position can change from the previous time step.
 Previous findings suggest three distinct dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, smoothly moving \cite{DavidsonHippocampalReplayExtended2009}, and fragmented / jumpy \cite{PfeifferAutoassociativedynamicsgeneration2015}, which could correspond to both spatially incoherent representations and representations that jump from one position to another.
@@
 One of our criteria for a more optimal method is that it use all of the available spiking data.
-Using only clustered spikes discards all of the spike events that cannot be uniquely assigned to a putative single neuron, reducing very substantially the quality of the resulting decoding.
+Using only clustered spikes discards all of the spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the quality of the resulting decoding.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}.
 We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding (see Methods).
@@
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content [REFS].
-We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see methods).
-We found that 22478 of 23382 or 96\% of classified SWRs had some spatially coherent structure.
+We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
+We found that 22478 of 23382 or 96\% of classified SWRs had time points with spatially coherent structure.
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented.
-We found that 3353 of 23382 or 14\% of classified SWRs had some spatially incoherent structure.
+We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure.
 This was consistent across animals (Figure \ref{3}F).
@@
 Here our findings were consistent with previous reports: in our datasets 4511 of  23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G).
-Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent but not continuous content.
+Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
 
@@
 
-Our model is specified in the context of a marked point process associated with different dynamics.
+Our model is specified in the context of a latent position associated with different dynamics.
 This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
@@
 An example of this change is shown in Figure \ref{4}D, where the cumulative HPD is small for most of the SWR until the end, where the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average cumulative HPD for each SWR, grouped by dynamic, we found a clear bi-modal split between spatially congruent dynamics and spatially incongruent dynamics (Figure \ref{4}E).
-For the spatially congruent dynamics, the cumulative HPD was much lower than the spatially incongruent dynamics (median 24 cm, spatially congruent vs.
-median 238 cm, spatially incongruent).
+Overall, when we examined the average cumulative HPD for each SWR, grouped by dynamic, we found a clear bi-modal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+For the spatially coherent dynamics, the cumulative HPD was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
+median 238 cm, spatially incoherent).
 
@@
 Our results indicate that most events with coherent spatial content are not continuous; that is, trajectories that move many times the running speed of the animal.
-We therefore examined these events in more details.
+We therefore examined these events in more detail.
 We first noted that most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic.
 Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at rates consistent with real-world movement (Fig.
+These events contain representations that move at rates more consistent with real-world movement (Fig.
 1G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
@@
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary trajectories most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}.
+These non-local stationary trajectories most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of these trajectories occurred at other locations of the track.
+This suggests that these representations can be flexibly deployed to represent discrete locations.
 
-Finally, we examined the overall activity levels across dynamics and found that all the dynamics had similar average multiunit spiking rates by tetrode (Figure \ref{5}E).
-This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures and importantly not just the continuous dynamics, were driven by sustained spiking information.
+Finally, we examined the overall activity levels across dynamics.
+We found that all the dynamics had similar average tetrode multiunit spiking rates (Figure \ref{5}E).
+This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
 
 \section*{Discussion}
+Here we developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
+We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
+We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would be difficult to characterize with the typical Bayesian decoder.
+Finally, we showed that this model is robust; it was relatively insensitive to transition parameters and depended strongly on the hippocampal place representation structure.
 
-Summary
+Using this model, we found that most SWRs contained spatially coherent content.
+Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed.
+Many of these events did not simply represent the animal's current location, but some other location on track.
 
-Challenges the notion that replays consist only of high speed events.
+Our results challenge the notion that hippocampal replays consist only of high speed events.
+Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions
 
-adjudicates/clarifies results of Pfiefer and Foster and Stella et al. Brownian motion but couldn't tell the speeds. Pfiefer and Foster didn't know the percentage of events that acted this way
+Our results complement those of Stella and colleagues \cite{StellaHippocampalReactivationRandom2019} who found that hippocampal replay trajectories followed Brownian like diffusion dynamics.
+We found that  We also clarify how these results can coexist with Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} who found that a subset of replay trajectories alternated between stationary and jumpy representations.
+We find that, for our task, these jumpy representations were relatively rare.
 
 Why our results are different from the Bayesian decoder
-seek to describe as many SWR events, clusterless decoding, more permissive model with different dynamics, explicit descriptive model instead of implicit noisy model.
+seek to describe as many SWR events, clusterless decoding, more permissive model with different dynamics, explicit descriptive model instead of implicit noisy model, smaller time bins
 
-relationship to Johnson and Redish 2008, Maboudi, model represents middle ground between HMM and typical Bayesian decoder.
+Our model is similar to the competing temporal models approach of Johnson and Redish 2008 \cite{JohnsonMeasuringdistributedproperties2008}, but allows for more .
+Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016}, which seek to be environment-agnostic discovers of sequential patterns but require a lot of data, and typical Bayesian decoder, which has very restrictive assumptions about the nature of the trajectories.
 allows SWR to quickly be identified because of the discrete states.
@@
 
-Can extend to incorporate other previously experienced environments. Add prior on dynamics of speed
+We believe our model can enable a richer set of analyses that takes advantage of all the data, 
+The model can easily be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
 
@@
 Animals were rewarded for performing the task correctly.
-Only run epochs with at least 9 hippocampal principal cells that fired at least 100 spikes were included in the analysis.
+Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
 
@@
 In order to decrease the time it takes to run the model, the 2D position of the animal was converted into a 1D position.
-This is done by first defining a 2D graph representation of the track, where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
+This is done by first defining a 2D graph representation of the track (herein referred to as the track graph), where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
 Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to classify the most likely track segment the animal is on.
@@
 To obtain the neural data used for decoding, hippocampus electrical potentials were recorded at 30 kHz, referenced to a corpus callosum tetrode, and then digitally filtered between 600 Hz and 6 kHz.
-Events were considered threshold crossing events if any one of the four tetrode wires exceeded a 60 mV threshold.
+Events were considered threshold crossing events if any one of the four tetrode wires exceeded a 60 $\mu V$ threshold.
 These multiunit threshold crossing events and the associated peak amplitudes of the tetrodes wires---defined as the electrical potential value at the time of the maximum amplitude at the threshold crossing---were used as data for the clusterless version of the model.
@@
 Lastly, we evaluate the likelihood of the observations $p(O_{k} \mid x_{k}, I_{k})$ based on an encoding model fit during the encoding period.
-We assume the likelihood is the same for each dynamic $I_k$, so we only need to estimate $p(O_{k} \mid x_{k})$.
-It has been shown \cite{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the likelihood with sorted spikes can be computed as:
+We assume the likelihood is the same for each dynamic $I_k$, so we only need to evaluate $p(O_{k} \mid x_{k})$.
+It has been shown \cite{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the Poisson likelihood with sorted spikes can be computed as:
 $$
@@
 where $N_{t_{k}}^{i}$ represents a spike at time $t_k$ from cell $i$, $\Delta_k$ is the time bin size, and $\lambda_{i}(t_{k} \mid x_{k})$ is the instantaneous firing rate of cell $i$ given position $x_{k}$.
-$\lambda_{i}(t_{k} \mid x_{k})$ is the "place field" of the cell, which can be estimated via generalized linear model during the encoding period (which we explain further in the Model Implementation section).
+$\lambda_{i}(t_{k} \mid x_{k})$ is the "place field" of the cell, which can be estimated by fitting a generalized linear model to each cell's spiking during the encoding period.
 
@@
 where $\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
-Both of these quantities can be estimated via kernel density estimation based on multiunit spikes and their associated waveform features observed in the encoding period.
-
-\subsection*{Model Implementation}
-In order to implement the model, we used a grid based approximation of the latent position $x_k$ that respected the geometry of the track.
-For 1D linearized positions, we discretized the position space based on the same track graph we used for linearization by finding bins less than or equal to the chosen position bin size for each edge of the graph.
-For 2D positions, we discretized the position space by binning 2D positions occupied by the animal with equal sized bins of the chosen position bin size, followed by morphological opening to get rid of any holes smaller than the bin size.
-This was done using the scipy image processing tools \cite{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
-This grid based approximation allows us to use Reimann sums to approximate the integrals in the causal filter and acausal smoother equations.
-We also used this grid in combination with the track graph in order to construct the appropriate 1D random walk transition matrices that respected the track geometry.
-To do this, we inserted the position bin centers as nodes in the track graph, and then computed the shortest path distance between all pairs of position bin centers.
-We then evaluated a zero mean Gaussian with a variance of 6 cm on these distances to get the appropriate transition matrix.
-
-\subsection*{Encoding - Sorted Spikes}
-In order to encode how each cell's spiking activity relates to position (the place field), we fit a generalized linear model (GLM) with a Poisson response distribution to each cell's spiking activity during movement (defined as time periods when the running speed is greater than 4 cm/s).
-We estimate the parameters $\beta$, which consist of $\beta_{0}$, the average firing rate over time, and $\beta_{i}$, weights for third degree B-splines basis functions $f_{i}$ over position (or tensor products of the B-splines when position is two dimensional).
-B-spline basis functions are used because place field firing activity is assumed to vary smoothly over position and this prior knowledge can be exploited to reduce the total number of model parameters needed.
-Each basis function is spaced every 5 cm over the range of the position and zero constrained so that the change encoded by the parameters is relative to the baseline firing rate.
-We use a log link function to convert the linear combination of parameters to an instantaneous firing rate over time $\lambda(t)$ to ensure the rate is always positive.
-
-$$log(\lambda(t)) = \beta_{0} + \sum_{i} f_{i}(position)\beta_{i}$$
-
-A small L2 penalization term $-\lambda\norm{\beta_{i}}_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low.
-We set this to 0.5 for all cells.
-Fitting is done by maximizing the penalized likelihood using a Newton-Raphson algorithm.
-The code used to fit the GLMs is available at \url{https://github.com/Eden-Kramer-Lab/regularized_glm}.
-
-\subsection*{Encoding - Clusterless}
-In order to encode how each tetrode's unsorted spiking activity and waveform features relate to position, For each multiunit spike during movement (defined as time periods when the running speed is greater than 4 cm/s), the associated waveform features and position are recorded in a (n-spikes, n-features + n-position-dimensions) array.
-In our case, the waveform features correspond to the max amplitude observed on each tetrode wire at the time of the multiunit spike.
-This array is used as the training samples in a kernel density estimator during the decoding step.
-
 Both of these rates can be defined as:
-
 $$
@@
 where $\mu_i$ is the mean firing rate for tetrode $i$, $\pi(x)$ is the spatial occupancy of the animal on the track, and $p_{i}(x_k, \vec{m}_k)$ and $p_{i}(x_k)$ are the joint probability of observing waveform features $\Vec{m}$ and position $x$ and the probability of observing position $x$, respectively.
-$\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$ can all be estimated using the kernel density estimator:
+$\pi(x)$, $p_{i}(x_k, \vec{m}_k)$, $p_{i}(x_k)$ can all be estimated by training a kernel density estimator on each tetrode's spike waveform features and corresponding position during the encoding period.
+
+\subsection*{Encoding - Sorted Spikes}
+In order to encode how each cell's spiking activity relates to position (the place field), we fit a generalized linear model (GLM) with a Poisson response distribution to each cell's spiking activity during the encoding period, which we define as all movement times (time periods when the running speed is greater than 4 cm/s).
+We estimate the parameters $\beta$, which consist of $\beta_{0}$, the average firing rate over time, and $\beta_{i}$, weights for third degree B-splines basis functions $f_{i}$ over position (or tensor products of the B-splines when position is two dimensional).
+B-spline basis functions are used because place field firing activity is assumed to vary smoothly over position and this prior knowledge can be exploited to reduce the total number of model parameters needed.
+Each basis function is spaced every 5 cm over the range of the position and zero constrained so that the change encoded by the parameters is relative to the baseline firing rate.
+We use a log link function to convert the linear combination of parameters to an instantaneous firing rate over time $\lambda(t)$ to ensure the rate is always positive.
+
+$$log(\lambda(t)) = \beta_{0} + \sum_{i} f_{i}(position)\beta_{i}$$
+
+A small L2 penalization term $-\lambda\|\beta_{i}\|_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low.
+We set this to 0.5 for all cells.
+Fitting is done by maximizing the penalized likelihood using a Newton-Raphson algorithm.
+The code used to fit the GLMs is available at \url{https://github.com/Eden-Kramer-Lab/regularized_glm}.
+
+\subsection*{Encoding - Clusterless}
+In order to encode how each tetrode's unsorted spiking activity and waveform features relate to position, we used kernel density estimation (KDE) to estimate the following distributions: $\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$.
+We used KDEs of the form:
 $$
@@
 $$
-where $y$ is the data with $D$ dimensions, $K$ is a 1 dimensional Gaussian kernel with bandwidth $h_d$, $y_{i, d}$ is the data observed during the encoding period, $N$ is the number of observations in the encoding period.
+where $y$ is the data with $D$ dimensions, $K$ is a 1 dimensional Gaussian kernel with bandwidth $h_d$, $y_{i, d}$ is the data observed during movement, $N$ is the number of observations during the movement period.
+For $\pi(x)$, $y_{i, d}$ is all positions observed during movement.
+For $p_{i}(x_k)$, $y_{i, d}$ is all positions at the time of multiunit spikes during movement.
+For $p_{i}(x_k, \vec{m}_k)$, $y_{i, d}$ is all positions at the time of multiunit spikes and their associated waveform features during movement.
+We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters minimized decoding error in \cite{KloostermanBayesiandecodingusing2014}.
+The mean firing rate $\mu_k$ was also estimated during movement.
+
 \subsection*{Decoding}
+In order to decode $p(x_{k}, I_{k} \mid O_{1:T})$, we used a grid based approximation of the latent position $x_k$ that respected the geometry of the track.
+For 1D linearized positions, we discretized the position space based on the same track graph we used for linearization by finding bins less than or equal to the chosen position bin size for each edge of the graph.
+For 2D positions, we discretized the position space by binning 2D positions occupied by the animal with equal sized bins of the chosen position bin size, followed by morphological opening to get rid of any holes smaller than the bin size.
+This was done using Scipy image processing tools \cite{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
+This grid based approximation allows us to use Reimann sums to approximate the integrals in the causal filter and acausal smoother equations.
+We chose a position grid with bins of 3 cm in width (and height if the model was computed with 2D positions) in order to get good resolution for the random walk transition matrices (which had 6 cm variance) as well as for the clusterless and sorted spikes decoding (which have 6 cm bandwidth for the KDE position dimensions and 5 cm spline knots for the GLM).
+For sorted spikes decoding, we evaluated the place field $\lambda_{i}(t_{k} \mid x_{k})$ on the midpoint of these bins.
+Likewise, for clusterless decoding, we evaluated the spike amplitudes observed during the decoding period by evaluating the KDE for $p_{i}(x_k, \vec{m}_k)$ for the midpoint of these bins.
 
-position bin size 3 cm
+We also used this grid bins in combination with the track graph in order to construct the appropriate 1D random walk transition matrices that respected the track geometry.
+To do this, we inserted the bin centers as nodes in the track graph, and then computed the shortest path distance between all pairs of position bin centers.
+We then evaluated a zero mean Gaussian with a variance of 6 cm on these distances to get the appropriate transition probability of moving from bin to bin.
 
-\subsection*{Decoding - Sorted Spikes}
+Finally, for our SWR analysis of $p(x_{k}, I_{k} \mid O_{1:T})$, we decoded each immobility time periods (times when the animal's speed was < 4 cm / s) in 2 ms time bins and extracted the SWR times.
 
-\subsection*{Decoding - Clusterless}
-24 mV smoothing for peak amplitude
-6.0 cm smoothing for position (1D and 2D)
 \subsection*{Mixture probability of the dynamics}
@@
 As in other calculations, the integral is approximated with a Reimann sum.
+
 \subsection*{Posterior probability of position}
-The posterior probability of position is . We estimate it by summing over the dynamics.
+The posterior probability of position is a quantity that indicates the most probable "mental" positions of the animal based on the data.
+We estimate it by marginalizing the joint probability over the dynamics.
 $$p(x_{k} \mid O_{1:T}) = \sum_{I_{k}} p(x_{k}, I_{k} \mid O_{1:T})$$
+
 \subsection*{Classification of Dynamics}
@@
 
+\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics}
+In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
+We fit the clusterless encoding model on the resampled data and then decoded the immobility periods.
+Like with the non-resampled data, we then extracted the SWR times and determined their classification based on our classification scheme.
+We repeated this shuffle analysis 50 times and compared this distribution to the real data for two epochs on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
+
 \subsection*{Highest Posterior Density}
-The 95\% highest posterior density (HPD) is a measurement of the uncertainty of the posterior and is defined as the region of the posterior that contains the top 95\% values of the posterior probability.
-It can be calculated by determining the maximum threshold value $h$ that fulfills the follow equality:
+The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% values of the posterior probability.
+In this manuscript, we use the HPD to evaluate the uncertainty of the posterior probability of position.
+We calculate the 95\% HPD by determining the maximum threshold value $h$ that fulfills the follow equality:
 $$
@@
 which we approximate with a Reimann sum.
-\subsection*{Shuffle}
 

### 2020-07-15 Article.tex
Update on Overleaf.

@@
 Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed.
-Many of these events did not simply represent the animal's current location, but some other location on track.
+Many of these events did not simply represent the animal's current location, but other locations on track.
 
-Our results challenge the notion that hippocampal replays consist only of high speed events.
+Our results challenge the notion that hippocampal replays are necessarily many times the speed of the animal.
 Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions
-
-Our results complement those of Stella and colleagues \cite{StellaHippocampalReactivationRandom2019} who found that hippocampal replay trajectories followed Brownian like diffusion dynamics.
-We found that  We also clarify how these results can coexist with Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} who found that a subset of replay trajectories alternated between stationary and jumpy representations.
-We find that, for our task, these jumpy representations were relatively rare.
 
@@
 
+Our results complement those of Stella and colleagues \cite{StellaHippocampalReactivationRandom2019} who found that hippocampal replay trajectories followed Brownian-like diffusion dynamics.
+We found that  We also clarify how these results can coexist with Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} who found that a subset of replay trajectories alternated between stationary and jumpy representations.
+We also found a log normal distribution of speeds, but it should be noted that they presented their results in terms of  at least for our task, these jumpy representations are relatively rare.
+
 Our model is similar to the competing temporal models approach of Johnson and Redish 2008 \cite{JohnsonMeasuringdistributedproperties2008}, but allows for more .
-Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016}, which seek to be environment-agnostic discovers of sequential patterns but require a lot of data, and typical Bayesian decoder, which has very restrictive assumptions about the nature of the trajectories.
+Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns but require a lot of data, and typical Bayesian decoder, which has very restrictive assumptions about the nature of the trajectories.
 allows SWR to quickly be identified because of the discrete states.

### 2020-07-15 Article.tex
Update on Overleaf.

@@
 continuous 42 cm, Figure \ref{5}C).
-This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position.
+This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position and cannot be explained by erroneously detected SWRs.
 
@@
 We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
-We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would be difficult to characterize with the typical Bayesian decoder.
-Finally, we showed that this model is robust; it was relatively insensitive to transition parameters and depended strongly on the hippocampal place representation structure.
+We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would not be found with the typical Bayesian decoder.
+Finally, we showed that this model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
 
@@
 Many of these events did not simply represent the animal's current location, but other locations on track.
-
-Our results challenge the notion that hippocampal replays are necessarily many times the speed of the animal.
+This result challenges the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999}.
 Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions

### 2020-07-16 Article.tex

@@
 \author[2, 3]{Michael E. Coulter}
+\author[5]{Marielena Sosa}
+\author[6]{Kenneth Kay}
 \author[4]{Uri T. Eden}
@@
 \affil[4]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
+\affil[5]{Department of Neurobiology, Stanford University, Palo Alto, California}
+\affil[6]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
 
@@
 \begin{keywords}
-Hippocampus | Replay | State Space
+Hippocampus | Replay | State Space | Dynamics
 \end{keywords}
@@
 
-Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the dynamics model--how the position can change over time.
+Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamics model--how the position can change over time.
 For the data model, our decoder is the same as the Bayesian decoder.
@@
 
-Next, we specify a dynamics model for how latent position---the "mental" position of the animal represented by the cells--- evolves over time (Figure \ref{1}C).
+Next, we specify a movement dynamics model for how latent position---the "mental" position of the animal represented by the cells--- evolves over time (Figure \ref{1}C).
 We do this by defining a movement state transition matrix, which defines how the latent position can change from the previous time step.
@@
 
-Finally, we specify how likely the dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
+Finally, we specify how likely each dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
 In order to be conservative, we assume that each dynamic is likely to dominate for the duration of the SWR, with a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
@@
 
-Our model is specified in the context of a latent position associated with different dynamics.
+Our model is specified in the context of a latent position associated with different movement dynamics.
 This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
@@
 We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
-We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would not be found with the typical Bayesian decoder.
+We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover both SWRs with continuous many times the speed of the animal and SWRs with non-constant speed and non-linear trajectories that would not be found with the typical Bayesian decoder.
 Finally, we showed that this model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
 
-Using this model, we found that most SWRs contained spatially coherent content.
+Using this model, we found that most SWRs contained spatially coherent trajectories.
 Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed.
-Many of these events did not simply represent the animal's current location, but other locations on track.
-This result challenges the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999}.
-Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions
+Many of these events did not represent the animal's current location, but other locations on track.
+These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
+Instead there is a much richer variety of speeds the animal can use to recall prior experience or generate possible future events, including stationary representations of single locations.
 
-Why our results are different from the Bayesian decoder
-seek to describe as many SWR events, clusterless decoding, more permissive model with different dynamics, explicit descriptive model instead of implicit noisy model, smaller time bins
+Our results also explain the seeming conflict between Stella and colleagues \cite{StellaHippocampalReactivationRandom2019}, who found that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015}, who found that replay trajectories contained jumps in position.
+We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
+Our model allows identification of both types of sequences.
 
-Our results complement those of Stella and colleagues \cite{StellaHippocampalReactivationRandom2019} who found that hippocampal replay trajectories followed Brownian-like diffusion dynamics.
-We found that  We also clarify how these results can coexist with Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} who found that a subset of replay trajectories alternated between stationary and jumpy representations.
-We also found a log normal distribution of speeds, but it should be noted that they presented their results in terms of  at least for our task, these jumpy representations are relatively rare.
+We were able to find these results for several key reasons.
+First, we sought to describe as many SWRs as possible, instead of throwing many events out.
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated.
+Third, we built an explicit, but flexible model that allowed us to interpret more of the SWRs.
+And fourth, we used small time bins and acausal estimation of latent position, which allows us to identify the specific timing of the dynamics.
 
-Our model is similar to the competing temporal models approach of Johnson and Redish 2008 \cite{JohnsonMeasuringdistributedproperties2008}, but allows for more .
-Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns but require a lot of data, and typical Bayesian decoder, which has very restrictive assumptions about the nature of the trajectories.
-allows SWR to quickly be identified because of the discrete states.
-Can work in 2D.
-Can precisely identify timing of dynamics because of acausal decoding
+Our model can be viewed as an extension to the multiple temporal models approach of Johnson and colleagues 2008 \cite{JohnsonMeasuringdistributedproperties2008}, which used model selection to determine the model with the most appropriate trajectory speed.
+Our model goes beyond the Johnson model in that it explicitly permits a mixture between the movement speeds, can work for arbitrary track geometries, and uses clusterless decoding.
+Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns and the typical Bayesian decoder, which either uses very restrictive assumptions about the nature of the trajectories or no assumptions.
+The Hidden Markov-style models require lots of data and still involve interpreting the sequences in terms of known variables, such as the environment.
+The Bayesian decoder is either quite restrictive, when assuming a constant velocity trajectory, or potentially quite noisy, depending on the data.
+Both of these styles of models are dependent on the time bin size.
+Our model leverages some of our prior knowledge about hippocampus and its relationship to space to give less noisy and more interpretable estimates of the replay trajectories.
 
-We believe our model can enable a richer set of analyses that takes advantage of all the data, 
+We believe our model can enable a richer set of analyses that takes advantage of all the data.
 The model can easily be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
@@
 \begin{acknowledgements}
+We thank Margaret Carr Larkin and Mattias P. Karlsson for use of their openly available datasets.
 \end{acknowledgements}

### 2020-07-18 Article.tex
Update on Overleaf.

@@
 
-\title{A state space model reveals the rich and varied dynamics of replay}
+\title{A state space model reveals the rich and varied dynamics of hippocampal replay}
 \shorttitle{Replay Dynamics}
@@
 
-Mental events may not occur in the same time span as actual events; memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
+Mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
 Consistent with this, generative representations can unfold rapidly.
@@
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter.
+Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter (cite Foster and Wilson, Diba and Buzsaki, Karlsson and Frank, appropriate Csicsvari paper).
 The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed.
+Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed (Foster and Wilson, Diba and Buzsaki, Karlsson and Frank,).
 A statistical test is then used to determining whether the fit is better than the fit to shuffled versions of the data.
@@
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which are often excluded.
+Approaches that focus on the order of cell activity within each even (Lee and Wilson, Gupta et.
+al?) allow for a relaxation of that linear assumption, but replace by either ignoring all but the first spike from each cell or some arbitrarily smoothed version of the spike train.
+These approaches also do not provide information about the dynamics of the underlying spatial representation.
+Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches expclude .
 In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
@@
 \cite{StellaHippocampalReactivationRandom2019} found that replays are more spatially continuous, following Brownian diffusion dynamics.
-Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
+Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
 
@@
 We apply this model to data from 10 animals, focusing on SWR events to permit a direct comparison to previous work.
+
 We find that the vast majority of SWRs contain spatially coherent content.
 Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at speeds consistent with real animal movement, suggesting that high speed replay is not the most common manifestation of generative activity.
-Our method provides a useful and powerful tool for understanding the dynamics of mental representations, and could be adapted for use in many brain areas.
+These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
 
@@
 Finally, we specify how likely each dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for the duration of the SWR, with a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
-Accordingly, we set the probability of staying in a state to 0.98 for each 2 ms time step.
-Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular state.
+In order to be conservative, we assume that each dynamic is likely to dominate for ~100 ms, which is approximately the duration of the SWR.
+There is however,  a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
+Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step.
+Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular dynamic.
+Importantly, the data drive the dynamic, so even if the probability of a particular dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of ~10 ms, drive a change to a different dynamic.
 
@@
 We note here that the choice of any particular threshold is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95 because the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are likely to be meaningful to the brain.
+Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95 because the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are meaningful in the brain.
 Nonetheless, we also verified that our central results hold with a higher threshold of 0.95.
@@
 As mentioned above, the model is robust to the choice of probability of staying in a state.
-We decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
+To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
 We found that for a large range of plausible probabilities of staying in one of the dynamics (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
@@
 
-Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content [REFS].
+Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content (Karlsson and Frank, Foster and Wilson Davidson et.
+al.
+etc.
+.
 We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 22478 of 23382 or 96\% of classified SWRs had time points with spatially coherent structure.
-We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented.
+We found that 22478 of 23382 or 96\% of classified SWRs (85\% of all SWRs) included spatially coherent structure .
+We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented dynamics.
 We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure.
 This was consistent across animals (Figure \ref{3}F).
-We then asked what percentage of classified SWR contained continuous content, like would typically be analyzed when using the Bayesian decoder.
+
+To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the Bayesian decoder.
 Here our findings were consistent with previous reports: in our datasets 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G).
@@
 We repeated this analysis with a higher threshold of 0.95 to ensure that our result was not dependent on this threshold.
-We found that, while this slightly reduced the total fraction of classified ripples (19180 / 25845 or 74\% SWRs), an even higher fraction of the classified ripples (19020 / 19180 or 99\% classified) had spatially continuous content.
+We found that, while this slightly reduced the total fraction of classified ripples (19180 / 25845 or 74\% SWRs), an even higher fraction of the classified ripples (19020 / 19180 or 99\% classified) had spatially coherent content.
 Classified SWRs containing spatially incoherent content similarly were a small fraction of the classified SWRs (spatially incoherent: 484 / 19180 or 3\%).
@@
 
-We found that spatially coherent events did indeed have lower cumulative HPDs.
+We found that spatially coherent events did indeed have lower cumulative HPDs than events with fragmented dynamics.
 Figure \ref{4}A and B show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
@@
 
-Our results indicate that most events with coherent spatial content are not continuous; that is, trajectories that move many times the running speed of the animal.
+Our results indicate that most events with coherent spatial content are not dominated by the continuous dynamic, and thus correspond to trajectories that move relatively slowly compared to their purely continuous counteparts.
 We therefore examined these events in more detail.
@@
 \section*{Discussion}
-Here we developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
-We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
-We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover both SWRs with continuous many times the speed of the animal and SWRs with non-constant speed and non-linear trajectories that would not be found with the typical Bayesian decoder.
-Finally, we showed that this model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
+We developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
+This model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
+We show that this model is interpretable: it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
+This model is also flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with representations that move more or less linearly at a constant, high speed , but also representations that move non-linearly  with slower speeds or are stationary.
+These slower moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical Bayesian decoder.
+The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and real-world speeds.
 
-Using this model, we found that most SWRs contained spatially coherent trajectories.
-Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed.
-Many of these events did not represent the animal's current location, but other locations on track.
+Previous word reported that at most half of SWRs or high-population-activity events contained replay (REFS from above).
+This could lead one to the assumption that most events are not spatially coherent or meaningful.
+Our results indicate that this is not the case: in our dataset the large majority of SWRs (85\%) contained spatially coherent content, defined as content with some combination of stationary or continuous dynamics.
+These events included decoded representations that were highly spatial concentrated, confirming their spatial coherence.
+This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
+
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with real-world movement speeds of <1 m/s.
+The next most common category were stationary events (Cite Yu etl al, Farooq et.
+al, perhaps Dupret et.
+al 2009?) that activated a representation of a single location, while the classical "replay" events made up only about ~19\% of classified events.
+
 These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
-Instead there is a much richer variety of speeds the animal can use to recall prior experience or generate possible future events, including stationary representations of single locations.
+Instead there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+Interestingly, Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
+Here, we show that they do not disappear, but are a very common occurrence with adult rats.
+Here we note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds of 0 m/s.
 
-Our results also explain the seeming conflict between Stella and colleagues \cite{StellaHippocampalReactivationRandom2019}, who found that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015}, who found that replay trajectories contained jumps in position.
+Our results may also explain the seeming conflict between Stella and colleagues \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
 We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
 Our model allows identification of both types of sequences.
+Our model also makes it possible to identify different dynamics within the same SWR.
+As the boundaries of each event are also arbitrary, we suggest that knowing the start and end of each dynamic may prove a powerful way to understand how inputs to the hippocampus influence these events and how these events in turn influence downstream structures.
 
-We were able to find these results for several key reasons.
-First, we sought to describe as many SWRs as possible, instead of throwing many events out.
-Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated.
-Third, we built an explicit, but flexible model that allowed us to interpret more of the SWRs.
-And fourth, we used small time bins and acausal estimation of latent position, which allows us to identify the specific timing of the dynamics.
+Our identification of rich and varied dynamics in SWRs depended on several factors.
+First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.
+This more permissive approch was also critical for the identification of events that activate specific, immobility-associated locations in previous work (Cite Yu et.
+al).
+More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
+Instead, the sizes of these events are well described by a unimodal, long-tailed distribution (cite cheng and frank 2008).
+High thresholds can therefore exclude large numbers of real events and potentially lead to erroneous conclusions.
+
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated (REFS).
+Third, we built an explicit, but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
+Fourth, our model avoids assumptions about temporal bins and does not make assumptions about the linear progression of events.
+Instead, the model can use any size of bin and allows for movement in any direction.
+Finally, we used an  acausal estimator of latent position, which allowed us to identify the specific timing of transitions in dynamics at a fine timescale.
 
@@
 Our model goes beyond the Johnson model in that it explicitly permits a mixture between the movement speeds, can work for arbitrary track geometries, and uses clusterless decoding.
-Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns and the typical Bayesian decoder, which either uses very restrictive assumptions about the nature of the trajectories or no assumptions.
-The Hidden Markov-style models require lots of data and still involve interpreting the sequences in terms of known variables, such as the environment.
-The Bayesian decoder is either quite restrictive, when assuming a constant velocity trajectory, or potentially quite noisy, depending on the data.
-Both of these styles of models are dependent on the time bin size.
-Our model leverages some of our prior knowledge about hippocampus and its relationship to space to give less noisy and more interpretable estimates of the replay trajectories.
+Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns and the typical Bayesian decoder, which typically use arbitrarily chosen bin sizes and  restrictive assumptions about the nature of the trajectories.
+In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results without the need to make assumptions about the right temporal bin size.
 
-We believe our model can enable a richer set of analyses that takes advantage of all the data.
-The model can easily be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
+State-space models like the one we use here can enable a richer set of analyses of generative events that takes advantage of all of the available data.
+These approaches can be applied to not just SWRs but all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.
+The model can be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
+It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+Finally, models can be built to estimate any covariate, including movement direction (cite Kay 2020).
+We therefore suggest that this approach has the potential to provide a much richer understanding of neural representations throughout the brain.
 
@@
 \subsection*{Simulated Data}
-Encoding data for Figure \ref{1} and Figure \ref{fig:Figure1-Figure supplement 1} was generated by simulating 15 traversals of a 180 cm linear track.
+Encoding data for Figure \ref{1} and Figure \ref{fig:Figure1-Figure supplement 1} were generated by simulating 15 traversals of a 180 cm linear track.
 The spiking of 19 neurons was simulated using an inhomogeneous Poisson process where the instantaneous firing rate of each neuron changes according to a Gaussian place field.
@@
 In brief, each day, animals alternated between 20 minute rest epochs in a rest box and 15 minutes run epochs in the W-track.
-On the W-track, animals alternated between two types of runs: inbound and outbound.
-On inbound runs, the animal had to run from an outer arm to the center arm.
-On outbound runs, the animals had to run from the center arm to the least recently visited outer arm.
-Animals were rewarded for performing the task correctly.
-Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
+On the W-track, animals were rewarded at the ends of an arm when that arm was the next correct arm in the sequence.
+Two rules determined the next correct arm.
+If the animal was in an outer arm, it had to run to the center arm.
+If the animal was in the center arm, it had to run from the center arm to the least recently visited outer arm.
+Animals were free to choose any arm at any time Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
 
@@
 where $\delta(x_{k-1})$ is an identity transition matrix where position cannot change from the previous time step, $\mathcal{N}(x_{k-1}, 6.0)$ is a random walk from the previous position with variance 6.0, and $\mathcal{U}(\min x, \max x)$ is a uniform transition that allows transitions to any possible position.
-As discussed in the Results, this says that when persisting in the same dynamic, the stationary, continuous, and fragmented dynamics are defined by the identity transition, the random walk, and the uniform transition, respectively.
+As discussed in the Results, this means that when persisting in the same dynamic, the stationary, continuous, and fragmented dynamics are defined by the identity transition, the random walk, and the uniform transition, respectively.
 When transitioning to or from the fragmented dynamic, we assume we do not have any information about the position, so the transition is uniform.
@@
 \subsection*{Encoding - Clusterless}
-In order to encode how each tetrode's unsorted spiking activity and waveform features relate to position, we used kernel density estimation (KDE) to estimate the following distributions: $\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$.
+In order to relate  each tetrode's unsorted spiking activity and waveform features  to position, we used kernel density estimation (KDE) to estimate the following distributions: $\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$.
 We used KDEs of the form:
@@
 We used a threshold of 0.8 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
-Time periods during sharp wave ripples are labeled as Stationary, Continuous, or Fragmented when the probability of each state is above 0.8.
+Time periods during sharp wave ripples are labeled as Stationary, Continuous, or Fragmented when the probability of each individual state is above 0.8.
 Time periods are labeled as Stationary-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Stationary and Continuous or Fragmented and Continuous are above 0.8, respectively.

### 2020-07-18 Article.tex
Update on Overleaf.

@@
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches expclude .
+Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches exclude.
 In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
@@
 
-Our results indicate that most events with coherent spatial content are not dominated by the continuous dynamic, and thus correspond to trajectories that move relatively slowly compared to their purely continuous counteparts.
+Our results indicate that most events with coherent spatial content are not dominated by the continuous dynamic, and thus correspond to trajectories that move relatively slowly compared to their purely continuous counterparts.
 We therefore examined these events in more detail.
@@
 
-Previous word reported that at most half of SWRs or high-population-activity events contained replay (REFS from above).
+Previous work reported that at most half of SWRs or high-population-activity events contained replay (REFS from above).
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
@@
 First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.
-This more permissive approch was also critical for the identification of events that activate specific, immobility-associated locations in previous work (Cite Yu et.
-al).
+This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \cite{JaiDistincthippocampalcorticalmemory2017}.
 More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
-Instead, the sizes of these events are well described by a unimodal, long-tailed distribution (cite cheng and frank 2008).
+Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \cite{ChengNewExperiencesEnhance2008}.
 High thresholds can therefore exclude large numbers of real events and potentially lead to erroneous conclusions.
@@
 It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
-Finally, models can be built to estimate any covariate, including movement direction (cite Kay 2020).
+Finally, models can be built to estimate any covariate, including movement direction \cite{KayConstantSubsecondCycling2020}.
 We therefore suggest that this approach has the potential to provide a much richer understanding of neural representations throughout the brain.

### 2020-07-18 Article.tex
Draft 1

@@
 Mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
-Consistent with this, generative representations can unfold rapidly.
+Consistent with this, generative representations can unfold rapidly \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 As a result, decoding generative representations often involves making assumptions about their timescale and temporal structure, and these assumptions can limit the interpretation of the data.
@@
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter (cite Foster and Wilson, Diba and Buzsaki, Karlsson and Frank, appropriate Csicsvari paper).
+Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
 The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed (Foster and Wilson, Diba and Buzsaki, Karlsson and Frank,).
+Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determining whether the fit is better than the fit to shuffled versions of the data.
@@
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Approaches that focus on the order of cell activity within each even (Lee and Wilson, Gupta et.
-al?) allow for a relaxation of that linear assumption, but replace by either ignoring all but the first spike from each cell or some arbitrarily smoothed version of the spike train.
+Approaches that focus on the order of cell activity within each even \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace by either ignoring all but the first spike from each cell or some arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
@@
 \subsection*{Overview of the model}
-We begin with an application of the model to simulated data, both to validate the model and to provide intuition.
-(Figure \ref{1}).
+We begin with an application of the model to simulated data, both to validate the model and to provide intuition (Figure \ref{1}).
 We simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
@@
 
-Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamics model--how the position can change over time.
+Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamics model---how the position can change over time.
 For the data model, our decoder is the same as the Bayesian decoder.
@@
 
-Next, we specify a movement dynamics model for how latent position---the "mental" position of the animal represented by the cells--- evolves over time (Figure \ref{1}C).
+Next, we specify a movement dynamics model for how latent position---the "mental" position of the animal represented by the cells---evolves over time (Figure \ref{1}C).
 We do this by defining a movement state transition matrix, which defines how the latent position can change from the previous time step.
@@
 In order to be conservative, we assume that each dynamic is likely to dominate for ~100 ms, which is approximately the duration of the SWR.
-There is however,  a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
+There is, however, a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
 Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step.
@@
 As mentioned above, the model is robust to the choice of probability of staying in a state.
-To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure 1-Supplemental Figure 1).
+To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure \ref{fig:Figure1-Figure supplement 1}).
 We found that for a large range of plausible probabilities of staying in one of the dynamics (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
@@
 We can also see this when we project the maximum of the posterior of this trajectory to 2D (Figure \ref{2}C) to better see the spatial trajectory.
-Moreover, when we apply the same model using the 2D position of the animal, we get a similar result as when we used the 1D linearized position (Figure 2-Supplemental 1A).
+Moreover, when we apply the same model using the 2D position of the animal, we get a similar result as when we used the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).
 
@@
 Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}).
-Applying a clusterless decoder to the same SWR, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure 2-Supplemental 1B).
+Applying a clusterless decoder to the same SWR, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
 The spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model, however.
@@
 
-After testing our model on a single SWR, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (tetrodes = [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020}).
+After testing our model on a single SWR, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}).
 One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see Methods) to ensure that we were included both the rare large events as well as the much more common small events.
 
-As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of SWR (Figure 2-Supplemental 2A-C).
+As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of SWR (Figure \ref{fig:Figure2-Figure supplement 2}A-C).
 However, we also observed many events with spatially coherent content that did not have this structure.
-For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure 3-Supplemental 1B), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure3-Supplemental 1A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure3-Supplemental1D).
+For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}1), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
 
@@
 We then decoded the same SWR events containing the same spikes.
-Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for epoch 1, p=0.02 for epoch 2, Figure 3-Supplemental 2).
+Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for epoch 1, p=0.02 for epoch 2, Figure \ref{fig:Figure3-Figure supplement 2}).
 
-Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content (Karlsson and Frank, Foster and Wilson Davidson et.
-al.
-etc.
-.
+Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 22478 of 23382 or 96\% of classified SWRs (85\% of all SWRs) included spatially coherent structure .
+We found that 22478 of 23382 or 96\% of classified SWRs (85\% of all SWRs) included spatially coherent structure.
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented dynamics.
@@
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the Bayesian decoder.
-Here our findings were consistent with previous reports: in our datasets 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G).
+Here our findings were consistent with previous reports: in our datasets 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G, 17\% of all SWRs).
 Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
@@
 Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at rates more consistent with real-world movement (Fig.
-1G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+These events contain representations that move at rates more consistent with real-world movement (Figure \ref{1}G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 
-Previous work reported that at most half of SWRs or high-population-activity events contained replay (REFS from above).
+Previous work reported that at most half of SWRs or high-population-activity events contained replay \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with real-world movement speeds of <1 m/s.
-The next most common category were stationary events (Cite Yu etl al, Farooq et.
-al, perhaps Dupret et.
-al 2009?) that activated a representation of a single location, while the classical "replay" events made up only about ~19\% of classified events.
+The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a representation of a single location, while the classical "replay" events made up only about ~19\% of classified events.
 
@@
 Here, we show that they do not disappear, but are a very common occurrence with adult rats.
-Here we note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds of 0 m/s.
+We note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds of 0 m/s.
 
@@
 
-Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated (REFS).
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \cite{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
 Third, we built an explicit, but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
@@
 If the animal was in the center arm, it had to run from the center arm to the least recently visited outer arm.
-Animals were free to choose any arm at any time Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
+Animals were free to choose any arm at any time.
+Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
 
@@
 Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
-\label{Figure3-Figure supplement 2}
+\label{fig:Figure3-Figure supplement 2}
 \end{figure}

### 2020-07-20 Article.tex
Update on Overleaf.

@@
 \author[6]{Kenneth Kay}
+\author[7]{Jason E. Chung}
 \author[4]{Uri T. Eden}
@@
 \affil[6]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
+\affil[7]{Department of Neurological Surgery, University of California, San Francisco, California}
 
@@
 These internal "generative" representations \cite{KayConstantSubsecondCycling2020} can be identified because the underlying neural activity patterns resemble those seen during actual events.
-In particular, an encoding model that relates neural firing to external covariates can be constructed during an experience, and this model can inverted to decode an internal representation of the covariates from the neural activity at a different time.
+In particular, an encoding model that relates neural firing to external covariates can be constructed during an experience, and this model can be inverted to decode an internal representation of the covariates from the neural activity at a different time.
 
@@
 Other panels have the same convention as(A).
-\textbf{(C)} 1D MAP estimate of the latent position in (A) projected back into 2D.
+\textbf{(C)} Maximum of the 1D decoded position (the most probable position) in (A) projected back into 2D.
 Color indicates time.
@@
 L, R, and C correspond to the left, right and center reward well as in (A).
-\textbf{(D)} 1D MAP estimate of the latent position in (B) projected back into 2D.
+\textbf{(D)} Maximum of the 1D decoded position (the most probable position) in (B) projected back into 2D.
 Figure conventions are the same as in (C).
@@
 Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
+Both code bases rely on the python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
 All code is open-source and licensed under the MIT Software License.

### 2020-07-21 Article.tex
Update on Overleaf.

@@
 Hippocampal replay events are a prototypical example of a generative representation.
-As animals move through space, neurons in the hippocampus preferentially fire in at specific locations in an environment, and thus sets of cells can fire in sequence as the animal moves through a series of locations.
+As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells can fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous or potential future experience, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed \cite{DavidsonHippocampalReplayExtended2009}.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience or likely to occur during potential future experiences, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-Identifying these events typically involves multiple steps and assumptions about the nature of the event.
+Identifying these events typically involves multiple steps and assumptions about their nature.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on an threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
+Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
 The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
 Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
-A statistical test is then used to determining whether the fit is better than the fit to shuffled versions of the data.
+A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
 
@@
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Approaches that focus on the order of cell activity within each even \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace by either ignoring all but the first spike from each cell or some arbitrarily smoothed version of the spike train.
+Approaches that focus on the order of cell activity within each even \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using some arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches exclude.
+Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches often exclude.
 In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the weighted mean or maximum of each posterior time bin and connecting each time bin with a line.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using an estimated represented location at each time bin and connecting each with a line.
 For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that replays can alternate between representing a single location and sequential spatial trajectories.
@@
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
-It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for the dynamic.
+It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for each category of dynamics.
 Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} and characterizes represented trajectories as a mixture of three underlying patterns of dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
-We show how this model can take advantage of clusterless decoding---which relates multiunit spike wave form features to position without spike sorting---giving us more information about the population spiking activity.
-We apply this model to data from 10 animals, focusing on SWR events to permit a direct comparison to previous work.
+Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} and characterizes represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
+We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
+We apply this model to data from 10 rats, focusing on SWR events to permit a direct comparison to previous work.
 
@@
 Finally, we specify how likely each dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for ~100 ms, which is approximately the duration of the SWR.
+In order to be conservative, we assume that each dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of the SWR.
 There is, however, a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
@@
 Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular dynamic.
-Importantly, the data drive the dynamic, so even if the probability of a particular dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of ~10 ms, drive a change to a different dynamic.
+Importantly, the data drives the estimated dynamic, so even if the probability of a particular dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
 
@@
 To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
+We found that for a large range of plausible probabilities of staying in one of the dynamics over a single 2 ms time step (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
 
@@
 However, we also observed many events with spatially coherent content that did not have this structure.
-For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}1), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
+For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
 
@@
 The most likely track segment the animal is on is computed using the Viterbi algorithm.
-After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the track segment.
+After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the nearest point of the track segment.
 This allows us to define a distance from the center well in terms of shortest path length on the track, where 0 cm represents the center well position.
@@
 \subsection*{The Model}
-Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and $I_{k}$ be a discrete latent variable that is an indicator for the dynamics we wish to characterize: stationary, continuous, and fragmented.
+Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and $I_{k}$ be a discrete latent variable that is an indicator for the dynamics we wish to movement characterize: stationary, continuous, and fragmented.
 The goal of the model is to simultaneously estimate the posterior probability of position and dynamics $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed spiking data from time 1 to time $T$.
@@
 For $p_{i}(x_k, \vec{m}_k)$, $y_{i, d}$ is all positions at the time of multiunit spikes and their associated waveform features during movement.
-We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters minimized decoding error in \cite{KloostermanBayesiandecodingusing2014}.
+We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters were found to minimize decoding error in Kloosterman et al.
+\cite{KloostermanBayesiandecodingusing2014}.
 The mean firing rate $\mu_k$ was also estimated during movement.
@@
 \subsection*{Highest Posterior Density}
-The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% values of the posterior probability.
+The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% values of the posterior probability \cite{CasellaStatisticalinference2001}.
 In this manuscript, we use the HPD to evaluate the uncertainty of the posterior probability of position.

### 2020-07-21 Article.tex
Update on Overleaf.

@@
 Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
-The shortest path is found using Dijkstra's algorithm.
+The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959}.
 
@@
 Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
-Both code bases rely on the python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
+Both code bases rely on the following python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
 All code is open-source and licensed under the MIT Software License.

### 2020-07-21 Article.tex
Update on Overleaf.

@@
 Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
-Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
+Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} ((doi: 10.5281/zenodo.3713412).
 Both code bases rely on the following python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).

### 2020-07-21 Article.tex
Update on Overleaf.

--- 2020-07-21 Article.tex
@@
 Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
-Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
+Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
 Both code bases rely on the following python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).

### 2020-07-21 Article.tex

@@
 Analysis of SWRs from all animals.
-\textbf{(A)} UpSet plot---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each ripple.
+\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each ripple.
 Each row represents a possible category present in the SWR.
@@
 Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
-The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959}.
+The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959} as implemented in NetworkX \cite{HagbergExploringNetworkStructure2008}.
 
@@
 \caption{The model is robust to changes in the discrete transition matrix.
-Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value--which governs the probability of remaining in that state.
-The off-diagonal values--the probability of switching to one of the other dynamics--are set to be equally likely with the remainder of the probability, as in Figure 1D.
+Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that state.
+The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in Figure 1D.
 The first plot (upper left) shows the case when all dynamics are equally likely.
-The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero--i.e.
+The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e.
 the case where there is no probability of switching to another state.}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}
-\caption{Decoding the same SWR in Figure 2 using 2D clusterless and sorted spikes.
+\caption{Decoding the same SWR in Figure \ref{2} with 2D position using sorted spikes decoding and clusterless decoding, respectively.
 \textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure 2.
@@
 Shaded regions indicate classification category as in Figure 1G and 2.
-The rightmost plot shows the MAP estimate of the latent position with color indicating time.
+The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
 The latent position posterior summed over time is shown in purple underneath.
@@
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing.}
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}
 \label{fig:Figure3-Figure supplement 2}

### 2020-07-29 Article.tex
Draft 2

@@
 
-\title{A state space model reveals the rich and varied dynamics of hippocampal replay}
+\title{Revealing the rich and varied dynamics of hippocampal replay}
 \shorttitle{Replay Dynamics}
@@
 \author[2, 3]{Michael E. Coulter}
+\author[6]{Kenneth Kay}
 \author[5]{Marielena Sosa}
-\author[6]{Kenneth Kay}
 \author[7]{Jason E. Chung}
@@
 \affil[4]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
-\affil[5]{Department of Neurobiology, Stanford University, Palo Alto, California}
+\affil[5]{Department of Neurobiology, Stanford University School of Medicine, Stanford, CA}
 \affil[6]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
@@
 \begin{abstract}
-Representations of past or possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus is a brain region known to express these representations across multiple states, including during sharp-wave ripple (SWR) events where sets of hippocampal place cells can be activated in time-compressed sequences consistent with a rapid mental traversal of a path.
-Efforts to understand these "replay" events typically analyze a small subset of events and assume rapid mental movement, and as a result we lack a clear picture of both the prevalence and the range of dynamics associated with replay.
-Here we develop a flexible state space model that uses a combination of discrete and latent states to capture the content and time evolution of SWR events.
-Application of this model reveals the large majority of SWR events contain coherent spatial content that evolves in accordance with a surprisingly wide range of mental movement speeds.
+Representations of past and possible future experiences play a critical role in memory and decision-making processes.
+The hippocampus is known to express such representations during sharp-wave ripple (SWR) events, when ensembles of hippocampal place cells can be activated in time-compressed sequences consistent with prior experiences.
+Efforts to understand these "replay" events have typically imposed stringent criteria on event significance, and thus analyze only a small subset of events.
+As a result, we lack a clear picture of both the prevalence and the range of dynamics associated with replay.
+Here, we develop a flexible state space model that uses a combination of discrete and latent states to capture the spatial content and time evolution of replay during SWR events.
+Application of this model reveals that the large majority of SWR events contain coherent spatial content that evolves in accordance with a surprisingly wide range of mental movement speeds.
 
@@
 \section*{Introduction}
+
 The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
@@
 Hippocampal replay events are a prototypical example of a generative representation.
-As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells can fire in sequence as the animal moves through a series of locations.
+As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 
-Identifying these events typically involves multiple steps and assumptions about their nature.
+Identifying these events typically involves multiple steps and assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
@@
 The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which instantiates the assumption that the representations progress at a constant speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations smoothly progress at a constant speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
 
-While this approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bin acts as a boxcar smoother that limits the potential movement speeds of the representation.
+While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bin acts as a boxcar smoother that limits the potential movement speeds of the representation.
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Approaches that focus on the order of cell activity within each even \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using some arbitrarily smoothed version of the spike train.
+Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using some arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, there is now evidence that a subset of these events have stationary representations of a single location, which both approaches often exclude.
+Moreover, there is now evidence that a subset of these events have stationary representations of a single location \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, which both approaches often exclude.
 In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
@@
 On the other hand, Stella et al.
-\cite{StellaHippocampalReactivationRandom2019} found that replays are more spatially continuous, following Brownian diffusion dynamics.
+\cite{StellaHippocampalReactivationRandom2019} found that replays are spatially continuous, following Brownian diffusion dynamics.
 Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} and characterizes represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 
-We find that the vast majority of SWRs contain spatially coherent content.
-Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at speeds consistent with real animal movement, suggesting that high speed replay is not the most common manifestation of generative activity.
+We find that the majority of SWRs contain spatially coherent content; that is, the replay trajectories have no large jumps.
+Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real world animal movement, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 \caption{Components of the state space model and demonstration on simulated data.
-\textbf{(A)} 19 simulated place cells that exhibit three different sequence patterns.
-For the first third of time, one cell fires repeatedly, representing one stationary location.
-For the second third of time, the cells fire in sequence, representing a smooth continuous trajectory through space.
-For the last third cells fire in an incoherent pattern, representing a non-smooth fragmented trajectory through space.
-\textbf{(B)} Like the Bayesian Decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
-\textbf{(C)} This is combined with an explicit model of each dynamic state--which determines how latent position can change based on the posterior from the previous time step.
-In the stationary state, this is modeled as the identity matrix because the position remains the same.
-In the continuous state, the dynamics are modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close.
-In the fragmented state, the dynamics are modeled as uniform, meaning the next position is equally likely to be anywhere.
-Note that the dynamics displayed only reflect the within-state dynamics.
-\textbf{(D)} The probability of remaining in a particular state versus switching to another state.
-This is modeled as having a high probability of remaining in a particular state with a small probability of switching to one of the other dynamics at each time step.
-\textbf{(F)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
-The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic state over time (lines).
-This can also be used to get an estimate of latent position over time (right panel).
-\textbf{(G)} The probability of each state given a particular constant speed trajectory.
-Each speed corresponds to a simulated spiking sequence at that speed.
-The lines correspond to the average probability of that state over the sequence.
-Dotted line represents the 0.8 classification threshold we use to classify each range of speeds and shaded regions correspond to the classification we have given the range of speeds.
+\textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics.
+For the first 60 ms, one cell fires repeatedly, representing one stationary location.
+For the second 60 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
+For the last 60 ms, cells fire randomly, out of their spatial order, representing an incoherent spatial sequence.
+\textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
+\textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the posterior from the previous time step.
+In the stationary movement dynamic, the transition matrix is modeled as the identity matrix because the position remains the same.
+In the continuous movement dynamic, the transition matrix is modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close to the previous position.
+In the fragmented movement dynamic, the transition matrix is modeled as uniform, meaning the next position is equally likely to be any possible position.
+Note that the movement dynamics displayed only reflect the within-state movement dynamics.
+\textbf{(D)} The probability of remaining in a particular dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
+\textbf{(E)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
+The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic state over time.
+The summed posterior across states also provides an estimate of latent position over time (right panel).
+\textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
+The lines correspond to the average probability of that dynamic for a given constant velocity sequence.
+We use a 0.8  threshold (dotted line) to classify each sequence based on the state or states which contribute maximally to the posterior.
 }
@@
 Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then apply our decoding algorithm to the spiking sequence in Figure \ref{1}A.
-For the first third of this sequence, a single place cell fires repeatedly, representing a single location.
-For the middle third of the sequence, the cells fire in sequential spatial order, representing a fast moving trajectory.
-For the last third, the cells fire in a incoherent spatial order.
-These firing patterns represent three different types of sequence dynamics that could be expressed during replay events, which we call stationary, continuous, and fragmented, respectively.
+We then construct an 180 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
+For the first 60 ms of this sequence, a single place cell fires repeatedly, resulting in the extended representation of a single location.
+For the second 60 ms of the sequence, the cells fire in sequential spatial order, representing a fast moving trajectory across the virtual linear track.
+For the last 60 ms of the sequence, the cells fire in an incoherent spatial order.
+These three firing patterns represent three different types of sequence dynamics that could be expressed during replay events, which we call stationary, continuous, and fragmented, respectively.
 The goal of our model is to characterize SWRs in terms of a mixture of these three dynamics at every time point.
@@
 Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamics model---how the position can change over time.
-For the data model, our decoder is the same as the Bayesian decoder.
-We compute an estimate of how each cell's firing rate varies over position (the place field, Figure \ref{1}B).
+For the data model, our decoder is the same as the standard ("Bayesian") decoder \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}.
+We compute an estimate of how each cell's firing rate varies over position during movement (i.e.
+the place field, Figure \ref{1}B).
 This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the Bayesian decoder, we can use arbitrarily small time bins (in our case 2 ms vs.
+In contrast to the standard decoder, we can use arbitrarily small time bins (in our case 2 ms vs.
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the Bayesian decoder.
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder.
 
 Next, we specify a movement dynamics model for how latent position---the "mental" position of the animal represented by the cells---evolves over time (Figure \ref{1}C).
-We do this by defining a movement state transition matrix, which defines how the latent position can change from the previous time step.
-Previous findings suggest three distinct dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, smoothly moving \cite{DavidsonHippocampalReplayExtended2009}, and fragmented / jumpy \cite{PfeifferAutoassociativedynamicsgeneration2015}, which could correspond to both spatially incoherent representations and representations that jump from one position to another.
-We therefore developed state transition models to capture each of these dynamics.
+We do this by defining a state transition matrix, which defines how the latent position can move from the previous time step.
+Previous findings suggest that replay may exhibit at least three distinct types of movement: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented \cite{PfeifferAutoassociativedynamicsgeneration2015}, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another.
+We therefore define state transition models to capture each of these dynamics.
 
-In the stationary dynamic, the latent position would not change between time steps, so we use an identity state transition matrix, which predicts the next position will be the same as the last position.
-In the continuous dynamic, the latent position would be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+In the stationary dynamic, the latent position does not change between time steps.
+The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position.
+In the continuous dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
 This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
-In our case, since we are interested in identifying replays that move at speeds much faster than the animal's speed, we set this to 6.0 cm.
-This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (corresponding to a latent speed of up to ~25 m/s).
-This is consistent with replay speeds observed in previous studies \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}.
-Finally, in the fragmented dynamic, the latent position could move to any available position instantaneously.
-We model this using a uniform state transition matrix, which makes all positions equally likely.
+In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm.
+This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to ~25 m/s are most likely).
+Finally, in the fragmented dynamic, the latent position can move to any available position instantaneously.
+We model this using a uniform state transition matrix, which makes transitions to all positions equally likely.
 
 Finally, we specify how likely each dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of the SWR.
-There is, however, a small probability of switching to one of the other dynamics (note that below we show that our results are relatively insensitive to this parameter).
+In order to be conservative, we assume that each dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
+There is, however, a small probability of switching to one of the other dynamics.
 Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step.
@@
 Importantly, the data drives the estimated dynamic, so even if the probability of a particular dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
+In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
 
 Once we have specified the data and dynamics model, we have fully specified the state space model.
-We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic state (see Methods).
-With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time and the probability of position over time, irrespective of dynamic (Figure \ref{1}F, left and right plot respectively).
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
+With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time (by marginalizing over position) and the probability of latent position over time, irrespective of dynamic (by marginalizing over dynamic; Figure \ref{1}F, left and right plot respectively).
 
-An examination of these two summaries reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
-The stable firing of the one active neuron indicates a stationary representation, and the probability of the stationary dynamic is high at the beginning of the simulated replay.
-The change in the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spiking from the population of cells.
-Subsequently, as the population activity becomes spatially incoherent, the fragmented dynamic dominates for the last third.
+An examination of the two summaries shown in Figure \ref{1}F reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+The stable firing of the one active neuron indicates a stationary representation, and accordingly, the probability of the stationary dynamic is high at the beginning of the simulated replay.
+A change in the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spatially sequential spiking from the population of cells.
+Subsequently, as the population activity becomes spatially incoherent, the fragmented dynamic dominates for the last third of the simulated event.
 Importantly, these dynamics also inform the estimate of latent position of the animal, and the correct assignment of dynamic reduces the variance of the latent position estimate.
@@
 To demonstrate this, we applied the model to simulated trajectories of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
-From this, we can see that not only are there regions of speed that correspond to our three dynamics being high probable (where we define highly probable to be over 0.8 probability), there are intermediate speeds where two of the dynamics dominate with high probability; that is, the sum of two of the dynamics' probabilities is over 0.8.
+From this, we can see that not only are there regions of speed that correspond to our three dynamics being highly probable (where we define highly probable to be over 0.8 probability), there are intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
 In this manuscript, we will refer to these as mixture dynamics.
@@
 
-We note here that the choice of any particular threshold is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95 because the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are meaningful in the brain.
+We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
+Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold risks giving the false conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
 Nonetheless, we also verified that our central results hold with a higher threshold of 0.95.
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v4}
-\caption{Demonstration of the state space model on real data using both sorted and clusterless spikes on the same SWR.
+\caption{Demonstration of the state space model on hippocampal data using both sorted and clusterless spikes on the same SWR event.
 \textbf{(A)} Decoding using sorted spikes.
 The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
-The second panel shows the probability of each dynamic over time as in Figure 1F.
-Shaded regions correspond to the speed classifications as in Figure 1G.
+The second panel shows the probability of each dynamic over time as in Figure \ref{1}F.
+Shaded regions correspond to the speed classifications as in Figure \ref{1}G.
 The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
@@
 The top panel shows multiunit spiking activity from each tetrode.
-Other panels have the same convention as(A).
-\textbf{(C)} Maximum of the 1D decoded position (the most probable position) in (A) projected back into 2D.
+Other panels have the same convention as (A).
+\textbf{(C)} Maximum of the 1D decoded position (the most probable position) in (A) projected back onto the 2D maze for the sorted decoding.
 Color indicates time.
 The animal's actual position is denoted by the red circle.
-Light grey lines show the animal's 2D position over the entire epoch.
+Light grey lines show the animal's 2D position over the entire recording session.
 L, R, and C correspond to the left, right and center reward well as in (A).
-\textbf{(D)} Maximum of the 1D decoded position (the most probable position) in (B) projected back into 2D.
+\textbf{(D)} Maximum of the 1D decoded position (the most probable position) in (B) projected back into 2D using the clusterless decoding.
 Figure conventions are the same as in (C).
@@
 \caption{A-D.
-Four examples of replays that would not be well-characterized by using the Bayesian decoder.
-Figure conventions are the same as in Figure 2.
-\textbf{(A)} A replay that starts down the center arm away from the animal's position at the center well, slows down, and returns back.
-\textbf{(B)} A replay that stays at the left well for the entirety of the SWR while the animal is in the center arm.
-\textbf{(C)} A replay that begins stationary at the left well and then jumps to the middle of the right arm and proceeds up the right arm to the right well.
-\textbf{(D)} A replay that begins stationary at the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
-\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
+Four examples of replays that would not be well-characterized by using the standard decoder.
+Figure conventions are the same as in Figure \ref{2}.
+\textbf{(A)} A replay that starts moving down the center arm away from the animal's position at the center well, slows down, and returns back.
+\textbf{(B)} A replay that persistently represents the left well for the entirety of the SWR while the animal remains in the center arm.
+\textbf{(C)} A replay that begins with stationary representation of the left well, then jumps to the middle of the right arm and proceeds up the right arm to the right well.
+\textbf{(D)} A replay that begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
+\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories of dynamics (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
 \textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Stationary, Stationary-Continuous-Mix, or Continuous content.
@@
 
-Our success with simulated data inspired a further validation of the model on real hippocampal data.
+We next sought to validate the model on real hippocampal data.
 We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with sequential population activity (Figure \ref{2}A, top panel).
-As expected, we observed that the probability of being in the continuous dynamic is high throughout this SWR, but the probability of being in a stationary state was noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary state was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
 Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
 This is also evident in the posterior probability of linear position over time.
-This shows that the replay travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
-We can also see this when we project the maximum of the posterior of this trajectory to 2D (Figure \ref{2}C) to better see the spatial trajectory.
-Moreover, when we apply the same model using the 2D position of the animal, we get a similar result as when we used the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).
+This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
+We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D (Figure \ref{2}C) to better see the spatial trajectory on the maze.
+Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).
 
-One of our criteria for a more optimal method is that it use all of the available spiking data.
-Using only clustered spikes discards all of the spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the quality of the resulting decoding.
-Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}.
+One of our criteria for a more optimal method is for it to use all of the available spiking data.
+Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
 We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding (see Methods).
 Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}).
-Applying a clusterless decoder to the same SWR, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
-The spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model, however.
-This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event, and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).
+Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
+As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model.
+This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).
 
@@
 
-After testing our model on a single SWR, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}).
-One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see Methods) to ensure that we were included both the rare large events as well as the much more common small events.
+After testing our model on a single SWR event, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}).
+One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see Methods) to ensure that we included both the rare large events as well as the much more common small events.
 
-As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of SWR (Figure \ref{fig:Figure2-Figure supplement 2}A-C).
+As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR (Figure \ref{fig:Figure2-Figure supplement 2}A-C).
 However, we also observed many events with spatially coherent content that did not have this structure.
-For example, there were trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
+For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
 
-Using our 0.8 threshold, we were able to classify 89\% ( 23382 of 26160) of SWRs as containing at least one of the five dynamics.
-To ensure that this depended on the spatial receptive fields of cells in the hippocampus, we resampled position with replacement for two epochs, shuffling the relationship between spiking and spatial information.
+Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three states or common state mixtures.
+To ensure that this depended on the sequential firing of spatially tuned cells in the hippocampus, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
 We then decoded the same SWR events containing the same spikes.
-Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for epoch 1, p=0.02 for epoch 2, Figure \ref{fig:Figure3-Figure supplement 2}).
+Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, Figure \ref{fig:Figure3-Figure supplement 2}).
 
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
-We therefore asked what fraction of classified events contained spatially coherent content in any form, defined as a SWR containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 22478 of 23382 or 96\% of classified SWRs (85\% of all SWRs) included spatially coherent structure.
-We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented-continuous mixture or fragmented dynamics.
+We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
+We found that 96\% (22478 of 23382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}F).
+We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
 We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure.
-This was consistent across animals (Figure \ref{3}F).
 
-To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the Bayesian decoder.
-Here our findings were consistent with previous reports: in our datasets 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G, 17\% of all SWRs).
-Thus, focusing on only continuous content misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
+To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
+Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had some continuously moving content (Figure \ref{3}G, 17\% of all SWRs).
+Thus, focusing on only spatially continuous trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
 
-We repeated this analysis with a higher threshold of 0.95 to ensure that our result was not dependent on this threshold.
-We found that, while this slightly reduced the total fraction of classified ripples (19180 / 25845 or 74\% SWRs), an even higher fraction of the classified ripples (19020 / 19180 or 99\% classified) had spatially coherent content.
-Classified SWRs containing spatially incoherent content similarly were a small fraction of the classified SWRs (spatially incoherent: 484 / 19180 or 3\%).
+We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
+We found that, while this change slightly reduced the total fraction of classified SWRs (19180 / 25845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19020 / 19180 or 99\% classified) had spatially coherent content.
+Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (spatially incoherent: 484 / 19180 or 3\%).
 
@@
 This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
-To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) values of the posterior linear position.
-Larger values of the cumulative HPD indicate the model is less certain about position, because the HPD values are spread out over more of the track at a given time point, whereas lower values indicate that the model is more certain about the estimate of position, because the HPD values are more concentrated and cover less of the track.
-Thus, the HPD provides a complementary measures of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
+To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) region of the posterior linear position.
+Larger values of the cumulative HPD extent indicate the model is less certain about position, because the HPD values are distributed over more of the track at a given time point.
+In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
+Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
 
-We found that spatially coherent events did indeed have lower cumulative HPDs than events with fragmented dynamics.
+We found that spatially coherent events did indeed have lower cumulative HPD sizes than events with fragmented dynamics.
 Figure \ref{4}A and B show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
-The bulk of the HPD values at each time step in these SWRs is concentrated in a small portion of the track and the cumulative HPD is low throughout the SWRs.
-In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD is much higher.
+The bulk of the HPD values at each time step in these SWRs is concentrated in a small portion of the track and the cumulative HPD size is low throughout the SWRs.
+In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD size is much higher.
 The HPD also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in Figure \ref{4}D, where the cumulative HPD is small for most of the SWR until the end, where the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average cumulative HPD for each SWR, grouped by dynamic, we found a clear bi-modal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+An example of this change is shown in Figure \ref{4}D, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
 For the spatially coherent dynamics, the cumulative HPD was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
@@
 \textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density.
-Top panel.
-Probability of dynamic over time.
-Shading and labels indicates dynamic categories.
-2nd panel.
-Posterior linear position over time.
+Top panel: Probability of dynamic over time.
+Shading and labels indicate dynamic categories.
+2nd panel: Posterior distribution of estimated linear position over time.
 Red triangle indicates animal's position.
-3rd panel.
-Black indicates the position bin is in the 95\% highest posterior density (HPD) values.
-4th panel.
-Cumulative HPD.
-\textbf{(E)} Average Cumulative 95\% HPD for each dynamic category.}
+3rd panel: Posterior distribution plot, where black designated the position bins that are included in the 95\% highest posterior density (HPD) region.
+4th panel: Cumulative HPD size.
+\textbf{(E)} Average Cumulative 95\% HPD size for each dynamic category.}
 \label{4}
@@
 Analysis of SWRs from all animals.
-\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each ripple.
-Each row represents a possible category present in the SWR.
-Each column represents a set of dynamics categories, where filled-in black dots with an edge between the dots indicates that multiple categories are present in the SWR (at temporally distinct times).
-The sets are ordered by how often they occurred as indicated by the bar plot above each category.
-The total number of each category is indicated by the rotated bar plot on the left.
+\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each SWR.
+Each row represents a possible dynamic and each column represents a set of dynamics categories, where filled-in black dots with an edge between the dots indicates that multiple categories are present in the SWR (at temporally distinct times).
+The sets are ordered by how often they occur as indicated by the bar plot above each category.
+The total number of each classification is indicated by the rotated bar plot on the left.
 \textbf{(B)} Distribution of the duration of each dynamic category within a SWR.
@@
 \textbf{(D)} Position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
-\textbf{(E)} Distribution of multiunit spike rates for each dynamic category within each SWR.
+\textbf{(E)} Distribution of average tetrode multiunit spike rates for each dynamic category within each SWR.
 }
@@
 
-Our results indicate that most events with coherent spatial content are not dominated by the continuous dynamic, and thus correspond to trajectories that move relatively slowly compared to their purely continuous counterparts.
+Surprisingly, our results indicate that most events with coherent spatial content are not dominated by the continuous movement dynamic, and thus correspond to trajectories that are stationary or that move relatively slowly compared to their purely continuous counterparts.
 We therefore examined these events in more detail.
-We first noted that most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic.
-Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at rates more consistent with real-world movement (Figure \ref{1}G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic.
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
+These events contain representations that move at rates more consistent with real-world movement (Figure \ref{1}G), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 
-The next most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\%).
-Unlike the stationary-continuous mixtures, most of these events represented a location close to the the animal's actual position (Figure \ref{5}C).
+The second most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\%).
+Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C).
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify "non-local stationary trajectories" and found these trajectories in 1585 of 23382 (~7\%) of classified SWRs.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such sequences in 1585 of 23382 (~7\%) of classified SWRs.
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary trajectories most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of these trajectories occurred at other locations of the track.
+These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
 
-Finally, we examined the overall activity levels across dynamics.
-We found that all the dynamics had similar average tetrode multiunit spiking rates (Figure \ref{5}E).
+Finally, we examined the overall levels of neuronal spiking across dynamics.
+We found that all the dynamics had similar average multiunit spiking rates (Figure \ref{5}E).
 This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
@@
 We developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
-This model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
-We show that this model is interpretable: it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before.
-This model is also flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with representations that move more or less linearly at a constant, high speed , but also representations that move non-linearly  with slower speeds or are stationary.
-These slower moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical Bayesian decoder.
+This model is robust; it is relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
+We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been explained before.
+This model is also flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move more or less linearly at a constant, high speed, but also representations that move non-linearly,  with slower speeds, or that are stationary.
+These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
 The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and real-world speeds.
 
-Previous work reported that at most half of SWRs or high-population-activity events contained replay \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work reported that less than half of SWRs or high-population-activity events contained replay \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
-Our results indicate that this is not the case: in our dataset the large majority of SWRs (85\%) contained spatially coherent content, defined as content with some combination of stationary or continuous dynamics.
-These events included decoded representations that were highly spatial concentrated, confirming their spatial coherence.
+Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
+These events included decoded representations that were highly spatially concentrated, confirming their spatial coherence.
 This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with real-world movement speeds of <1 m/s.
-The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a representation of a single location, while the classical "replay" events made up only about ~19\% of classified events.
+The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the classical "replay" events made up only about 19\% of classified events.
 
 These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
-Instead there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
 Interestingly, Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
@@
 Our model also makes it possible to identify different dynamics within the same SWR.
-As the boundaries of each event are also arbitrary, we suggest that knowing the start and end of each dynamic may prove a powerful way to understand how inputs to the hippocampus influence these events and how these events in turn influence downstream structures.
+As the start and end time of each SWR event are also dependent on a selection of arbitrary parameters, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
+Examination of such periods may shed light on when and why the hippocampus may be representing current position, past experience, or possible future locations, to understand how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
 
@@
 Our model goes beyond the Johnson model in that it explicitly permits a mixture between the movement speeds, can work for arbitrary track geometries, and uses clusterless decoding.
-Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic discovers of sequential patterns and the typical Bayesian decoder, which typically use arbitrarily chosen bin sizes and  restrictive assumptions about the nature of the trajectories.
+Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
 In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results without the need to make assumptions about the right temporal bin size.
 
-State-space models like the one we use here can enable a richer set of analyses of generative events that takes advantage of all of the available data.
-These approaches can be applied to not just SWRs but all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.
+State-space models like the one we use here can enable a richer set of analyses of generative events that take advantage of all of the available data.
+These approaches can be applied not just to SWRs but to all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.
 The model can be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
-It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+It can also be extended to account for trajectory task context and forward/reverse sequences as in Deng et al.
+\cite{DengRapidclassificationhippocampal2016}.
 Finally, models can be built to estimate any covariate, including movement direction \cite{KayConstantSubsecondCycling2020}.
@@
 \begin{acknowledgements}
-We thank Margaret Carr Larkin and Mattias P. Karlsson for use of their openly available datasets.
+We thank Margaret Carr Larkin and Mattias P.
+Karlsson for use of their openly available datasets.
+We also thank ...
+for feedback on the manuscript.
+This work was supported by the Simons Foundation for the Global Brain Grants 542971 (U.T.E.) and 542981 (L.M.F.) and the Howard Hughes Medical Institute (L.M.F.).
 \end{acknowledgements}
@@
 Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
-We only used hippocampal tetrodes (CA1, CA2, CA3) in this study.
+We only used tetrodes located in the CA1, CA2, and CA3 subregions in this study.
 
@@
 All animals performed a W-track spatial alternation task, which is described in detail in \cite{KarlssonAwakereplayremote2009}.
-In brief, each day, animals alternated between 20 minute rest epochs in a rest box and 15 minutes run epochs in the W-track.
-On the W-track, animals were rewarded at the ends of an arm when that arm was the next correct arm in the sequence.
-Two rules determined the next correct arm.
-If the animal was in an outer arm, it had to run to the center arm.
-If the animal was in the center arm, it had to run from the center arm to the least recently visited outer arm.
+In brief, each day, animals alternate between 20 minute rest sessions in a rest box and 15 minutes run sessions on the W-shaped track equipped with a reward well at each arm end.
+On the W-track, animals are rewarded at the ends of an arm when that arm is the next correct arm in the sequence.
+Two rules determine the next correct arm.
+If the animal is in an outer arm, it must next visit the center arm.
+If the animal is in the center arm, it has to next visit the less recently visited outer arm.
+Correct performance of these rules result in the visit sequence: center, left, center, right, center, left, etc.
 Animals were free to choose any arm at any time.
@@
 \subsection*{Position of the Animal and Linearization}
-The animal's 2D position was estimated from digital video (30 Hz) of two infrared diodes placed on the head stage preamplifiers using a semi-automated analysis.
+The animal's 2D position was estimated from digital video (30 Hz) of two infrared diodes placed on the headstage preamplifiers using a semi-automated analysis.
 In order to decrease the time it takes to run the model, the 2D position of the animal was converted into a 1D position.
 This is done by first defining a 2D graph representation of the track (herein referred to as the track graph), where edges correspond to segments of the W-track and nodes represent intersection points between those segments.
-Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to classify the most likely track segment the animal is on.
+Then, based on the algorithm in \cite{NewsonHiddenMarkovmap2009}, we use a Hidden Markov Model (HMM) to assign the position detected in each video frame to the most likely track segment.
 Using the HMM takes into account the time dependent nature of the data and helps prevents sudden jumps from one track segment to another, which is particularly important near intersections.
 The observation model of the HMM is Gaussian and it models the likelihood of being on a track segment as the Gaussian distance to that segment with a 5 cm standard deviation.
-The state transition model is an empirically estimated state transition that changes with each time point that tries to ensure that the Euclidean distance between successive position estimates is similar to the shortest path distance along the graph between successive position estimates.
-A slight bias of 0.1 is given to the diagonal to encourage staying on the same track segment.
-The most likely track segment the animal is on is computed using the Viterbi algorithm.
+The state transition model is empirically estimated, and changes with each time point to ensure that the Euclidean distance between successive position estimates is similar to the shortest path distance along the graph between successive position estimates.
+A slight bias of 0.1 is given to the diagonal of the state transition model to encourage staying on the same track segment.
+The most likely track segment the animal is on is computed using the Viterbi algorithm(cite?).
 After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the nearest point of the track segment.
@@
 The linear distance can then be converted into a linear position by assigning each track segment a position in 1D space.
-15 cm gaps were placed between the center arm, left arm, and right arms in 1D space to prevent any smoothing done in the model from bleeding over inappropriately.
+15 cm gaps were placed between the center arm, left arm, and right arms in 1D space to prevent any smoothing done in the model from influencing neighboring segments inappropriately.
 The code used for linearization can be found at \url{https://github.com/Eden-Kramer-Lab/loren_frank_data_processing}.
@@
 \subsection*{Sorted Spikes, Multiunit Spikes and Waveform Features}
-To obtain the neural data used for decoding, hippocampus electrical potentials were recorded at 30 kHz, referenced to a corpus callosum tetrode, and then digitally filtered between 600 Hz and 6 kHz.
-Events were considered threshold crossing events if any one of the four tetrode wires exceeded a 60 $\mu V$ threshold.
-These multiunit threshold crossing events and the associated peak amplitudes of the tetrodes wires---defined as the electrical potential value at the time of the maximum amplitude at the threshold crossing---were used as data for the clusterless version of the model.
+To obtain the neural spiking data used for decoding, electrical potentials from rat hippocampus were recorded at 30 kHz, referenced to a tetrode located in the corpus callosum, and then digitally filtered between 600 Hz and 6 kHz.
+Spiking events were detected as any potential exceeding a 60 $\mu V$ threshold on any one of the four tetrode wires of a tetrode.
+The electrical potential value on each wire of the tetrode at the time of maximum potential of any of the four wires was used as the waveform feature in the clusterless decoding.
 
@@
 Sharp wave ripples were detected using the same method as in \cite{Kayhippocampalnetworkspatial2016}.
-Each CA1 LFP was obtained by downsampling the original 30 kHz electrical potential to 1.5 kHz, bandpass filtering between 0.5 Hz and 400 Hz.
-This was further bandpass filtered between 150-250 Hz for ripple band frequencies, squared and then summed across tetrodes---forming a single population trace over time.
+Each CA1 LFP was obtained by downsampling the original 30 kHz electrical potential to 1.5 kHz and bandpass filtering between 0.5 Hz and 400 Hz.
+This was further bandpass filtered for the ripple band (150-250 Hz), squared, and then summed across tetrodes---forming a single population trace over time.
 This trace was smoothed with a Gaussian with a 4 ms standard deviation and the square root of this trace was taken to get an estimate of the population ripple band power.
-Candidate SWR times were found by z-scoring the population power trace and finding times when the z-score exceeded 2 standard deviations for a minimum of 15 ms and the speed of the animal was less than 4 cm/s.
-The SWR times were then defined as times when the z-score was greater than the mean and contained the candidate SWR times.
+Candidate SWR times were found by z-scoring the population power trace of an entire recording session and finding times when the z-score exceeded 2 standard deviations for a minimum of 15 ms and the speed of the animal was less than 4 cm/s.
+The SWR times were then extended before and after the threshold crossings to include the time until the population trace returned to the mean value.
 The code used for ripple detection can be found at \url{https://github.com/Eden-Kramer-Lab/ripple_detection}.
@@
 \subsection*{The Model}
-Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and $I_{k}$ be a discrete latent variable that is an indicator for the dynamics we wish to movement characterize: stationary, continuous, and fragmented.
-The goal of the model is to simultaneously estimate the posterior probability of position and dynamics $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed spiking data from time 1 to time $T$.
-The observed data can be either spike trains $\Delta N_{1:T}^{(1:C)}$ from $C$ putative cells when decoding with sorted spikes or multiunit spikes $\Delta N_{1:T}^{(1:E)}$ and their associated wave form features $\Vec{m}$ from each tetrode $E$ when decoding with clusterless spikes.
+Let $x_{k}$ be a continuous latent variable that corresponds to the position represented by the population of cells at time $t_k$ and let $I_{k}$ be a discrete latent variable that is an indicator for the movement dynamics we wish to characterize: stationary, continuous, and fragmented.
+The goal of the model is to estimate simultaneously the posterior probability of position and dynamics $p(x_k, I_k \mid O_{1:T})$, where $O_{1:T}$ corresponds to the observed spiking data from time 1 to time $T$.
+The observed data can be either spike trains $\Delta N_{1:T}^{(1:C)}$ from $C$ putative cells when decoding with sorted spikes or multiunit spikes $\Delta N_{1:T}^{(1:E)}$ and their associated wave form features $\Vec{m}^i_{k,j}$ from each tetrode $E$ when decoding with clusterless spikes, where $i \in 1:E$, $k \in 1:T$, $j \in 1:\Delta N_k^i$.
 
-We have previously shown \cite{DenovellisCharacterizinghippocampalreplay2019} that the posterior probability $p(x_k, I_k \mid O_{1:T})$ can be estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0} \mid O_{0})$ and iterating to time T:
+We have previously shown \cite{DenovellisCharacterizinghippocampalreplay2019} that the posterior probability $p(x_k, I_k \mid O_{1:T})$ can be estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0})$ and iterating to time $T$:
 
@@
 $$
-and then applying the acausal smoother equation, starting from the last estimate of the casual filter $p(x_{T}, I_{T} \mid O_{1:k})$ and recursively iterating backwards to time 1:
+and then applying the acausal smoother equation, starting from the last estimate of the casual filter $p(x_{T}, I_{T} \mid O_{1:T})$ and recursively iterating backwards to time $1$:
 $$
@@
 
-For the initial conditions $p(x_{0}, I_{0})$, we set each dynamic $I_0$ to have uniform probability over all possible positions $\mathcal{U}(\min x, \max x)$, reflecting the fact that we do not have any prior knowledge about which dynamic or position is more likely:
+For the initial conditions $p(x_{0}, I_{0})$, we set each dynamic $I_0$ to have uniform $1/3$ probability and each initial latent position to have uniform probability density over all possible positions $\mathcal{U}(\min x, \max x)$, reflecting the fact that we do not have any prior knowledge about which dynamic or position is more likely:
 
 \begin{equation*}
-  p(x_{0}, I_{0}) = 
-  \begin{blockarray}{*{1}{c} l}
-    \begin{block}{[*{1}{c}]>{$\footnotesize}l<{$}}
-      \mathcal{U}(\min x, \max x) \bigstrut[t]& $I_{0} = Stationary$ \\
-      \mathcal{U}(\min x, \max x) & $I_{0} = Continuous$ \\
-      \mathcal{U}(\min x, \max x) & $I_{0} = Fragmented$ \\
-    \end{block}
-  \end{blockarray}
+  p(x_{0}, I_{0}) = \frac{1}{3} \mathcal{U}(\min x, \max x)
 \end{equation*}
@@
 $$
-p(O_{k} \mid x_{k}) = p(\Delta N_{k}^{(1:E)}, \vec{m_k} \mid x_{k}) \propto
-\prod^{E}_{i=1} [\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k})\Delta_{k}]^{N_{t_{k}}^{i}} \\
+p(O_{k} \mid x_{k}) = p(\Delta N_{k}^{(1:E)}, \{\vec{m}^i_{k,j}\}^{i=1:E}_{j=1:\Delta N_{k}^i} \mid x_{k}) \propto
+\prod^{E}_{i=1} \prod^{\Delta N_{k}^i}_{j=1} [\lambda_{i}(t_{k}, \vec{m}^i_{k,j} \mid x_{k})\Delta_{k}] \\
     \exp[-\Lambda_{i}(t_{k} \mid x_{k})\Delta_{k}]
 $$
-where $\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
+where $\lambda_{i}(t_{k}, \vec{m}^i_{k,j} \mid x_{k})$ is now a generalized firing rate that depends on an associated wave form features $\Vec{m}$ and $\Lambda_{i}(t_{k} \mid x_{k})$ is a marginal firing rate that is equivalent to a place field estimated on multiunit spikes.
 Both of these rates can be defined as:
 $$
-\lambda_{i}(t_{k}, \vec{m}_k \mid x_{k}) = \mu_i \frac{p_{i}(x_k, \vec{m}_k)}{\pi(x)}
+\lambda_{i}(t_{k}, \vec{m}^i_{k,j} \mid x_{k}) = \mu_i \frac{p_{i}(x_k, \vec{m}^i_{k,j})}{\pi(x)}
 $$
@@
 $$
-where $\mu_i$ is the mean firing rate for tetrode $i$, $\pi(x)$ is the spatial occupancy of the animal on the track, and $p_{i}(x_k, \vec{m}_k)$ and $p_{i}(x_k)$ are the joint probability of observing waveform features $\Vec{m}$ and position $x$ and the probability of observing position $x$, respectively.
-$\pi(x)$, $p_{i}(x_k, \vec{m}_k)$, $p_{i}(x_k)$ can all be estimated by training a kernel density estimator on each tetrode's spike waveform features and corresponding position during the encoding period.
+where $\mu_i$ is the mean firing rate for tetrode $i$, $\pi(x)$ is the smoothed spatial occupancy of the animal on the track, $p_{i}(x_k)$ is the smoothed spatial occupancy only at times when spikes occur in tetrode $i$, and $p_{i}(x_k, \vec{m}^i_{k,j})$ is the smoothed occupancy in the space of both space and waveform features.
+$\pi(x)$, $p_{i}(x_k, \vec{m}^i_{k,j})$, $p_{i}(x_k)$ can all be estimated by training a kernel density estimator on each tetrode's spike waveform features and corresponding position during the encoding period.
 
@@
 In order to encode how each cell's spiking activity relates to position (the place field), we fit a generalized linear model (GLM) with a Poisson response distribution to each cell's spiking activity during the encoding period, which we define as all movement times (time periods when the running speed is greater than 4 cm/s).
-We estimate the parameters $\beta$, which consist of $\beta_{0}$, the average firing rate over time, and $\beta_{i}$, weights for third degree B-splines basis functions $f_{i}$ over position (or tensor products of the B-splines when position is two dimensional).
+We estimate the parameters $\beta$, which consist of $\beta_{0}$, the baseline firing rate over time, and $\beta_{i}$, weights for third degree B-spline basis functions $f_{i}$ over position (or tensor products of the B-splines when position is two dimensional).
 B-spline basis functions are used because place field firing activity is assumed to vary smoothly over position and this prior knowledge can be exploited to reduce the total number of model parameters needed.
 Each basis function is spaced every 5 cm over the range of the position and zero constrained so that the change encoded by the parameters is relative to the baseline firing rate.
-We use a log link function to convert the linear combination of parameters to an instantaneous firing rate over time $\lambda(t)$ to ensure the rate is always positive.
+We use a log link function to convert the linear combination of parameters to an instantaneous firing rate over time $\lambda(t_k)$ to ensure the rate is always positive.
 
-$$log(\lambda(t)) = \beta_{0} + \sum_{i} f_{i}(position)\beta_{i}$$
+$$log(\lambda(t_k)) = \beta_{0} + \sum_{i} f_{i}(x_k)\beta_{i}$$
 
-A small L2 penalization term $-\lambda\|\beta_{i}\|_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low.
+A small L2 penalization term $-\lambda\|\beta_{i}\|_{2}^{2}$ is used to prevent model fitting instability when spiking activity is very low.
 We set this to 0.5 for all cells.
@@
 $$
-kde(y) = \frac{1}{N h_1...h_D} \sum^{N}_{i=1} \prod^{D}_{d=1} K_{d}\left(\frac{y_d - y_{i,d}}{h_d}\right)
+kde(y) = \frac{1}{N h_1\cdot\cdot\cdot h_D} \sum^{N}_{i=1} \prod^{D}_{d=1} K_{d}\left(\frac{y_d - y_{i,d}}{h_d}\right)
 $$
@@
 For $p_{i}(x_k)$, $y_{i, d}$ is all positions at the time of multiunit spikes during movement.
-For $p_{i}(x_k, \vec{m}_k)$, $y_{i, d}$ is all positions at the time of multiunit spikes and their associated waveform features during movement.
+For $p_{i}(x_k, \vec{m}^i_{k,j})$, $y_{i, d}$ is all positions at the time of multiunit spikes and their associated waveform features during movement.
 We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters were found to minimize decoding error in Kloosterman et al.
@@
 This was done using Scipy image processing tools \cite{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
-This grid based approximation allows us to use Reimann sums to approximate the integrals in the causal filter and acausal smoother equations.
+This grid based approximation allows us to use Riemann sums to approximate the integrals in the causal filter and acausal smoother equations.
 We chose a position grid with bins of 3 cm in width (and height if the model was computed with 2D positions) in order to get good resolution for the random walk transition matrices (which had 6 cm variance) as well as for the clusterless and sorted spikes decoding (which have 6 cm bandwidth for the KDE position dimensions and 5 cm spline knots for the GLM).
 For sorted spikes decoding, we evaluated the place field $\lambda_{i}(t_{k} \mid x_{k})$ on the midpoint of these bins.
-Likewise, for clusterless decoding, we evaluated the spike amplitudes observed during the decoding period by evaluating the KDE for $p_{i}(x_k, \vec{m}_k)$ for the midpoint of these bins.
+Likewise, for clusterless decoding, we evaluated the spike amplitudes observed during the decoding period by evaluating the KDE for $p_{i}(x_k, \vec{m}^i_{k,j})$ for the midpoint of these bins.
 
-We also used this grid bins in combination with the track graph in order to construct the appropriate 1D random walk transition matrices that respected the track geometry.
+We also used these grid bins in combination with the track graph in order to construct the appropriate 1D random walk transition matrices that respected the track geometry.
 To do this, we inserted the bin centers as nodes in the track graph, and then computed the shortest path distance between all pairs of position bin centers.
@@
 
-Finally, for our SWR analysis of $p(x_{k}, I_{k} \mid O_{1:T})$, we decoded each immobility time periods (times when the animal's speed was < 4 cm / s) in 2 ms time bins and extracted the SWR times.
+Finally, for our SWR analysis of $p(x_{k}, I_{k} \mid O_{1:T})$, we decoded each immobility time period (times when the animal's speed was < 4 cm / s) in 2 ms time bins and extracted the SWR times.
 
@@
 The mixture probability of the dynamics is a quantity that indicates how much that dynamic is contributing to the posterior at a given time.
-We estimate the mixture probability of the dynamics by integrating out position the joint posterior:
+We estimate the mixture probability of the dynamics by integrating out position from the joint posterior:
 $$Pr(I_{k} \mid O_{1:T}) = \int p(x_{k}, I_{k} \mid O_{1:T}) dx_{k}$$
-As in other calculations, the integral is approximated with a Reimann sum.
+As in other calculations, the integral is approximated with a Riemann sum.
 
@@
 \subsection*{Highest Posterior Density}
-The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% values of the posterior probability \cite{CasellaStatisticalinference2001}.
+The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \cite{CasellaStatisticalinference2001}.
 In this manuscript, we use the HPD to evaluate the uncertainty of the posterior probability of position.
@@
 $$
-which we approximate with a Reimann sum.
+which we approximate with a Riemann sum.
 
 \subsection*{Replay distance from animal}
-Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
 The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959} as implemented in NetworkX \cite{HagbergExploringNetworkStructure2008}.
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v2_supplemental1}
-\caption{The model is robust to changes in the discrete transition matrix.
+\caption{The model is robust to change of the probability of persisting in the same dynamic.
 Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that state.
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}
-\caption{Decoding the same SWR in Figure \ref{2} with 2D position using sorted spikes decoding and clusterless decoding, respectively.
+\caption{Decoding the same SWR in Figure \ref{2} with 2D position using sorted spikes and clusterless decoding, respectively.
 \textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure 2.
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v3_supplemental1}
-\caption{More examples of replays that would not be well-characterized by using the Bayesian decoder.
+\caption{More examples of replays that would not be well-characterized by using the standard decoder.
 Conventions are the same as in Figure 3.
-\textbf{(A)} A continuous replay that starts in the left arm back toward the center well, jumps to the right arm and continuous back toward the center well.
+\textbf{(A)} A continuous replay that starts in the left arm back toward the center well, jumps to the right arm and continues back toward the center well.
 \textbf{(B)} A continuous replay that travels down the right arm toward the center well, proceeds past the choice point toward the left well and then returns back to the choice point.
@@
 position shuffled data for two epochs from different animals.
-Line represents the percent of SWR classified in that epoch for real data.
+Line represents the percent of SWR events classified in that epoch for real data.
 The histogram represents the distribution after 50 shuffles of the position data.

### 2020-07-29 src/figure1.py
Update titles of subplots

@@ -23,7 +23,7 @@ def plot_spikes(test_spikes, replay_time, ax=None, cmap=plt.get_cmap("tab20")):
     ax.set_xlabel("Time [ms]")
     ax.set_xlim((replay_time.min(), replay_time.max() + 1))
     ax.set_xticks((replay_time.min(), replay_time.max() + 1))
-    ax.set_title("Spike Trains", fontsize=16)
+    ax.set_title("Synthetic Replay Spike Train", fontsize=16)
     sns.despine()
 
 
@@ -95,7 +95,7 @@ def plot_state_transition(classifier):
     axes[-1].set_yticks(dist_lim)
     axes[1].set_xlabel("Current Position [cm]")
     axes[0].set_ylabel("Previous\nPosition [cm]")
-    plt.suptitle("Dynamics Model", y=1.1, fontsize=16)
+    plt.suptitle("Movement Dynamics Model", y=1.1, fontsize=16)
 
     sns.despine()
 
@@ -181,7 +181,7 @@ def plot_probabilities(results, ax=None):
     ax.set_ylabel("Probability")
     ax.set_yticks([0, 1])
     ax.set_ylim([0, 1.05])
-    ax.set_title("Probability of\nDynamic Mode", fontsize=16)
+    ax.set_title("Probability of\nMovement Dynamic", fontsize=16)
     sns.despine()
 
 
@@ -217,7 +217,7 @@ def plot_posterior(
     cbar.ax.set_yticklabels(["0", "Max"])  # vertically oriented colorbar
     cbar.outline.set_visible(False)
 
-    ax.set_title("Estimate of\nMental Position", fontsize=16)
+    ax.set_title("Estimate of\nLatent Position", fontsize=16)
 
     xlim = results.time.min(), results.time.max() + 1
     ax.set_xticks(xlim)
@@ -270,7 +270,8 @@ def plot_classification_of_speeds(probabilities, is_classified, ax=None):
     ax.set_xscale("log")
     ax.set_xticks((1e0, 1e1, 1e2, 1e3, 1e4))
     ax.set_xlim((1e0, 1e4))
-    ax.set_title("Classification of Speeds", fontsize=16, pad=10)
+    ax.set_title("Classification of Speeds by Movement Dynamic Probability",
+                 fontsize=16, pad=10)
 
     ax.set_xlim((probabilities.Speed.min(), probabilities.Speed.max()))
     sns.despine()

### 2020-09-02 src/parameters.py
Add documentation about bandwidth parameter

@@ -43,6 +43,10 @@ In one time step (2 ms),
 movement_var = 6.0
 replay_speed = 1
 model = NumbaKDE
+'''Bandwith is the standard deviation of the gaussian kernel. The first four
+entries correspond to the bandwidth for the marks. The last two for the
+position dimensions.
+'''
 model_kwargs = {
     'bandwidth': np.array([24.0, 24.0, 24.0, 24.0, 6.0, 6.0])
 }

### 2020-09-02 src/visualization.py
Change default movie colors

@@ -316,13 +316,13 @@ def make_movie(position, posterior_density, position_info, map_position,
     ax.set_ylim(position_info.y_position.min() + 1,
                 position_info.y_position.max() + 1)
 
-    position_dot = plt.scatter([], [], s=200, zorder=102, color='black',
+    position_dot = plt.scatter([], [], s=200, zorder=102, color='magenta',
                                label='Actual position')
-    position_line, = plt.plot([], [], '-', linewidth=3, color='black')
+    position_line, = plt.plot([], [], '-', linewidth=3, color='magenta')
 
-    map_dot = plt.scatter([], [], s=200, zorder=102, color='r',
+    map_dot = plt.scatter([], [], s=200, zorder=102, color='green',
                           label='Decoded position')
-    map_line, = plt.plot([], [], 'r-', linewidth=3)
+    map_line, = plt.plot([], [], 'green-', linewidth=3)
     # spikes_dot = plt.scatter([], [], s=40, zorder=104, color='k',
     #                          label='spikes')
     vmax = np.percentile(posterior_density.values, 99)

### 2020-09-02 Article.tex

@@
 \author[2, 3]{Michael E. Coulter}
-\author[6]{Kenneth Kay}
+\author[4]{Kenneth Kay}
 \author[5]{Marielena Sosa}
-\author[7]{Jason E. Chung}
-\author[4]{Uri T. Eden}
+\author[6]{Jason E. Chung}
+\author[7]{Uri T. Eden}
 \author[1, 2, 3, \Letter]{Loren M. Frank}
@@
 \affil[1]{Howard Hughes Medical Institute, University of California, San Francisco, San Francisco, California}
-\affil[2]{Department of Physiology, University of California, San Francisco, San Francisco, California}
+\affil[2]{Departments of Physiology and Psychiatry, University of California, San Francisco, San Francisco, California}
 \affil[3]{Kavli Institute for Fundamental Neuroscience, University of California, San Francisco, San Francisco, California}
-\affil[4]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
+\affil[4]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
 \affil[5]{Department of Neurobiology, Stanford University School of Medicine, Stanford, CA}
-\affil[6]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
-\affil[7]{Department of Neurological Surgery, University of California, San Francisco, California}
+\affil[6]{Department of Neurological Surgery, University of California, San Francisco, California}
+\affil[7]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
 
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v4}
-\caption{Components of the state space model and demonstration on simulated data.
+\caption{Verifying the model can capture different sequence dynamics on simulated data and showing how it can classify different trajectory speeds.
 \textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics.
@@
 We do this by defining a state transition matrix, which defines how the latent position can move from the previous time step.
-Previous findings suggest that replay may exhibit at least three distinct types of movement: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented \cite{PfeifferAutoassociativedynamicsgeneration2015}, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another.
+Previous findings suggest that replay may exhibit at least three distinct types of movement: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
 We therefore define state transition models to capture each of these dynamics.
@@
 \textbf{(D)} A replay that begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
-\textbf{(E)} Percentage of SWRs in each epoch that has some content that falls into one of the categories of dynamics (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
-\textbf{(F)} Of the classified content, the percentage of SWRs in each epoch that have some Stationary, Stationary-Continuous-Mix, or Continuous content.
-\textbf{(G)} Of the classified content, the percentage of SWR in each epoch that has some Continuous content.
+\textbf{(E)} Percentage of SWRs in each day that has some content that falls into one of the categories of dynamics (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
+\textbf{(F)} Of the classified content, the percentage of SWRs in each day that have some Stationary, Stationary-Continuous-Mix, or Continuous content.
+\textbf{(G)} Of the classified content, the percentage of SWR in each day that has some Continuous content.
 }
@@
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had some continuously moving content (Figure \ref{3}G, 17\% of all SWRs).
-Thus, focusing on only spatially continuous trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
+Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G, 17\% of all SWRs).
+Thus, focusing on only high speed trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
 
@@
 Our model is specified in the context of a latent position associated with different movement dynamics.
-This specification allows us to not only decode events, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
+This specification allows us to not only to classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
 To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) region of the posterior linear position.
@@
 Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
+Although this is certainly influenced by the dynamics imposed on the model via the movement state transition matrix, because the model is driven by the data, it does not necessarily have to be the case that the
 
@@
 An example of this change is shown in Figure \ref{4}D, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E,).
 For the spatially coherent dynamics, the cumulative HPD was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
-median 238 cm, spatially incoherent).
+median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
 
@@
 Our model also makes it possible to identify different dynamics within the same SWR.
-As the start and end time of each SWR event are also dependent on a selection of arbitrary parameters, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
+As the start and end time of each SWR event are also parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
 Examination of such periods may shed light on when and why the hippocampus may be representing current position, past experience, or possible future locations, to understand how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
@@
 Third, we built an explicit, but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
+And we use these dynamics to describe the SWR events, rather than declaring the events as significant or not.
 Fourth, our model avoids assumptions about temporal bins and does not make assumptions about the linear progression of events.
@@
 Finally, we used an acausal estimator of latent position, which allowed us to identify the specific timing of transitions in dynamics at a fine timescale.
+Our method allows us to identify periods of slower moving events that
 

### 2020-09-02 Article.tex
Update on Overleaf.

@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v4}
+\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v5}
 \caption{Demonstration of the state space model on hippocampal data using both sorted and clusterless spikes on the same SWR event.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v4}
+\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v5}
 \caption{A-D.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v5}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
 \caption{95 \% Highest Posterior Density of the Dynamics.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v3-supplemental2}
+\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}
 \caption{A-C.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v3_supplemental1}
+\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}
 \caption{More examples of replays that would not be well-characterized by using the standard decoder.

### 2020-09-02 Article.tex
Update on Overleaf.

@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental2/Figure3_v1-supplemental2}
-\caption{Comparison of real vs.
+\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}
+\caption{Comparison of speed classification on real vs.
 position shuffled data for two epochs from different animals.
@@
 
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}
+\caption{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data for two epochs from different animals.
+Line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}
+\label{fig:Figure4-Figure supplement 1}
+\end{figure*}
+
 %TC:endignore

### 2020-09-15 src/analysis.py
Fix speed calculation
Use distance from center well as speed reference because the animal can move and that changes the speed. Also remove smoothing.

@@ -270,12 +270,6 @@ def get_replay_distance_metrics(results, ripple_position_info, ripple_spikes,
         track_graph, map_estimate, actual_positions,
         actual_track_segment_ids, position_info)
 
-    SMOOTH_SIGMA = 0.001
-    replay_distance_from_actual_position = gaussian_smooth(
-        replay_distance_from_actual_position, SMOOTH_SIGMA, sampling_frequency)
-    replay_distance_from_center_well = gaussian_smooth(
-        replay_distance_from_center_well, SMOOTH_SIGMA, sampling_frequency)
-
     try:
         replay_total_displacement = np.abs(
             replay_distance_from_actual_position[-1] -
@@ -286,7 +280,7 @@ def get_replay_distance_metrics(results, ripple_position_info, ripple_spikes,
     time = np.asarray(posterior.time)
     map_estimate = map_estimate.squeeze()
     replay_speed = np.abs(np.gradient(
-        replay_distance_from_actual_position, time))
+        replay_distance_from_center_well, time))
     replay_velocity_actual_position = np.gradient(
         replay_distance_from_actual_position, time)
     replay_velocity_center_well = np.gradient(

### 2020-09-15 src/analysis.py
Measure displacement and total distance using the distance from the center well

@@ -272,8 +272,8 @@ def get_replay_distance_metrics(results, ripple_position_info, ripple_spikes,
 
     try:
         replay_total_displacement = np.abs(
-            replay_distance_from_actual_position[-1] -
-            replay_distance_from_actual_position[0])
+            replay_distance_from_center_well[-1] -
+            replay_distance_from_center_well[0])
     except IndexError:
         replay_total_displacement = np.nan
 
@@ -307,7 +307,7 @@ def get_replay_distance_metrics(results, ripple_position_info, ripple_spikes,
             replay_distance_from_center_well),
         'replay_linear_position': np.mean(map_estimate),
         'replay_total_distance': np.sum(
-            np.abs(np.diff(replay_distance_from_actual_position))),
+            np.abs(np.diff(replay_distance_from_center_well))),
         'replay_total_displacement': replay_total_displacement,
         'state_order': get_state_order(is_classified),
         'spatial_coverage': np.mean(spatial_coverage),

### 2020-09-24 scripts/run_by_epoch.py
Handle case where there are two instances of a stationary state in a ripple

@@ -411,7 +411,7 @@ def clusterless_analysis_1D(epoch_key, plot_ripple_figures=False):
                     group=f'/{data_type}/{dim}/classifier/ripples/')
 
     logging.info('Saving replay_info...')
-    replay_info = get_replay_info(
+    replay_info, replay_linear_position_hover = get_replay_info(
         results, ripple_spikes, data['ripple_times'], data['position_info'],
         track_graph, SAMPLING_FREQUENCY, PROBABILITY_THRESHOLD, epoch_key,
         classifier)
@@ -420,6 +420,12 @@ def clusterless_analysis_1D(epoch_key, plot_ripple_figures=False):
     replay_info_filename = os.path.join(
         PROCESSED_DATA_DIR, f'{epoch_identifier}_replay_info_{prob:02d}.csv')
     replay_info.to_csv(replay_info_filename)
+
+    replay_linear_position_hover_filename = os.path.join(
+        PROCESSED_DATA_DIR,
+        f'{epoch_identifier}_replay_linear_position_hover_{prob:02d}.npy')
+    np.save(replay_linear_position_hover_filename,
+            replay_linear_position_hover)
 
     logging.info('Plotting ripple figures...')
 

### 2020-09-24 src/analysis.py

@@ -1,6 +1,9 @@
+import itertools
+
 import networkx as nx
 import numpy as np
 import pandas as pd
+import scipy
 import xarray as xr
 from loren_frank_data_processing.track_segment_classification import (
     get_track_segments_from_graph, project_points_to_segment)
@@ -90,7 +93,18 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
     replay_info['max_linear_distance'] = list(
         classifier.distance_between_nodes_[center_well_id].values())[-1]
 
-    return replay_info
+    replay_linear_position_hover = [
+        get_replay_linear_position_by_state(
+            ripple_number, is_classified, results, state="Hover"
+        )
+        for ripple_number in ripple_times.index
+    ]
+
+    replay_linear_position_hover = np.asarray(
+        list(itertools.chain.from_iterable(replay_linear_position_hover))
+    ) / min_max.loc[3].linear_position.max()
+
+    return replay_info, replay_linear_position_hover
 
 
 def get_sleep_replay_info(results, ripple_spikes, ripple_times, position_info,
@@ -639,3 +653,23 @@ def gaussian_smooth(data, sigma, sampling_frequency, axis=0):
     '''
     return gaussian_filter1d(
         data, sigma * sampling_frequency, axis=axis)
+
+
+def get_replay_linear_position_by_state(
+    ripple_number, is_classified, results, state="Hover"
+):
+    posterior = (
+        results.sel(ripple_number=ripple_number)
+        .acausal_posterior.dropna("time", how="all")
+        .assign_coords(time=lambda ds: ds.time / np.timedelta64(1, "s"))
+    )
+    is_classified = (
+        is_classified.sel(ripple_number=ripple_number)
+        .dropna("time", how="all")
+        .assign_coords(time=lambda ds: ds.time / np.timedelta64(1, "s"))
+    )
+    map_estimate = maximum_a_posteriori_estimate(posterior.sum("state"))
+
+    labels, n_labels = scipy.ndimage.label(is_classified.sel(state=state))
+
+    return [np.mean(map_estimate[labels == l]) for l in range(1, n_labels + 1)]

### 2020-09-24 src/analysis.py
Exclude local hovers

@@ -97,7 +97,9 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
         get_replay_linear_position_by_state(
             ripple_number, is_classified, results, state="Hover"
         )
-        for ripple_number in ripple_times.index
+        for ripple_number in ripple_times.index[
+            replay_info.Hover_replay_distance_from_actual_position > 30
+        ]
     ]
 
     replay_linear_position_hover = np.asarray(

### 2020-09-24 Article.tex
Update on Overleaf.

@@
 
-\title{Revealing the rich and varied dynamics of hippocampal replay}
+\title{Revealing the movement dynamics of hippocampal replay using state space modeling}
 \shorttitle{Replay Dynamics}
@@
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous experience or likely to occur during potential future experiences, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters / second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience or likely to occur during potential future experiences, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
@@
 The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
-Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using some arbitrarily smoothed version of the spike train.
+Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v4}
-\caption{Verifying the model can capture different sequence dynamics on simulated data and showing how it can classify different trajectory speeds.
+\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v6}
+\caption{Verifying the model can capture different sequence dynamics on simulated data.
 \textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics.
@@
 For the second 60 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
-For the last 60 ms, cells fire randomly, out of their spatial order, representing an incoherent spatial sequence.
+For the last 60 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
 \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
-\textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the posterior from the previous time step.
-In the stationary movement dynamic, the transition matrix is modeled as the identity matrix because the position remains the same.
-In the continuous movement dynamic, the transition matrix is modeled as a Gaussian random walk, meaning the next possible position is likely to be spatially close to the previous position.
-In the fragmented movement dynamic, the transition matrix is modeled as uniform, meaning the next position is equally likely to be any possible position.
-Note that the movement dynamics displayed only reflect the within-state movement dynamics.
-\textbf{(D)} The probability of remaining in a particular dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
+\textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the position in the previous time step.
+We show the probability of the next position bin for each movement dynamic model (shading).
+Zero here represents the previous position.
+\textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
 \textbf{(E)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
@@
 
-Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamics model---how the position can change over time.
+Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time given a movement dynamic.
 For the data model, our decoder is the same as the standard ("Bayesian") decoder \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}.
@@
 This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the standard decoder, we can use arbitrarily small time bins (in our case 2 ms vs.
+In contrast to the standard decoder, we can use small time bins (in our case 2 ms vs.
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
@@
 
-Next, we specify a movement dynamics model for how latent position---the "mental" position of the animal represented by the cells---evolves over time (Figure \ref{1}C).
+Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C).
 We do this by defining a state transition matrix, which defines how the latent position can move from the previous time step.
-Previous findings suggest that replay may exhibit at least three distinct types of movement: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
-We therefore define state transition models to capture each of these dynamics.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
+We therefore define state transition models to capture each of these movement dynamics.
 
-In the stationary dynamic, the latent position does not change between time steps.
+In the stationary movement dynamic, the latent position does not change between time steps.
 The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position.
-In the continuous dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
+In the continuous movement dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
 This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
@@
 This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to ~25 m/s are most likely).
-Finally, in the fragmented dynamic, the latent position can move to any available position instantaneously.
+Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
 We model this using a uniform state transition matrix, which makes transitions to all positions equally likely.
 
-Finally, we specify how likely each dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D).
-In order to be conservative, we assume that each dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
-There is, however, a small probability of switching to one of the other dynamics.
-Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step.
-Because the probability of changing dynamics follows a geometric distribution, this corresponds to an expected duration of 100 ms for staying in a particular dynamic.
-Importantly, the data drives the estimated dynamic, so even if the probability of a particular dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
+Finally, we specify how likely each movement dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
+In order to be conservative, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
+There is, however, a small probability of switching to one of the other movement dynamics.
+Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic.
+Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
 In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
 
-Once we have specified the data and dynamics model, we have fully specified the state space model.
+Once we have specified the data and the movement dynamics models, we have fully specified the state space model.
 We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior with two quantities: the mixture probability of each dynamic over time (by marginalizing over position) and the probability of latent position over time, irrespective of dynamic (by marginalizing over dynamic; Figure \ref{1}F, left and right plot respectively).
+With this, we can summarize the resulting posterior with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; Figure \ref{1}F, left and right plot respectively).
 
 An examination of the two summaries shown in Figure \ref{1}F reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
-The stable firing of the one active neuron indicates a stationary representation, and accordingly, the probability of the stationary dynamic is high at the beginning of the simulated replay.
-A change in the data then drives a rapid transition to the continuous dynamic, reflecting the trajectory-like spatially sequential spiking from the population of cells.
-Subsequently, as the population activity becomes spatially incoherent, the fragmented dynamic dominates for the last third of the simulated event.
-Importantly, these dynamics also inform the estimate of latent position of the animal, and the correct assignment of dynamic reduces the variance of the latent position estimate.
+The stable firing of the one active neuron indicates a stationary representation, and accordingly, the probability of the stationary movement dynamic is high at the beginning of the simulated replay.
+A change in the data then drives a rapid transition to the continuous movement dynamic, reflecting the trajectory-like spatially sequential spiking from the population of cells.
+Subsequently, as the population activity becomes spatially incoherent, the fragmented movement dynamic dominates for the last third of the simulated event.
 
-The specification of these dynamics allows the model to capture a wide range of movement speeds for the latent position.
-To demonstrate this, we applied the model to simulated trajectories of increasing speed (Figure \ref{1}G), from 1 cm/s to 10,000 m/s.
-From this, we can see that not only are there regions of speed that correspond to our three dynamics being highly probable (where we define highly probable to be over 0.8 probability), there are intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
+The specification of these movement dynamics allows the model to capture a wide range of movement speeds for the latent position.
+To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}G) from 1 cm/s to 10,000 m/s.
+From this, we can see that not only are there regions of speed that correspond to our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), there are intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
 In this manuscript, we will refer to these as mixture dynamics.
 For example, when the stationary state has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
-By using this classification scheme, we can characterize the speed or set of speeds in a replay by its dynamics.
+By using this classification scheme, we can characterize the speed or set of speeds in a replay by its movement dynamics.
+To summarize, there are three movement dynamic models, each of which is assigned a probability over time.
+We classify these probabilities as being part of one of five speed categories: Stationary, Stationary-Continuous-Mixtures, Continuous, Fragmented-Continuous Mixtures, and Fragmented.
 
@@
 We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding (see Methods).
-Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus (\cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}).
+Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
 Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
@@
 
-Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three states or common state mixtures.
+Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three dynamics or common state mixtures.
 To ensure that this depended on the sequential firing of spatially tuned cells in the hippocampus, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
@@
 \end{acknowledgements}
+
+The authors declare that no competing interests exist.
 
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1-supplemental1/Figure1_v2_supplemental1}
+\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}
 \caption{The model is robust to change of the probability of persisting in the same dynamic.

### 2020-09-29 Article.tex

@@
 The hippocampus is known to express such representations during sharp-wave ripple (SWR) events, when ensembles of hippocampal place cells can be activated in time-compressed sequences consistent with prior experiences.
-Efforts to understand these "replay" events have typically imposed stringent criteria on event significance, and thus analyze only a small subset of events.
-As a result, we lack a clear picture of both the prevalence and the range of dynamics associated with replay.
-Here, we develop a flexible state space model that uses a combination of discrete and latent states to capture the spatial content and time evolution of replay during SWR events.
+Efforts to understand these "replay" events have typically imposed stringent criteria on event significance, and thus analyze only a small subset of events (~10-45\%).
+As a result, we lack a clear picture of both the prevalence and the range of representational dynamics associated with replay.
+Here, we develop a state space model that uses a combination of movement dynamics to capture the spatial content and time evolution of replay during SWR events.
 Application of this model reveals that the large majority of SWR events contain coherent spatial content that evolves in accordance with a surprisingly wide range of mental movement speeds.
@@
 We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold risks giving the false conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
+Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are meaningful in the brain.
 Nonetheless, we also verified that our central results hold with a higher threshold of 0.95.

### 2020-09-29 src/visualization.py
Increase font size on movie and change legend position

@@ -1116,17 +1116,17 @@ def make_classifier_movie(
     writer = Writer(fps=frame_rate, metadata=dict(artist="Me"), bitrate=1800)
 
     fig, axes = plt.subplots(
-        1,
-        3,
-        figsize=(14, 7),
-        gridspec_kw={"width_ratios": [10, 10, 1]},
+        2,
+        2,
+        figsize=(12, 8),
+        gridspec_kw={"height_ratios": [5, 2]},
         constrained_layout=False,
     )
 
     # Plot 1
-    axes[0].set_facecolor("black")
+    axes[0, 0].set_facecolor("black")
     position_2d = data["position_info"].loc[:, ["x_position", "y_position"]]
-    axes[0].plot(
+    axes[0, 0].plot(
         position_2d.values[:, 0],
         position_2d.values[:, 1],
         color="lightgrey",
@@ -1134,48 +1134,59 @@ def make_classifier_movie(
         zorder=1,
     )
 
-    axes[0].set_xlim(
+    axes[0, 0].set_xlim(
         data["position_info"].x_position.min() - 1,
         data["position_info"].x_position.max() + 1,
     )
-    axes[0].set_ylim(
+    axes[0, 0].set_ylim(
         data["position_info"].y_position.min() + 1,
         data["position_info"].y_position.max() + 1,
     )
-    axes[0].set_xlabel("x-position")
-    axes[0].set_ylabel("y-position")
+    axes[0, 0].set_xlabel("X-Position [cm]", fontsize=18)
+    axes[0, 0].set_ylabel("Y-Position [cm]", fontsize=18)
+    axes[0, 0].tick_params(labelsize=16)
+    axes[0, 0].set_title('Decoded Position', fontsize=20)
 
     position = np.asarray(position)
-    position_dot = axes[0].scatter(
-        [], [], s=80, zorder=102, color="b", label="Actual")
-    (position_line,) = axes[0].plot([], [], "b-", linewidth=3)
+    position_dot = axes[0, 0].scatter(
+        [], [], s=100, zorder=102, color="magenta", label="Actual")
+    (position_line,) = axes[0, 0].plot([], [], linewidth=3, color="magenta")
 
-    map_dot = axes[0].scatter([], [], s=80, zorder=102,
-                              color="r", label="Decoded")
-    (map_line,) = axes[0].plot([], [], "r-", linewidth=3)
[…]

### 2020-09-29 src/figure2_supplemental1.py
Update plotting

@@ -39,6 +39,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
         )
         axes[0].eventplot(
             times, lineoffsets=tetrode_ind + 1, linewidth=1, color="black",
+            clip_on=False
         )
 
     axes[0].set_xlim(
@@ -52,7 +53,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
     axes[0].set_xlabel("Time [ms]")
 
     axes[0].set_yticks((1, n_tetrodes))
-    axes[0].set_ylim((0, n_tetrodes + 0.4))
+    axes[0].set_ylim((1, n_tetrodes))
     if data_type == "sorted_spikes":
         axes[0].set_ylabel("Cells")
     else:
@@ -81,7 +82,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
     )
     axes[1].set_xlabel("Time [ms]")
 
-    axes[1].set_ylim((0, 1.05))
+    axes[1].set_ylim((0, 1))
     axes[1].set_yticks((0, 1))
     axes[1].set_ylabel("Probability")
 
@@ -178,7 +179,7 @@ def plot_2D_results(spike_times, data, results_2D, classifier_2D,
         zorder=200,
     )
     plt.suptitle(data_type.replace("_", " ").title(), y=1.05)
-    sns.despine()
+    sns.despine(offset=5)
 
     animal, day, epoch = epoch_key
     fig_name = (

### 2020-09-29 src/figure4.py

@@ -6,7 +6,7 @@ import numpy as np
 import seaborn as sns
 from src.analysis import (get_is_classified, get_probability,
                           highest_posterior_density)
-from src.figure_utilities import ONE_COLUMN, PAGE_HEIGHT, save_figure
+from src.figure_utilities import TWO_COLUMN, PAGE_HEIGHT, save_figure
 from src.parameters import PROBABILITY_THRESHOLD, STATE_COLORS
 from src.visualization import plot_1D_wtrack_landmarks
 
@@ -38,7 +38,7 @@ def plot_clusterless_1D_results_hpd(
         3,
         1,
         constrained_layout=True,
-        figsize=(0.6 * ONE_COLUMN, 0.9 * PAGE_HEIGHT / 3),
+        figsize=(TWO_COLUMN / 4, 0.8 * PAGE_HEIGHT / 3),
         gridspec_kw={"height_ratios": [1, 3, 1]},
     )
 

### 2020-09-29 src/sorted_spikes_simulation.py
Update simulation options

@@ -4,16 +4,16 @@ from replay_trajectory_classification.simulate import (
     simulate_place_field_firing_rate, simulate_time)
 
 SAMPLING_FREQUENCY = 1000
-TRACK_HEIGHT = 180
+TRACK_LENGTH = 180
 RUNNING_SPEED = 15
 PLACE_FIELD_VARIANCE = 6.0 ** 2
-PLACE_FIELD_MEANS = np.arange(0, TRACK_HEIGHT + 10, 10)
+PLACE_FIELD_MEANS = np.arange(0, TRACK_LENGTH + 10, 10)
 N_RUNS = 15
-REPLAY_SPEEDUP = 120.0
+REPLAY_SPEEDUP = 60.0
 
 
 def make_simulated_run_data(sampling_frequency=SAMPLING_FREQUENCY,
-                            track_height=TRACK_HEIGHT,
+                            track_length=TRACK_LENGTH,
                             running_speed=RUNNING_SPEED, n_runs=N_RUNS,
                             place_field_variance=PLACE_FIELD_VARIANCE,
                             place_field_means=PLACE_FIELD_MEANS):
@@ -23,7 +23,7 @@ def make_simulated_run_data(sampling_frequency=SAMPLING_FREQUENCY,
     Parameters
     ----------
     sampling_frequency : float, optional
-    track_height : float, optional
+    track_length : float, optional
     running_speed : float, optional
     n_runs : int, optional
     place_field_variance : float, optional
@@ -39,11 +39,11 @@ def make_simulated_run_data(sampling_frequency=SAMPLING_FREQUENCY,
 
     '''
     n_samples = int(n_runs * sampling_frequency *
-                    2 * track_height / running_speed)
+                    2 * track_length / running_speed)
 
     time = simulate_time(n_samples, sampling_frequency)
     linear_distance = simulate_linear_distance(
-        time, track_height, running_speed)
+        time, track_length, running_speed)
 
     place_fields = np.stack(
         [simulate_place_field_firing_rate(place_field_mean, linear_distance,
@@ -59,17 +59,17 @@ def make_simulated_run_data(sampling_frequency=SAMPLING_FREQUENCY,
 
 
 def make_continuous_replay(sampling_frequency=SAMPLING_FREQUENCY,
-                           track_height=TRACK_HEIGHT,
+                           track_length=TRACK_LENGTH,
                            running_speed=RUNNING_SPEED,
[…]

### 2020-09-30 src/analysis.py
Add option to pick max state by setting threshold to 1

@@ -230,18 +230,25 @@ def get_is_classified(probability, probablity_threshold):
     is_classified : xarray.DataArray
 
     '''
-    is_classified = probability > probablity_threshold
-    is_classified.loc[dict(state='Hover-Continuous-Mix')] = (
-        is_classified.sel(state='Hover-Continuous-Mix') &
-        ~is_classified.sel(state='Hover') &
-        ~is_classified.sel(state='Continuous') &
-        (probability.sel(state='Fragmented') < (1 - probablity_threshold) / 2))
+    if probablity_threshold < 1.00:
+        is_classified = probability > probablity_threshold
+        is_classified.loc[dict(state='Hover-Continuous-Mix')] = (
+            is_classified.sel(state='Hover-Continuous-Mix') &
+            ~is_classified.sel(state='Hover') &
+            ~is_classified.sel(state='Continuous') &
+            (probability.sel(state='Fragmented') <
+             (1 - probablity_threshold) / 2))
 
-    is_classified.loc[dict(state='Fragmented-Continuous-Mix')] = (
-        is_classified.sel(state='Fragmented-Continuous-Mix') &
-        ~is_classified.sel(state='Fragmented') &
-        ~is_classified.sel(state='Continuous') &
-        (probability.sel(state='Hover') < (1 - probablity_threshold) / 2))
+        is_classified.loc[dict(state='Fragmented-Continuous-Mix')] = (
+            is_classified.sel(state='Fragmented-Continuous-Mix') &
+            ~is_classified.sel(state='Fragmented') &
+            ~is_classified.sel(state='Continuous') &
+            (probability.sel(state='Hover') < (1 - probablity_threshold) / 2))
+    else:
+        is_classified = ((probability.copy() * 0.0).fillna(False)).astype(bool)
+        A = probability.sel(state=["Hover", "Continuous", "Fragmented"]).values
+        A = A.argmax(axis=-1)[..., None] == np.arange(A.shape[-1])
+        is_classified.values = np.concatenate((A, np.zeros((*A.shape[:2], 2), dtype=bool)), axis=-1)
     is_classified = is_classified.rename('is_classified')
     is_classified = is_classified.where(~np.isnan(probability))
 

### 2020-09-30 Article.tex

@@
 The hippocampus is known to express such representations during sharp-wave ripple (SWR) events, when ensembles of hippocampal place cells can be activated in time-compressed sequences consistent with prior experiences.
-Efforts to understand these "replay" events have typically imposed stringent criteria on event significance, and thus analyze only a small subset of events (~10-45\%).
-As a result, we lack a clear picture of both the prevalence and the range of representational dynamics associated with replay.
+Efforts to understand these "replay" events have typically imposed stringent criteria on SWR event significance, and thus analyze only a small subset of events (\texttildelow10-45\%).
+As a result, we lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay.
 Here, we develop a state space model that uses a combination of movement dynamics to capture the spatial content and time evolution of replay during SWR events.
@@
 \centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v3}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
 \caption{
@@
 A slight bias of 0.1 is given to the diagonal of the state transition model to encourage staying on the same track segment.
-The most likely track segment the animal is on is computed using the Viterbi algorithm(cite?).
+The most likely track segment the animal is on is computed using the Viterbi algorithm \cite{ViterbiErrorboundsconvolutional1967}.
 After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the nearest point of the track segment.

### 2020-09-30 Article.tex
Update on Overleaf.

@@
 \includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v6}
-\caption{Verifying the model can capture different sequence dynamics on simulated data.
+\caption{The model can capture different sequence dynamics on simulated data.
 \textbf{(A)} We construct a firing sequence of 19 simulated place cells that exhibits three different movement dynamics.
 For the first 60 ms, one cell fires repeatedly, representing one stationary location.
-For the second 60 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
-For the last 60 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
+For next 190 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
+Finally, for the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
 \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
 \textbf{(C)} The prediction from the neural data is then combined with an explicit model of each movement dynamic, which determines how latent position can change based on the position in the previous time step.
-We show the probability of the next position bin for each movement dynamic model (shading).
+We show the probability of the next position bin for each movement dynamic model (color scale).
 Zero here represents the previous position.
@@
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
-The lines correspond to the average probability of that dynamic for a given constant velocity sequence.
-We use a 0.8  threshold (dotted line) to classify each sequence based on the state or states which contribute maximally to the posterior.
+Each dot corresponds to the average probability of that dynamic for a given constant speed sequence.
+We use a 0.8 threshold (dotted line) to classify each sequence based on the dynamic or dynamics which contribute maximally to the posterior (shaded colors).
 }
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v5}
-\caption{Demonstration of the state space model on hippocampal data using both sorted and clusterless spikes on the same SWR event.
+\caption{The model can decode hippocampal replay trajectories using either sorted and clusterless spikes from the same SWR event.
 \textbf{(A)} Decoding using sorted spikes.
@@
 L, R, C correspond to the position of the left, right and center reward wells respectively.
-The animal's actual position is indicated by the red triangle.
+The animal's actual position is indicated by the the magenta dashed line.
+Underneath is the maximum of the 1D decoded position (the most probable position) projected back onto the 2D track for the sorted decoding.
+Color indicates time.
+The animal's actual position is denoted by the pink dot.
+Light grey lines show the animal's 2D position over the entire recording session.
 \textbf{(B)} Decoding using clusterless spikes.
@@
 Other panels have the same convention as (A).
-\textbf{(C)} Maximum of the 1D decoded position (the most probable position) in (A) projected back onto the 2D maze for the sorted decoding.
-Color indicates time.
-The animal's actual position is denoted by the red circle.
-Light grey lines show the animal's 2D position over the entire recording session.
-L, R, and C correspond to the left, right and center reward well as in (A).
-\textbf{(D)} Maximum of the 1D decoded position (the most probable position) in (B) projected back into 2D using the clusterless decoding.
-Figure conventions are the same as in (C).
+Underneath is the maximum of the 1D decoded position (the most probable position) projected back into 2D using the clusterless decoding.
 }
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v5}
-\caption{A-D.
-Four examples of replays that would not be well-characterized by using the standard decoder.
+\caption{A-F.
+Six examples of SWRs that would not be well-characterized by using the standard decoder.
 Figure conventions are the same as in Figure \ref{2}.
-\textbf{(A)} A replay that starts moving down the center arm away from the animal's position at the center well, slows down, and returns back.
-\textbf{(B)} A replay that persistently represents the left well for the entirety of the SWR while the animal remains in the center arm.
-\textbf{(C)} A replay that begins with stationary representation of the left well, then jumps to the middle of the right arm and proceeds up the right arm to the right well.
-\textbf{(D)} A replay that begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
-\textbf{(E)} Percentage of SWRs in each day that has some content that falls into one of the categories of dynamics (Stationary, Stationary-Continuous-Mix, Continuous, Fragmented-Continuous-Mix, or Fragmented).
-\textbf{(F)} Of the classified content, the percentage of SWRs in each day that have some Stationary, Stationary-Continuous-Mix, or Continuous content.
-\textbf{(G)} Of the classified content, the percentage of SWR in each day that has some Continuous content.
+Filtered SWR ripple (150-250 Hz) trace from the tetrode with the maximum amplitude displayed above each example.
+\textbf{(A)} A SWR where the decoded position starts moving down the center arm away from the animal's position at the center well, slows down, and returns back.
+\textbf{(B)} A SWR where the decoded position persistently stays at the choice point (open circle) while the animal remains at the left well.
+\textbf{(C)} A SWR where the decoded position begins with stationary representation of the left well, then jumps to the middle of the right arm and proceeds up the right arm to the right well.
+\textbf{(D)} A SWR where the decoded position begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
+\textbf{(E)} A SWR where the decoded position begins in the left arm and persists at the choice point.
+\textbf{(F)} A SWR where the decoded position starts in the left arm toward the choice point, jumps to the right arm and proceeds back toward the choice point.
+\textbf{(G)} Each dot represents the percentage of SWRs for each day for that animal.
+A SWR is included in the numerator of the percentage if at any time it includes the classifications listed below the column.
+The denominator is listed in the x-axis label.
 }
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
-\caption{95 \% Highest Posterior Density of the Dynamics.
+\caption{Validation of the Dynamics using the 95\% Highest Posterior Density.
 \textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density.
@@
 \caption{
-Analysis of SWRs from all animals.
-\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of dynamics within each SWR.
-Each row represents a possible dynamic and each column represents a set of dynamics categories, where filled-in black dots with an edge between the dots indicates that multiple categories are present in the SWR (at temporally distinct times).
+Quantification of the SWRs by dynamic.
+\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
+Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times).
 The sets are ordered by how often they occur as indicated by the bar plot above each category.
-The total number of each classification is indicated by the rotated bar plot on the left.
-\textbf{(B)} Distribution of the duration of each dynamic category within a SWR.
-\textbf{(C)} The distance of the average latent position from the animal's position for each dynamic category within each SWR.
-\textbf{(D)} Position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
-\textbf{(E)} Distribution of average tetrode multiunit spike rates for each dynamic category within each SWR.
+The total percentage of each classification is indicated by the rotated bar plot on the left.
+\textbf{(B)} Box plot of the duration of each dynamic category within a SWR.
+The box shows the interquartile range (25-75\%) of the data and the whiskers are 1.5 times the interquartile range.
+Points outside of this are labeled as diamonds.
+\textbf{(C)} Box plot of the average distance between the latent position to the animal's position for each classification within the SWR.
+\textbf{(D)} Box plot of the average speed of the classification within the SWR (excluding classifications < 20 ms).\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
+The shaded region represents the density estimate while the black hatch marks represent the observed non-local stationary positions.
+\textbf{(F)} Box plot of the average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
 }
@@
 \includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}
-\caption{The model is robust to change of the probability of persisting in the same dynamic.
-Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that state.
-The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in Figure 1D.
-The first plot (upper left) shows the case when all dynamics are equally likely.
+\caption{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
+\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that state.
+The corresponding expected duration of staying in the state is listed as duration.
+The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in Figure \ref{1}D.
 The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e.
-the case where there is no probability of switching to another state.}
+the case where there is no probability of switching to another state.
+Shaded regions correspond to the classification as in Figure \ref{1}.
+\textbf{(B)} The probability of position over time for each diagonal value.
+Conventions the same as in (A).}
 \label{fig:Figure1-Figure supplement 1}
@@
 The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
-The latent position posterior summed over time is shown in purple underneath.
-The light grey lines represent the position of the animal over the entire epoch and the red dot represents the animal's position.
-\textbf{(B)} Same as in A, but with clusterless decoding.}
+The latent position posterior summed over time is shown in the purple shading.
+The light grey lines represent the position of the animal over the entire epoch and the magenta dot represents the animal's position.
+\textbf{(B)} Same as in (A), but with clusterless decoding.}
 \label{fig:Figure2-Figure supplement 1}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}
-\caption{A-C.
-Examples of continuous replays.
-Left panel uses the same conventions as Figure 2A and 2B.
-Right panel shows the 1D MAP estimate projected back to 2D as in Figure 2C.
+\caption{A-F.
+More examples of SWRs that have continuous trajectories.
+Left panel uses the same conventions as Figure \ref{2}A and \ref{2}B.
+Right panel shows the 1D MAP estimate projected back to 2D as in Figure \ref{2}C.
 Color indicates time.
-Light grey lines indicate the animal's position over the entire epoch.}
+Light grey lines indicate the animal's position over the entire epoch.
+Magenta dot represents the animal's position during the SWR.}
 \label{fig:Figure2-Figure supplement 2}
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}
-\caption{More examples of replays that would not be well-characterized by using the standard decoder.
-Conventions are the same as in Figure 3.
-\textbf{(A)} A continuous replay that starts in the left arm back toward the center well, jumps to the right arm and continues back toward the center well.
-\textbf{(B)} A continuous replay that travels down the right arm toward the center well, proceeds past the choice point toward the left well and then returns back to the choice point.
-\textbf{(C)} A replay that starts down the left arm toward the center arm and turns into a stationary trajectory near the choice point.
-\textbf{(D)} The first half is a fragmented replay and the second half is unclassified.
-\textbf{(E)} A stationary trajectory in the first half of the SWR on the right arm and then a stationary trajectory in the SWR at the left well.
-\textbf{(F)} A stationary trajectory that starts at the animal's position at the center well, then jumps to the left well and comes back toward the center well.
+\caption{A-I. More examples of SWRs that would not be well-characterized by using the standard decoder. Conventions are the same as in Figure \ref{3}.
 }
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}
-\caption{Comparison of speed classification on real vs.
+\caption{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
 position shuffled data for two epochs from different animals.
-Line represents the percent of SWR events classified in that epoch for real data.
+Red line represents the percent of SWR events classified in that epoch for real data.
 The histogram represents the distribution after 50 shuffles of the position data.
@@
 position shuffled data for two epochs from different animals.
-Line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
+Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
 The histogram represents the distribution after 50 shuffles of the position data.

### 2020-10-01 Article.tex
Update on Overleaf.

@@
 
-\title{Revealing the movement dynamics of hippocampal replay using state space modeling}
+\title{Modeling the movement dynamics of hippocampal replay reveals slower spatially coherent trajectories}
 \shorttitle{Replay Dynamics}
@@
 As a result, we lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay.
-Here, we develop a state space model that uses a combination of movement dynamics to capture the spatial content and time evolution of replay during SWR events.
-Application of this model reveals that the large majority of SWR events contain coherent spatial content that evolves in accordance with a surprisingly wide range of mental movement speeds.
+Here, we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWR events.
+Application of this model reveals that over 60\% of SWR events contain slower spatially coherent trajectories or representations of single locations.
 
@@
 We find that the majority of SWRs contain spatially coherent content; that is, the replay trajectories have no large jumps.
-Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real world animal movement, suggesting that high speed replay is not the most common manifestation of generative activity.
+Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with stationary or slower moving trajectories, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 \textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
-\textbf{(E)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic state.
-The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic state over time.
-The summed posterior across states also provides an estimate of latent position over time (right panel).
+\textbf{(E)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic.
+The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
+The summed posterior across dynamics also provides an estimate of latent position over time (right panel).
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
@@
 Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then construct an 180 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
+We then construct an 280 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
 For the first 60 ms of this sequence, a single place cell fires repeatedly, resulting in the extended representation of a single location.
-For the second 60 ms of the sequence, the cells fire in sequential spatial order, representing a fast moving trajectory across the virtual linear track.
-For the last 60 ms of the sequence, the cells fire in an incoherent spatial order.
+For the second 190 ms of the sequence, the cells fire in sequential spatial order, representing a fast moving trajectory across the virtual linear track.
+For the last 30 ms of the sequence, the cells fire in an incoherent spatial order.
 These three firing patterns represent three different types of sequence dynamics that could be expressed during replay events, which we call stationary, continuous, and fragmented, respectively.
@@
 We do this by defining a state transition matrix, which defines how the latent position can move from the previous time step.
-Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, rapid continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
 We therefore define state transition models to capture each of these movement dynamics.
@@
 Once we have specified the data and the movement dynamics models, we have fully specified the state space model.
-We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; Figure \ref{1}F, left and right plot respectively).
+We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamics (see Methods).
+With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; Figure \ref{1}E, left and right plot respectively).
 
-An examination of the two summaries shown in Figure \ref{1}F reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+An examination of the two summaries shown in Figure \ref{1}E reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
 The stable firing of the one active neuron indicates a stationary representation, and accordingly, the probability of the stationary movement dynamic is high at the beginning of the simulated replay.
@@
 The specification of these movement dynamics allows the model to capture a wide range of movement speeds for the latent position.
-To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}G) from 1 cm/s to 10,000 m/s.
+To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}F) from 1 cm/s to 10,000 m/s.
 From this, we can see that not only are there regions of speed that correspond to our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), there are intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
 In this manuscript, we will refer to these as mixture dynamics.
-For example, when the stationary state has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
+For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
 By using this classification scheme, we can characterize the speed or set of speeds in a replay by its movement dynamics.
 To summarize, there are three movement dynamic models, each of which is assigned a probability over time.
-We classify these probabilities as being part of one of five speed categories: Stationary, Stationary-Continuous-Mixtures, Continuous, Fragmented-Continuous Mixtures, and Fragmented.
+We classify these probabilities as being part of one of five speed categories: stationary, stationary-continuous-Mixtures, continuous, fragmented-continuous Mixtures, and fragmented.
 
@@
 
-As mentioned above, the model is robust to the choice of probability of staying in a state.
-To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same state versus switching to another state (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics over a single 2 ms time step (0.9 - 0.9999), the model still correctly identified the dynamic with high probability.
+As mentioned above, the model is robust to the choice of probability of staying in a dynamic.
+To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
+We found that for a large range of plausible probabilities of staying in one of the dynamics over a single 2 ms time step (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the dynamic with high probability.
 
@@
 \caption{A-F.
-Six examples of SWRs that would not be well-characterized by using the standard decoder.
+Six examples of SWRs with non-constant speed trajectories.
 Figure conventions are the same as in Figure \ref{2}.
@@
 We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with sequential population activity (Figure \ref{2}A, top panel).
-As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary state was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
 Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
@@
 
-As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR (Figure \ref{fig:Figure2-Figure supplement 2}A-C).
+As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR as with the standard decoder(Figure \ref{fig:Figure2-Figure supplement 2}A-F).
 However, we also observed many events with spatially coherent content that did not have this structure.
-For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}), trajectories that remained fixed in one position (Figure \ref{3}B), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-D, Figure \ref{fig:Figure3-Figure supplement 1}A, E, F), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}D).
+For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remained fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
 
-Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three dynamics or common state mixtures.
+Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
 To ensure that this depended on the sequential firing of spatially tuned cells in the hippocampus, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
@@
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had some continuous content (Figure \ref{3}G, 17\% of all SWRs).
+Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
 Thus, focusing on only high speed trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
@@
 To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) region of the posterior linear position.
-Larger values of the cumulative HPD extent indicate the model is less certain about position, because the HPD values are distributed over more of the track at a given time point.
+Larger values of the cumulative HPD extent indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
 In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
 Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
-Although this is certainly influenced by the dynamics imposed on the model via the movement state transition matrix, because the model is driven by the data, it does not necessarily have to be the case that the
+Although this is certainly influenced by the dynamics imposed on the model via the movement state transition matrix, because the model is primarily driven by the data, it does not necessarily have to be the case that the two coincide.
+To show this, as before, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+The shuffled decodes had much larger spatial coverage than the real data (XX cm vs XX cm, p=2.2e-16, one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).
 
@@
 \textbf{(D)} Box plot of the average speed of the classification within the SWR (excluding classifications < 20 ms).\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
-The shaded region represents the density estimate while the black hatch marks represent the observed non-local stationary positions.
+The shaded region represents the density estimate while the black ticks represent the observed non-local stationary positions.
 \textbf{(F)} Box plot of the average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
@@
 Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
-These events included decoded representations that were highly spatially concentrated, confirming their spatial coherence.
+These events included decoded representations that were highly spatially concentrated.
 This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
@@
 Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \cite{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort \cite{ChungFullyAutomatedApproach2017}) in combination wiht better recording technology may eliminate some of this gap between sorted spikes and clusterless decoding, by removing experimenter subjectivity involved in spike sorting and by more accurately identifying neurons.
+
 Third, we built an explicit, but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
@@
 
-\subsection*{Mixture probability of the dynamics}
-The mixture probability of the dynamics is a quantity that indicates how much that dynamic is contributing to the posterior at a given time.
-We estimate the mixture probability of the dynamics by integrating out position from the joint posterior:
+\subsection*{Probability of the dynamics}
+The probability of the dynamics is a quantity that indicates how much that dynamic is contributing to the posterior at a given time.
+We estimate the probability of the dynamics by integrating out position from the joint posterior:
 $$Pr(I_{k} \mid O_{1:T}) = \int p(x_{k}, I_{k} \mid O_{1:T}) dx_{k}$$
@@
 We used a threshold of 0.8 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
-Time periods during sharp wave ripples are labeled as Stationary, Continuous, or Fragmented when the probability of each individual state is above 0.8.
-Time periods are labeled as Stationary-Continuous-Mix or Fragmented-Continuous-Mix when the sum of Stationary and Continuous or Fragmented and Continuous are above 0.8, respectively.
+Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.8.
+Time periods are labeled as stationary-continuous-Mix or fragmented-continuous-Mix when the sum of stationary and continuous or fragmented and continuous are above 0.8, respectively.
 Time periods where none of these criterion are met are considered unclassified.
@@
 
-\subsection*{Replay distance from animal}
-Replay distance from the animal is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+\subsection*{Distance from animal}
+The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
 The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959} as implemented in NetworkX \cite{HagbergExploringNetworkStructure2008}.
+
+\subsection*{Estimation of Speed}
+In order to estimate the speed over time, we first found the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+We then computed the first derivative using the Python library Numpy's gradient function, which computes the difference in the forward and backward time direction and then averages them for all points except the boundaries.
+We then smoothed this quantity with a small Gaussian (2.5 ms standard deviation) and then averaged this for the classified time points.
 
@@
 \caption{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
-\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that state.
-The corresponding expected duration of staying in the state is listed as duration.
+\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that dynamic.
+The corresponding expected duration of staying in the dynamic is listed as duration.
 The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in Figure \ref{1}D.
 The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e.
-the case where there is no probability of switching to another state.
+the case where there is no probability of switching to another dynamic.
 Shaded regions correspond to the classification as in Figure \ref{1}.

### 2020-10-02 Article.tex

@@
 \textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
-\textbf{(E)} The state space model uses the components in A-D over all time to decode the joint probability of latent position and dynamic.
+\textbf{(E)} The model uses the components in A-D over all time to decode the joint probability of latent position and dynamic.
 The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
+The shaded colors indicate the category of the speed of the trajectory at that time, which is determined from the probability.
 The summed posterior across dynamics also provides an estimate of latent position over time (right panel).
@@
 There is, however, a small probability of switching to one of the other movement dynamics.
-Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic.
+Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (because the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution).
 Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
@@
 A change in the data then drives a rapid transition to the continuous movement dynamic, reflecting the trajectory-like spatially sequential spiking from the population of cells.
-Subsequently, as the population activity becomes spatially incoherent, the fragmented movement dynamic dominates for the last third of the simulated event.
+Subsequently, as the population activity becomes spatially incoherent, the fragmented movement dynamic dominates for the last 30 ms of the simulated event.
 
-The specification of these movement dynamics allows the model to capture a wide range of movement speeds for the latent position.
+Because the model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, this allows the model to capture a wide range of movement speeds for the latent position.
 To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}F) from 1 cm/s to 10,000 m/s.
@@
 In this manuscript, we will refer to these as mixture dynamics.
-For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G).
+For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G) and this indicates that the trajectory is moving slowly.
+Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates.
 By using this classification scheme, we can characterize the speed or set of speeds in a replay by its movement dynamics.
-To summarize, there are three movement dynamic models, each of which is assigned a probability over time.
-We classify these probabilities as being part of one of five speed categories: stationary, stationary-continuous-Mixtures, continuous, fragmented-continuous Mixtures, and fragmented.
+To summarize, there are three movement dynamics, each of which is assigned a probability over time.
+We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.
 
-We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part in the ability to assign a confidence value for each dynamic or combinations of dynamics to each moment in time.
+We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part in the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
 Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are meaningful in the brain.
@@
 
-As mentioned above, the model is robust to the choice of probability of staying in a dynamic.
-To demonstrate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics over a single 2 ms time step (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the dynamic with high probability.
+Finally, as mentioned above, we wanted to test the robustness of the model to the choice of probability of staying in a dynamic, because our choice of 0.98 or an expected duration of 100 ms is only based on the expected duration of a SWR.
+To investigate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A).
+Furthermore, the probability of latent position remains relatively consistent for these probabilities as well ((Figure \ref{fig:Figure1-Figure supplement 1}B).
 

### 2020-10-02 Article.tex

@@
 Here, we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWR events.
-Application of this model reveals that over 60\% of SWR events contain slower spatially coherent trajectories or representations of single locations.
+Application of this model reveals that 85\% of events contained spatially coherent, interpretable content.
+Furthermore, over 60\% of SWR events contain slower spatially coherent trajectories or representations of single locations, challenging the notion that SWRs contain primarily replay of high speed trajectories.
 
@@
 
-Our identification of rich and varied dynamics in SWRs depended on several factors.
+Our identification of the rich and varied dynamics in SWRs depended on several factors.
 First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.

### 2020-10-05 Article.tex
Update on Overleaf.

@@
 
-\title{Modeling the movement dynamics of hippocampal replay reveals slower spatially coherent trajectories}
+\title{Hippocampal replay of experience at real-world speeds}
 \shorttitle{Replay Dynamics}
@@
 Representations of past and possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus is known to express such representations during sharp-wave ripple (SWR) events, when ensembles of hippocampal place cells can be activated in time-compressed sequences consistent with prior experiences.
-Efforts to understand these "replay" events have typically imposed stringent criteria on SWR event significance, and thus analyze only a small subset of events (\texttildelow10-45\%).
-As a result, we lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay.
-Here, we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWR events.
-Application of this model reveals that 85\% of events contained spatially coherent, interpretable content.
-Furthermore, over 60\% of SWR events contain slower spatially coherent trajectories or representations of single locations, challenging the notion that SWRs contain primarily replay of high speed trajectories.
+The hippocampus expresses these sorts of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds.
+Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay, however.
+We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay.
+Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs.
+We find that the large majority of replay events contain spatially coherent, interpretable content.
+Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
 
@@
 The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
-These internal "generative" representations \cite{KayConstantSubsecondCycling2020} can be identified because the underlying neural activity patterns resemble those seen during actual events.
-In particular, an encoding model that relates neural firing to external covariates can be constructed during an experience, and this model can be inverted to decode an internal representation of the covariates from the neural activity at a different time.
+These "generative" \cite{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
+Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience
 
-Mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
-Consistent with this, generative representations can unfold rapidly \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
-As a result, decoding generative representations often involves making assumptions about their timescale and temporal structure, and these assumptions can limit the interpretation of the data.
-
-Hippocampal replay events are a prototypical example of a generative representation.
+Hippocampal "replay" events are well suited to that role \cite{CarrHippocampalreplayawake2011}.
 As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous experience or likely to occur during potential future experiences, and previous work has reported that these sequential firing events proceed at an average speed of ~10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences.
+Importantly, previous work has reported that these sequential firing events proceed at an average speed of ~10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-Identifying these events typically involves multiple steps and assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder.
+While the existence of these sequential events is well established, the current consensus is that only a minority (perhaps ~10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay.
+Of the remaining events, some contain spiking of neurons associate with single locations where animals are immobile (Cite Yu et.
+al.), but another report suggested that events that replay a single location are only seen in young animals (Cite Stella).
+Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that sort of replay in awake animals remains unclear.
+
+Our uncertainty stems in part from the dominant approaches used to identify the content of replay events.
+These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
 Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
-The spikes within these events decode the probability of position in overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations smoothly progress at a constant speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+A decoding algorithm is then applied to the  spikes within these events, yielding a set of probability distributions.
+Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
+Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
@@
 While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bin acts as a boxcar smoother that limits the potential movement speeds of the representation.
-The linear fit is also problematic, because it has the potential to reject real events that do not move at constant speeds.
+The linear fit is also problematic: it has the potential to reject real events that do not move at constant speeds.
 Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, there is now evidence that a subset of these events have stationary representations of a single location \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, which both approaches often exclude.
-In addition, these methods identify only a subset of events as containing content (typically somewhere between 10 and 45\% of the chosen events), and thereby provide little insight into the remaining majority of the events.
+Moreover, these approaches exclude events that  have stationary representations of a single location \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
 
 Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using an estimated represented location at each time bin and connecting each with a line.
-For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that replays can alternate between representing a single location and sequential spatial trajectories.
+For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
 On the other hand, Stella et al.
-\cite{StellaHippocampalReactivationRandom2019} found that replays are spatially continuous, following Brownian diffusion dynamics.
+\cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
 Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
@@
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
-It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for each category of dynamics.
+It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for each category of dynamic.
 Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that are many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 
-We find that the majority of SWRs contain spatially coherent content; that is, the replay trajectories have no large jumps.
-Surprisingly, while the expected high speed replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with stationary or slower moving trajectories, suggesting that high speed replay is not the most common manifestation of generative activity.
+We find that the large majority of SWRs contain spatially coherent content; that is, the replay trajectories that are spatially concentrated at each moment in time and have no large jumps.
+Surprisingly, while the expected high speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C).
-We do this by defining a state transition matrix, which defines how the latent position can move from the previous time step.
+We do this by defining a state transition matrix that defines how the latent position can move from the previous time step.
 Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
@@
 
-Because the model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, this allows the model to capture a wide range of movement speeds for the latent position.
+Importantly, this approach allows the model to to capture a wide range of movement speeds for the latent position.
+The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the movement of the underlying position representation.
 To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}F) from 1 cm/s to 10,000 m/s.
@@
 The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
-The second panel shows the probability of each dynamic over time as in Figure \ref{1}F.
+The middle panel shows the probability of each dynamic over time as in Figure \ref{1}F.
 Shaded regions correspond to the speed classifications as in Figure \ref{1}G.
-The third panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
+The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
 L, R, C correspond to the position of the left, right and center reward wells respectively.
@@
 \textbf{(D)} A SWR where the decoded position begins with stationary representation of the left well, jumps to the center arm, proceeds away from the center well, jumps to the right arm, proceeds back toward the center well, and then becomes fragmented.
-\textbf{(E)} A SWR where the decoded position begins in the left arm and persists at the choice point.
+\textbf{(E)} A SWR where the decoded position begins in the left arm and persists at the end of the center arm.
 \textbf{(F)} A SWR where the decoded position starts in the left arm toward the choice point, jumps to the right arm and proceeds back toward the choice point.
-\textbf{(G)} Each dot represents the percentage of SWRs for each day for that animal.
+\textbf{(G)} Classification of SWRs from multiple animals and datasets.
+Each dot represents the percentage of SWRs for each day for that animal.
 A SWR is included in the numerator of the percentage if at any time it includes the classifications listed below the column.
@@
 
-\subsection*{Replays with non-constant speed}
+\subsection*{Replays with changing speed}
 
-After testing our model on a single SWR event, we applied our decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}).
-One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a relatively permissive threshold (see Methods) to ensure that we included both the rare large events as well as the much more common small events.
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a permissive threshold (see Methods) to ensure that we included both the rare large events as well as the much more common small events.
 
@@
 In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is high certainty around the position estimates.
+Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is most often high certainty around the position estimates.
+
 Although this is certainly influenced by the dynamics imposed on the model via the movement state transition matrix, because the model is primarily driven by the data, it does not necessarily have to be the case that the two coincide.
@@
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
+We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Fig.
+4E), indicating less certainty for those specific events.
 
@@
 \includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
-\caption{Validation of the Dynamics using the 95\% Highest Posterior Density.
+\caption{Validation of classification using the 95\% Highest Posterior Density.
 \textbf{(A-D)}  Examples of the 95 \% Highest Posterior Density.
-Top panel: Probability of dynamic over time.
+In each column: top panel: Probability of dynamic over time.
 Shading and labels indicate dynamic categories.
-2nd panel: Posterior distribution of estimated linear position over time.
+Middle panel: Posterior distribution of estimated linear position over time.
 Red triangle indicates animal's position.
-3rd panel: Posterior distribution plot, where black designated the position bins that are included in the 95\% highest posterior density (HPD) region.
+Bottom panel: Posterior distribution plot, where black designated the position bins that are included in the 95\% highest posterior density (HPD) region.
 4th panel: Cumulative HPD size.
-\textbf{(E)} Average Cumulative 95\% HPD size for each dynamic category.}
+\textbf{(E)} Average Cumulative 95\% HPD size for each dynamic category.
+Spatially coherent events (stationary, stationary-continuous, and continuous) each had much smaller median HPDs than spatially incoherent events (fragmented-continuous, fragmented, and unclassified), p < XX, ranksum tests.}
 \label{4}
@@
 \caption{
-Quantification of the SWRs by dynamic.
+Prevalence of classifications.
 \textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
@@
 The total percentage of each classification is indicated by the rotated bar plot on the left.
-\textbf{(B)} Box plot of the duration of each dynamic category within a SWR.
+\textbf{(B)} Duration of each dynamic category within a SWR.
 The box shows the interquartile range (25-75\%) of the data and the whiskers are 1.5 times the interquartile range.
 Points outside of this are labeled as diamonds.
-\textbf{(C)} Box plot of the average distance between the latent position to the animal's position for each classification within the SWR.
-\textbf{(D)} Box plot of the average speed of the classification within the SWR (excluding classifications < 20 ms).\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
+\textbf{(C)} Average distance between the latent position and the animal's actual position for each classification within the SWR.
+\textbf{(D)} Average speed of the classification within the SWR (excluding classifications < 20 ms).\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
 The shaded region represents the density estimate while the black ticks represent the observed non-local stationary positions.
-\textbf{(F)} Box plot of the average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
+\textbf{(F)} Average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
 }
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic.
+We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic (Figure \ref{5}A).
 Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at rates more consistent with real-world movement (Figure \ref{1}G), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+These events contain representations that move at slower speeds (Figure \ref{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such sequences in 1585 of 23382 (~7\%) of classified SWRs.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such sequences in 1551 of 23382 (~7\%) of classified SWRs.
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}D), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of occurred at other locations of the track.
+These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
@@
 Finally, we examined the overall levels of neuronal spiking across dynamics.
-We found that all the dynamics had similar average multiunit spiking rates (Figure \ref{5}E).
+We found that all the dynamics had similar average multiunit spiking rates (Figure \ref{5}F).
 This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
@@
 These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
-The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and real-world speeds.
+The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and closer to real-world speeds.
 
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with real-world movement speeds of <1 m/s.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
 The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the classical "replay" events made up only about 19\% of classified events.
@@
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
-Interestingly, Farooq and Dragoi \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
+Interestingly, a previous paper \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
 Here, we show that they do not disappear, but are a very common occurrence with adult rats.
-We note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds of 0 m/s.
+We believe that the difference in these findings is a result of challenge of defining "chance" levels of occurrence of these events.
+We also note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
 
@@
 Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \cite{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
-We note that more fully automated modern spike sorting algorithms (such as MountainSort \cite{ChungFullyAutomatedApproach2017}) in combination wiht better recording technology may eliminate some of this gap between sorted spikes and clusterless decoding, by removing experimenter subjectivity involved in spike sorting and by more accurately identifying neurons.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort \cite{ChungFullyAutomatedApproach2017}) in combination with better recording technology may reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
 
-Third, we built an explicit, but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
-And we use these dynamics to describe the SWR events, rather than declaring the events as significant or not.
-Fourth, our model avoids assumptions about temporal bins and does not make assumptions about the linear progression of events.
+Third, we built an explicit but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
+We used these dynamics to describe the SWR events, rather than declaring the events as significant or not.
+Fourth, our model avoids large temporal bins and does not make assumptions about the linear progression of events.
 Instead, the model can use any size of bin and allows for movement in any direction.
 Finally, we used an acausal estimator of latent position, which allowed us to identify the specific timing of transitions in dynamics at a fine timescale.
-Our method allows us to identify periods of slower moving events that
 
@@
 In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results without the need to make assumptions about the right temporal bin size.
+
+Code and algorithms for decoding hippocampal replay have not typically been made accessible or easy to use.
+We believe this is not ideal because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
+Accordingly, we have made code for this model publicly available as a python package at the following URL \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
+It is easily installable using the pip or conda package installer and contains tutorial Jupyter notebooks in order to facilitate reuse.
 
@@
 Karlsson for use of their openly available datasets.
-We also thank ...
-for feedback on the manuscript.
+We also thank Abhilasha Joshi for feedback on the manuscript.
 This work was supported by the Simons Foundation for the Global Brain Grants 542971 (U.T.E.) and 542981 (L.M.F.) and the Howard Hughes Medical Institute (L.M.F.).
@@
 
+\begin{interests}
 The authors declare that no competing interests exist.
+\end{interests}
+
+\begin{contributions}
+
+\end{contributions}
 
@@
 Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.8.
-Time periods are labeled as stationary-continuous-Mix or fragmented-continuous-Mix when the sum of stationary and continuous or fragmented and continuous are above 0.8, respectively.
+Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.8, respectively.
 Time periods where none of these criterion are met are considered unclassified.
@@
 Light grey lines indicate the animal's position over the entire epoch.
-Magenta dot represents the animal's position during the SWR.}
+Magenta dashed line represents the animal's position during the SWR.}
 \label{fig:Figure2-Figure supplement 2}

### 2020-10-05 Article.tex

@@
 Representations of past and possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus expresses these sorts of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds.
-Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay, however.
+The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds.
+Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.
 We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay.
@@
 A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences.
-Importantly, previous work has reported that these sequential firing events proceed at an average speed of ~10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (perhaps ~10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay.
-Of the remaining events, some contain spiking of neurons associate with single locations where animals are immobile (Cite Yu et.
-al.), but another report suggested that events that replay a single location are only seen in young animals (Cite Stella).
-Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that sort of replay in awake animals remains unclear.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay.
+Of the remaining events, some contain spiking of neurons associated with single locations where animals are immobile \cite{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \cite{StellaHippocampalReactivationRandom2019}.
+Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
 
@@
 Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
-A decoding algorithm is then applied to the  spikes within these events, yielding a set of probability distributions.
+A decoding algorithm is then applied to the spikes within these events, yielding a set of position probability distributions for each time bin.
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.

### 2020-10-07 Article.tex

@@
 Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
-
 \end {abstract}
@@
 There is, however, a small probability of switching to one of the other movement dynamics.
-Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (because the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution).
-Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic.
+Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution).
+Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
 In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
@@
 To investigate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A).
-Furthermore, the probability of latent position remains relatively consistent for these probabilities as well ((Figure \ref{fig:Figure1-Figure supplement 1}B).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A), demonstrating that data is the most important element in the model.
+Furthermore, the probability of latent position remains relatively consistent for these probabilities as well (Figure \ref{fig:Figure1-Figure supplement 1}B).
 
@@
 This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
-We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D (Figure \ref{2}C) to better see the spatial trajectory on the maze.
+We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (Figure \ref{2}C).
 Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).
@@
 
-As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR as with the standard decoder(Figure \ref{fig:Figure2-Figure supplement 2}A-F).
+As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR as with the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
 However, we also observed many events with spatially coherent content that did not have this structure.
-For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remained fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), trajectories that jumped between arms and between dynamics (Figure \ref{3}C-F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I), and trajectories that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
+For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remained fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), trajectories that jumped between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I).
+We also observed SWRs that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
 
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 96\% (22478 of 23382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}F).
+We found that 96\% (22478 of 23382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure.
+We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure (Figure \ref{3}G).
 
@@
 
-Our model is specified in the context of a latent position associated with different movement dynamics.
-This specification allows us to not only to classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
-To do so we can compute the cumulative spatial extent of the 95\% highest posterior density (HPD) region of the posterior linear position.
-Larger values of the cumulative HPD extent indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
+Because our model is specified in the context of a latent position associated with different movement dynamics, it allows us to not only to classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
+To do so we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the posterior linear position.
+Larger values of the HPD size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
 In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
@@
 
-Although this is certainly influenced by the dynamics imposed on the model via the movement state transition matrix, because the model is primarily driven by the data, it does not necessarily have to be the case that the two coincide.
-To show this, as before, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-The shuffled decodes had much larger spatial coverage than the real data (XX cm vs XX cm, p=2.2e-16, one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).
-
 We found that spatially coherent events did indeed have lower cumulative HPD sizes than events with fragmented dynamics.
-Figure \ref{4}A and B show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
-The bulk of the HPD values at each time step in these SWRs is concentrated in a small portion of the track and the cumulative HPD size is low throughout the SWRs.
+Figure \ref{4}A and D show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
+The majority of the positions in the HPD at each time step in these SWRs is concentrated in a small portion of the track and the cumulative HPD size is low throughout the SWRs.
 In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD size is much higher.
 The HPD also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in Figure \ref{4}D, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E,).
-For the spatially coherent dynamics, the cumulative HPD was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
+An example of this change is shown in Figure \ref{4}B, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+For the spatially coherent dynamics, the cumulative HPD size was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
-We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Fig.
-4E), indicating less certainty for those specific events.
+Additionally, the HPD size for the unclassified SWRs was also We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
 
@@
 Middle panel: Posterior distribution of estimated linear position over time.
-Red triangle indicates animal's position.
-Bottom panel: Posterior distribution plot, where black designated the position bins that are included in the 95\% highest posterior density (HPD) region.
-4th panel: Cumulative HPD size.
+Magenta dashed line indicates animal's position.
+Bottom panel: Cumulative HPD size---the integral of the position bins with the top 95\% probable values.
 \textbf{(E)} Average Cumulative 95\% HPD size for each dynamic category.
-Spatially coherent events (stationary, stationary-continuous, and continuous) each had much smaller median HPDs than spatially incoherent events (fragmented-continuous, fragmented, and unclassified), p < XX, ranksum tests.}
+Median 24 cm for spatially coherent vs.
+median 238 cm for spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U}
 \label{4}
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic (Figure \ref{5}A).
+We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
 Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at slower speeds (Figure \ref{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+These events contain representations that move at slower speeds (Figure \ref{5}D; though note these speeds are calculated from the MAP estimate and thus noisy), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
-These events included decoded representations that were highly spatially concentrated.
+These events included decoded representations where the most probable positions were highly spatially concentrated.
 This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
@@
 The Gaussian place fields had a variance of 36 cm and were spaced at 10 cm intervals over the 180 cm track.
-The decoding data for Figure \ref{1}G was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
+The decoding data for Figure \ref{1}F was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
 Each simulated neuron "spiked" when the traversal passed through the peak location of its place field.
@@
 \caption{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
-\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure 1 with a different diagonal value---which governs the probability of remaining in that dynamic.
+\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure \ref{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
 The corresponding expected duration of staying in the dynamic is listed as duration.
@@
 \caption{Decoding the same SWR in Figure \ref{2} with 2D position using sorted spikes and clusterless decoding, respectively.
-\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure 2.
+\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure \ref{2}.
 The middle plot shows the probability of each dynamic over time from the 2D decode.
-Shaded regions indicate classification category as in Figure 1G and 2.
+Shaded regions indicate classification category as in Figure \ref{1}G and \ref{2}.
 The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.

### 2020-10-08 src/visualization.py
New line in xlabel

@@ -616,7 +616,7 @@ def plot_replay_distance_from_actual_position(replay_info, kind='strip',
         ax = plt.gca()
     _plot_category(replay_info, 'replay_distance_from_actual_position',
                    kind=kind, ax=ax, **kwargs)
-    ax.set_xlabel('Average distance from animal position [cm]')
+    ax.set_xlabel('Average distance\nfrom animal position [cm]')
 
 
 def plot_replay_distance_from_center_well(replay_info, kind='strip', ax=None,

### 2020-10-08 Article.tex

@@
 We find that the large majority of SWRs contain spatially coherent content; that is, the replay trajectories that are spatially concentrated at each moment in time and have no large jumps.
-Surprisingly, while the expected high speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
+Surprisingly, while the expected high speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 Figure \ref{4}A and D show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
-The majority of the positions in the HPD at each time step in these SWRs is concentrated in a small portion of the track and the cumulative HPD size is low throughout the SWRs.
-In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the HPD is much more spatially diffuse and the cumulative HPD size is much higher.
-The HPD also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
+The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the cumulative HPD size is small throughout the SWRs.
+In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the cumulative HPD size is much larger.
+The HPD size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
 An example of this change is shown in Figure \ref{4}B, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
 Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
-For the spatially coherent dynamics, the cumulative HPD size was much lower than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
+For the spatially coherent dynamics, the average cumulative HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
-Additionally, the HPD size for the unclassified SWRs was also We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
+Additionally, the HPD size for the unclassified SWRs was similarly large.
+We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
+To show this, as before, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+The shuffled decodes had much larger HPD size than the real data (epoch \#1: 55 cm vs.
+229 cm, p=0.02, real vs.
+shuffled; epoch \#2: 72 cm vs.
+231 cm, p=0.02, real vs.
+shuffled; one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).
 
@@
 \textbf{(C)} Average distance between the latent position and the animal's actual position for each classification within the SWR.
-\textbf{(D)} Average speed of the classification within the SWR (excluding classifications < 20 ms).\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
+\textbf{(D)} Average speed of the classification within the SWR (excluding classifications with durations < 20 ms).
+Note that these speeds are calculated using the most probable position (MAP estimate) which can be noisy when the probability of position is flat or multimodal.\textbf{(E)} Kernel density estimate of the position of stationary trajectories on the W-track at least 30 cm away from the animal's position.
 The shaded region represents the density estimate while the black ticks represent the observed non-local stationary positions.
@@
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+These representation do not move far in the span of a SWR.
 Interestingly, a previous paper \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
 Here, we show that they do not disappear, but are a very common occurrence with adult rats.
-We believe that the difference in these findings is a result of challenge of defining "chance" levels of occurrence of these events.
+We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events, using large time bins, and using a restrictive decoding model.
 We also note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.

### 2020-10-08 Article.tex
Update on Overleaf.

@@
 We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A, Figure\ref{fig:Figure5-Figure supplement 1}).
 These events contain representations that move at slower speeds (Figure \ref{5}D; though note these speeds are calculated from the MAP estimate and thus noisy), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
@@
 
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5-supplemental1/Figure5_v1-supplemental1}
+\caption{More examples of stationary-continuous-mixtures.}
+\label{fig:Figure5-Figure supplement 1}
+\end{figure*}
 %TC:endignore

### 2020-10-08 Article.tex
Update on Overleaf.
Draft3

--- 2020-10-08 Article.tex
@@
 
-Using a 0.8 threshold, we were able to classify 89\% (23382 of 26160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
+Using a 0.8 threshold, we were able to classify 89\% (23,382 of 26,160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
 To ensure that this depended on the sequential firing of spatially tuned cells in the hippocampus, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 96\% (22478 of 23382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}G).
+We found that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We found that only 3353 of 23382 or 14\% of classified SWRs had time points with spatially incoherent structure (Figure \ref{3}G).
+We found that only 3353 of 23,382 or 14\% of classified SWRs had time points with spatially incoherent structure (Figure \ref{3}G).
 
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4511 of 23382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
+Here our findings were consistent with previous reports: in our datasets, 4511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
 Thus, focusing on only high speed trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
@@
 We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We found that, while this change slightly reduced the total fraction of classified SWRs (19180 / 25845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19020 / 19180 or 99\% classified) had spatially coherent content.
-Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (spatially incoherent: 484 / 19180 or 3\%).
+We found that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) had spatially coherent content.
+Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (spatially incoherent: 484 of 19,180 or 3\% classified SWRs).
 
@@
 
-We found that spatially coherent events did indeed have lower cumulative HPD sizes than events with fragmented dynamics.
+We found that spatially coherent events did indeed have lower HPD sizes than events with fragmented dynamics.
 Figure \ref{4}A and D show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
-The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the cumulative HPD size is small throughout the SWRs.
-In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the cumulative HPD size is much larger.
+The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD size is small throughout the SWRs.
+In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD size is much larger.
 The HPD size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in Figure \ref{4}B, where the cumulative HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average cumulative HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
-For the spatially coherent dynamics, the average cumulative HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
+An example of this change is shown in Figure \ref{4}B, where the HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examined the average HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+For the spatially coherent dynamics, the average HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
@@
 We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
-To show this, as before, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+In general, if the data does not exhibit strong place specific firing, the HPD size will be large, no matter the influence of the dynamics.
+To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
 The shuffled decodes had much larger HPD size than the real data (epoch \#1: 55 cm vs.
@@
 Magenta dashed line indicates animal's position.
-Bottom panel: Cumulative HPD size---the integral of the position bins with the top 95\% probable values.
-\textbf{(E)} Average Cumulative 95\% HPD size for each dynamic category.
+Bottom panel: HPD size---the integral of the position bins with the top 95\% probable values.
+\textbf{(E)} Average 95\% HPD size for each dynamic category.
 Median 24 cm for spatially coherent vs.
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 classified SWRs or 70\%) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A).
-These events contain representations that move at slower speeds (Figure \ref{5}D; though note these speeds are calculated from the MAP estimate and thus noisy), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+We first note that most of the SWRs (16454 of 23382 or 70\% classified SWRs) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% classified SWRs, Figure \ref{5}A, Figure\ref{fig:Figure5-Figure supplement 1}).
+These events contain representations that move at slower speeds (Figure \ref{5}D; though note these speeds are calculated from the MAP estimate and thus noisy when the estimated position is more uncertain i.e.
+has a large HPD size), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
@@
 
-The second most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\%).
+The second most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\% classified SWRs).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C).
@@
 Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
-These events included decoded representations where the most probable positions were highly spatially concentrated.
+These events included decoded representations where the most probable positions were highly spatially concentrated (i.e.
+had small HPD size).
 This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
@@
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
-These representation do not move far in the span of a SWR.
 Interestingly, a previous paper \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
@@
 \begin{contributions}
-
+E.L.D.
+conceived of the model and did the analysis with help from U.T.E, A.K.G., and M.E.C.
+E.L.D and L.M.F wrote the manuscript with help from A.K.G.
+and M.E.C.
+A.K.G, K.K., M.S., J.E.C.
+contributed data to this manuscript.
 \end{contributions}
@@
 The 95\% HPD is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
-The cumulative spatial extent of the HPD is calculated by taking the integral of the members of this set:
+The HPD size is calculated by taking the integral of the members of this set:
 $$
@@
 We then smoothed this quantity with a small Gaussian (2.5 ms standard deviation) and then averaged this for the classified time points.
+Because our position bin size of 3 cm makes it hard to distinguish between slower speeds in a 2 ms time step, we only analyzed classifications longer than 20 ms.
 
@@
 
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=0.80\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}
+\caption{More examples of stationary-continuous-mixtures.}
+\label{fig:Figure5-Figure supplement 1}
+\end{figure*}
 %TC:endignore

### 2020-10-19 Article.tex

@@
 Representations of past and possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds.
+The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x real-world speeds.
 Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.
@@
 These "generative" \cite{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
-Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience
+Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience.
 
@@
 While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay.
-Of the remaining events, some contain spiking of neurons associated with single locations where animals are immobile \cite{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \cite{StellaHippocampalReactivationRandom2019}.
+One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
+Our subjective experience is that we can retrieve memories of locations without having to mentally traverse a long distance.
+If the rodent hippocampus is capable of that sort of retrieval, then one might expect to see SWRs where the neural activity corresponds not to a very rapid trajectory through space but instead to either a stable pattern associated with a single location or perhaps a pattern more similar to one seen during a real experience.
+Importantly, this sort of activity could still be useful, as it would take place during the short duration of an SWR (\texttildelow100 ms) and could thus support rapid retrieval and use of representations.
+
+Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \cite{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \cite{StellaHippocampalReactivationRandom2019}.
 Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
@@
 Our uncertainty stems in part from the dominant approaches used to identify the content of replay events.
-These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder.
+These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are most commonly characterized using the standard "Bayesian" decoder.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
@@
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
 
-While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bin acts as a boxcar smoother that limits the potential movement speeds of the representation.
-The linear fit is also problematic: it has the potential to reject real events that do not move at constant speeds.
+While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
+The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds.
 Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
@@
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
-It would be able to capture different types of movement dynamics, ranging from stationary to rapidly evolving to disorganized, and provide a statistical assessment of confidence for each category of dynamic.
+It would be able to capture a range of movement dynamics -- how the position representation evolves over time.
+An ideal model would capture key dynamics of interest including stationary or unchanging representations, trajectories that progress through space at constant or variable speeds, and disorganized, spatially incoherent events.
+It would provide a robust statistical assessment of confidence for each dynamic.
 Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
@@
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
-We apply this model to data from 10 rats, focusing on SWR events to permit a direct comparison to previous work.
+We apply this model to spiking data during SWR events from 10 rats, enabling a direct comparison to previous work.
 
-We find that the large majority of SWRs contain spatially coherent content; that is, the replay trajectories that are spatially concentrated at each moment in time and have no large jumps.
-Surprisingly, while the expected high speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
+We find that the large majority of SWRs contain spatially coherent content; that is, trajectories that are spatially concentrated at each moment in time and have no large discontinuities in position.
+Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 \textbf{(D)} The probability of remaining in a particular movement dynamic versus switching to another dynamic is modeled as having a high probability of remaining in a particular dynamic with a small probability of switching to one of the other dynamics at each time step.
-\textbf{(E)} The model uses the components in A-D over all time to decode the joint probability of latent position and dynamic.
-The joint probability can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
+\textbf{(E)} The model uses the components in A-D over all time to decode the joint posterior probability of latent position and dynamic.
+This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
 The shaded colors indicate the category of the speed of the trajectory at that time, which is determined from the probability.
-The summed posterior across dynamics also provides an estimate of latent position over time (right panel).
+Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel).
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
@@
 We begin with an application of the model to simulated data, both to validate the model and to provide intuition (Figure \ref{1}).
-We simulated 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
+We simulate 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
 Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then construct an 280 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
+We then construct a 280 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
 For the first 60 ms of this sequence, a single place cell fires repeatedly, resulting in the extended representation of a single location.
@@
 For the last 30 ms of the sequence, the cells fire in an incoherent spatial order.
-These three firing patterns represent three different types of sequence dynamics that could be expressed during replay events, which we call stationary, continuous, and fragmented, respectively.
+These three firing patterns represent three different types of movement dynamics that could be expressed during SWRs, which we call stationary, continuous, and fragmented, respectively.
 The goal of our model is to characterize SWRs in terms of a mixture of these three dynamics at every time point.
 
-Decoding the spiking sequence dynamics requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time given a movement dynamic.
+Decoding the spiking sequence requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time in each movement dynamic.
 For the data model, our decoder is the same as the standard ("Bayesian") decoder \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}.
@@
 This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
-In contrast to the standard decoder, we can use small time bins (in our case 2 ms vs.
+In contrast to the standard decoding approaches, we can use small time bins (in our case 2 ms vs.
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
@@
 In the stationary movement dynamic, the latent position does not change between time steps.
-The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position.
+The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the size of the position bin).
 In the continuous movement dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
@@
 Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
-We model this using a uniform state transition matrix, which makes transitions to all positions equally likely.
+We model this using a uniform state transition matrix, which makes transitions to each position equally likely.
 
@@
 
-Once we have specified the data and the movement dynamics models, we have fully specified the state space model.
-We use acausal decoding, meaning that we use all information from the past and future spikes, to estimate the joint posterior probability of position and dynamics (see Methods).
+Once we have specified the data and the movement dynamic models, we have fully specified the state space model.
+We use acausal decoding, meaning that we use all information from both past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
 With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; Figure \ref{1}E, left and right plot respectively).
@@
 To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}F) from 1 cm/s to 10,000 m/s.
-From this, we can see that not only are there regions of speed that correspond to our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), there are intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
-In this manuscript, we will refer to these as mixture dynamics.
-For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}G) and this indicates that the trajectory is moving slowly.
+From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
+In this manuscript, we will refer to these intermediate speeds as mixture dynamics.
+For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}F) and this indicates that the trajectory is moving slowly.
 Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates.
-By using this classification scheme, we can characterize the speed or set of speeds in a replay by its movement dynamics.
-To summarize, there are three movement dynamics, each of which is assigned a probability over time.
+In summary, we can characterize the speed or set of speeds in a SWR based on the probability of each of the three movement dynamics over time.
 We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.
 
-We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part in the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads people to conclude that only events that are statistically significant at the 0.05 level are meaningful in the brain.
-Nonetheless, we also verified that our central results hold with a higher threshold of 0.95.
+We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
+Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
+Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.
 
@@
 To investigate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A), demonstrating that data is the most important element in the model.
-Furthermore, the probability of latent position remains relatively consistent for these probabilities as well (Figure \ref{fig:Figure1-Figure supplement 1}B).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A), demonstrating that data itself is the most influential element in the model.
+Furthermore, the probability of latent position remains relatively consistent across the range of these probabilities as well (Figure \ref{fig:Figure1-Figure supplement 1}B).
 
@@
 
-We next sought to validate the model on real hippocampal data.
-We fit the place fields of 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track, and then applied the decoding algorithm to a SWR with sequential population activity (Figure \ref{2}A, top panel).
+We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
+To ensure that we could capture rapid trajectories many times the speed of the animal, as described by most studies of hippocampal replay, we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track.
 As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
@@
 Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
-We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to position without spike sorting, which we call clusterless decoding (see Methods).
-Clusterless decoding previously has been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
+We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
+Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
 Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
@@
 
-\subsection*{Replays with changing speed}
+\subsection*{Most replays contain coherent spatial content}
 
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 animals performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-One major goal of our work was to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a permissive threshold (see Methods) to ensure that we included both the rare large events as well as the much more common small events.
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+One major goal of our work is to assess the overall prevalence of spatial content across SWRs, so we detect SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
 
-As expected, and providing further support for the validity of the method, we observed many replays that were classified as continuous throughout the entirety of the SWR as with the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
-However, we also observed many events with spatially coherent content that did not have this structure.
-For example, we observed trajectories that started in one direction and reversed back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remained fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), trajectories that jumped between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I).
-We also observed SWRs that were spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
+As expected, and providing further support for the validity of the method, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
+However, we also observe many events with spatially coherent content that do not have this structure.
+For example, we observe trajectories that start in one direction and reverse back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remain fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), and trajectories that jump between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I).
+We also observe SWRs that are spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
 
-Using a 0.8 threshold, we were able to classify 89\% (23,382 of 26,160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
-To ensure that this depended on the sequential firing of spatially tuned cells in the hippocampus, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-We then decoded the same SWR events containing the same spikes.
-Only 9\% of the SWRs were classified in the shuffled datasets, a value that was significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, Figure \ref{fig:Figure3-Figure supplement 2}).
+Using a 0.8 threshold, we are able to classify 89\% (23,382 of 26,160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
+To ensure that this depends on the sequential firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+We then decode the same SWR events containing the same spikes.
+Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, Figure \ref{fig:Figure3-Figure supplement 2}).
 
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
-We found that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs included spatially coherent structure and the prevalence of spatially coherent structure was consistent across animals (Figure \ref{3}G).
+We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (Figure \ref{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We found that only 3353 of 23,382 or 14\% of classified SWRs had time points with spatially incoherent structure (Figure \ref{3}G).
+We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (Figure \ref{3}G).
 
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
-Thus, focusing on only high speed trajectories misses a large fraction of events where there is evidence for spatially coherent, but not continuous content.
+Here our findings were consistent with previous reports: in our datasets, 4,511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
+Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
 
 We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We found that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) had spatially coherent content.
-Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (spatially incoherent: 484 of 19,180 or 3\% classified SWRs).
+We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) had spatially coherent content.
+Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
 
-Because our model is specified in the context of a latent position associated with different movement dynamics, it allows us to not only to classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
-To do so we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the posterior linear position.
+Because our model is specified in the context of a latent position associated with different movement dynamics, it allows us to not only classify events in terms of their dynamics, but also to quantify the model's certainty in each position estimate at each moment in time given the model parameters.
+To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate.
 Larger values of the HPD size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
 In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent correspond to events where there is most often high certainty around the position estimates.
+Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
 
-We found that spatially coherent events did indeed have lower HPD sizes than events with fragmented dynamics.
-Figure \ref{4}A and D show two example SWRs that were classified as having stationary and continuous dynamics, respectively.
+We find that spatially coherent events indeed have lower HPD sizes than events with fragmented dynamics.
+Figure \ref{4}A and D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
 The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD size is small throughout the SWRs.
@@
 An example of this change is shown in Figure \ref{4}B, where the HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examined the average HPD size for each SWR, grouped by dynamic, we found a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+Overall, when we examine the average HPD size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
 For the spatially coherent dynamics, the average HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
@@
 Additionally, the HPD size for the unclassified SWRs was similarly large.
-We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains the case that a continuous or stationary dynamic can correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
-In general, if the data does not exhibit strong place specific firing, the HPD size will be large, no matter the influence of the dynamics.
+We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
+In general, if the data does not exhibit strong place specific firing, the HPD size will be large, regardless of dynamic classification.
 To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-The shuffled decodes had much larger HPD size than the real data (epoch \#1: 55 cm vs.
+The shuffled decodes have much larger HPD size than the real data (epoch \#1: 55 cm vs.
 229 cm, p=0.02, real vs.
@@
 
-\subsection*{Many replay trajectories have slower movement speeds}
+\subsection*{Many replay events are slow or stationary}
 \begin{figure*}%[tbhp]
@@
 
-Surprisingly, our results indicate that most events with coherent spatial content are not dominated by the continuous movement dynamic, and thus correspond to trajectories that are stationary or that move relatively slowly compared to their purely continuous counterparts.
+Surprisingly, our results indicate that most events with coherent spatial content are not dominated by the continuous movement dynamic, but instead correspond to trajectories that are stationary or that move relatively slowly compared to their purely continuous counterparts.
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 or 70\% classified SWRs) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% classified SWRs, Figure \ref{5}A, Figure\ref{fig:Figure5-Figure supplement 1}).
-These events contain representations that move at slower speeds (Figure \ref{5}D; though note these speeds are calculated from the MAP estimate and thus noisy when the estimated position is more uncertain i.e.
-has a large HPD size), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+We first note that most of the SWRs (16454 of 23382 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% of classified SWRs, Figure \ref{5}A, Figure\ref{fig:Figure5-Figure supplement 1}).
+These events contain representations that move at slower speeds (Figure \ref{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
 continuous 95 ms, Figure \ref{5}B).
-Nonetheless, both these slow-moving events and continuous events represented locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
+Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
 continuous 42 cm, Figure \ref{5}C).
-This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position and cannot be explained by erroneously detected SWRs.
+This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
 
-The second most common category is those SWRs that were exclusively stationary (5149 of 23382 or 22\% classified SWRs).
+The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C).
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such sequences in 1551 of 23382 (~7\%) of classified SWRs.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (~7\%) of classified SWRs.
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of occurred at other locations of the track.
+These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
@@
 We found that all the dynamics had similar average multiunit spiking rates (Figure \ref{5}F).
-This indicates that all dynamics, including the slower dynamics such as stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
+This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
 
@@
 We developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
-This model is robust; it is relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.
-We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been explained before.
-This model is also flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move more or less linearly at a constant, high speed, but also representations that move non-linearly, with slower speeds, or that are stationary.
+This model is robust: it is relatively insensitive to a range of plausible transition parameters between dynamics and instead strongly reflects the hippocampal place representation structure.
+We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been previously captured.
+This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move more or less linearly at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
 These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
-The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the classical "replay" events made up only about 19\% of classified events.
+The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events ---continuous, extended trajectories through space---made up only about 19\% of classified events.
+Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
+Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location (\cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009})
 
@@
 
+We hypothesize that his large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \cite{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \cite{YuHippocampalcorticalinteraction2015}.
+
 Our results may also explain the seeming conflict between Stella and colleagues \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
 We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
-Our model allows identification of both types of sequences.
-Our model also makes it possible to identify different dynamics within the same SWR.
-As the start and end time of each SWR event are also parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
+Because our model makes it possible to identify different dynamics within the same SWR, we are able to identify both types of sequences.
+
+Additionally, as the start and end time of each SWR event are also detection method- and parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
 Examination of such periods may shed light on when and why the hippocampus may be representing current position, past experience, or possible future locations, to understand how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
@@
 Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
-In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results without the need to make assumptions about the right temporal bin size.
+In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms)
 
@@
 These approaches can be applied not just to SWRs but to all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.
-The model can be extend to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
+The model can be extended to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
 It can also be extended to account for trajectory task context and forward/reverse sequences as in Deng et al.
@@
 
-\section*{Methods}
+\section*{Materials and methods}
 

### 2020-10-19 Article.tex

@@
 In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm.
-This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to ~25 m/s are most likely).
+This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).
 Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
@@
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (~7\%) of classified SWRs.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (\texttildelow7\%) of classified SWRs.
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
@@
 Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
-Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location (\cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009})
+Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 

### 2020-10-19 Article.tex
Update on Overleaf.

@@
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \cite{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009}.
 One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
@@
 \subsection*{Overview of the model}
-We begin with an application of the model to simulated data, both to validate the model and to provide intuition (Figure \ref{1}).
+We begin with an application of the state space model to simulated data, both to validate the model and to provide intuition (Figure \ref{1}).
 We simulate 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
 Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then construct a 280 ms spiking sequence and apply our decoding algorithm in Figure \ref{1}A.
+We then construct a 280 ms spiking sequence (Figure \ref{1}A) and apply our model to the sequence.
 For the first 60 ms of this sequence, a single place cell fires repeatedly, resulting in the extended representation of a single location.
@@
 Additionally, the HPD size for the unclassified SWRs was similarly large.
-We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less certainty for those specific events.
+We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events.
 In general, if the data does not exhibit strong place specific firing, the HPD size will be large, regardless of dynamic classification.

### 2020-10-20 Article.tex
Biorvix Draft 1

@@
 Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs.
-We find that the large majority of replay events contain spatially coherent, interpretable content.
+Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content.
 Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
@@
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
-It would be able to capture a range of movement dynamics -- how the position representation evolves over time.
-An ideal model would capture key dynamics of interest including stationary or unchanging representations, trajectories that progress through space at constant or variable speeds, and disorganized, spatially incoherent events.
+It would be able to capture a range of movement dynamics from stationary or unchanging representations to trajectories that progress through space at constant or variable speeds to disorganized, spatially incoherent events.
 It would provide a robust statistical assessment of confidence for each dynamic.
@@
 Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C).
-We do this by defining a state transition matrix that defines how the latent position can move from the previous time step.
+We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
 Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
@@
 In the stationary movement dynamic, the latent position does not change between time steps.
-The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the size of the position bin).
+The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the resolution of the position bin).
 In the continuous movement dynamic, the latent position is most likely to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
@@
 
-Finally, we specify how likely each movement dynamic is to persist in time versus changing to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
 In order to be conservative, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
@@
 \end{figure*}
-\subsection*{Identification of continuous replays with sorted and clusterless spikes}
+\subsection*{Identification of continuous dynamics in a SWR with sorted or clusterless spikes}
 
@@
 We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
-To ensure that we could capture rapid trajectories many times the speed of the animal, as described by most studies of hippocampal replay, we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track.
+To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
 As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
 Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
-This is also evident in the posterior probability of linear position over time.
+This is also evident in the posterior probability of latent linear position over time.
 This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
@@
 Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
-As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of position is narrower for the clusterless model.
+As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
 This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).
@@
 
-As expected, and providing further support for the validity of the method, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
+As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
 However, we also observe many events with spatially coherent content that do not have this structure.
@@
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
-We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods).
+We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and Figure \ref{1}F).
 We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (Figure \ref{3}G).
@@
 To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate.
-Larger values of the HPD size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
-In contrast, lower values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
+Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
+In contrast, smaller values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
+Thus, the HPD region provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
 
-We find that spatially coherent events indeed have lower HPD sizes than events with fragmented dynamics.
-Figure \ref{4}A and D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
-The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD size is small throughout the SWRs.
-In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD size is much larger.
-The HPD size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in Figure \ref{4}B, where the HPD size is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examine the average HPD size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
-For the spatially coherent dynamics, the average HPD size for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
+We find that spatially coherent events indeed have smaller HPD regions than events with fragmented dynamics.
+Figure \ref{4}A and Figure \ref{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
+The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD region is small throughout the SWRs.
+In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD region is much larger.
+The HPD region size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
+An example of this change is shown in Figure \ref{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+For the spatially coherent dynamics, the average HPD region for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
-Additionally, the HPD size for the unclassified SWRs was similarly large.
-We note here that while the size of the HPD will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events.
-In general, if the data does not exhibit strong place specific firing, the HPD size will be large, regardless of dynamic classification.
+Additionally, the HPD region for the unclassified SWRs was similarly large.
+We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events.
+In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification.
 To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-The shuffled decodes have much larger HPD size than the real data (epoch \#1: 55 cm vs.
+The shuffled decodes have much larger HPD regions than the real data (epoch \#1: 55 cm vs.
 229 cm, p=0.02, real vs.
@@
 Magenta dashed line indicates animal's position.
-Bottom panel: HPD size---the integral of the position bins with the top 95\% probable values.
-\textbf{(E)} Average 95\% HPD size for each dynamic category.
+Bottom panel: HPD region size---the integral of the position bins with the top 95\% probable values.
+\textbf{(E)} Average 95\% HPD region size for each dynamic category.
 Median 24 cm for spatially coherent vs.
@@
 The box shows the interquartile range (25-75\%) of the data and the whiskers are 1.5 times the interquartile range.
-Points outside of this are labeled as diamonds.
+Outlier points outside of this are labeled as diamonds.
 \textbf{(C)} Average distance between the latent position and the animal's actual position for each classification within the SWR.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
-The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events ---continuous, extended trajectories through space---made up only about 19\% of classified events.
+The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
 Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
@@
 The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \cite{CasellaStatisticalinference2001}.
-In this manuscript, we use the HPD to evaluate the uncertainty of the posterior probability of position.
-We calculate the 95\% HPD by determining the maximum threshold value $h$ that fulfills the follow equality:
+In this manuscript, we use the HPD size to evaluate the uncertainty of the posterior probability of position.
+We calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
 $$
@@
 $$
-The 95\% HPD is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
+The 95\% HPD region is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
 The HPD size is calculated by taking the integral of the members of this set:

### 2020-12-11 Article.tex
Biorvix Draft 2

@@
 \author[2, 3]{Michael E. Coulter}
-\author[4]{Kenneth Kay}
-\author[5]{Marielena Sosa}
-\author[6]{Jason E. Chung}
-\author[7]{Uri T. Eden}
+\author[4]{Marielena Sosa}
+\author[5]{Jason E. Chung}
+\author[6]{Uri T. Eden}
 \author[1, 2, 3, \Letter]{Loren M. Frank}
@@
 \affil[3]{Kavli Institute for Fundamental Neuroscience, University of California, San Francisco, San Francisco, California}
-\affil[4]{Mortimer B. Zuckerman Mind Brain Behavior Institute, Columbia University, New York, New York}
-\affil[5]{Department of Neurobiology, Stanford University School of Medicine, Stanford, CA}
-\affil[6]{Department of Neurological Surgery, University of California, San Francisco, California}
-\affil[7]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
+\affil[4]{Department of Neurobiology, Stanford University School of Medicine, Stanford, CA}
+\affil[5]{Department of Neurological Surgery, University of California, San Francisco, California}
+\affil[6]{Department of Mathematics and Statistics, Boston University, Boston, Massachusetts}
 
@@
 Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
-In order to be conservative, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
+In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
 There is, however, a small probability of switching to one of the other movement dynamics.
@@
 
-We hypothesize that his large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
+We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
 Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \cite{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
@@
 and M.E.C.
-A.K.G, K.K., M.S., J.E.C.
+A.K.G, M.S., J.E.C.
 contributed data to this manuscript.
@@
 \end{document}
+
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
+In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
+There is, however,

### 2020-12-12 src/load_data.py
Add ripple zscore

@@ -12,15 +12,16 @@ from loren_frank_data_processing import (get_all_multiunit_indicators,
                                          make_tetrode_dataframe)
 from ripple_detection import (Kay_ripple_detector,
                               get_multiunit_population_firing_rate)
-from ripple_detection.core import _get_ripplefilter_kernel
-from scipy.signal import filtfilt
-
-from .parameters import _BRAIN_AREAS, _MARKS, ANIMALS, SAMPLING_FREQUENCY
+from ripple_detection.core import _get_ripplefilter_kernel, gaussian_smooth
+from scipy.fftpack import next_fast_len
+from scipy.signal import filtfilt, hilbert
+from scipy.stats import zscore
+from src.parameters import _BRAIN_AREAS, ANIMALS, SAMPLING_FREQUENCY
 
 logger = getLogger(__name__)
 
 
-def filter_ripple_band(data, sampling_frequency=1500):
+def filter_ripple_band(data):
     '''Returns a bandpass filtered signal between 150-250 Hz
 
     Parameters
@@ -33,11 +34,32 @@ def filter_ripple_band(data, sampling_frequency=1500):
 
     '''
     filter_numerator, filter_denominator = _get_ripplefilter_kernel()
-    is_nan = np.isnan(data)
+    is_nan = np.any(np.isnan(data), axis=-1)
     filtered_data = np.full_like(data, np.nan)
     filtered_data[~is_nan] = filtfilt(
         filter_numerator, filter_denominator, data[~is_nan], axis=0)
     return filtered_data
+
+
+def get_envelope(data, axis=0):
+    '''Extracts the instantaneous amplitude (envelope) of an analytic
+    signal using the Hilbert transform'''
+    n_samples = data.shape[axis]
+    instantaneous_amplitude = np.abs(
+        hilbert(data, N=next_fast_len(n_samples), axis=axis))
+    return np.take(instantaneous_amplitude, np.arange(n_samples), axis=axis)
+
+
+def get_ripple_consensus_trace(ripple_filtered_lfps, sampling_frequency):
+    SMOOTHING_SIGMA = 0.004
+    ripple_consensus_trace = np.full_like(ripple_filtered_lfps, np.nan)
+    not_null = np.all(pd.notnull(ripple_filtered_lfps), axis=1)
+    ripple_consensus_trace[not_null] = get_envelope(
[…]

### 2020-12-12 src/load_data.py
Don't pre-filter marks

@@ -145,7 +145,6 @@ def load_data(epoch_key, brain_areas=None):
     tetrode_info = tetrode_info.loc[is_brain_areas]
     multiunit = (get_all_multiunit_indicators(
         tetrode_info.index, ANIMALS, _time_function)
-        .sel(features=_MARKS)
         .reindex({'time': time}))
     multiunit_spikes = (np.any(~np.isnan(multiunit.values), axis=1)
                         ).astype(np.float)
@@ -283,7 +282,6 @@ def load_sleep_data(epoch_key, brain_areas=None):
     multiunit = (
         get_all_multiunit_indicators(
             tetrode_info.index, ANIMALS, _time_function)
-        .sel(features=_MARKS)
         .reindex({"time": time})
     )
     multiunit_spikes = (

### 2020-12-12 src/load_data.py
Add some logging statements

@@ -116,6 +116,7 @@ def load_data(epoch_key, brain_areas=None):
     def _time_function(*args, **kwargs):
         return time
 
+    logger.info('Loading position info...')
     position_info = (
         get_interpolated_position_dataframe(
             epoch_key, ANIMALS, _time_function)
@@ -130,6 +131,7 @@ def load_data(epoch_key, brain_areas=None):
     lfps = get_LFPs(tetrode_keys, ANIMALS)
     lfps = lfps.resample('2ms').mean().fillna(method='pad').reindex(time)
 
+    logger.info('Loading spikes...')
     try:
         neuron_info = make_neuron_dataframe(ANIMALS).xs(
             epoch_key, drop_level=False)
@@ -142,6 +144,7 @@ def load_data(epoch_key, brain_areas=None):
     except KeyError:
         spikes = None
 
+    logger.info('Loading multiunit...')
     tetrode_info = tetrode_info.loc[is_brain_areas]
     multiunit = (get_all_multiunit_indicators(
         tetrode_info.index, ANIMALS, _time_function)

### 2020-12-12 Article.tex
Reformat to eLife format

@@
-\documentclass[times, twoside]{zHenriquesLab-StyleBioRxiv}
+\documentclass[9pt,lineno]{elife}
 \usepackage{blindtext}
@@
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{Denovellis}
-
 \begin{document}
@@
 \title{Hippocampal replay of experience at real-world speeds}
-\shorttitle{Replay Dynamics}
 
@@
 \author[6]{Uri T. Eden}
-\author[1, 2, 3, \Letter]{Loren M. Frank}
+\author[1, 2, 3]{Loren M. Frank}
 
@@
 
+\corr{loren@phy.ucsf.edu}{LMF}
+
 \maketitle
 
-%TC:break Abstract
-%the command above serves to have a word count for the abstract
 \begin{abstract}
@@
 \end {abstract}
-%TC:break main
-%the command above serves to have a word count for the abstract
-
-\begin{keywords}
-Hippocampus | Replay | State Space | Dynamics
-\end{keywords}
-
-\begin{corrauthor}
-%\texttt{loren{@}phy.ucsf.edu}
-loren\at phy.ucsf.edu
-\end{corrauthor}
 
@@
 The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
-These "generative" \cite{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
+These "generative" \citep{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
 Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience.
 
-Hippocampal "replay" events are well suited to that role \cite{CarrHippocampalreplayawake2011}.
+Hippocampal "replay" events are well suited to that role \citep{CarrHippocampalreplayawake2011}.
 As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations.
@@
 A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences.
-Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \cite{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009}.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009}.
 One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
@@
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \cite{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \cite{StellaHippocampalReactivationRandom2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
 Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
@@
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
+Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
 A decoding algorithm is then applied to the spikes within these events, yielding a set of position probability distributions for each time bin.
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
@@
 The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds.
-Approaches that focus on the order of cell activity within each event \cite{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
+Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, these approaches exclude events that  have stationary representations of a single location \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
+Moreover, these approaches exclude events that  have stationary representations of a single location \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
 
 Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using an estimated represented location at each time bin and connecting each with a line.
-For example, using this approach, Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
-On the other hand, Stella et al.
-\cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
+For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
+On the other hand, \cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
 Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \cite{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 \section*{Results}
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure1/Figure1_v6}
+
+% Figure 1
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v6}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 }
-\label{1}
-\end{figure*}
+\label{fig:1}
+% Figure 1-Supplemental 1
+\figsupp[The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms)]{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
+\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
+The corresponding expected duration of staying in the dynamic is listed as duration.
+The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in \FIG{1}D.
+The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e.
+the case where there is no probability of switching to another dynamic.
+Shaded regions correspond to the classification as in \FIG{1}.
+\textbf{(B)} The probability of position over time for each diagonal value.
+Conventions the same as in (A).}{\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}}\label{figsupp:fig1sup1}
+\end{figure}
+
 \subsection*{Overview of the model}
-We begin with an application of the state space model to simulated data, both to validate the model and to provide intuition (Figure \ref{1}).
+We begin with an application of the state space model to simulated data, both to validate the model and to provide intuition (\FIG{1}).
 We simulate 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
 Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
-We then construct a 280 ms spiking sequence (Figure \ref{1}A) and apply our model to the sequence.
+We then construct a 280 ms spiking sequence (\FIG{1}A) and apply our model to the sequence.
 For the first 60 ms of this sequence, a single place cell fires repeatedly, resulting in the extended representation of a single location.
@@
 Decoding the spiking sequence requires specifying two elements: the data model---how the spikes relate to position---and the movement dynamic models---how the position can change over time in each movement dynamic.
-For the data model, our decoder is the same as the standard ("Bayesian") decoder \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}.
+For the data model, our decoder is the same as the standard ("Bayesian") decoder \citep{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019}.
 We compute an estimate of how each cell's firing rate varies over position during movement (i.e.
-the place field, Figure \ref{1}B).
+the place field, \FIG{1}B).
 This is used during decoding to compute the Poisson likelihood of position over time given the spiking sequence of interest.
@@
 
-Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (Figure \ref{1}C).
+Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (\FIG{1}C).
 We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
-Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \cite{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \cite{PfeifferAutoassociativedynamicsgeneration2015}.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
 We therefore define state transition models to capture each of these movement dynamics.
@@
 This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian.
-In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm.
+In our case, since replay has been shown to move at speeds much faster than the animal's speed \citep{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm.
 This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).
@@
 
-Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via a state transition matrix between dynamics.
 In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
@@
 We use acausal decoding, meaning that we use all information from both past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; Figure \ref{1}E, left and right plot respectively).
+With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, left and right plot respectively).
 
-An examination of the two summaries shown in Figure \ref{1}E reveals that that the model successfully captures the dynamics of the population spiking activity in Figure \ref{1}A.
+An examination of the two summaries shown in \FIG{1}E reveals that that the model successfully captures the dynamics of the population spiking activity in \FIG{1}A.
 The stable firing of the one active neuron indicates a stationary representation, and accordingly, the probability of the stationary movement dynamic is high at the beginning of the simulated replay.
@@
 The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the movement of the underlying position representation.
-To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (Figure \ref{1}F) from 1 cm/s to 10,000 m/s.
+To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 m/s.
 From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
 In this manuscript, we will refer to these intermediate speeds as mixture dynamics.
-For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, Figure \ref{1}F) and this indicates that the trajectory is moving slowly.
+For example, when the stationary dynamic has a probability of 0.6 and the continuous has a probability of 0.4, we call this a stationary-continuous mixture (light blue, \FIG{1}F) and this indicates that the trajectory is moving slowly.
 Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates.
@@
 Finally, as mentioned above, we wanted to test the robustness of the model to the choice of probability of staying in a dynamic, because our choice of 0.98 or an expected duration of 100 ms is only based on the expected duration of a SWR.
-To investigate this we decoded the spiking sequence in Figure \ref{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (Figure \ref{fig:Figure1-Figure supplement 1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (Figure \ref{fig:Figure1-Figure supplement 1}A), demonstrating that data itself is the most influential element in the model.
-Furthermore, the probability of latent position remains relatively consistent across the range of these probabilities as well (Figure \ref{fig:Figure1-Figure supplement 1}B).
+To investigate this we decoded the spiking sequence in \FIG{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (\FIGSUPP[1]{fig1sup1}).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
+Furthermore, the probability of latent position remains relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
 
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2/Figure2_v5}
+% Figure 2
+\begin{figure}
+\includegraphics[width=1.0\linewidth]{figures/Figure2/Figure2_v5}
 \caption{The model can decode hippocampal replay trajectories using either sorted and clusterless spikes from the same SWR event.
@@
 The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
-The middle panel shows the probability of each dynamic over time as in Figure \ref{1}F.
-Shaded regions correspond to the speed classifications as in Figure \ref{1}G.
+The middle panel shows the probability of each dynamic over time as in \FIG{1}F.
+Shaded regions correspond to the speed classifications as in \FIG{1}G.
 The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
@@
 }
-\label{2}
-\end{figure*}
+\label{fig:2}
+
+% Figure 2-Supplemental 1
+\figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.
+\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
+The middle plot shows the probability of each dynamic over time from the 2D decode.
+Shaded regions indicate classification category as in \FIG{1}G and \FIG{2}.
+The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
+The latent position posterior summed over time is shown in the purple shading.
+The light grey lines represent the position of the animal over the entire epoch and the magenta dot represents the animal's position.
+\textbf{(B)} Same as in (A), but with clusterless decoding.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}}\label{figsupp:fig2sup1}
+
+% Figure 2-Supplemental 2
+\figsupp[More examples of SWRs that have continuous trajectories.]{A-F.
+More examples of SWRs that have continuous trajectories.
+Left panel uses the same conventions as \FIG{2}A and \FIG{2}B.
+Right panel shows the 1D MAP estimate projected back to 2D as in \FIG{2}C.
+Color indicates time.
+Light grey lines indicate the animal's position over the entire epoch.
+Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}}\label{figsupp:fig2sup2}
+
+\end{figure}
+
 \subsection*{Identification of continuous dynamics in a SWR with sorted or clusterless spikes}
 
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3/Figure3_v5}
+% Figure 3
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_v5}
 \caption{A-F.
 Six examples of SWRs with non-constant speed trajectories.
-Figure conventions are the same as in Figure \ref{2}.
+Figure conventions are the same as in \FIG{2}.
 Filtered SWR ripple (150-250 Hz) trace from the tetrode with the maximum amplitude displayed above each example.
@@
 }
-\label{3}
-\end{figure*}
+\label{fig:3}
+
+% Figure 3-Supplemental 1
+\figsupp[More examples of SWRs that would not be well-characterized by using the standard decoder]{A-I.
+More examples of SWRs that would not be well-characterized by using the standard decoder.
+Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}}\label{figsupp:fig3sup1}
+
+% Figure 3-Supplemental 2
+\figsupp[Comparison of percentage of SWRs classified on real vs.
+position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
+position shuffled data for two epochs from different animals.
+Red line represents the percent of SWR events classified in that epoch for real data.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
+\end{figure}
 
 We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
-To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (Figure \ref{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
-As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (Figure \ref{2}A, middle panel).
+To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
 Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
 This is also evident in the posterior probability of latent linear position over time.
-This shows that the replay most likely travels down the center arm and up the right arm (Figure \ref{2}A, bottom panel).
-We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (Figure \ref{2}C).
-Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (Figure \ref{fig:Figure2-Figure supplement 1}A).
+This shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
+We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
+Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (\FIGSUPP[2]{fig2sup1}A).
 
@@
 Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
-Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \cite{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
 We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
-Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \cite{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
-Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (Figure \ref{2}B, D), both with 1D linearized position and 2D position (Figure \ref{fig:Figure2-Figure supplement 1}B).
+Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
+Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
 As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
-This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (Figure \ref{2}D vs C).
+This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (\FIG{2}D vs C).
 
@@
 
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
 One major goal of our work is to assess the overall prevalence of spatial content across SWRs, so we detect SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
 
-As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (Figure \ref{fig:Figure2-Figure supplement 2}A-F).
+As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (\FIGSUPP[2]{fig2sup2}A-F).
 However, we also observe many events with spatially coherent content that do not have this structure.
-For example, we observe trajectories that start in one direction and reverse back to the original position (Figure \ref{3}A, Figure \ref{fig:Figure3-Figure supplement 1}B), trajectories that remain fixed in one position (Figure \ref{3}B, Figure \ref{fig:Figure3-Figure supplement 1}G), and trajectories that jump between arms and between dynamics (Figure \ref{3}C, D, F, Figure \ref{fig:Figure3-Figure supplement 1}F, H, I).
-We also observe SWRs that are spatially incoherent throughout the SWR (Figure \ref{fig:Figure3-Figure supplement 1}A, D).
+For example, we observe trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
+We also observe SWRs that are spatially incoherent throughout the SWR (\FIGSUPP[3]{fig3sup1}A, D).
 
@@
 We then decode the same SWR events containing the same spikes.
-Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, Figure \ref{fig:Figure3-Figure supplement 2}).
+Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
 
-Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
-We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and Figure \ref{1}F).
-We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (Figure \ref{3}G).
+Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
+We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (Figure \ref{3}G).
+We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
 
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4,511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (Figure \ref{3}G, 17\% of all SWRs).
+Here our findings were consistent with previous reports: in our datasets, 4,511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
 Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
@@
 We find that spatially coherent events indeed have smaller HPD regions than events with fragmented dynamics.
-Figure \ref{4}A and Figure \ref{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
+\FIG{4}A and \FIG{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
 The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD region is small throughout the SWRs.
-In contrast, Figure \ref{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD region is much larger.
+In contrast, \FIG{4}C shows a SWR where the dynamics are fragmented and correspondingly, the most probable positions are more spatially diffuse and the HPD region is much larger.
 The HPD region size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in Figure \ref{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
-Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (Figure \ref{4}E).
+An example of this change is shown in \FIG{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (\FIG{4}E).
 For the spatially coherent dynamics, the average HPD region for each SWR was much smaller than the spatially incoherent dynamics (median 24 cm, spatially coherent vs.
@@
 Additionally, the HPD region for the unclassified SWRs was similarly large.
-We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in Figure \ref{4}E), indicating less spatial certainty for those specific events.
+We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in \FIG{4}E), indicating less spatial certainty for those specific events.
 In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification.
@@
 231 cm, p=0.02, real vs.
-shuffled; one-sided Mann-Whitney-U, \ref{fig:Figure4-Figure supplement 1}).
+shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}.
 
-\begin{figure*}%[tb2]
-\centering
+% Figure 4
+\begin{figure}
 \includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
@@
 median 238 cm for spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U}
-\label{4}
-\end{figure*}
+\label{fig:4}
+
+% Figure 4-Supplemental 1
+\figsupp[Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data for two epochs from different animals.
+Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
+\end{figure}
 
 \subsection*{Many replay events are slow or stationary}
-\begin{figure*}%[tbhp]
-\centering
+
+% Figure 5
+\begin{figure}
 \includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
@@
 Prevalence of classifications.
-\textbf{(A)} UpSet plot \cite{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
+\textbf{(A)} UpSet plot \citep{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
 Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times).
@@
 }
-\label{5}
-\end{figure*}
+\label{fig:5}
+
+% Figure 5-Supplemental 1
+\figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
+
+\end{figure}
 
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (Figure \ref{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% of classified SWRs, Figure \ref{5}A, Figure\ref{fig:Figure5-Figure supplement 1}).
-These events contain representations that move at slower speeds (Figure \ref{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
-continuous 95 ms, Figure \ref{5}B).
+We first note that most of the SWRs (16454 of 23382 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (\FIG{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
+These events contain representations that move at slower speeds (\FIG{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+continuous 95 ms, \FIG{5}B).
 Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
-continuous 42 cm, Figure \ref{5}C).
+continuous 42 cm, \FIG{5}C).
 This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
@@
 The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs).
-Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C).
+Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C).
 There were, however, a number of stationary events that represented positions far from the animal.
@@
 About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
+These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
@@
 Finally, we examined the overall levels of neuronal spiking across dynamics.
-We found that all the dynamics had similar average multiunit spiking rates (Figure \ref{5}F).
+We found that all the dynamics had similar average multiunit spiking rates (\FIG{5}F).
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
@@
 
-Previous work reported that less than half of SWRs or high-population-activity events contained replay \cite{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work reported that less than half of SWRs or high-population-activity events contained replay \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
-The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
+The next most common category were stationary events \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
 Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
-Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
+These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
-Interestingly, a previous paper \cite{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
+Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
 Here, we show that they do not disappear, but are a very common occurrence with adult rats.
@@
 We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
-Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \cite{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
-Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \cite{YuHippocampalcorticalinteraction2015}.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \citep{YuHippocampalcorticalinteraction2015}.
 
-Our results may also explain the seeming conflict between Stella and colleagues \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and Pfeiffer and Foster \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
+Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
 We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
@@
 First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.
-This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \cite{JaiDistincthippocampalcorticalmemory2017}.
+This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \citep{JaiDistincthippocampalcorticalmemory2017}.
 More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
-Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \cite{ChengNewExperiencesEnhance2008}.
+Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \citep{ChengNewExperiencesEnhance2008}.
 High thresholds can therefore exclude large numbers of real events and potentially lead to erroneous conclusions.
 
-Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \cite{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
-We note that more fully automated modern spike sorting algorithms (such as MountainSort \cite{ChungFullyAutomatedApproach2017}) in combination with better recording technology may reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology may reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
 
@@
 
-Our model can be viewed as an extension to the multiple temporal models approach of Johnson and colleagues 2008 \cite{JohnsonMeasuringdistributedproperties2008}, which used model selection to determine the model with the most appropriate trajectory speed.
+Our model can be viewed as an extension to the multiple temporal models approach of \cite{JohnsonMeasuringdistributedproperties2008}, which used model selection to determine the model with the most appropriate trajectory speed.
 Our model goes beyond the Johnson model in that it explicitly permits a mixture between the movement speeds, can work for arbitrary track geometries, and uses clusterless decoding.
-Our model also represents a middle ground between Hidden Markov-style models \cite{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
+Our model also represents a middle ground between Hidden Markov-style models \citep{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
 In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms)
@@
 The model can be extended to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
-It can also be extended to account for trajectory task context and forward/reverse sequences as in Deng et al.
-\cite{DengRapidclassificationhippocampal2016}.
-Finally, models can be built to estimate any covariate, including movement direction \cite{KayConstantSubsecondCycling2020}.
+It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+Finally, models can be built to estimate any covariate, including movement direction \citep{KayConstantSubsecondCycling2020}.
 We therefore suggest that this approach has the potential to provide a much richer understanding of neural representations throughout the brain.
 
-\begin{acknowledgements}
-We thank Margaret Carr Larkin and Mattias P.
-Karlsson for use of their openly available datasets.
-We also thank Abhilasha Joshi for feedback on the manuscript.
-This work was supported by the Simons Foundation for the Global Brain Grants 542971 (U.T.E.) and 542981 (L.M.F.) and the Howard Hughes Medical Institute (L.M.F.).
-\end{acknowledgements}
-
-\begin{interests}
-The authors declare that no competing interests exist.
-\end{interests}
-
-\begin{contributions}
-E.L.D.
-conceived of the model and did the analysis with help from U.T.E, A.K.G., and M.E.C.
-E.L.D and L.M.F wrote the manuscript with help from A.K.G.
-and M.E.C.
-A.K.G, M.S., J.E.C.
-contributed data to this manuscript.
-\end{contributions}
-
-\section*{References}
-\bibliography{refs}
-
-\onecolumn
-\newpage
-
-\section*{Materials and methods}
+\section*{Materials and Methods}
 
 \subsection*{Simulated Data}
-Encoding data for Figure \ref{1} and Figure \ref{fig:Figure1-Figure supplement 1} were generated by simulating 15 traversals of a 180 cm linear track.
+Encoding data for \FIG{1} and \FIGSUPP[1]{fig1sup1} were generated by simulating 15 traversals of a 180 cm linear track.
 The spiking of 19 neurons was simulated using an inhomogeneous Poisson process where the instantaneous firing rate of each neuron changes according to a Gaussian place field.
 The Gaussian place fields had a variance of 36 cm and were spaced at 10 cm intervals over the 180 cm track.
-The decoding data for Figure \ref{1}F was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
+The decoding data for \FIG{1}F was generated by simulating 20,000 linear traversals of the 180 cm track, each with a unique constant speed, starting at 0.5 cm/s and increasing by 0.5 cm/s up to 10,000 cm/s.
 Each simulated neuron "spiked" when the traversal passed through the peak location of its place field.
@@
 Ten male Long Evans rats (500-700 g, 4-9 months old) were trained on a W-track spatial alternation task.
-9 rats contributed to previous studies \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, Kayhippocampalnetworkspatial2016, CarrTransientSlowGamma2012}.
+9 rats contributed to previous studies \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, Kayhippocampalnetworkspatial2016, CarrTransientSlowGamma2012}.
 Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
@@
 A slight bias of 0.1 is given to the diagonal of the state transition model to encourage staying on the same track segment.
-The most likely track segment the animal is on is computed using the Viterbi algorithm \cite{ViterbiErrorboundsconvolutional1967}.
+The most likely track segment the animal is on is computed using the Viterbi algorithm \citep{ViterbiErrorboundsconvolutional1967}.
 After finding the track segment that corresponds to each 2D position, the 2D position is projected onto the nearest point of the track segment.
@@
 
-We have previously shown \cite{DenovellisCharacterizinghippocampalreplay2019} that the posterior probability $p(x_k, I_k \mid O_{1:T})$ can be estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0})$ and iterating to time $T$:
+We have previously shown \citep{DenovellisCharacterizinghippocampalreplay2019} that the posterior probability $p(x_k, I_k \mid O_{1:T})$ can be estimated by applying the following recursive causal filter equation, starting with initial conditions $p(x_{0}, I_{0})$ and iterating to time $T$:
 
@@
 We assume the likelihood is the same for each dynamic $I_k$, so we only need to evaluate $p(O_{k} \mid x_{k})$.
-It has been shown \cite{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the Poisson likelihood with sorted spikes can be computed as:
+It has been shown \citep{ZhangInterpretingNeuronalPopulation1998, BrownStatisticalParadigmNeural1998} that the Poisson likelihood with sorted spikes can be computed as:
 $$
@@
 
-Likewise, it has been shown \cite{ChenTransductiveneuraldecoding2012, DengRapidclassificationhippocampal2016, KloostermanBayesiandecodingusing2014} that the clusterless likelihood can be computed as: 
+Likewise, it has been shown \citep{ChenTransductiveneuraldecoding2012, DengRapidclassificationhippocampal2016, KloostermanBayesiandecodingusing2014} that the clusterless likelihood can be computed as:
 $$
@@
 For $p_{i}(x_k, \vec{m}^i_{k,j})$, $y_{i, d}$ is all positions at the time of multiunit spikes and their associated waveform features during movement.
-We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters were found to minimize decoding error in Kloosterman et al.
-\cite{KloostermanBayesiandecodingusing2014}.
+We choose the bandwidth $h_d$ to be 6.0 cm for any position dimension and 24.0 $\mu V$ for any spike amplitude dimension, because these parameters were found to minimize decoding error in \cite{KloostermanBayesiandecodingusing2014}.
 The mean firing rate $\mu_k$ was also estimated during movement.
@@
 For 2D positions, we discretized the position space by binning 2D positions occupied by the animal with equal sized bins of the chosen position bin size, followed by morphological opening to get rid of any holes smaller than the bin size.
-This was done using Scipy image processing tools \cite{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
+This was done using Scipy image processing tools \citep{SciPy1.0ContributorsSciPyfundamentalalgorithms2020}.
 This grid based approximation allows us to use Riemann sums to approximate the integrals in the causal filter and acausal smoother equations.
@@
 \subsection*{Highest Posterior Density}
-The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \cite{CasellaStatisticalinference2001}.
+The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \citep{CasellaStatisticalinference2001}.
 In this manuscript, we use the HPD size to evaluate the uncertainty of the posterior probability of position.
@@
 The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
-The shortest path is found using Dijkstra's algorithm \cite{Dijkstranotetwoproblems1959} as implemented in NetworkX \cite{HagbergExploringNetworkStructure2008}.
+The shortest path is found using Dijkstra's algorithm \citep{Dijkstranotetwoproblems1959} as implemented in NetworkX \citep{HagbergExploringNetworkStructure2008}.
 
@@
 Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
-Both code bases rely on the following python packages: Numpy \cite{vanderWaltNumPyArrayStructure2011}, Numba \cite{LamNumbaLLVMbasedPython2015}, Matplotlib \cite{HunterMatplotlib2DGraphics2007}, xarray \cite{HoyerxarrayNDlabeled2017}, NetworkX \cite{HagbergExploringNetworkStructure2008}, Pandas \cite{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
+Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
 All code is open-source and licensed under the MIT Software License.
@@
 
-\newpage
+\section{Acknowledgments}
+We thank Margaret Carr Larkin and Mattias P.
+Karlsson for use of their openly available datasets.
+We also thank Abhilasha Joshi for feedback on the manuscript.
+This work was supported by the Simons Foundation for the Global Brain Grants 542971 (U.T.E.) and 542981 (L.M.F.) and the Howard Hughes Medical Institute (L.M.F.).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Supplementary Information %
-%%%%%%%%%%%%%%%%%%%%x%%%%%%%%%
-\beginsupplement
-\captionsetup*{format=largeformat}
+\section{Interests}
+The authors declare that no competing interests exist.
 
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}
-\caption{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
-\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from Figure \ref{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
-The corresponding expected duration of staying in the dynamic is listed as duration.
-The off-diagonal values---the probability of switching to one of the other dynamics---are set to be equally likely with the remainder of the probability, as in Figure \ref{1}D.
-The diagonal increases from left to right, top to bottom, until the case where the diagonal is one and the off-diagonal is zero---i.e.
-the case where there is no probability of switching to another dynamic.
-Shaded regions correspond to the classification as in Figure \ref{1}.
-\textbf{(B)} The probability of position over time for each diagonal value.
-Conventions the same as in (A).}
-\label{fig:Figure1-Figure supplement 1}
-\end{figure*}
+\section{Contributions}
+E.L.D.
+conceived of the model and did the analysis with help from U.T.E, A.K.G., and M.E.C.
+E.L.D and L.M.F wrote the manuscript with help from A.K.G.
+and M.E.C.
+A.K.G, M.S., J.E.C.
+contributed data to this manuscript.
 
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}
-\caption{Decoding the same SWR in Figure \ref{2} with 2D position using sorted spikes and clusterless decoding, respectively.
-\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in Figure \ref{2}.
-The middle plot shows the probability of each dynamic over time from the 2D decode.
-Shaded regions indicate classification category as in Figure \ref{1}G and \ref{2}.
-The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
-The latent position posterior summed over time is shown in the purple shading.
-The light grey lines represent the position of the animal over the entire epoch and the magenta dot represents the animal's position.
-\textbf{(B)} Same as in (A), but with clusterless decoding.}
-\label{fig:Figure2-Figure supplement 1}
-\end{figure*}
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}
-\caption{A-F.
-More examples of SWRs that have continuous trajectories.
-Left panel uses the same conventions as Figure \ref{2}A and \ref{2}B.
-Right panel shows the 1D MAP estimate projected back to 2D as in Figure \ref{2}C.
-Color indicates time.
-Light grey lines indicate the animal's position over the entire epoch.
-Magenta dashed line represents the animal's position during the SWR.}
-\label{fig:Figure2-Figure supplement 2}
-\end{figure*}
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}
-\caption{A-I. More examples of SWRs that would not be well-characterized by using the standard decoder. Conventions are the same as in Figure \ref{3}.
-}
-\label{fig:Figure3-Figure supplement 1}
-\end{figure*}
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}
-\caption{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
-position shuffled data for two epochs from different animals.
-Red line represents the percent of SWR events classified in that epoch for real data.
-The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}
-\label{fig:Figure3-Figure supplement 2}
-\end{figure}
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}
-\caption{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data for two epochs from different animals.
-Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
-The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}
-\label{fig:Figure4-Figure supplement 1}
-\end{figure*}
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=0.80\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}
-\caption{More examples of stationary-continuous-mixtures.}
-\label{fig:Figure5-Figure supplement 1}
-\end{figure*}
-%TC:endignore
-%the command above ignores this section for word count
-
+\bibliography{refs}
 \end{document}
-
-Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (Figure \ref{1}D) via a state transition matrix between dynamics.
-In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
-There is, however,

### 2020-12-16 Article.tex

@@
 
+% Figure 2-video1
+\videosupp{Example of a SWR with continuous content.
+Magenta dot represents the animal's position.
+Green dot represents the most likely decoded position projected from 1D back to the linearized 2D position.
+Green line represents the decoded positions in the last 10 ms.
+Grey lines represent the position of the animal over the course of the epoch.}\label{videosupp:fig2supv1}
+
 \end{figure}
@@
 Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
+
+% Figure 3-video1
+\videosupp{Example of a SWR with that is not purely continuous. Conventions the same as Figure 2-video1}\label{videosupp:fig3supv1}
+
 \end{figure}
@@
 231 cm, p=0.02, real vs.
-shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}.
+shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}).
 
@@
 \figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
+
+% Figure 5-video1
+\videosupp{Example of a stationary-continuous-mixture. Conventions the same as Figure 2-video1}\label{videosupp:fig5supv1}
 

### 2020-12-17 scripts/run_by_epoch.py
Add new statistics
+ ripple power
+ unclassified statistics
+ spikes per bin
+ proportion of spikes under 6 ms ISI

@@ -420,7 +420,7 @@ def clusterless_analysis_1D(epoch_key, plot_ripple_figures=False):
     replay_info, replay_linear_position_hover = get_replay_info(
         results, ripple_spikes, data['ripple_times'], data['position_info'],
         track_graph, SAMPLING_FREQUENCY, PROBABILITY_THRESHOLD, epoch_key,
-        classifier)
+        classifier, data["ripple_consensus_trace_zscore"])
     epoch_identifier = f'{animal}_{day:02d}_{epoch:02d}_{data_type}_{dim}'
     prob = int(PROBABILITY_THRESHOLD * 100)
     replay_info_filename = os.path.join(

### 2020-12-17 src/analysis.py

@@ -12,7 +12,7 @@ from scipy.ndimage.filters import gaussian_filter1d
 
 def get_replay_info(results, ripple_spikes, ripple_times, position_info,
                     track_graph, sampling_frequency, probability_threshold,
-                    epoch_key, classifier):
+                    epoch_key, classifier, ripple_consensus_trace_zscore):
     '''
 
     Parameters
@@ -34,6 +34,17 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
     '''
     probability = get_probability(results)
     is_classified = get_is_classified(probability, probability_threshold)
+
+    new_index = pd.Index(np.unique(np.concatenate(
+        (ripple_consensus_trace_zscore.index, position_info.index))),
+        name='time')
+    ripple_consensus_trace_zscore = (ripple_consensus_trace_zscore
+                                     .reindex(index=new_index)
+                                     .interpolate(method='linear')
+                                     .reindex(index=position_info.index)
+                                     )
+    ripple_consensus_trace_zscore = reshape_to_segments(
+        ripple_consensus_trace_zscore, ripple_times)
 
     duration = (is_classified.sum('time') / sampling_frequency)
     duration = duration.to_dataframe().unstack(level=1)
@@ -66,7 +77,7 @@ def get_replay_info(results, ripple_spikes, ripple_times, position_info,
         [get_replay_distance_metrics(
             results, ripple_position_info, ripple_spikes, ripple_number,
             position_info, track_graph, sampling_frequency, is_classified,
-            probability)
+            probability, ripple_consensus_trace_zscore)
          for ripple_number in ripple_times.index], index=ripple_times.index)
     replay_info = pd.concat((ripple_times, duration, metrics), axis=1)
     animal, day, epoch = epoch_key
@@ -248,7 +259,8 @@ def get_is_classified(probability, probablity_threshold):
         is_classified = ((probability.copy() * 0.0).fillna(False)).astype(bool)
         A = probability.sel(state=["Hover", "Continuous", "Fragmented"]).values
         A = A.argmax(axis=-1)[..., None] == np.arange(A.shape[-1])
-        is_classified.values = np.concatenate((A, np.zeros((*A.shape[:2], 2), dtype=bool)), axis=-1)
+        is_classified.values = np.concatenate(
[…]

### 2020-12-17 src/load_data.py

@@ -163,6 +163,14 @@ def load_data(epoch_key, brain_areas=None):
     ripple_times = ripple_times.assign(
         duration=lambda df: (df.end_time - df.start_time).dt.total_seconds())
 
+    new_index = pd.Index(np.unique(np.concatenate(
+        (ripple_consensus_trace_zscore.index, time))), name='time')
+    ripple_consensus_trace_zscore = (ripple_consensus_trace_zscore
+                                     .reindex(index=new_index)
+                                     .interpolate(method='linear')
+                                     .reindex(index=time)
+                                     )
+
     return {
         'position_info': position_info,
         'ripple_times': ripple_times,

### 2021-02-01 src/load_data.py
Lower threshold on interneuron spike width

@@ -168,7 +168,7 @@ def load_data(epoch_key, brain_areas=None,
             SPIKE_SAMPLING_RATE)
 
     if exclude_interneuron_spikes:
-        INTERNEURON_SPIKE_WIDTH_MAX = 0.4  # ms
+        INTERNEURON_SPIKE_WIDTH_MAX = 0.3  # ms
         is_interneuron_spike = (
             multiunit.sel(features='max_width') < INTERNEURON_SPIKE_WIDTH_MAX)
         multiunit = multiunit.where(~is_interneuron_spike)

### 2021-02-01 src/standard_decoder.py
Add map estimate and weighted correlation

@@ -272,3 +272,34 @@ def detect_line_with_radon(likelihood, time, place_bin_edges):
     score = np.nanmax(sinogram) / n_position_bins
 
     return estimated_velocity, estimated_position, score
+
+
+def map_estimate(likelihood, place_bin_centers):
+    likelihood[np.isnan(likelihood)] = 0.0
+    likelihood = likelihood / likelihood.sum(axis=1, keepdims=True)
+
+    return place_bin_centers[likelihood.argmax(axis=1)].squeeze()
+
+
+def m(x, w):
+    """Weighted Mean"""
+    return np.sum(x * w) / np.sum(w)
+
+
+def cov(x, y, w):
+    """Weighted Covariance"""
+    return np.sum(w * (x - m(x, w)) * (y - m(y, w))) / np.sum(w)
+
+
+def corr(x, y, w):
+    """Weighted Correlation"""
+    return cov(x, y, w) / np.sqrt(cov(x, x, w) * cov(y, y, w))
+
+
+def get_weighted_correlation(likelihood, time, place_bin_centers):
+    place_bin_centers = place_bin_centers.squeeze()
+    likelihood[np.isnan(likelihood)] = 0.0
+    likelihood = likelihood / likelihood.sum(axis=1, keepdims=True)
+
+    return corr(time[:, np.newaxis],
+                place_bin_centers[np.newaxis, :], likelihood)

### 2021-02-01 Article.tex

@@
 While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
-The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds.
+For example with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps.
+The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds such as a discontinuous or curved trajectory \citep{LiuMethodsAssessmentMemory2018}.
 Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
@@
 % Figure 5-Supplemental 1
-\figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
+\figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.
+Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
 
@@
 \end{document}
+\label{fig:5}

### 2021-02-02 src/standard_decoder.py
Add isotonic regression

@@ -303,3 +303,29 @@ def get_weighted_correlation(likelihood, time, place_bin_centers):
 
     return corr(time[:, np.newaxis],
                 place_bin_centers[np.newaxis, :], likelihood)
+
+
+def isotonic_regression(likelihood, time, place_bin_centers):
+    place_bin_centers = place_bin_centers.squeeze()
+    likelihood[np.isnan(likelihood)] = 0.0
+    likelihood = likelihood / likelihood.sum(axis=1, keepdims=True)
+
+    map = map_estimate(likelihood, place_bin_centers)
+    map_probabilities = np.max(likelihood, axis=1)
+
+    regression = IsotonicRegression(increasing='auto').fit(
+        X=time,
+        y=map,
+        sample_weight=map_probabilities,
+    )
+
+    score = regression.score(
+        X=time,
+        y=map,
+        sample_weight=map_probabilities,
+    )
+    score *= np.mean(map)
+
+    prediction = regression.predict(time)
+
+    return prediction, score

### 2021-02-12 Article.tex
Update on Overleaf.

@@
 
-\title{Hippocampal replay of experience at real-world speeds}
+\title{Hippocampal replay of experience at slow speeds}
 
@@
 Representations of past and possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x real-world speeds.
+The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
 Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.
@@
 Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content.
-Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
+Furthermore, most events progress at slow, rather than accelerated, movement speeds, consistent with actual experiences.
 \end {abstract}
@@
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009}.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009, Tingleymethodsreactivationreplay2020}.
 One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
@@
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
+Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
 A decoding algorithm is then applied to the spikes within these events, yielding a set of position probability distributions for each time bin.
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
@@
 While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
-For example with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps.
-The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds such as a discontinuous or curved trajectory \citep{LiuMethodsAssessmentMemory2018}.
+For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps.
+The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds such as those that have curved trajectories or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
+The standard decoder is also limited by the choice of event boundaries, meaning the result is also influenced by the estimation of the start and end times of a sharp wave ripple or high multiunit synchrony event.
 Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
@@
 After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-One major goal of our work is to assess the overall prevalence of spatial content across SWRs, so we detect SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
+We first confirmed that our decoding algorithm could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sweeps and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
+We found, similar to other studies, that the average median difference is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation), but importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run and we were accurately able to decode immobility times as well, demonstrating the power of our algorithm.
 
+Having established the accuracy of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
 As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (\FIGSUPP[2]{fig2sup2}A-F).
@@
 These events included decoded representations where the most probable positions were highly spatially concentrated (i.e.
-had small HPD size).
+had small HPD size), indicating that they were driven by spatially specific firing.
 This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 100 cm/s, which is consistent with the running speed of rats (citation?).
 The next most common category were stationary events \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events, using large time bins, and using a restrictive decoding model.
-We also note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
+We also note that we chose to include a stationary dynamic to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
 
@@
 
-Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
+Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that a subset of replay trajectories contained jumps in position.
 We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
@@
 We used these dynamics to describe the SWR events, rather than declaring the events as significant or not.
-Fourth, our model avoids large temporal bins and does not make assumptions about the linear progression of events.
+Fourth, our model avoids large temporal bins and does not make assumptions about the linear progression of events while still taking into account the uncertainty of the data.
 Instead, the model can use any size of bin and allows for movement in any direction.

### 2021-02-26 Article.tex

@@
 Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content.
-Furthermore, most events progress at slow, rather than accelerated, movement speeds, consistent with actual experiences.
+Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
 \end {abstract}
@@
 Hippocampal "replay" events are well suited to that role \citep{CarrHippocampalreplayawake2011}.
-As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations.
+As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 Our uncertainty stems in part from the dominant approaches used to identify the content of replay events.
-These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are most commonly characterized using the standard "Bayesian" decoder.
+These approaches typically involve multiple steps and assumptions about the nature of replay, which are most commonly characterized using the standard "Bayesian" decoder.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
@@
 Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
-A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
+A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data.
 
-While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
-For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps.
-The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds such as those that have curved trajectories or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
-The standard decoder is also limited by the choice of event boundaries, meaning the result is also influenced by the estimation of the start and end times of a sharp wave ripple or high multiunit synchrony event.
+While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test.
+This is problematic because it has the potential to mischaracterize real events that do not move at constant speeds such as those that change direction or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
+Furthermore, the use of large, fixed-size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
+For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps and all spikes within the 20 ms bins are combined.
+The linear fit is also highly dependent on the estimation of event boundaries, such as the start and end times of a SWR, because the fit depends on all the data over the course of the event.
 Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
@@
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using an estimated represented location at each time bin and connecting each with a line.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin.
 For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
 On the other hand, \cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
-Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
+However, both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
 
@@
 It would use all of the available spiking data to yield the most accurate decoded positions.
-It would be applicable to either thresholded events or to all of the data to permit an unbiased assessment of the prevalence and nature of generative activity.
+It would provide a moment-by-moment estimate of position and dynamics that would not be dependent on the estimation of SWR start and end times and could rapidly adjust to changes in dynamics.
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
@@
 Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via a state transition matrix between dynamics.
-In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
+In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms on average, which is approximately the duration of a SWR event.
 There is, however, a small probability of switching to one of the other movement dynamics.
@@
 We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
-Furthermore, the probability of latent position remains relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
+Furthermore, the most probable latent positions remain relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
 
@@
 
-Using a 0.8 threshold, we are able to classify 89\% (23,382 of 26,160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
+Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
 To ensure that this depends on the sequential firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
-We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
+We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
+We find that only 14\% (3,295 of 23,071) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
 
 To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4,511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
+Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
 Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v5}
 \caption{
@@
 
+% Figure 5-Supplemental 2
+\figsupp[Distribution of dynamics is insensitive to parameters]{Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}
+
+% Figure 5-Supplemental 3
+\figsupp[Further quantification of spiking properties]{\textbf{(A)} Dynamics.
+\textbf{(B)} Ripples  }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5_v1_supplemental3}}\label{figsupp:fig5sup3}
+
 % Figure 5-video1
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (16454 of 23382 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (\FIG{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
-These events contain representations that move at slower speeds (\FIG{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
-continuous 95 ms, \FIG{5}B).
-Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
-continuous 42 cm, \FIG{5}C).
+We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (\FIG{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
+These events contain representations that move at slower speeds (\FIG{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
+continuous 94 ms, \FIG{5}B).
+Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 52 cm vs.
+continuous 43 cm, \FIG{5}C).
 This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
 
-The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs).
+The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C).
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (\texttildelow7\%) of classified SWRs.
-About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1,437 of 21,433 (\texttildelow7\%) of classified SWRs.
+46\% of these (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
 These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
@@
 We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been previously captured.
-This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move more or less linearly at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
+This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
 These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
@@
 Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
-These events included decoded representations where the most probable positions were highly spatially concentrated (i.e.
+These events were mostly decoded representations where the most probable positions were highly spatially concentrated (i.e.
 had small HPD size), indicating that they were driven by spatially specific firing.
-This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
+This was in contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 100 cm/s, which is consistent with the running speed of rats (citation?).
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (100 cm/s), which is consistent with the running speed of rats (citation?).
 The next most common category were stationary events \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
-Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
+Events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
 Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
+These results further challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
@@
 Here, we show that they do not disappear, but are a very common occurrence with adult rats.
-We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events, using large time bins, and using a restrictive decoding model.
-We also note that we chose to include a stationary dynamic to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
+We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events and using a restrictive decoding model which requires the decoded position to be stationary for the entire event.
+We also note that we chose to include a stationary dynamic to explicitly evaluate the prevalence of stationary events in adult animals, but in principle, a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
 
@@
 Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \citep{YuHippocampalcorticalinteraction2015}.
+
+Our results showed that 19\% of SWRs contained continuous trajectories---meaning that portions of the SWR had spatially coherent trajectories that moved at high velocities---but only 4\% were continuous throughout the entire SWR.
+This, \textit{prima facie}, may seem at odds with previous results, which found between 5\% and 45\% of SWRs were "significant", meaning that they were well fit by a line compared to various shuffled versions of the dataset.
+Furthermore, these studies found trajectories on average move at much higher speeds (20x normal rat movement speed) than we report here.
+While the plurality of methods and parameters makes it hard to directly compare, we believe there are several reasons for this difference.
+We use a model with smaller time bins that allows for changes of speeds within SWRs, whereas methods using a linear fit use larger time bins and cannot accommodate rapid change.
+Studies using a linear fit typically incorporate position bins within 15 cm of the best fit line in their best fit score in order to compensate for the rigidness of this restriction, but this allows for a wide range of speeds when using a 20 ms time bin (up to 3.75 m/s).
+The linear fit also depends on the start and end event boundaries.
+Many studies explicitly excluded slow events.
 
@@
 
-Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
-We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology may reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenTransductiveneuraldecoding2012, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology could reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
 
-Third, we built an explicit but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
+Third, we built an explicit and flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
 We used these dynamics to describe the SWR events, rather than declaring the events as significant or not.
-Fourth, our model avoids large temporal bins and does not make assumptions about the linear progression of events while still taking into account the uncertainty of the data.
-Instead, the model can use any size of bin and allows for movement in any direction.
-Finally, we used an acausal estimator of latent position, which allowed us to identify the specific timing of transitions in dynamics at a fine timescale.
+
+Fourth, our model avoided using large temporal bins and does not make assumptions about the linear progression of events while still taking into account the uncertainty of the data.
+Instead, the model allows for movement in any direction allowed by geometry of the environment.
+Because the model gives a moment-by-moment estimate of the position and dynamic, this also reduces the effect of misidentifying the SWR start and end times.
+Finally, combined with an acausal estimation of latent position, this approach allows us to identify the specific timing of transitions in dynamics at a fine timescale.
 

### 2021-02-27 Article.tex

@@
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
 Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
@@
 Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
-A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data.
+A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent some definition of a "random" sequence.
 
@@
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, these approaches exclude events that have stationary representations of a single location \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
+Moreover, these approaches often exclude events that have stationary representations of a single location \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
 
@@
 We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
-Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
 We therefore define state transition models to capture each of these movement dynamics.
@@
 
-Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via a state transition matrix between dynamics.
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via another state transition matrix between the movement dynamics.
 In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms on average, which is approximately the duration of a SWR event.
@@
 46\% of these (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
+These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (100 cm/s), which is consistent with the running speed of rats (citation?).
-The next most common category were stationary events \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
+The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
 Events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
@@
 We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
-Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
 Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \citep{YuHippocampalcorticalinteraction2015}.
@@
 First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.
-This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \citep{JaiDistincthippocampalcorticalmemory2017}.
+This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \citep{YuDistincthippocampalcorticalmemory2017}.
 More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
@@
 The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \citep{CasellaStatisticalinference2001}.
-In this manuscript, we use the HPD size to evaluate the uncertainty of the posterior probability of position.
-We calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
+By using only the top values, this measurement of spread is not influenced by multimodal distributions (whereas an alternative measure like the quantiles of the distribution would be).
+In this manuscript, we use the HPD size---the total area of the track covered by the 95\% HPD region---to evaluate the uncertainty of the posterior probability of position.
+
+In order to calculate the HPD size, we first calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
 $$
@@
 \subsection*{Distance from animal}
-The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is the center of the position bin with the greatest posterior value.
 The shortest path is found using Dijkstra's algorithm \citep{Dijkstranotetwoproblems1959} as implemented in NetworkX \citep{HagbergExploringNetworkStructure2008}.
@@
 \subsection*{Estimation of Speed}
-In order to estimate the speed over time, we first found the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
+In order to estimate the speed over time, we first found the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is the center of the position bin with the greatest posterior value.
 We then computed the first derivative using the Python library Numpy's gradient function, which computes the difference in the forward and backward time direction and then averages them for all points except the boundaries.

### 2021-03-10 Article.tex
Update on Overleaf.

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[1]{Eric L. Denovellis}
+\author[1, 2, 3]{Eric L. Denovellis}
 \author[2, 3]{Anna K. Gillespie}
@@
 \begin{figure}
-\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v6}
+\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v7}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel).
+Red dotted line in the right panel is the best fit line from the standard decoder.
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
@@
 After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-We first confirmed that our decoding algorithm could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sweeps and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
+We confirmed that our decoding algorithm could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sweeps and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
 We found, similar to other studies, that the average median difference is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation), but importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run and we were accurately able to decode immobility times as well, demonstrating the power of our algorithm.
 
-Having established the accuracy of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs, so we detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
+Having established the accuracy of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs.
+We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
 As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (\FIGSUPP[2]{fig2sup2}A-F).
@@
 Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
-To ensure that this depends on the sequential firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
 We then decode the same SWR events containing the same spikes.
@@
 Surprisingly, the most common category is stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
-These events contain representations that move at slower speeds (\FIG{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
+These events contain representations that move at slower speeds (\FIG{5}D), and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
 continuous 94 ms, \FIG{5}B).
@@
 This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
+We note that roughly a third of these events persisted for the entire duration of the SWR, but the other two thirds had some portion of time where the model was uncertain about the dynamic.
+This shows the power of our classifier because it allows us to characterize the content of a SWR even if the model is only certain about a portion of the SWR time.
 
@@
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
+
+We performed several controls to see if our results were influenced by the firing of interneurons, which typically do not project to other brain areas and therefore cannot be read by a downstream region.
+First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active in the main results) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
+This had little effect on the overall proportions, with many stationary and stationary-continuous mixtures still present (\FIGSUPP[5]{fig5sup2}A).
+Indeed, most of our events included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}B), including within each dynamic (\FIGSUPP[5]{fig5sup3}A).
+Next, we tried removing spikes that had spike widths of less than 0.3 ms because interneurons typically have narrow spike widths.
+Note that this is a rather imprecise threshold, because typically this is computed on the average waveform for a putative cell, not on a single spike, and combined with a requirement for the putative cell to have a high firing rate, as interneurons typically exhibit high firing rates.
+Nevertheless, even with this imprecise criteria, we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
+We also looked at the 9 of 10 animals in our dataset that already had clustered cells, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
+Although there were many fewer events to examine, we still observed many stationary and stationary-continuoux mixtures in SWRs compared to continuous sequences (\FIGSUPP[5]{fig5sup2}D).
+This provides further evidence that our results were not the result of the firing of interneurons.
+
+Another possible concern is that many of these slow spatially coherent events occur when the animal is moving very slowly, meaning the stationary-continuous mixtures potentially reflect a slow theta sweep rather than a replay event.
+We do not believe this to be the case because the majority of our events occur when the animal is moving at speeds < 2 cm/s (\FIGSUPP[5]{fig5sup3}B) and when we restrict ourselves to SWRs where the animal was moving at speeds < 1 cm/s, we find similar proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}C).
 
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (100 cm/s), which is consistent with the running speed of rats (citation?).
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the running speed of rats.
 The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 
-These results further challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
+These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
 Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
@@
 While the plurality of methods and parameters makes it hard to directly compare, we believe there are several reasons for this difference.
+First, our model is much more flexible because it allows the trajectory to change speed during the SWR, whereas fitting a line assumes a constant speed trajectory.
+So many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics (see for example \FIG{1}).
+Second, the line has to be fit to the whole SWR event, meaning that it is sensitive to the estimation of the start end of the SWR event.
+In contrast, our model makes an estimate for each time point and only depends on the previous and future time step, which minimizes the dependence of the event on those boundaries and any possible "noise" events within the SWR.
+
 We use a model with smaller time bins that allows for changes of speeds within SWRs, whereas methods using a linear fit use larger time bins and cannot accommodate rapid change.

### 2021-03-10 src/standard_decoder.py
Better names

@@ -297,11 +297,13 @@ def detect_line_with_radon(
 ):
 
     if incorporate_nearby_positions:
-        n_bins = int(nearby_positions_max // dp)
-        filt = np.ones(2 * n_bins + 1)
+        n_nearby_bins = int(nearby_positions_max // dp)
+        filt = np.ones(2 * n_nearby_bins + 1)
         posterior = np.apply_along_axis(
             lambda m: np.convolve(m, filt, mode="same"), axis=1, arr=posterior
         )
+    else:
+        n_nearby_bins = 1
     # Sinogram is shape (pixels_from_center, projection_angles)
     sinogram = radon(
         posterior.T, theta=projection_angles, circle=False,

### 2021-03-10 src/standard_decoder.py
Normalize the radon score to be less than one

@@ -335,7 +335,7 @@ def detect_line_with_radon(
     n_pixels_from_center = pixels_from_center[n_pixels_from_center_ind]
 
     # Normalized score based on the integrated projection
-    score = np.max(sinogram) / n_time
+    score = np.max(sinogram) / n_time * n_nearby_bins
 
     # Convert from polar form to slope-intercept form
     start_position, velocity = convert_polar_to_slope_intercept(

### 2021-03-10 scripts/queue_cluster_jobs_standard_decoder.py
Add script to run standard decoder

@@ -1,0 +1,90 @@
+'''Script for executing run_by_epoch on the cluster
+'''
+from argparse import ArgumentParser
+from os import environ, getcwd, makedirs
+from os.path import join
+from subprocess import run
+from sys import exit
+
+from loren_frank_data_processing import (make_epochs_dataframe,
+                                         make_neuron_dataframe)
+from src.parameters import _BRAIN_AREAS, ANIMALS
+
+
+def get_command_line_arguments():
+    parser = ArgumentParser()
+    parser.add_argument('--Animal', type=str, help='Short name of animal')
+    parser.add_argument('--Day', type=int, help='Day of recording session')
+    parser.add_argument('--Epoch', type=int,
+                        help='Epoch number of recording session')
+    parser.add_argument('--n_cores', type=int, default=16)
+    parser.add_argument('--wall_time', type=str, default='6:00:00')
+    parser.add_argument('--n_workers', type=int, default=16)
+    parser.add_argument('--threads_per_worker', type=int, default=16)
+
+    return parser.parse_args()
+
+
+def queue_job(python_cmd, directives=None, log_file='log.log',
+              job_name='job'):
+    queue_cmd = f'qsub {directives} -j y -o {log_file} -N {job_name}'
+    cmd_line_script = f'echo python {python_cmd}  | {queue_cmd}'
+    run(cmd_line_script, shell=True)
+
+
+def main():
+
+    args = get_command_line_arguments()
+
+    # Set the maximum number of threads for openBLAS to use.
+    num_threads = str(args.threads_per_worker)
+    environ['OPENBLAS_NUM_THREADS'] = num_threads
+    environ['MKL_NUM_THREADS'] = num_threads
+    environ['NUMBA_NUM_THREADS'] = num_threads
+    environ['OMP_NUM_THREADS'] = num_threads
+    LOG_DIRECTORY = join(getcwd(), 'logs')
+    makedirs(LOG_DIRECTORY,  exist_ok=True)
+
+    python_function = 'run_standard_decoder.py'
+    directives = ' '.join(
+        [f'-l h_rt={args.wall_time}', f'-pe omp {args.n_cores}',
+         '-P braincom', '-notify',
+         '-v OPENBLAS_NUM_THREADS', '-v NUMBA_NUM_THREADS',
+         '-v OMP_NUM_THREADS'])
+
+    if args.Animal is None or args.Day is None or args.Epoch is None:
[…]

### 2021-03-10 scripts/run_standard_decoder.py

@@ -1,0 +1,122 @@
+import logging
+import sys
+from argparse import ArgumentParser
+from signal import SIGUSR1, SIGUSR2, signal
+from subprocess import PIPE, run
+
+import xarray as xr
+from loren_frank_data_processing import save_xarray
+from src.parameters import PROCESSED_DATA_DIR
+from src.standard_decoder import (fit_mark_likelihood, load_data,
+                                  predict_clusterless_radon_wtrack)
+
+FORMAT = '%(asctime)s %(message)s'
+
+logging.basicConfig(level='INFO', format=FORMAT, datefmt='%d-%b-%y %H:%M:%S')
+
+
+def clusterless_analysis_1D(epoch_key, dt=0.020):
+    (
+        linear_position,
+        multiunit_dfs,
+        _,
+        ripple_times,
+        track_graph,
+        center_well_id,
+    ) = load_data(epoch_key)
+    (
+        occupancy,
+        joint_pdf_models,
+        multiunit_dfs,
+        ground_process_intensities,
+        mean_rates,
+        is_track_interior,
+        place_bin_centers,
+        place_bin_edges,
+        edges,
+    ) = fit_mark_likelihood(
+        linear_position, multiunit_dfs, track_graph, center_well_id)
+
+    results = []
+    for ripple_number in ripple_times.index:
+        (
+            time,
+            radon_estimated_velocity,
+            radon_prediction,
+            radon_score,
+            likelihood,
+        ) = predict_clusterless_radon_wtrack(
+            ripple_times,
+            ripple_number,
+            place_bin_centers,
+            occupancy,
+            joint_pdf_models,
+            multiunit_dfs,
+            ground_process_intensities,
+            mean_rates,
+            is_track_interior,
+            place_bin_edges,
+            dt=dt,
+        )
+        results.append(
+            xr.Dataset(
+                data_vars={
+                    "radon_prediction": (["time"], radon_prediction),
+                    "radon_score": radon_score,
+                    "radon_estimated_velocity": radon_estimated_velocity,
+                },
+                coords={"time": time + dt / 2},
+            ))
+
+    results = xr.concat(results, dim=ripple_times.index)
+    save_xarray(PROCESSED_DATA_DIR,
+                epoch_key,
+                results,
+                group='/clusterless/1D/standard_decoder/ripples/')
+
+
+def get_command_line_arguments():
[…]

### 2021-03-10 src/analysis.py
Add function to load all replay info

@@ -1,10 +1,16 @@
+import os
+from glob import glob
+
 import numpy as np
 import pandas as pd
 import scipy
 import xarray as xr
+from loren_frank_data_processing import make_tetrode_dataframe
 from loren_frank_data_processing.track_segment_classification import \
     project_points_to_segment
 from scipy.ndimage.filters import gaussian_filter1d
+from src.parameters import (_BRAIN_AREAS, ANIMALS, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR)
 from trajectory_analysis_tools import (get_ahead_behind_distance,
                                        get_trajectory_data)
 
@@ -547,3 +553,52 @@ def gaussian_smooth(data, sigma, sampling_frequency, axis=0):
     '''
     return gaussian_filter1d(
         data, sigma * sampling_frequency, axis=axis)
+
+
+def load_all_replay_info(
+    n_unique_spiking=2,
+    data_type="clusterless",
+    dim="1D",
+    probability_threshold=PROBABILITY_THRESHOLD,
+    speed_threshold=4,
+    exclude_interneuron_spikes=False,
+):
+    tetrode_info = make_tetrode_dataframe(ANIMALS)
+    prob = int(probability_threshold * 100)
+    if exclude_interneuron_spikes:
+        interneuron = 'no_interneuron_'
+    else:
+        interneuron = ''
+    file_regex = f"*_{data_type}_{dim}_{interneuron}replay_info_{prob:02d}.csv"
+    file_paths = glob(os.path.join(PROCESSED_DATA_DIR, file_regex))
+    replay_info = pd.concat(
+        [pd.read_csv(file_path) for file_path in file_paths], axis=0,
+    ).set_index(["animal", "day", "epoch", "ripple_number"])
+    replay_info["fraction_unclassified"] = (
+        replay_info.Unclassified_duration
+        / replay_info.duration
+    )
+    replay_info["duration_classified"] = (
+        replay_info.duration - replay_info.Unclassified_duration)
+    replay_info = replay_info.loc[
+        (replay_info.n_unique_spiking >= n_unique_spiking) &
+        (replay_info.actual_speed <= speed_threshold)
+    ].sort_index()
+
+    is_brain_areas = tetrode_info.area.astype(
+        str).str.upper().isin(_BRAIN_AREAS)
+    n_tetrodes = (
+        tetrode_info.loc[is_brain_areas]
[…]

### 2021-03-10 src/parameters.py
Add unclassified category

@@ -78,22 +78,26 @@ TRANSITION_TO_CATEGORY = {
 PROBABILITY_THRESHOLD = 0.8
 
 STATE_ORDER = ['Hover', 'Hover-Continuous-Mix', 'Continuous',
-               'Fragmented-Continuous-Mix', 'Fragmented']
+               'Fragmented-Continuous-Mix', 'Fragmented', 'Unclassified']
 
-SHORT_STATE_ORDER = ['Hover', 'Hover-Cont.-Mix', 'Cont.',
-                     'Frag.-Cont.-Mix', 'Frag.']
+SHORT_STATE_ORDER = ['Stat.', 'Stat.-Cont.-Mix', 'Cont.',
+                     'Frag.-Cont.-Mix', 'Frag.', 'Unclass.']
 
 # Plotting Colors
 STATE_COLORS = {
     'Hover': '#9f043a',
+    'Stat.': '#9f043a',
+    'Stationary': '#9f043a',
     'Fragmented': '#ff6944',
     'Frag.': '#ff6944',
     'Continuous': '#521b65',
     'Cont.': '#521b65',
     'Hover-Continuous-Mix': '#61c5e6',
-    'Hover-Cont.-Mix': '#61c5e6',
+    'Stationary-Continuous-Mix': '#61c5e6',
+    'Stat.-Cont.-Mix': '#61c5e6',
     'Fragmented-Continuous-Mix': '#2a586a',
     'Frag.-Cont.-Mix': '#2a586a',
     '': '#c7c7c7',
     'Unclassified': '#c7c7c7',
+    'Unclass.': '#c7c7c7',
 }

### 2021-03-10 src/visualization.py

@@ -31,11 +31,12 @@ MILLISECONDS_TO_SECONDS = 1000
 
 
 SHORT_STATE_NAMES = {
-    "Hover": "Hover",
-    "Hover-Continuous-Mix": "Hover-Cont.-Mix",
+    "Hover": "Stat.",
+    "Hover-Continuous-Mix": "Stat.-Cont.-Mix",
     "Continuous": "Cont.",
     "Fragmented-Continuous-Mix": "Frag.-Cont.-Mix",
-    "Fragmented": "Frag."
+    "Fragmented": "Frag.",
+    "Unclassified": "Unclass."
 }
 
 

### 2021-03-10 scripts/run_standard_decoder.py
Add logging statements

@@ -16,6 +16,7 @@ logging.basicConfig(level='INFO', format=FORMAT, datefmt='%d-%b-%y %H:%M:%S')
 
 
 def clusterless_analysis_1D(epoch_key, dt=0.020):
+    logging.info('Loading data...')
     (
         linear_position,
         multiunit_dfs,
@@ -24,6 +25,7 @@ def clusterless_analysis_1D(epoch_key, dt=0.020):
         track_graph,
         center_well_id,
     ) = load_data(epoch_key)
+    logging.info('Fitting model...')
     (
         occupancy,
         joint_pdf_models,
@@ -37,6 +39,7 @@ def clusterless_analysis_1D(epoch_key, dt=0.020):
     ) = fit_mark_likelihood(
         linear_position, multiunit_dfs, track_graph, center_well_id)
 
+    logging.info('Predicting with radon...')
     results = []
     for ripple_number in ripple_times.index:
         (
@@ -68,6 +71,7 @@ def clusterless_analysis_1D(epoch_key, dt=0.020):
                 coords={"time": time + dt / 2},
             ))
 
+    logging.info('Saving results...')
     results = xr.concat(results, dim=ripple_times.index)
     save_xarray(PROCESSED_DATA_DIR,
                 epoch_key,

### 2021-04-22 Article.tex

@@
 This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).
-Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
+Last, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
 We model this using a uniform state transition matrix, which makes transitions to each position equally likely.
@@
 Subsequently, as the population activity becomes spatially incoherent, the fragmented movement dynamic dominates for the last 30 ms of the simulated event.
+This also shows that the fragmented dynamic gives us a way to test for spatially incoherent content directly for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder.
 
@@
 Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}}\label{figsupp:fig2sup2}
+
+% Figure 2-Supplemental 3
+\figsupp[Spatial coverage of the track for each animal.]{}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
 
@@
 After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-We confirmed that our decoding algorithm could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sweeps and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
-We found, similar to other studies, that the average median difference is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation), but importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run and we were accurately able to decode immobility times as well, demonstrating the power of our algorithm.
+We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
+We found, similar to other studies, that the average median difference is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
+Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
+We also verified that the population spiking for all positions on the track was consistent for each animal \FIGSUPP[2]{fig2sup3}.
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v5}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v6}
 \caption{
@@
 
-Finally, we examined the overall levels of neuronal spiking across dynamics.
+We examined the overall levels of neuronal spiking across dynamics.
 We found that all the dynamics had similar average multiunit spiking rates (\FIG{5}F).
@@
 
-We performed several controls to see if our results were influenced by the firing of interneurons, which typically do not project to other brain areas and therefore cannot be read by a downstream region.
-First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active in the main results) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
+We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (although note, their firing still may reflect relevant position information \ref{}).
+First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
 This had little effect on the overall proportions, with many stationary and stationary-continuous mixtures still present (\FIGSUPP[5]{fig5sup2}A).
@@
 Nevertheless, even with this imprecise criteria, we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
-We also looked at the 9 of 10 animals in our dataset that already had clustered cells, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
-Although there were many fewer events to examine, we still observed many stationary and stationary-continuoux mixtures in SWRs compared to continuous sequences (\FIGSUPP[5]{fig5sup2}D).
-This provides further evidence that our results were not the result of the firing of interneurons.
+We also examined the 9 of 10 animals in our dataset that already had clustered cells, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
+Although there were many fewer events to examine, we still observed many stationary and stationary-continuous mixtures in SWRs compared to continuous sequences (\FIGSUPP[5]{fig5sup2}D).
+Finally,  if interneurons were not spatially specific and driving the slower dynamics, we would expect to see the posterior be less spatially concentrated.
+Instead, we found that these dynamics tended to have narrow posteriors (as characterized by 95\% HPD region size in \FIG{4}E).
+This provides further evidence that our results were not the solely the result of the firing of interneurons.
 
@@
 We do not believe this to be the case because the majority of our events occur when the animal is moving at speeds < 2 cm/s (\FIGSUPP[5]{fig5sup3}B) and when we restrict ourselves to SWRs where the animal was moving at speeds < 1 cm/s, we find similar proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}C).
+
+A third possible concern is that burst firing from a single cell is driving the stationary dynamics.
+There are several lines of evidence against this.
+First, as mentioned above, most of our SWRs had spikes from multiple tetrodes, including during stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell is driving these dynamics.
+Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
+bursts which have a duration of 6-24 ms).
+Third, only ~50\% of the spikes had ISIs of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), so if bursting occurs, it only occurs on half of the SWRs we have analyzed.
+It should be noted that this metric in particular should be interpreted with extreme caution, because with clusterless decoding, the ISI of the multiunit spikes is not the same as if we had an individual unit.
+The spikes could be from multiple neurons firing, which would not tell us if there was bursting or not.
+Nevertheless, even with these caveats, it seems unlikely that burst firing from cells is driving the stationary dynamics.
 

### 2021-04-26 Article.tex
Update on Overleaf.

@@
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the standard decoder.
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder, incorporate information about times when there is no spiking, and make fewer assumptions about the amount of time a downstream neuron can integrate spikes from multiple neurons (because place estimates from spikes within a single bin are combined).
 
@@
 
+% Figure 3-Supplemental 3
+\figsupp[Comparison of percentage of SWRs classified on real vs.
+position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
+position shuffled data for two epochs from different animals.
+Red line represents the percent of SWR events classified in that epoch for real data.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1-supplemental3}}\label{figsupp:fig3sup3}
+
 % Figure 3-video1
@@
 
-Having established the accuracy of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs.
+Having established the validity of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs.
 We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
@@
 
-Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
-To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-We then decode the same SWR events containing the same spikes.
+Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or dynamic mixtures.
+To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement, which shuffles the relationship between spiking and position, for two recording sessions , fit the encoding model and then decode the same SWR events containing the same spikes.
 Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
+This shows that our model will not classify dynamics when there is no spatial information present in the data.
 
@@
 Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
+Note that here we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by the standard decoder in the case of fitting a line, because our decoder is much more flexible in terms of allowing different speeds during the SWR event.
 
@@
 Surprisingly, the most common category is stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
-These events contain representations that move at slower speeds (\FIG{5}D), and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
+These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
 continuous 94 ms, \FIG{5}B).
@@
 
-We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (although note, their firing still may reflect relevant position information \ref{}).
+We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (although note, interneuron firing still may have relevant position information and it is not clear that they should be excluded \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
 First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
@@
 
+One last concern is that we have combined spiking information from several hippocampal subfields (CA1, CA2, and CA3), which may influence the types of dynamics we observed.
+However, CA1 and CA3 spiking is largely coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
+Therefore, we feel our combination of spiking information from multiple subfields is appropriate.
+
 \section*{Discussion}
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the running speed of rats.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats experience the environment.
 The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 
-We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
+We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of experience.
 Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
-Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \citep{YuHippocampalcorticalinteraction2015}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to recall and evaluate the value of specific locations or short movement trajectories without having to generate full trajectories \citep{YuHippocampalcorticalinteraction2015}.
+This is consistent with our own subjective experience of memory, where we do not necessarily have to recall an entire path in detail for it to be useful, particularly if it is a well traveled route.
 
@@
 The code used for ripple detection can be found at \url{https://github.com/Eden-Kramer-Lab/ripple_detection}.
+We only analyzed SWRs with spikes from at least 2 tetrodes.
 

### 2021-04-27 Article.tex

@@
 % Figure 3-Supplemental 1
-\figsupp[More examples of SWRs that would not be well-characterized by using the standard decoder]{A-I.
+\figsupp[More examples of SWRs with non-constant speed trajectories]{A-I.
 More examples of SWRs that would not be well-characterized by using the standard decoder.
@@
 % Figure 3-Supplemental 3
-\figsupp[Comparison of percentage of SWRs classified on real vs.
-position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
-position shuffled data for two epochs from different animals.
-Red line represents the percent of SWR events classified in that epoch for real data.
-The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1-supplemental3}}\label{figsupp:fig3sup3}
+\figsupp[Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
+run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
+run-swapped, circularly shuffled position.
+The red line represents the percent of spatially coherent or incoherent SWRs in that epoch for actual data.
+The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
+This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1_supplemental3}}\label{figsupp:fig3sup3}
 
@@
 
-Having established the validity of our decoder, we next wanted to assess the overall prevalence of spatial content across SWRs.
+Having established the validity of our decoder in tracking the animal's position, we next wanted to assess the overall prevalence of spatial content across SWRs.
 We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
@@
 Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or dynamic mixtures.
-To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement, which shuffles the relationship between spiking and position, for two recording sessions , fit the encoding model and then decode the same SWR events containing the same spikes.
+To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement, which shuffles the relationship between spiking and position, for two recording sessions, fit the encoding model and then decode the same SWR events containing the same spikes.
 Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
@@
 We find that only 14\% (3,295 of 23,071) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
+To further validate this result, we performed a shuffle that randomized the order of the reward well to reward well runs and then circularly permuted the run.
+This, in contrast to our previous shuffle, should preserve correlation between spikes locally but destroy global spatial structure.
+Using this shuffle, we found that there was more shuffled spatially incoherent content compared to the real data and there was less spatially coherent content compared to the real data for two epochs (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
 
@@
 % Figure 5-Supplemental 3
-\figsupp[Further quantification of spiking properties]{\textbf{(A)} Dynamics.
-\textbf{(B)} Ripples  }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5_v1_supplemental3}}\label{figsupp:fig5sup3}
+\figsupp[Further quantification of spiking and ripple properties by dynamic]{\textbf{(A)} Further quantification of the dynamics with respect to the spiking and ripple properties.
+\textbf{(B)} Speed and number of tetrodes with spikes for entire SWR }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5_v1_supplemental3}}\label{figsupp:fig5sup3}
 
@@
 Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to recall and evaluate the value of specific locations or short movement trajectories without having to generate full trajectories \citep{YuHippocampalcorticalinteraction2015}.
-This is consistent with our own subjective experience of memory, where we do not necessarily have to recall an entire path in detail for it to be useful, particularly if it is a well traveled route.
+
+One possibility is that these events are supported by different subcircuits within the hippocampus.
+For example, \cite{OlivaRoleHippocampalCA22016} showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs We did not have enough simultaneously recorded tetrodes in each area to attribute the different dynamics to particular subregions.
 
@@
 While the plurality of methods and parameters makes it hard to directly compare, we believe there are several reasons for this difference.
-First, our model is much more flexible because it allows the trajectory to change speed during the SWR, whereas fitting a line assumes a constant speed trajectory.
-So many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics (see for example \FIG{1}).
+First, our model is much more flexible because it allows the trajectory to change speed during the SWR, whereas fitting a line assumes a constant speed trajectory for the entire SWR.
+Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics (see for example \FIG{1}).
 Second, the line has to be fit to the whole SWR event, meaning that it is sensitive to the estimation of the start end of the SWR event.
 In contrast, our model makes an estimate for each time point and only depends on the previous and future time step, which minimizes the dependence of the event on those boundaries and any possible "noise" events within the SWR.
-
-We use a model with smaller time bins that allows for changes of speeds within SWRs, whereas methods using a linear fit use larger time bins and cannot accommodate rapid change.
-Studies using a linear fit typically incorporate position bins within 15 cm of the best fit line in their best fit score in order to compensate for the rigidness of this restriction, but this allows for a wide range of speeds when using a 20 ms time bin (up to 3.75 m/s).
-The linear fit also depends on the start and end event boundaries.
-Many studies explicitly excluded slow events.
+This, along with our more permissive threshold for SWRs, allows us to characterize more SWRs than previous studies, rather than not analyze them.
 

### 2021-04-28 Article.tex
Update on Overleaf.

@@
 
+Finally, one important question is how our method compares to the standard decoder.
+We would like to emphasize, as we have stated in the introduction, that this comparison is complicated due to the variety of parameters and inference methods used and referred to as the standard "Bayesian" decoder.
+
 \section*{Discussion}
@@
 Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
-Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to recall and evaluate the value of specific locations or short movement trajectories without having to generate full trajectories \citep{YuHippocampalcorticalinteraction2015}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to store and evaluate the value of specific locations or short movement trajectories without having to generate full trajectories \citep{YuHippocampalcorticalinteraction2015}.
 
-One possibility is that these events are supported by different subcircuits within the hippocampus.
-For example, \cite{OlivaRoleHippocampalCA22016} showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs We did not have enough simultaneously recorded tetrodes in each area to attribute the different dynamics to particular subregions.
+One possibility is that these events are supported by different subcircuits within the hippocampus.For example, \cite{OlivaRoleHippocampalCA22016} showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs.
+We did not have enough simultaneously recorded tetrodes in each area to attribute the different dynamics to particular subregions.
 

### 2021-04-28 Article.tex

@@
 
-One possibility is that these events are supported by different subcircuits within the hippocampus.For example, \cite{OlivaRoleHippocampalCA22016} showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs.
+One possibility is that these events are supported by different subcircuits within the hippocampus.
+For example, \cite{OlivaOriginGammaFrequency2018a} showed that layer 2/3 neurons in the entorhinal cortex increased their firing 50 ms before the start of longer duration SWRs (>100 ms), but not before the start of shorter duration SWRs, suggesting that
+
+showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs.
 We did not have enough simultaneously recorded tetrodes in each area to attribute the different dynamics to particular subregions.
+However, we did show that our result holds even if restricting the dataset only to CA1 tetrodes.
 

### 2021-04-28 Article.tex
Update on Overleaf.

@@
 One possibility is that these events are supported by different subcircuits within the hippocampus.
-For example, \cite{OlivaOriginGammaFrequency2018a} showed that layer 2/3 neurons in the entorhinal cortex increased their firing 50 ms before the start of longer duration SWRs (>100 ms), but not before the start of shorter duration SWRs, suggesting that
+For example, \cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
+This raises the possibility that these slower replays are driven by CA3-CA1 interaction while the faster trajectories are more externally driven by MEC-CA1.
+Our
 

### 2021-04-30 Article.tex
Update on Overleaf.

@@
 
+% Figure 6
+\begin{figure}
+\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_v2}
+\caption{
+Comparison of the state space decoder to the standard "Bayesian" decoder. \textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder
+}
+\label{fig:6}
+
+\end{figure}
+
 We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (although note, interneuron firing still may have relevant position information and it is not clear that they should be excluded \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
@@
 
-Finally, one important question is how our method compares to the standard decoder.
-We would like to emphasize, as we have stated in the introduction, that this comparison is complicated due to the variety of parameters and inference methods used and referred to as the standard "Bayesian" decoder.
+% Table 1
+\begin{wraptable}{l}{0.5\textwidth}{
+\begin{tabular}{lll}
+\textbf{Dynamic}                       & \textbf{\begin{tabular}[c]{@{}l@{}}Percent.\\ Sig.\end{tabular}} & \textbf{Number Sig.}               \\ \hline
+\multicolumn{1}{|l|}{Stat.}            & \multicolumn{1}{l|}{8\%}                                         & \multicolumn{1}{l|}{542 of 7176}   \\ \hline
+\multicolumn{1}{|l|}{Stat.-Cont.-Mix.} & \multicolumn{1}{l|}{14\%}                                        & \multicolumn{1}{l|}{2253 of 16106} \\ \hline
+\multicolumn{1}{|l|}{Cont.}            & \multicolumn{1}{l|}{30\%}                                        & \multicolumn{1}{l|}{1368 of 4511}  \\ \hline
+\multicolumn{1}{|l|}{Cont.-Frag.-Mix.} & \multicolumn{1}{l|}{31\%}                                        & \multicolumn{1}{l|}{799 of 2562}   \\ \hline
+\multicolumn{1}{|l|}{Frag.}            & \multicolumn{1}{l|}{15\%}                                        & \multicolumn{1}{l|}{244 of 1598}   \\ \hline
+\end{tabular}
+}
+  \caption{Number of significant SWRs that contain each dynamic. SWRs were fit using the random transform with 1000 shuffles of the posterior in each time bin.}\label{tab:supp1}
+\end{wraptable}
+
+Finally, an important question is how our method compares to the standard "Bayesian" decoder.
+This comparison is complicated due to the variety of parameters and inference methods used and referred to as the standard "Bayesian" decoder.
+For the sake of comparison, we used clusterless decoding in 20 ms time bins and used three common methods of inferring the trajectory: the radon transform \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, linear regression \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, and the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+We then compared how the estimated trajectory speeds varied by dynamic for each method.
+This reveals two key insights.
+First, the choice of method can give very different estimated speeds distributions for the SWRs.
+In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the SWRs proceed at faster speeds.
 

### 2021-05-03 Article.tex

@@
 Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
+
+% Figure 4-Supplemental 2
+\figsupp[Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data for two epochs from different animals.
+Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
+The histogram represents the distribution after 50 shuffles of the position data.
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
+
 \end{figure}
@@
 \caption{
-Comparison of the state space decoder to the standard "Bayesian" decoder. \textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder
+Comparison of the state space decoder to the standard "Bayesian" decoder.
+For each panel, the top five rows are the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
+Note that these categories are not mutually exclusive since a SWR can have more than one dynamic or combination of dynamics.
+The sixth row is all SWRs containing unclassified dynamics.
+The final row is the estimated speed for all SWRs.
+\textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder presented in this manuscript.
 }
@@
 This comparison is complicated due to the variety of parameters and inference methods used and referred to as the standard "Bayesian" decoder.
-For the sake of comparison, we used clusterless decoding in 20 ms time bins and used three common methods of inferring the trajectory: the radon transform \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, linear regression \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, and the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
-We then compared how the estimated trajectory speeds varied by dynamic for each method.
+We chose to use clusterless decoding in 20 ms time bins and three common methods of inferring the trajectory of the replay: the radon transform \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, linear regression \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, and the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+We then compared the estimated the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
 This reveals two key insights.
 First, the choice of method can give very different estimated speeds distributions for the SWRs.
-In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the SWRs proceed at faster speeds.
+In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C vs.
+\FIG{6}B,D).
+Second, the
 

### 2021-05-05 Article.tex
Update on Overleaf.

@@
 The latent position posterior summed over time is shown in the purple shading.
-The light grey lines represent the position of the animal over the entire epoch and the magenta dot represents the animal's position.
+The light grey lines represent the position of the animal over the entire recording session and the magenta dot represents the animal's position.
 \textbf{(B)} Same as in (A), but with clusterless decoding.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}}\label{figsupp:fig2sup1}
@@
 Color indicates time.
-Light grey lines indicate the animal's position over the entire epoch.
+Light grey lines indicate the animal's position over the entire recording session.
 Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}}\label{figsupp:fig2sup2}
@@
 Green line represents the decoded positions in the last 10 ms.
-Grey lines represent the position of the animal over the course of the epoch.}\label{videosupp:fig2supv1}
+Grey lines represent the position of the animal over the course of the recording session.}\label{videosupp:fig2supv1}
 
@@
 % Figure 3-Supplemental 2
-\figsupp[Comparison of percentage of SWRs classified on real vs.
-position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
-position shuffled data for two epochs from different animals.
-Red line represents the percent of SWR events classified in that epoch for real data.
+\figsupp[Percentage of SWRs classified decreases on position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
+position shuffled data for two recording sessions from different animals.
+Red line represents the percent of SWR events classified in that recording session for real data.
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
+Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
 
 % Figure 3-Supplemental 3
-\figsupp[Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
-run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
+\figsupp[Percentage of spatially incoherent SWRs increases for run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
 run-swapped, circularly shuffled position.
-The red line represents the percent of spatially coherent or incoherent SWRs in that epoch for actual data.
+The red line represents the percent of spatially coherent or incoherent SWRs in that recording session for actual data.
 The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
@@
 This, in contrast to our previous shuffle, should preserve correlation between spikes locally but destroy global spatial structure.
-Using this shuffle, we found that there was more shuffled spatially incoherent content compared to the real data and there was less spatially coherent content compared to the real data for two epochs (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
+Using this shuffle, we found that there was more shuffled spatially incoherent content compared to the real data and there was less spatially coherent content compared to the real data for two recording sessions (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
 
@@
 To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
-The shuffled decodes have much larger HPD regions than the real data (epoch \#1: 55 cm vs.
+The shuffled decodes have much larger HPD regions than the real data (recording session \#1: 55 cm vs.
 229 cm, p=0.02, real vs.
-shuffled; epoch \#2: 72 cm vs.
+shuffled; recording session \#2: 72 cm vs.
 231 cm, p=0.02, real vs.
@@
 % Figure 4-Supplemental 1
-\figsupp[Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data for two epochs from different animals.
+\figsupp[Average 95\% HPD size increases on the 95\% highest posterior density on position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data for two recording sessions from different animals.
 Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
+Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
 
 % Figure 4-Supplemental 2
-\figsupp[Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data for two epochs from different animals.
+\figsupp[Average 95\% HPD size increases on run-swapped, circularly shuffled position data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+run-swapped, circularly shuffled position data for two recording sessions from different animals.
 Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
+The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
+This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
 
@@
 We chose to use clusterless decoding in 20 ms time bins and three common methods of inferring the trajectory of the replay: the radon transform \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, linear regression \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, and the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
-We then compared the estimated the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
+We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
 This reveals two key insights.
 First, the choice of method can give very different estimated speeds distributions for the SWRs.
-In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C vs.
-\FIG{6}B,D).
-Second, the
+In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, has a mixture of both fast and slow speeds.
+Second, the speeds from the MAP method are quite similar to the speeds estimated by our decoder (\FIG{6}D).
+This is to be expected because the MAP method is imposing the least amount of smoothing out of the other methods (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
+Our decoder imposes more smoothing than the MAP estimate but much less than the line fitting methods and therefore is closer to the MAP estimates.
 
@@
 
-One possibility is that these events are supported by different subcircuits within the hippocampus.
-For example, \cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
+One possibility is that slow and fast dynamics are supported by different hippocampal subcircuits.
+\cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
 This raises the possibility that these slower replays are driven by CA3-CA1 interaction while the faster trajectories are more externally driven by MEC-CA1.
-Our
-
-showed that CA2 neurons tend to lead CA1 and CA3 neurons during SWRs.
-We did not have enough simultaneously recorded tetrodes in each area to attribute the different dynamics to particular subregions.
-However, we did show that our result holds even if restricting the dataset only to CA1 tetrodes.
 
@@
 Animals were free to choose any arm at any time.
-Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
+Only run recording sessions with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
 
@@
 
-\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics}
-In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
-We fit the clusterless encoding model on the resampled data and then decoded the immobility periods.
-Like with the non-resampled data, we then extracted the SWR times and determined their classification based on our classification scheme.
-We repeated this shuffle analysis 50 times and compared this distribution to the real data for two epochs on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
-
 \subsection*{Highest Posterior Density}
@@
 
+\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics and HPD Size}
+In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
+We fit the clusterless encoding model on the resampled data and then decoded the immobility periods.
+Like with the non-resampled data, we then extracted the SWR times and determined their classification based on our classification scheme.
+We repeated this shuffle analysis 50 times and compared this distribution to the real data for two recording sessions on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
+
+We also performed another shuffle that preserved more of the local correlation structure of spiking by shuffling the well-to-well runs and then circularly shuffling the position within the runs.
+A well-to-well run starts when the animal moves more than 5 cm from a well and ends when the animal moves back within 5 cm of a well.
+This mostly captures runs to separate wells (e.g.
+from the center well to the left well), but for a small number of well-to-well runs, the animal did turn back to the same well (e.g.
+went back to the center well after leaving the center well).
+As above, after permuting the order of the well-to-well runs and circularly shuffling the position, we fit the clusterless encoding model and decoded SWRs.
+We repeated this shuffle analysis 50 times and compared this distribution to the real data for two recording sessions on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
+
 \subsection*{Distance from animal}
@@
 
+\subsection*{Standard "Bayesian" Decoding}
+For the standard decoder analysis, we first fit a clusterless encoding model in the same manner as described in the subsection Encoding - Clusterless.
+We then decoded the posterior probability of each position, using 20 ms time bins, using the clusterless likelihood equation as described in the subsection The Model and assuming a uniform prior probability.
+Finally, to get an estimate of speed of the replay, we used either the radon transform method, the linear regression method, or the MAP estimate method.
+
+For the radon transform method, we densely sampled potential line trajectories and picked the line that maximized the probability along that line.
+The sum of the probability along the line is called the radon score.
+We picked the maximum radon score on lines fit on the combination of center arm and left arm posterior versus center arm and right arm posteriors.
+P-values were computed by circularly shuffling the posterior probabilities 1000 times and comparing the real data vs.
+the shuffled distribution.
+
+For the linear regression method, we draw 1000 position samples from each time bin according to the probability of position in that time bin.
+A linear regression is then fit to the sampled positions using time as a covariate.
+Fits were made separately for the posterior probabilities corresponding to the center arm with the left arm and the center arm with the right arm.
+Quality of the fit was assessed using the R-squared and the best fit between the arms was selected based on this metric.
+
+For the MAP estimate, we pick the maximum position bin for each time bin, which corresponds to the most probable position.
+
 \subsection*{Software and Code Availability}
@@
 Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
-Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
+Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn \citep{Waskomseabornstatisticaldata2021}.
 All code is open-source and licensed under the MIT Software License.

### 2021-05-05 Article.tex

@@
 Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
-Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn \citep{Waskomseabornstatisticaldata2021}.
+Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
 All code is open-source and licensed under the MIT Software License.

### 2021-05-05 Article.tex
Update on Overleaf.

--- 2021-05-05 Article.tex
@@
 
-\title{Hippocampal replay of experience at slow speeds}
+\title{Hippocampal replay of experience at real world speeds}
 
@@
 Representations of past and possible future experiences play a critical role in memory and decision-making processes.
-The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
+The rat hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
 Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.
@@
 Hippocampal "replay" events are well suited to that role \citep{CarrHippocampalreplayawake2011}.
-As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+As rats move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel).
-Red dotted line in the right panel is the best fit line from the standard decoder.
+Red dotted line in the right panel is the best fit line from the standard decoder using the Radon transform.
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
@@
 Subsequently, as the population activity becomes spatially incoherent, the fragmented movement dynamic dominates for the last 30 ms of the simulated event.
-This also shows that the fragmented dynamic gives us a way to test for spatially incoherent content directly for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder.
 
-Importantly, this approach allows the model to to capture a wide range of movement speeds for the latent position.
+This shows off two key features of the model.
+First, the fragmented dynamic gives us a way to directly test for spatially incoherent content for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder, because we have a model for spatial incoherent content.
+Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E).
 The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the movement of the underlying position representation.
@@
 % Figure 2-Supplemental 3
-\figsupp[Spatial coverage of the track for each animal.]{}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
+\figsupp[Spatial coverage of the track for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
 
@@
 
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
@@
 Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
-We also verified that the population spiking for all positions on the track was consistent for each animal \FIGSUPP[2]{fig2sup3}.
+We also verified that the population spiking for all positions on the track was consistent for each animal (\FIGSUPP[2]{fig2sup3}).
 
@@
 To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate.
+The 95\% HPD region corresponds to the most probable latent positions as estimated by the model.
 Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
@@
 Additionally, the HPD region for the unclassified SWRs was similarly large.
+
 We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in \FIG{4}E), indicating less spatial certainty for those specific events.
@@
 shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}).
+We also performed a different shuffle that preserved more of the local spatial structure while destroying global spatial structure by permuting the well-to-well runs and circularly shifting the position within the runs.
+This also resulted in larger HPD regions than the real data (recording session \#1: p=0.02, recording session \#2: p=0.02, \FIGSUPP[4]{fig4sup1}).
 
@@
 
-We examined the overall levels of neuronal spiking across dynamics.
+We wondered if our results could be driven by the overall levels of neuronal spiking across dynamics.
 We found that all the dynamics had similar average multiunit spiking rates (\FIG{5}F).
@@
 
-We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (although note, interneuron firing still may have relevant position information and it is not clear that they should be excluded \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
+We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (However it is important to note that interneuron firing still may have relevant position information and it is not clear that interneurons should be excluded \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
 First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
@@
 In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, has a mixture of both fast and slow speeds.
-Second, the speeds from the MAP method are quite similar to the speeds estimated by our decoder (\FIG{6}D).
+Second, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
 This is to be expected because the MAP method is imposing the least amount of smoothing out of the other methods (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.

### 2021-05-13 Article.tex

@@
 \begin{abstract}
-Representations of past and possible future experiences play a critical role in memory and decision-making processes.
+Representations related to past experiences play a critical role in memory and decision-making processes.
 The rat hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
@@
 Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content.
-Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
+Furthermore, many events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
 \end {abstract}
@@
 
-The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
-These "generative" \citep{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
-Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience.
+The brain has the remarkable ability to store and retrieve representations of past events, allowing memories of the past to influence the future.
+These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events (Cite Eichenbaum and Cohen?).
+Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits (REFs).
+Experiences can take place over many seconds, minutes, hours or days, and to be useful in guiding behavior, memories of these events must be stored in such a way that they can be retrieved quickly to inform upcoming choices.
+Consistent with this need for rapid storage, hippocampal replay events can express time compressed representations of past experience, activating regions outside the hippocampus and presumably driving plasticity in those regions to promote memory storage.
+\citep{CarrHippocampalreplayawake2011}.
 
-Hippocampal "replay" events are well suited to that role \citep{CarrHippocampalreplayawake2011}.
-As rats move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+In the context a rodent navigating through space preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience, and thus "replay" these previous experiences.
 Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
@@
 While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009, Tingleymethodsreactivationreplay2020}.
-One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
-Our subjective experience is that we can retrieve memories of locations without having to mentally traverse a long distance.
-If the rodent hippocampus is capable of that sort of retrieval, then one might expect to see SWRs where the neural activity corresponds not to a very rapid trajectory through space but instead to either a stable pattern associated with a single location or perhaps a pattern more similar to one seen during a real experience.
-Importantly, this sort of activity could still be useful, as it would take place during the short duration of an SWR (\texttildelow100 ms) and could thus support rapid retrieval and use of representations.
+One might therefore conclude that all such events are sequential in nature, and thus that memories that are stored or updated by SWRs are those corresponding sequences of locations.
+Our subjective experience is that we can retrieve memories of locations without having to mentally traverse a long distance, however.
+Such memories would seem to be useful, in that they could encode the stimuli and values associated with a given place irrespective of the path used to get there.
+If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a very rapid trajectory through space but instead to either a stable pattern associated with a single location or perhaps a brief pattern  more similar to one seen during a real experience.
 
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
-Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
+Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
 
@@
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
-A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent some definition of a "random" sequence.
+Finally, the most commonly used approaches for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
+A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent an implicit definition of a "random" sequence.
 
-While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test.
+While the standard approach identifies constant speed events, it does not consider events that are rejected by the statistical test.
 This is problematic because it has the potential to mischaracterize real events that do not move at constant speeds such as those that change direction or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
 Furthermore, the use of large, fixed-size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
-For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments between time steps and all spikes within the 20 ms bins are combined.
-The linear fit is also highly dependent on the estimation of event boundaries, such as the start and end times of a SWR, because the fit depends on all the data over the course of the event.
+For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments (1 or more bins in 20 ms) between time steps.
+The linear fit is also highly dependent on the estimation of event boundaries, such as the start and end times of the SWR, because the fit depends on all the data over the course of the event.
 Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
@@
 
-We therefore developed a state space model of generative activity that achieves all of those goals.
+We therefore developed a state space model that achieves all of those goals.
 State space models are a well-understood, well-known statistical solution to the problems described above.
@@
 We find that the large majority of SWRs contain spatially coherent content; that is, trajectories that are spatially concentrated at each moment in time and have no large discontinuities in position.
-Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
+Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 We simulate 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
-Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
+Each place field has a 36 cm variance and a 15 Hz peak firing rate, and fields are spaced every 10 cm along the virtual track.
 We then construct a 280 ms spiking sequence (\FIG{1}A) and apply our model to the sequence.
@@
 In the stationary movement dynamic, the latent position does not change between time steps.
-The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the resolution of the position bin).
+The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (at the resolution of the position bin).
 In the continuous movement dynamic, the latent position is most likely to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
@@
 We model this using a uniform state transition matrix, which makes transitions to each position equally likely.
+Importantly, the fragmented dynamic provides a well-specified definition of what we mean by random, or unstructured activity.
 
@@
 
-This shows off two key features of the model.
-First, the fragmented dynamic gives us a way to directly test for spatially incoherent content for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder, because we have a model for spatial incoherent content.
+This illustrates two key features of the model.
+First, as mentioned above, the fragmented dynamic gives us a way to directly test for spatially incoherent content for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder.
 Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E).
@@
 We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
-To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
+To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel).
+Here the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
 As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
-Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+Using our speed classification scheme, this means that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
 This is also evident in the posterior probability of latent linear position over time.
@@
 We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
-Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (\FIGSUPP[2]{fig2sup1}A).
+Importantly, when we apply the same model using the 2D position of the animal we get a similar result (\FIGSUPP[2]{fig2sup1}A).
 
@@
 Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
-Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019} and often requires inclusion decisions that can vary widely across experimenters.
 We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
@@
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
-We found, similar to other studies, that the average median difference is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
+We found, similar to other studies, that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
 Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
@@
 
-Having established the validity of our decoder in tracking the animal's position, we next wanted to assess the overall prevalence of spatial content across SWRs.
+Having established the validity of our decoder in tracking the animal's position, we next assessed the prevalence of spatial content across SWRs.
 We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
-As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (\FIGSUPP[2]{fig2sup2}A-F).
-However, we also observe many events with spatially coherent content that do not have this structure.
-For example, we observe trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
-We also observe SWRs that are spatially incoherent throughout the SWR (\FIGSUPP[3]{fig3sup1}A, D).
+As expected, and providing further support for the validity of the model, we observed replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using standard approaches (\FIGSUPP[2]{fig2sup2}A-F).
+However, we also observed many events with spatially coherent content that do not have this structure.
+For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
+We also observed SWRs where the content is spatially incoherent throughout(\FIGSUPP[3]{fig3sup1}A, D).
 
-Using a 0.8 threshold, we are able to classify 89\% (23,071 of 25,844) of SWRs as containing at least one of the three dynamics or dynamic mixtures.
-To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resample position with replacement, which shuffles the relationship between spiking and position, for two recording sessions, fit the encoding model and then decode the same SWR events containing the same spikes.
+Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
+To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resampled position with replacement, which shuffles the relationship between spiking and position, for two recording sessions.
+We then fit the encoding model and then decoded the same SWR events containing the same spikes.
 Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
-This shows that our model will not classify dynamics when there is no spatial information present in the data.
+This shows that our model does not identify dynamics when there is no spatial information present in the data.
 
@@
 We find that only 14\% (3,295 of 23,071) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
-To further validate this result, we performed a shuffle that randomized the order of the reward well to reward well runs and then circularly permuted the run.
-This, in contrast to our previous shuffle, should preserve correlation between spikes locally but destroy global spatial structure.
-Using this shuffle, we found that there was more shuffled spatially incoherent content compared to the real data and there was less spatially coherent content compared to the real data for two recording sessions (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
+To further validate this result, we performed a shuffle that randomized the order of the reward well to reward well runs and then circularly permuted the resulting data.
+This, in contrast to our previous shuffle, will preserve the local correlations between spikes but reduce global spatial structure.
+Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
 
-To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
+To more directly compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
 Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
@@
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
-Additionally, the HPD region for the unclassified SWRs was similarly large.
+The HPD region for the unclassified SWRs was similarly large.
 
@@
 shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}).
-We also performed a different shuffle that preserved more of the local spatial structure while destroying global spatial structure by permuting the well-to-well runs and circularly shifting the position within the runs.
+We also performed a different shuffle that preserved more of the local spatial structure while destroying global spatial structure by permuting the order of the well-to-well runs and circularly shifting the position within the runs.
 This also resulted in larger HPD regions than the real data (recording session \#1: p=0.02, recording session \#2: p=0.02, \FIGSUPP[4]{fig4sup1}).
@@
 \textbf{(A)} UpSet plot \citep{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
-Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times).
+Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at different times).
 The sets are ordered by how often they occur as indicated by the bar plot above each category.
@@
 \textbf{(F)} Average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
+\textbf{(G)} Probability density of animal movement speeds, illustrating prevalence of slower speed real-world movement consistent with stationary and stationary-continuous mixture replay events.
 }
@@
 
-The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs).
+The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C).
@@
 We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1,437 of 21,433 (\texttildelow7\%) of classified SWRs.
-46\% of these (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
+Forty-six percent of these (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
 These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
-
-We wondered if our results could be driven by the overall levels of neuronal spiking across dynamics.
-We found that all the dynamics had similar average multiunit spiking rates (\FIG{5}F).
-This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
 
@@
 \end{figure}
+\subsection*{Control analyses}
 
-We performed several controls to see if our results were primarily influenced by the firing of interneurons, which typically do not project to other brain areas and may not be spatially specific in their firing (However it is important to note that interneuron firing still may have relevant position information and it is not clear that interneurons should be excluded \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
-First, we checked if requiring five or more tetrodes to have spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high firing interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
-This had little effect on the overall proportions, with many stationary and stationary-continuous mixtures still present (\FIGSUPP[5]{fig5sup2}A).
-Indeed, most of our events included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}B), including within each dynamic (\FIGSUPP[5]{fig5sup3}A).
-Next, we tried removing spikes that had spike widths of less than 0.3 ms because interneurons typically have narrow spike widths.
-Note that this is a rather imprecise threshold, because typically this is computed on the average waveform for a putative cell, not on a single spike, and combined with a requirement for the putative cell to have a high firing rate, as interneurons typically exhibit high firing rates.
-Nevertheless, even with this imprecise criteria, we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
-We also examined the 9 of 10 animals in our dataset that already had clustered cells, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
-Although there were many fewer events to examine, we still observed many stationary and stationary-continuous mixtures in SWRs compared to continuous sequences (\FIGSUPP[5]{fig5sup2}D).
+We then carried out a series of control analyses to determine whether our results could be do to differences in activity across dynamics or interneuron-specific activity patterns.
+We first calculated the multiunit spiking rates and found that these were similar across the different dynamics.
+(\FIG{5}F).
+This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
+
+We also verified that our results were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
+First, we checked if requiring five or more tetrodes to have detected spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high rate interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
+We found that our results were robust to this increase in threshold (\FIGSUPP[5]{fig5sup2}A).
+Indeed, most of our events and most periods of dynamics within each event included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}B, \FIGSUPP[5]{fig5sup3}A).
+Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons (Fox and Ranck https://pubmed.ncbi.nlm.nih.gov/1183529/).
+This had little effect: we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
+We then repeated this analysis on clustered data for 9 of the 10 animals, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
+Although there were many fewer events to examine, we still observed many stationary and stationary-continuous mixtures in SWRs (\FIGSUPP[5]{fig5sup2}D).
 Finally, if interneurons were not spatially specific and driving the slower dynamics, we would expect to see the posterior be less spatially concentrated.
@@
 
-Another possible concern is that many of these slow spatially coherent events occur when the animal is moving very slowly, meaning the stationary-continuous mixtures potentially reflect a slow theta sweep rather than a replay event.
-We do not believe this to be the case because the majority of our events occur when the animal is moving at speeds < 2 cm/s (\FIGSUPP[5]{fig5sup3}B) and when we restrict ourselves to SWRs where the animal was moving at speeds < 1 cm/s, we find similar proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}C).
+We then asked whether the slow spatially coherent events (stationary-continuous mixtures) potentially reflect a slow theta sweep rather than a replay event.
+We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would only be seen during rapid theta sequences (median 52 cm \FIG{5}C).
+In addition, the majority of our events occur when the animal is moving at speeds < 2 cm/s, as expected for SWRs (\FIGSUPP[5]{fig5sup3}B).
+Further, when we restrict the analyses to SWRs where the animal was moving at speeds < 1 cm/s, we find similar proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}C).
 
@@
 There are several lines of evidence against this.
-First, as mentioned above, most of our SWRs had spikes from multiple tetrodes, including during stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell is driving these dynamics.
+First, as mentioned above, most of our SWRs had spikes from multiple tetrodes, including those with stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell is driving these dynamics.
 Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
-bursts which have a duration of 6-24 ms).
-Third, only ~50\% of the spikes had ISIs of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), so if bursting occurs, it only occurs on half of the SWRs we have analyzed.
-It should be noted that this metric in particular should be interpreted with extreme caution, because with clusterless decoding, the ISI of the multiunit spikes is not the same as if we had an individual unit.
-The spikes could be from multiple neurons firing, which would not tell us if there was bursting or not.
-Nevertheless, even with these caveats, it seems unlikely that burst firing from cells is driving the stationary dynamics.
+bursts which have a typical duration of 6-24 ms).
+Third, only ~50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode.
+.
 
 One last concern is that we have combined spiking information from several hippocampal subfields (CA1, CA2, and CA3), which may influence the types of dynamics we observed.
-However, CA1 and CA3 spiking is largely coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
+However, CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 (CITE DIBA , KArlsson).
+Finally, we also verified that when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
 Therefore, we feel our combination of spiking information from multiple subfields is appropriate.
 
-% Table 1
-\begin{wraptable}{l}{0.5\textwidth}{
-\begin{tabular}{lll}
-\textbf{Dynamic}                       & \textbf{\begin{tabular}[c]{@{}l@{}}Percent.\\ Sig.\end{tabular}} & \textbf{Number Sig.}               \\ \hline
-\multicolumn{1}{|l|}{Stat.}            & \multicolumn{1}{l|}{8\%}                                         & \multicolumn{1}{l|}{542 of 7176}   \\ \hline
-\multicolumn{1}{|l|}{Stat.-Cont.-Mix.} & \multicolumn{1}{l|}{14\%}                                        & \multicolumn{1}{l|}{2253 of 16106} \\ \hline
-\multicolumn{1}{|l|}{Cont.}            & \multicolumn{1}{l|}{30\%}                                        & \multicolumn{1}{l|}{1368 of 4511}  \\ \hline
-\multicolumn{1}{|l|}{Cont.-Frag.-Mix.} & \multicolumn{1}{l|}{31\%}                                        & \multicolumn{1}{l|}{799 of 2562}   \\ \hline
-\multicolumn{1}{|l|}{Frag.}            & \multicolumn{1}{l|}{15\%}                                        & \multicolumn{1}{l|}{244 of 1598}   \\ \hline
-\end{tabular}
-}
-  \caption{Number of significant SWRs that contain each dynamic. SWRs were fit using the random transform with 1000 shuffles of the posterior in each time bin.}\label{tab:supp1}
-\end{wraptable}
+\subsection*{Comparison to standard approaches}
+Finally, we asked how our method compares to the standard approaches to identifying significant events and characterizing their content.
+This comparison is complicated due to the variety of parameters and inference methods used previously, so here we chose three commonly used approaches and a reasonable parameter set for the comparison.
+We began with an examination of events based on a measure of statistical significance, which has been the standard in the field.
+We applied the clusterless decoder in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
+We then used the radon transform method \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019} to calculate, for each SWR, a measure of significance based on the assumptions that the probability density moves at a constant speed throughout the event.
+We found that ~12\% of the classified events (using the full state-space model) were identified as significant.
+Not surprisingly, when events were classified based on which dynamic they contained, we found substantial variability: of 7176 SWRs that contained a period of stationary dynamics, 542 (8\%) were significant while of 4511 SWRs that contained a period of continuous dynamics, 1386 (~30\%) were significant.
+And when we restricted the analysis to stationary events that were >100 ms long, we found that 28\% were significant, and increase that is not surprising given that the shuffle used for the radon method penalizes shorter events.
+We also note that the linear regression method \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, detects only non-zero slopes as significant, and thus cannot identify stationary events as significant.
+The substantial variability in the proportions of "significant" events indicates that conclusions about which events are meaningful to the brain could be very different depending on the specific method and parameters used.
 
-Finally, an important question is how our method compares to the standard "Bayesian" decoder.
-This comparison is complicated due to the variety of parameters and inference methods used and referred to as the standard "Bayesian" decoder.
-We chose to use clusterless decoding in 20 ms time bins and three common methods of inferring the trajectory of the replay: the radon transform \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, linear regression \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, and the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+Conclusions about the representational content of events were similarly variable across methods.
+In addition to the radon and linear regression methods, we also computed  the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
-This reveals two key insights.
-First, the choice of method can give very different estimated speeds distributions for the SWRs.
-In particular, the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, has a mixture of both fast and slow speeds.
-Second, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
+We found that the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, has a mixture of both fast and slow speeds.
+In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
 This is to be expected because the MAP method is imposing the least amount of smoothing out of the other methods (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
 Our decoder imposes more smoothing than the MAP estimate but much less than the line fitting methods and therefore is closer to the MAP estimates.
+These results indicate that the methods that made more assumptions about the structure of replay (radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
 
@@
 This model is robust: it is relatively insensitive to a range of plausible transition parameters between dynamics and instead strongly reflects the hippocampal place representation structure.
-We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been previously captured.
+We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far more than have previously associated with identifiable activity patterns.
 This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
-These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
+These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical approaches that assume that event progress at constant speed.
 The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and closer to real-world speeds.
@@
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
-Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
+Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary and continuous dynamics.
 These events were mostly decoded representations where the most probable positions were highly spatially concentrated (i.e.
@@
 These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
-Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+Instead, there is a much richer variety of speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
 Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
-Here, we show that they do not disappear, but are a very common occurrence with adult rats.
+Our results strongly suggest that they do not disappear, but are a very common occurrence with adult rats.
 We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events and using a restrictive decoding model which requires the decoded position to be stationary for the entire event.
@@
 
-We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of experience.
-Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
-Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to store and evaluate the value of specific locations or short movement trajectories without having to generate full trajectories \citep{YuHippocampalcorticalinteraction2015}.
+The differences between our finding and previous approaches also highlight the importance of identifying the specific question that an analysis seeks to answer.
+Initial studies of replay (CITE LEE Wilson, Foster Wilson, Davidson, Karlsson) focused on the question of whether the observed patterns of spiking seen during replay reflected structured content or "random" spiking as defined by a comparison to one or more shuffles.
+These and subsequent studies answered the question "are replay events the result of random spiking" with a clear "no".
+These studies did not, however, establish that the significance criteria used to show non-random structure has anything to do with the presence of spatial structure or events that are potentially useful for memory processes
 
-One possibility is that slow and fast dynamics are supported by different hippocampal subcircuits.
-\cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
-This raises the possibility that these slower replays are driven by CA3-CA1 interaction while the faster trajectories are more externally driven by MEC-CA1.
+Our method aims to identify structure within events and to capture their complex dynamics to determine what sort of information replay events could provide in service of memory functions.
+We define precisely patterns of spatial representation that are, and are not, interpretable in terms of the structure of a given environment and the spatial activity seen during movement.
+In so doing, we find that most, but not all, events contain spiking that corresponds to a location or a trajectory through the animal's environment.
 
+The different questions addressed by our method compare to previous methods can also explain the apparent differences in findings.
 Our results showed that 19\% of SWRs contained continuous trajectories---meaning that portions of the SWR had spatially coherent trajectories that moved at high velocities---but only 4\% were continuous throughout the entire SWR.
@@
 Furthermore, these studies found trajectories on average move at much higher speeds (20x normal rat movement speed) than we report here.
-While the plurality of methods and parameters makes it hard to directly compare, we believe there are several reasons for this difference.
-First, our model is much more flexible because it allows the trajectory to change speed during the SWR, whereas fitting a line assumes a constant speed trajectory for the entire SWR.
+These previous findings were based largely on linear fits to the data, while our model allows the trajectory to change speed during the SWR.
 Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics (see for example \FIG{1}).
@@
 
+We hypothesize that this large set of events supports the ability to store and update  representations of individual places and snippets of experience.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to store or update the value of specific locations or short movement trajectories without having to generate full trajectories representing one of multiple ways to arrive at a location \citep{YuHippocampalcorticalinteraction2015}.
+Events that contain multiple dynamics (e.g.
+two continuous representations separated by a brief fragmented representation) might also help bind together experiences that occur in the same environment but at different places.
+Alternatively, it might be that downstream structures are attuned to these dynamics, and that each period of coherent hippocamapal dynamics drives its own set of responses elsewhere in the brain.
+
+Simiarly, it is possible that slow and fast dynamics are supported by different subcircuits.
+\cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
+This raises the possibility that these slower replays are driven by CA3-CA1 interaction while the faster trajectories are more externally driven by MEC-CA1, and that these different events could also engage different circuits outside of the hippocampus and entorhinal cortex.
+
 Additionally, as the start and end time of each SWR event are also detection method- and parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
@@
 More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
-Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \citep{ChengNewExperiencesEnhance2008}.
+Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \citep{YuDistincthippocampalcorticalmemory2017}.
 High thresholds can therefore exclude large numbers of real events and potentially lead to erroneous conclusions.
@@
 Because the model gives a moment-by-moment estimate of the position and dynamic, this also reduces the effect of misidentifying the SWR start and end times.
-Finally, combined with an acausal estimation of latent position, this approach allows us to identify the specific timing of transitions in dynamics at a fine timescale.
+Combined with an acausal estimation of latent position, this approach allows us to identify the specific timing of transitions in dynamics at a fine timescale.
+This approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
 
@@
 
-We also performed another shuffle that preserved more of the local correlation structure of spiking by shuffling the well-to-well runs and then circularly shuffling the position within the runs.
+We also performed another shuffle that preserved more of the local correlation structure of spiking by shuffling the order of the well-to-well runs and then circularly shuffling the position within the runs.
 A well-to-well run starts when the animal moves more than 5 cm from a well and ends when the animal moves back within 5 cm of a well.
-This mostly captures runs to separate wells (e.g.
+This largely reflects runs from one well to another (e.g.
 from the center well to the left well), but for a small number of well-to-well runs, the animal did turn back to the same well (e.g.
@@
 
-\subsection*{Standard "Bayesian" Decoding}
+\subsection*{Standard "Bayesian" Decoding and Significance Estimation}
 For the standard decoder analysis, we first fit a clusterless encoding model in the same manner as described in the subsection Encoding - Clusterless.
@@
 
-For the linear regression method, we draw 1000 position samples from each time bin according to the probability of position in that time bin.
-A linear regression is then fit to the sampled positions using time as a covariate.
+For the linear regression method, we drew 1000 position samples from each time bin according to the probability of position in that time bin.
+A linear regression was then fit to the sampled positions using time as a covariate.
 Fits were made separately for the posterior probabilities corresponding to the center arm with the left arm and the center arm with the right arm.
@@
 
-For the MAP estimate, we pick the maximum position bin for each time bin, which corresponds to the most probable position.
+For the MAP estimate, we pick the position bin with the maximum value for each time bin, which corresponds to the most probable position.
 

### 2021-05-13 Article.tex

@@
 The brain has the remarkable ability to store and retrieve representations of past events, allowing memories of the past to influence the future.
-These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events (Cite Eichenbaum and Cohen?).
-Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits (REFs).
+These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumCanWeReconcile2014}.
+Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Joohippocampalsharpwave2018}(REFs).
 Experiences can take place over many seconds, minutes, hours or days, and to be useful in guiding behavior, memories of these events must be stored in such a way that they can be retrieved quickly to inform upcoming choices.
@@
 We find that spatially coherent events indeed have smaller HPD regions than events with fragmented dynamics.
-\FIG{4}A and \FIG{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
+\FIG{4}A and \FIG{4}B show two example SWRs that are classified as having continuous and stationary dynamics, respectively.
 The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD region is small throughout the SWRs.
@@
 The HPD region size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in \FIG{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+An example of this change is shown in \FIG{4}D, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
 Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (\FIG{4}E).
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v7}
 \caption{Validation of classification using the 95\% Highest Posterior Density.
@@
 
-We also verified that our results were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007}).
+We also verified that our results were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007}).
 First, we checked if requiring five or more tetrodes to have detected spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high rate interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
@@
 Indeed, most of our events and most periods of dynamics within each event included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}B, \FIGSUPP[5]{fig5sup3}A).
-Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons (Fox and Ranck https://pubmed.ncbi.nlm.nih.gov/1183529/).
+Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons \citep{FoxLocalizationanatomicalidentification1975}.
 This had little effect: we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
@@
 Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
-bursts which have a typical duration of 6-24 ms).
+bursts which have a typical duration of 6-24 ms, \citep{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
 Third, only ~50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode.
@@
 One last concern is that we have combined spiking information from several hippocampal subfields (CA1, CA2, and CA3), which may influence the types of dynamics we observed.
-However, CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 (CITE DIBA , KArlsson).
+However, CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 \citep{DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
 Finally, we also verified that when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
@@
 Alternatively, it might be that downstream structures are attuned to these dynamics, and that each period of coherent hippocamapal dynamics drives its own set of responses elsewhere in the brain.
+We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well (CITE Pfeiffer paper), suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
 

### 2021-05-13 Article.tex
Update on Overleaf.

@@
 These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumCanWeReconcile2014}.
-Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Joohippocampalsharpwave2018}(REFs).
+Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018}.
 Experiences can take place over many seconds, minutes, hours or days, and to be useful in guiding behavior, memories of these events must be stored in such a way that they can be retrieved quickly to inform upcoming choices.
-Consistent with this need for rapid storage, hippocampal replay events can express time compressed representations of past experience, activating regions outside the hippocampus and presumably driving plasticity in those regions to promote memory storage.
-\citep{CarrHippocampalreplayawake2011}.
+Consistent with this need for rapid storage, hippocampal replay events can express time compressed representations of past experience, activating regions outside the hippocampus and presumably driving plasticity in those regions to promote memory storage \citep{CarrHippocampalreplayawake2011}.
 
@@
 
+We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
+To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel).
+Here the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
+Using our speed classification scheme, this means that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+This is also evident in the posterior probability of latent linear position over time.
+This shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
+We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
+Importantly, when we apply the same model using the 2D position of the animal we get a similar result (\FIGSUPP[2]{fig2sup1}A).
+
+One of our criteria for a more optimal method is for it to use all of the available spiking data.
+Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019} and often requires inclusion decisions that can vary widely across experimenters.
+We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
+Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
+Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
+As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
+This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (\FIG{2}D vs C).
+
+\subsection*{Most replays contain coherent spatial content}
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
+We found, similar to other studies, that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
+Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
+We also verified that the population spiking for all positions on the track was consistent for each animal (\FIGSUPP[2]{fig2sup3}).
+
+Having established the validity of our decoder in tracking the animal's position, we next assessed the prevalence of spatial content across SWRs.
+We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
+As expected, and providing further support for the validity of the model, we observed replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using standard approaches (\FIGSUPP[2]{fig2sup2}A-F).
+However, we also observed many events with spatially coherent content that do not have this structure.
+For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
+We also observed SWRs where the content is spatially incoherent throughout(\FIGSUPP[3]{fig3sup1}A, D).
+
+Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
+To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resampled position with replacement, which shuffles the relationship between spiking and position, for two recording sessions.
+We then fit the encoding model and then decoded the same SWR events containing the same spikes.
+Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
+This shows that our model does not identify dynamics when there is no spatial information present in the data.
+
 % Figure 3
@@
 \end{figure}
-
-We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
-To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel).
-Here the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
-As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
-Using our speed classification scheme, this means that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
-This is also evident in the posterior probability of latent linear position over time.
-This shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
-We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
-Importantly, when we apply the same model using the 2D position of the animal we get a similar result (\FIGSUPP[2]{fig2sup1}A).
-
-One of our criteria for a more optimal method is for it to use all of the available spiking data.
-Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
-Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019} and often requires inclusion decisions that can vary widely across experimenters.
-We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
-Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
-Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
-As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
-This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (\FIG{2}D vs C).
-
-\subsection*{Most replays contain coherent spatial content}
-
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
-We found, similar to other studies, that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
-Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
-We also verified that the population spiking for all positions on the track was consistent for each animal (\FIGSUPP[2]{fig2sup3}).
-
-Having established the validity of our decoder in tracking the animal's position, we next assessed the prevalence of spatial content across SWRs.
-We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
-As expected, and providing further support for the validity of the model, we observed replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using standard approaches (\FIGSUPP[2]{fig2sup2}A-F).
-However, we also observed many events with spatially coherent content that do not have this structure.
-For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
-We also observed SWRs where the content is spatially incoherent throughout(\FIGSUPP[3]{fig3sup1}A, D).
-
-Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
-To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resampled position with replacement, which shuffles the relationship between spiking and position, for two recording sessions.
-We then fit the encoding model and then decoded the same SWR events containing the same spikes.
-Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
-This shows that our model does not identify dynamics when there is no spatial information present in the data.
 

### 2021-05-13 Article.tex
Update on Overleaf.

@@
 Alternatively, it might be that downstream structures are attuned to these dynamics, and that each period of coherent hippocamapal dynamics drives its own set of responses elsewhere in the brain.
-We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well (CITE Pfeiffer paper), suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
+We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020}, suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
 
-Simiarly, it is possible that slow and fast dynamics are supported by different subcircuits.
+Similarly, it is possible that slow and fast dynamics are supported by different subcircuits.
 \cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).

### 2021-05-13 Article.tex

@@
 
-In the context a rodent navigating through space preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+In the context of rodents navigating through space preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approaches for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent an implicit definition of a "random" sequence.
@@
 This, in contrast to our previous shuffle, will preserve the local correlations between spikes but reduce global spatial structure.
-Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup3}).
+Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (recording session \#1, spatially coherent: 99\% vs.
+71\%, p=0.02, real vs.
+shuffled; recording session \#1, spatially incoherent: 22\% vs.
+61\%, p=0.02, real vs.
+shuffled; recording session \#2, spatially coherent: 95\% vs.
+63\%, p=0.02, real vs.
+shuffled; recording session \#2, spatially incoherent: 16\% vs.
+79\%, p=0.02, real vs.
+shuffled; \FIGSUPP[3]{fig3sup3}).
 
@@
 In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification.
-To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+To show this, we used the same shuffle as above, resampling position with replacement for two recording sessions and shuffling the relationship between spiking and spatial information and then fitting the encoding model.
 The shuffled decodes have much larger HPD regions than the real data (recording session \#1: 55 cm vs.
@@
 231 cm, p=0.02, real vs.
-shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}).
-We also performed a different shuffle that preserved more of the local spatial structure while destroying global spatial structure by permuting the order of the well-to-well runs and circularly shifting the position within the runs.
-This also resulted in larger HPD regions than the real data (recording session \#1: p=0.02, recording session \#2: p=0.02, \FIGSUPP[4]{fig4sup1}).
+shuffled, one-sided; \FIGSUPP[4]{fig4sup1}).
+We also compared the HPD regions between the real data and the shuffle that preserved more of the local spatial structure while destroying global spatial structure.
+This also resulted in larger HPD regions than the real data for spatially coherent and incoherent SWRs (recording session \#1, spatially coherent: 29 cm vs.
+52 cm, p=0.02, real vs.
+shuffled; recording session \#1, spatially incoherent: 181 cm vs.
+243 cm, p=0.02, real vs.
+shuffled; recording session \#2, spatially coherent: 26 cm vs.
+53 cm, p=0.02, real vs.
+shuffled; recording session \#2, spatially incoherent: 210 cm vs.
+263 cm, p=0.02, real vs.
+shuffled; \FIGSUPP[4]{fig4sup2}).
 
@@
 
-% Figure 6
-\begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_v2}
-\caption{
-Comparison of the state space decoder to the standard "Bayesian" decoder.
-For each panel, the top five rows are the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
-Note that these categories are not mutually exclusive since a SWR can have more than one dynamic or combination of dynamics.
-The sixth row is all SWRs containing unclassified dynamics.
-The final row is the estimated speed for all SWRs.
-\textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder presented in this manuscript.
-}
-\label{fig:6}
-
-\end{figure}
 \subsection*{Control analyses}
@@
 We then asked whether the slow spatially coherent events (stationary-continuous mixtures) potentially reflect a slow theta sweep rather than a replay event.
-We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would only be seen during rapid theta sequences (median 52 cm \FIG{5}C).
+We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would only be seen during rapid theta sequences (median 50 cm \FIG{5}C).
 In addition, the majority of our events occur when the animal is moving at speeds < 2 cm/s, as expected for SWRs (\FIGSUPP[5]{fig5sup3}B).
@@
 \subsection*{Comparison to standard approaches}
+
+% Figure 6
+\begin{figure}
+\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_v2}
+\caption{
+Comparison of the state space decoder to the standard "Bayesian" decoder.
+For each panel, the top five rows are the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
+Note that these categories are not mutually exclusive since a SWR can have more than one dynamic or combination of dynamics.
+The sixth row is all SWRs containing unclassified dynamics.
+The final row is the estimated speed for all SWRs.
+\textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder presented in this manuscript.
+}
+\label{fig:6}
+\end{figure}
+
 Finally, we asked how our method compares to the standard approaches to identifying significant events and characterizing their content.
@@
 Alternatively, it might be that downstream structures are attuned to these dynamics, and that each period of coherent hippocamapal dynamics drives its own set of responses elsewhere in the brain.
-We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020}, suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
+We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020a}, suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
 

### 2021-05-13 Article.tex
Update on Overleaf.

@@
 The brain has the remarkable ability to store and retrieve representations of past events, allowing memories of the past to influence the future.
-These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumCanWeReconcile2014}.
+These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{Eichenbaumconditioningconsciousrecollection2004}.
 Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018}.
@@
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009, TingleyRoutingHippocampalRipples2020, BhattaraiDistincteffectsreward2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DrieuNestedsequenceshippocampal2018, OlafsdottirTaskDemandsPredict2017, TangHippocampalPrefrontalReactivationLearning2017, CarrTransientSlowGamma2012, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent an implicit definition of a "random" sequence.
@@
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{}.
 For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.

### 2021-05-13 Article.tex

@@
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{}.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{KaeferReplayBehavioralSequences2020, CareyRewardrevaluationbiases2019, GrosmarkDiversityneuralfiring2016, WuHippocampalReplayCaptures2014, PfeifferHippocampalplacecellsequences2013}.
 For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.

### 2021-05-27 Article.tex
Update on Overleaf.

@@
 The brain has the remarkable ability to store and retrieve representations of past events, allowing memories of the past to influence the future.
-These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{Eichenbaumconditioningconsciousrecollection2004}.
-Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking contributes to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018}.
-Experiences can take place over many seconds, minutes, hours or days, and to be useful in guiding behavior, memories of these events must be stored in such a way that they can be retrieved quickly to inform upcoming choices.
-Consistent with this need for rapid storage, hippocampal replay events can express time compressed representations of past experience, activating regions outside the hippocampus and presumably driving plasticity in those regions to promote memory storage \citep{CarrHippocampalreplayawake2011}.
-
-In the context of rodents navigating through space preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumConditioningConsciousRecollection2004}.
+Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking is thought to contribute to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018, CarrHippocampalreplayawake2011}.
+
+The canonical example of "replay" is in the rodent hippocampus, where cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous experience, and thus "replay" these previous experiences.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience, and thus are thought to "replay" these previous experiences.
 Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
@@
 While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009, Tingleymethodsreactivationreplay2020}.
-One might therefore conclude that all such events are sequential in nature, and thus that memories that are stored or updated by SWRs are those corresponding sequences of locations.
-Our subjective experience is that we can retrieve memories of locations without having to mentally traverse a long distance, however.
+One might therefore conclude that all memories that are stored or updated by SWRs correspond to sequences of locations.
+Our subjective experience, however, suggests that we can retrieve memories of individual locations without having to mentally traverse a long distance.
 Such memories would seem to be useful, in that they could encode the stimuli and values associated with a given place irrespective of the path used to get there.
-If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a very rapid trajectory through space but instead to either a stable pattern associated with a single location or perhaps a brief pattern  more similar to one seen during a real experience.
+If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a rapid trajectory through space, but instead to either a stable pattern associated with a single location or perhaps a brief pattern  more similar to one seen during a real experience.
 
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
 Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
@@
 While the standard approach identifies constant speed events, it does not consider events that are rejected by the statistical test.
-This is problematic because it has the potential to mischaracterize real events that do not move at constant speeds such as those that change direction or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
+This is problematic because it has the potential to mischaracterize real events that do not move at constant speeds, such as those that change direction or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
 Furthermore, the use of large, fixed-size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
@@
 Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{KaeferReplayBehavioralSequences2020, CareyRewardrevaluationbiases2019, GrosmarkDiversityneuralfiring2016, WuHippocampalReplayCaptures2014, PfeifferHippocampalplacecellsequences2013}.
-For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
+For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found awake replay events containing alternation between representations of a single location and sequential spatial trajectories.
 On the other hand, \cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
@@
 
-An ideal approach to identifying and quantifying the dynamics of generative representations would circumvent these problems.
+An ideal approach to identifying and quantifying the dynamics of replay would circumvent these problems.
 It would use all of the available spiking data to yield the most accurate decoded positions.
 It would provide a moment-by-moment estimate of position and dynamics that would not be dependent on the estimation of SWR start and end times and could rapidly adjust to changes in dynamics.
-It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
+It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and would provide information about the certainty of the decoded estimates.
 It would be able to capture a range of movement dynamics from stationary or unchanging representations to trajectories that progress through space at constant or variable speeds to disorganized, spatially incoherent events.
 It would provide a robust statistical assessment of confidence for each dynamic.
-Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
+Finally, where assumptions are made, it would provide well-defined parameters whose values could be explored systematically to understand their influence on the results.
 
@@
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing the spatial representations during SWRs as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
-The shaded colors indicate the category of the speed of the trajectory at that time, which is determined from the probability.
+The shaded colors indicate the category of the speed of the trajectory at that time (Classification: Stat.
+= Stationary, S-C-M = Stationary-Continuous-Mixture, Cont.
+= Continuous, F-C-M = Fragmented-Continuous-Mixture, Frag.
+= Fragmented), which is determined from the probability.
 Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel).
@@
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the standard decoder, incorporate information about times when there is no spiking, and make fewer assumptions about the amount of time a downstream neuron can integrate spikes from multiple neurons (because place estimates from spikes within a single bin are combined).
-
-Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (\FIG{1}C).
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder, and incorporate information about times when there is no spiking.
+Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.
+
+Next, we specify movement dynamic models that describe a variety of ways that the latent position---the "mental" position of the animal represented by the cells--- could evolve over time.
+(\FIG{1}C).
 We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
 Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
-We therefore define state transition models to capture each of these movement dynamics.
+We therefore define movement dynamic models to capture each of these possibilities.
 
 In the stationary movement dynamic, the latent position does not change between time steps.
-The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (at the resolution of the position bin).
+The state transition matrix can thus be defined as an identity matrix, which predicts that the next position will be the same as the last position (at the resolution of the position bin).
 In the continuous movement dynamic, the latent position is most likely to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
@@
 
-Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via another state transition matrix between the movement dynamics.
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic via another state transition matrix between the movement dynamics (\FIG{1}D).
 In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms on average, which is approximately the duration of a SWR event.
@@
 Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution).
-Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
+Importantly, the data drives the estimated dynamic, so even if the probability of staying in a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within a time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
 In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
@@
 This illustrates two key features of the model.
-First, as mentioned above, the fragmented dynamic gives us a way to directly test for spatially incoherent content for each time, rather than by the use of non-parametric shuffle, as is commonly used with the standard decoder.
+First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle.
 Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E).
-The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the movement of the underlying position representation.
-To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 m/s.
+The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation.
+To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 cm/s.
 From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
@@
 Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates.
-In summary, we can characterize the speed or set of speeds in a SWR based on the probability of each of the three movement dynamics over time.
+In summary, we can characterize the speed or set of speeds that occur within a SWR based on the probability of each of the three movement dynamics over time.
 We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.
 
-We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
+We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
 Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
@@
 To investigate this we decoded the spiking sequence in \FIG{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (\FIGSUPP[1]{fig1sup1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
-Furthermore, the most probable latent positions remain relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993, corresponding to an expected duration between 25 and 150 ms), the model identified the movement dynamics consistently, with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
+Furthermore, the most probable latent positions remained relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
 
@@
 The middle plot shows the probability of each dynamic over time from the 2D decode.
-Shaded regions indicate classification category as in \FIG{1}G and \FIG{2}.
+Shaded regions indicate classification category as in \FIG{1}F and \FIG{2}.
 The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
@@
 % Figure 2-Supplemental 3
-\figsupp[Spatial coverage of the track for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
+\figsupp[Spatial coverage of the track for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.
+Each grey line represents a recording session for that animal.
+}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
 
@@
 Here the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
-As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
-Using our speed classification scheme, this means that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
-This is also evident in the posterior probability of latent linear position over time.
-This shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
+As expected, we observed that the probability of the continuous dynamic is high throughout this event, but the probability of the stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
+Using our speed classification scheme, this indicates that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+This is also evident in the posterior probability of latent linear position over time, which shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
 We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
@@
 After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
-We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal (with the caveat that theta sequences and replay sequences will represent positions other than where the animal is and our resolution is limited by the size of our 3 cm position bins).
-We found, similar to other studies, that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
-Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during run, demonstrating the power of our algorithm.
-We also verified that the population spiking for all positions on the track was consistent for each animal (\FIGSUPP[2]{fig2sup3}).
+We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
+Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
+This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies [cite].
+Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during movement, demonstrating the power of our algorithm.
+For each animal, we also verified that the population spiking rate was relatively consistent over all positions on the track, suggesting that we have comparable sampling of all locations in the environment (\FIGSUPP[2]{fig2sup3}).
 
 Having established the validity of our decoder in tracking the animal's position, we next assessed the prevalence of spatial content across SWRs.
-We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
+We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large-amplitude events as well as the much more common small-amplitude events.
 As expected, and providing further support for the validity of the model, we observed replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using standard approaches (\FIGSUPP[2]{fig2sup2}A-F).
@@
 For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
-We also observed SWRs where the content is spatially incoherent throughout(\FIGSUPP[3]{fig3sup1}A, D).
+We also observed SWRs where the content is spatially incoherent throughout (\FIGSUPP[3]{fig3sup1}A, D).
 
 Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
-To ensure that this depends on the firing of spatially tuned cells in the hippocampus, we resampled position with replacement, which shuffles the relationship between spiking and position, for two recording sessions.
-We then fit the encoding model and then decoded the same SWR events containing the same spikes.
+To ensure that this reflects the spatially tuned firing of the neurons, we trained the encoding model with positions resampled with replacement, a shuffling procedure which disrupts the relationship between spiking and position, for two recording sessions.
+We then decoded the same SWR events, containing the original spikes.
 Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
-This shows that our model does not identify dynamics when there is no spatial information present in the data.
+This shows that our model does not impose movement dynamics in the absence of spatial information in the data.
 
@@
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
-We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
+We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
 We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We find that only 14\% (3,295 of 23,071) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
-To further validate this result, we performed a shuffle that randomized the order of the reward well to reward well runs and then circularly permuted the resulting data.
-This, in contrast to our previous shuffle, will preserve the local correlations between spikes but reduce global spatial structure.
+We find that only 14\% (3,295 of 23,071) of classified SWRs include any times with spatially incoherent structure (\FIG{3}G).
+To further validate this result, we performed a shuffling procedure that, in contrast to our previous shuffle, preserves the local correlations between spikes but reduces global spatial structure.
+This controls for dynamics that may have been induced by bursts or other local features of the spike train while randomizing the position to spike relationship.
+To do this, we randomized the order of runs (from one reward well to another) and then circularly permuted the resulting segments of data across all tetrodes uniformly.
 Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (recording session \#1, spatially coherent: 99\% vs.
 71\%, p=0.02, real vs.
-shuffled; recording session \#1, spatially incoherent: 22\% vs.
+shuffled; spatially incoherent: 22\% vs.
 61\%, p=0.02, real vs.
@@
 63\%, p=0.02, real vs.
-shuffled; recording session \#2, spatially incoherent: 16\% vs.
+shuffled; spatially incoherent: 16\% vs.
 79\%, p=0.02, real vs.
@@
 To more directly compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
-Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
-Note that here we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by the standard decoder in the case of fitting a line, because our decoder is much more flexible in terms of allowing different speeds during the SWR event.
+Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs included time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all detected SWRs).
+Thus, focusing on only high-speed, linear-fit trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
+Note that here we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by line-fitting approach of the standard decoder, to take advantage of the ability of our decoder to capture a range different speeds within each SWR event.
 
 We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) had spatially coherent content.
-Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
+We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content.
+Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
 
@@
 In contrast, smaller values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD region provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
+Thus, the HPD region provides a complementary measure of spatial information, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
 
@@
 In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification.
-To show this, we used the same shuffle as above, resampling position with replacement for two recording sessions and shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+To show this, we used the same shuffle as above, resampling position with replacement for two recording sessions and shuffling the relationship between spiking and spatial information when fitting the encoding model.
 The shuffled decodes have much larger HPD regions than the real data (recording session \#1: 55 cm vs.
@@
 shuffled, one-sided; \FIGSUPP[4]{fig4sup1}).
-We also compared the HPD regions between the real data and the shuffle that preserved more of the local spatial structure while destroying global spatial structure.
+We also compared the HPD regions between the real data and the shuffle that preserved more of the local temporal structure while destroying global spatial structure.
 This also resulted in larger HPD regions than the real data for spatially coherent and incoherent SWRs (recording session \#1, spatially coherent: 29 cm vs.
 52 cm, p=0.02, real vs.
-shuffled; recording session \#1, spatially incoherent: 181 cm vs.
+shuffled; spatially incoherent: 181 cm vs.
 243 cm, p=0.02, real vs.
@@
 53 cm, p=0.02, real vs.
-shuffled; recording session \#2, spatially incoherent: 210 cm vs.
+shuffled; spatially incoherent: 210 cm vs.
 263 cm, p=0.02, real vs.
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (\FIG{5}A).
-Surprisingly, the most common category is stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
+We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) were classified as having only a single dynamic or dynamic mixture (\FIG{5}A, for specific example SWR see \FIG{3}B), whereas SWRs with multiple dynamics (such as those in \FIG{3}A, C, D, E, and F) were less common.
+Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
 These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
 continuous 94 ms, \FIG{5}B).
-Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 52 cm vs.
+Nonetheless, both the slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 52 cm vs.
 continuous 43 cm, \FIG{5}C).
 This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
-We note that roughly a third of these events persisted for the entire duration of the SWR, but the other two thirds had some portion of time where the model was uncertain about the dynamic.
-This shows the power of our classifier because it allows us to characterize the content of a SWR even if the model is only certain about a portion of the SWR time.
+We note that roughly a third of these events persisted for the entire duration of the SWR, but the other two thirds included some time where the model was uncertain about the dynamic.
+This shows the power of our classifier because it allows us identify periods of time when the model has high confidence with each SWR, and use these periods to characterize the SWR overall or to focus on specifically for further analysis.
 
@@
 We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1,437 of 21,433 (\texttildelow7\%) of classified SWRs.
-Forty-six percent of these (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
+Of these, 46\% (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
 These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
@@
 (\FIG{5}F).
-This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
-
-We also verified that our results were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007}).
-First, we checked if requiring five or more tetrodes to have detected spikes during the SWR (up from requiring two tetrodes be active) would influence the proportions of dynamics, with the thought that a single tetrode with a high rate interneuron might result in stationary or stationary-continuous mixture dynamics during SWRs.
+This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
+
+We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007}).
+First, we increased our tetrode participation threshold from two up to five tetrodes active during the SWR, to eliminate the possibility that events with a single tetrode with a high rate interneuron might drive the classification of stationary or stationary-continuous mixture dynamics during SWRs.
 We found that our results were robust to this increase in threshold (\FIGSUPP[5]{fig5sup2}A).
@@
 Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons \citep{FoxLocalizationanatomicalidentification1975}.
-This had little effect: we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
-We then repeated this analysis on clustered data for 9 of the 10 animals, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
+Similarly, this had little effect: we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
+Using clustered data, we can more robustly categorize units as putative interneurons; thus, we then repeated this analysis on clustered data for 9 of the 10 animals, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
 Although there were many fewer events to examine, we still observed many stationary and stationary-continuous mixtures in SWRs (\FIGSUPP[5]{fig5sup2}D).
-Finally, if interneurons were not spatially specific and driving the slower dynamics, we would expect to see the posterior be less spatially concentrated.
-Instead, we found that these dynamics tended to have narrow posteriors (as characterized by 95\% HPD region size in \FIG{4}E).
-This provides further evidence that our results were not the solely the result of the firing of interneurons.
+Finally, if interneurons with less spatial specificity were driving the classification of slower dynamics, we would expect to see the posterior be less spatially concentrated during these events.
+Instead, we found that stationary and stationary-continuous mixtures tended to have narrow posteriors (as characterized by 95\% HPD region size in \FIG{4}E).
+This provides further evidence that our large fraction of slow-speed events were not the solely the result of the firing of interneurons.
 
 We then asked whether the slow spatially coherent events (stationary-continuous mixtures) potentially reflect a slow theta sweep rather than a replay event.
-We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would only be seen during rapid theta sequences (median 50 cm \FIG{5}C).
+We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would require rapid, long-distance theta sequences (median 50 cm \FIG{5}C).
 In addition, the majority of our events occur when the animal is moving at speeds < 2 cm/s, as expected for SWRs (\FIGSUPP[5]{fig5sup3}B).
@@
 There are several lines of evidence against this.
-First, as mentioned above, most of our SWRs had spikes from multiple tetrodes, including those with stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell is driving these dynamics.
+First, as mentioned above, most of our SWRs contained spikes from multiple tetrodes, including those with stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell could drive these dynamics.
 Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
 bursts which have a typical duration of 6-24 ms, \citep{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
-Third, only ~50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode.
-.
-
-One last concern is that we have combined spiking information from several hippocampal subfields (CA1, CA2, and CA3), which may influence the types of dynamics we observed.
-However, CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 \citep{DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
-Finally, we also verified that when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
-Therefore, we feel our combination of spiking information from multiple subfields is appropriate.
+Third, only ~50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode, rather than single units.
+
+Finally, we wished to assess whether the spiking information from different hippocampal subfields (CA1, CA2, and CA3), could influence the types of dynamics we observed.
+This seemed unlikely, given that CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 \citep{DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
+However, we verified that when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
 
@@
 \caption{
-Comparison of the state space decoder to the standard "Bayesian" decoder.
-For each panel, the top five rows are the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
-Note that these categories are not mutually exclusive since a SWR can have more than one dynamic or combination of dynamics.
-The sixth row is all SWRs containing unclassified dynamics.
+Event speeds calculated using three common "Bayesian" decoder approaches (A-C) compared to using the state space model (D).
+For each panel, the top five rows show the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
+Note that these categories are not mutually exclusive since a SWR can include more than one dynamic or combination of dynamics.
+The sixth row shows all SWRs containing unclassified dynamics.
 The final row is the estimated speed for all SWRs.
@@
 This comparison is complicated due to the variety of parameters and inference methods used previously, so here we chose three commonly used approaches and a reasonable parameter set for the comparison.
-We began with an examination of events based on a measure of statistical significance, which has been the standard in the field.
+We began with an examination of events based on statistical significance, which has been the standard in the field.
 We applied the clusterless decoder in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
-We then used the radon transform method \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019} to calculate, for each SWR, a measure of significance based on the assumptions that the probability density moves at a constant speed throughout the event.
+We then used the radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
 We found that ~12\% of the classified events (using the full state-space model) were identified as significant.
-Not surprisingly, when events were classified based on which dynamic they contained, we found substantial variability: of 7176 SWRs that contained a period of stationary dynamics, 542 (8\%) were significant while of 4511 SWRs that contained a period of continuous dynamics, 1386 (~30\%) were significant.
-And when we restricted the analysis to stationary events that were >100 ms long, we found that 28\% were significant, and increase that is not surprising given that the shuffle used for the radon method penalizes shorter events.
-We also note that the linear regression method \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, detects only non-zero slopes as significant, and thus cannot identify stationary events as significant.
-The substantial variability in the proportions of "significant" events indicates that conclusions about which events are meaningful to the brain could be very different depending on the specific method and parameters used.
+Not surprisingly, the fraction of significant events was strongly related to which dynamic they contained.
+Of the SWRs that contained a period of continuous dynamics, ~30\% (1386/4511) were significant.
+Of all SWRs that contained a period of stationary dynamics, 8\% (542/7176) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
+This increase is not surprising, since the shuffle used for the radon method penalizes shorter events.
+An alternative approach, the linear regression method, which samples from the posterior probability and uses linear regression to fit a line to the samples \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, detects only non-zero slopes as significant, and thus by definition cannot identify stationary events as significant.
 
 Conclusions about the representational content of events were similarly variable across methods.
-In addition to the radon and linear regression methods, we also computed  the MAP estimate (the maximum probability position of each time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+In addition to the radon and linear regression methods, we also computed  the MAP estimate (the maximum probability position of each 20 ms time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
-We found that the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, has a mixture of both fast and slow speeds.
+We found that the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
 In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
-This is to be expected because the MAP method is imposing the least amount of smoothing out of the other methods (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
-Our decoder imposes more smoothing than the MAP estimate but much less than the line fitting methods and therefore is closer to the MAP estimates.
+This is to be expected because the MAP method imposes minimal smoothing (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
+Our decoder imposes more smoothing than the MAP estimate but much less than the linear methods and therefore is closer to the MAP estimates.
 These results indicate that the methods that made more assumptions about the structure of replay (radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
+The substantial variability in both the proportions of "significant" events and the speed of the events demonstrates that key conclusions about the events could be very different depending on the specific method and parameters used.
 
@@
 This model is robust: it is relatively insensitive to a range of plausible transition parameters between dynamics and instead strongly reflects the hippocampal place representation structure.
-We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far more than have previously associated with identifiable activity patterns.
+We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far more than have been previously associated with identifiable activity patterns.
 This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
-These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical approaches that assume that event progress at constant speed.
+In fact, these slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical approaches that assume that events progress at constant speed.
 The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and closer to real-world speeds.
@@
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
-Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary and continuous dynamics.
+Our results indicate that this is not the case: in our dataset, the large majority of SWRs (86\%) contained spatially coherent content, as defined by some combination of stationary and continuous dynamics.
 These events were mostly decoded representations where the most probable positions were highly spatially concentrated (i.e.
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats experience the environment.
-The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
-Events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
-Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
+Events with slower dynamics tended to persist for approximately the same length of time as events  with faster continuous dynamics (\texttildelow50-100 ms), consistent with a brief activation of a representation of a location or a short trajectory through space.
+Events with slower dynamics most often represented the animal's current location, similar to continuous events identified in previous work \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}; however, slower events were also capable of representing locations far from the animal.
 
 These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
-Instead, there is a much richer variety of speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+Instead, replay events encompass a much richer variety of event speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
 Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
@@
 
-The differences between our finding and previous approaches also highlight the importance of identifying the specific question that an analysis seeks to answer.
-Initial studies of replay (CITE LEE Wilson, Foster Wilson, Davidson, Karlsson) focused on the question of whether the observed patterns of spiking seen during replay reflected structured content or "random" spiking as defined by a comparison to one or more shuffles.
-These and subsequent studies answered the question "are replay events the result of random spiking" with a clear "no".
-These studies did not, however, establish that the significance criteria used to show non-random structure has anything to do with the presence of spatial structure or events that are potentially useful for memory processes
+The differences between our findings and previous approaches also highlight the importance of identifying the specific question that an analysis seeks to answer.
+Initial studies of replay \citep{LeeMemorySequentialExperience2002, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009} focused on the question of whether the observed patterns of spiking seen during replay reflected structured content or "random" spiking as defined by a comparison to one or more shuffles.
+These and subsequent studies answered the question "are replay events the result of random spiking?" with a clear "no".
+These studies did not, however, establish that the significance criteria used to show non-random structure has anything to do with the presence of spatial structure or events that are potentially useful for memory processes.
 
 Our method aims to identify structure within events and to capture their complex dynamics to determine what sort of information replay events could provide in service of memory functions.
-We define precisely patterns of spatial representation that are, and are not, interpretable in terms of the structure of a given environment and the spatial activity seen during movement.
+We precisely define patterns of spatial representation that are, and are not, interpretable in terms of the structure of a given environment and the spatial activity seen during movement.
 In so doing, we find that most, but not all, events contain spiking that corresponds to a location or a trajectory through the animal's environment.
 
-The different questions addressed by our method compare to previous methods can also explain the apparent differences in findings.
+The different questions addressed by our method compared to previous methods can also explain the apparent differences in findings.
 Our results showed that 19\% of SWRs contained continuous trajectories---meaning that portions of the SWR had spatially coherent trajectories that moved at high velocities---but only 4\% were continuous throughout the entire SWR.
@@
 Furthermore, these studies found trajectories on average move at much higher speeds (20x normal rat movement speed) than we report here.
-These previous findings were based largely on linear fits to the data, while our model allows the trajectory to change speed during the SWR.
-Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics (see for example \FIG{1}).
-Second, the line has to be fit to the whole SWR event, meaning that it is sensitive to the estimation of the start end of the SWR event.
-In contrast, our model makes an estimate for each time point and only depends on the previous and future time step, which minimizes the dependence of the event on those boundaries and any possible "noise" events within the SWR.
-This, along with our more permissive threshold for SWRs, allows us to characterize more SWRs than previous studies, rather than not analyze them.
-
-Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that a subset of replay trajectories contained jumps in position.
-We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
+These previous findings were based largely on a single linear fits to the data for each event, while our model allows the trajectory to change speed during the SWR.
+Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics for some portion of the event(see for example \FIG{1}).
+Further, because the linear fit includes the whole SWR event, it is highly sensitive to the definition of the start and end of the SWR event, an often arbitrary definition which varies across research groups.
+In contrast, our model makes an estimate for each time point and only depends on the previous and future time step, which minimizes the dependence of the event on those boundaries and any possible "noise" activity within the SWR.
+This, along with our more permissive threshold for SWRs, allows us to characterize more SWRs than previous studies, rather than exclude them from analysis.
+
+Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that a subset of replay trajectories contained systematic jumps in position.
+We found that while replay speeds are, on average, distributed log normally, a small subset of the replays include definitive "jumps" from one position to another.
 Because our model makes it possible to identify different dynamics within the same SWR, we are able to identify both types of sequences.
@@
 Events that contain multiple dynamics (e.g.
-two continuous representations separated by a brief fragmented representation) might also help bind together experiences that occur in the same environment but at different places.
-Alternatively, it might be that downstream structures are attuned to these dynamics, and that each period of coherent hippocamapal dynamics drives its own set of responses elsewhere in the brain.
-We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020a}, suggesting that the underlying sequences seen during replay and theta may overlap in their content and dynamics.
+two continuous representations separated by a brief fragmented representation) might also help bind together experiences that occur in the same environment but at different places, whereas sustained spatially incoherent representations might represent coherent representations in other spatial environments \citep{KarlssonAwakereplayremote2009} or non-spatial experiences.
+Alternatively, downstream structures may be are attuned to the shifts in dynamics, and that each period of coherent hippocampal dynamics drives its own set of responses elsewhere in the brain.
+We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020a}, suggesting that the underlying sequences seen during replay and theta may overlap in their dynamics as well as their content.
 
@@
 Additionally, as the start and end time of each SWR event are also detection method- and parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
-Examination of such periods may shed light on when and why the hippocampus may be representing current position, past experience, or possible future locations, to understand how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
+Examination of such periods may advance our understanding of when and why the hippocampus may be representing current position, past experience, or possible future locations, how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
 
@@
 
-Fourth, our model avoided using large temporal bins and does not make assumptions about the linear progression of events while still taking into account the uncertainty of the data.
-Instead, the model allows for movement in any direction allowed by geometry of the environment.
-Because the model gives a moment-by-moment estimate of the position and dynamic, this also reduces the effect of misidentifying the SWR start and end times.
+Fourth, our model avoided using large temporal bins and does not make assumptions about the linear progression of events.
+Instead, the model allows for movement in any direction allowed by geometry of the environment, and further, still accounts for the uncertainty of the data.
+Because the model gives a moment-by-moment estimate of the position and dynamic, this also minimizes the effect of misidentifying the SWR start and end times.
 Combined with an acausal estimation of latent position, this approach allows us to identify the specific timing of transitions in dynamics at a fine timescale.
-This approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
+In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
 
@@
 Our model also represents a middle ground between Hidden Markov-style models \citep{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
-In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms)
+In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms).
 
 Code and algorithms for decoding hippocampal replay have not typically been made accessible or easy to use.
-We believe this is not ideal because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
+We believe this is suboptimal because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
 Accordingly, we have made code for this model publicly available as a python package at the following URL \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
@@
 The model can be extended to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
-It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+It can also be extended to account for task and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
 Finally, models can be built to estimate any covariate, including movement direction \citep{KayConstantSubsecondCycling2020}.

### 2021-06-01 Article.tex
Update on Overleaf.

@@
 
-The brain has the remarkable ability to store and retrieve representations of past events, allowing memories of the past to influence the future.
+The brain has the remarkable ability to store and retrieve representations of past events, enabling memories of the past to influence future behavior.
 These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumConditioningConsciousRecollection2004}.
@@
 
-The canonical example of "replay" is in the rodent hippocampus, where cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+The canonical example of "replay" is seen in the rodent hippocampus, where cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 Such memories would seem to be useful, in that they could encode the stimuli and values associated with a given place irrespective of the path used to get there.
-If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a rapid trajectory through space, but instead to either a stable pattern associated with a single location or perhaps a brief pattern  more similar to one seen during a real experience.
+If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a rapid trajectory through space, but instead to either a stable pattern associated with a single location or perhaps a brief pattern more similar to one that occurs during a real experience.
 
@@
 We find that the large majority of SWRs contain spatially coherent content; that is, trajectories that are spatially concentrated at each moment in time and have no large discontinuities in position.
-Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences.
+Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at slower speeds, more consistent with real-world experiences.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
-The shaded colors indicate the category of the speed of the trajectory at that time (Classification: Stat.
+The shaded colors indicate the category of the speed of the trajectory at that time (Stat.
 = Stationary, S-C-M = Stationary-Continuous-Mixture, Cont.
@@
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the standard decoder, and incorporate information about times when there is no spiking.
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \cite{DengClusterlessDecodingPosition2015}.
 Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.
@@
 Importantly, the data drives the estimated dynamic, so even if the probability of staying in a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within a time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
-In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
+In line with this, we show below that our results are relatively insensitive to the value of this parameter.
 
@@
 We use acausal decoding, meaning that we use all information from both past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, left and right plot respectively).
+With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position; \FIG{1}E, left plot) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, right plot).
 
@@
 First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle.
-Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E).
+Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the radon transform).
 The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation.
@@
 The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
-The middle panel shows the probability of each dynamic over time as in \FIG{1}F.
-Shaded regions correspond to the speed classifications as in \FIG{1}G.
+The middle panel shows the probability of each dynamic over time as in \FIG{1}E, left panel.
+Shaded regions correspond to the speed classifications as in \FIG{1}F.
 The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
@@
 % Figure 2-Supplemental 1
-\figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.
+\figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.
 \textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
@@
 % Figure 2-Supplemental 3
-\figsupp[Spatial coverage of the track for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.
+\figsupp[Population firing rate on the track is spatially uniform and consistent for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.
 Each grey line represents a recording session for that animal.
@@
 Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
-This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies [cite].
+This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}.
 Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during movement, demonstrating the power of our algorithm.
@@
 However, we also observed many events with spatially coherent content that do not have this structure.
-For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
+For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), representations that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
 We also observed SWRs where the content is spatially incoherent throughout (\FIGSUPP[3]{fig3sup1}A, D).
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
-We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
+We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
@@
 To more directly compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs included time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all detected SWRs).
+Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs included time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
 Thus, focusing on only high-speed, linear-fit trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
-Note that here we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by line-fitting approach of the standard decoder, to take advantage of the ability of our decoder to capture a range different speeds within each SWR event.
+We emphasize here that we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by line-fitting approach of the standard decoder.
+The consideration of this broader class of SWRs allows us to take advantage of the ability of our decoder to capture a range different speeds within each SWR event.
 
 We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content.
+We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content.
 Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
@@
 To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate.
-The 95\% HPD region corresponds to the most probable latent positions as estimated by the model.
+The 95\% HPD region corresponds to the set of most probable latent positions as estimated by the model.
 Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
@@
 shuffled, one-sided; \FIGSUPP[4]{fig4sup1}).
-We also compared the HPD regions between the real data and the shuffle that preserved more of the local temporal structure while destroying global spatial structure.
+We also compared the HPD regions between the real data and a shuffle that swapped the runs and circularly shuffled position, as above.
 This also resulted in larger HPD regions than the real data for spatially coherent and incoherent SWRs (recording session \#1, spatially coherent: 29 cm vs.
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) were classified as having only a single dynamic or dynamic mixture (\FIG{5}A, for specific example SWR see \FIG{3}B), whereas SWRs with multiple dynamics (such as those in \FIG{3}A, C, D, E, and F) were less common.
+We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) were classified as having only a single dynamic or dynamic mixture (\FIG{5}A, for specific example SWR see \FIG{3}B), whereas SWRs with multiple dynamics or dynamic mixtures (such as those in \FIG{3}A, C, D, E, and F) were less common.
 Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
@@
 We note that roughly a third of these events persisted for the entire duration of the SWR, but the other two thirds included some time where the model was uncertain about the dynamic.
-This shows the power of our classifier because it allows us identify periods of time when the model has high confidence with each SWR, and use these periods to characterize the SWR overall or to focus on specifically for further analysis.
+This shows the power of our state space decoder because it allows us identify periods of time when the model has high confidence with each SWR, and use these periods to characterize the SWR overall or to focus on specifically for further analysis.
 
@@
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1,437 of 21,433 (\texttildelow7\%) of classified SWRs.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in \texttildelow7\% of classified SWRs (1,437 of 21,433 classified SWRs).
 Of these, 46\% (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
@@
 
-We then carried out a series of control analyses to determine whether our results could be do to differences in activity across dynamics or interneuron-specific activity patterns.
+We then carried out a series of control analyses to determine whether our results could be due to differences in spiking statistics across dynamics or interneuron-specific activity patterns.
 We first calculated the multiunit spiking rates and found that these were similar across the different dynamics.
@@
 We found that our results were robust to this increase in threshold (\FIGSUPP[5]{fig5sup2}A).
-Indeed, most of our events and most periods of dynamics within each event included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}B, \FIGSUPP[5]{fig5sup3}A).
+Indeed, most of our events and most periods of dynamics within each event included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}).
 Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons \citep{FoxLocalizationanatomicalidentification1975}.
@@
 bursts which have a typical duration of 6-24 ms, \citep{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
-Third, only ~50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode, rather than single units.
+Third, only \texttildelow50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode, rather than single units.
 
@@
 
-Finally, we asked how our method compares to the standard approaches to identifying significant events and characterizing their content.
+Finally, we asked how our method compares to the standard approaches to identifying significant replay events via a non-parametric shuffle test and characterizing their content.
 This comparison is complicated due to the variety of parameters and inference methods used previously, so here we chose three commonly used approaches and a reasonable parameter set for the comparison.
-We began with an examination of events based on statistical significance, which has been the standard in the field.
-We applied the clusterless decoder in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
+We began with an examination of events based on statistical significance of a line fit to a circularly permuted spatial distribution, which has been the standard in the field.
+We applied the clusterless algorithm in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
 We then used the radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
-We found that ~12\% of the classified events (using the full state-space model) were identified as significant.
+We found that \texttildelow12\% of the classified events (as defined by our state-space decoder) were identified as significant.
 Not surprisingly, the fraction of significant events was strongly related to which dynamic they contained.
-Of the SWRs that contained a period of continuous dynamics, ~30\% (1386/4511) were significant.
-Of all SWRs that contained a period of stationary dynamics, 8\% (542/7176) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
+Of the SWRs that contained a period of continuous dynamics, 31\% (1386 of 4511 SWRs) were significant.
+Of all SWRs that contained a period of stationary dynamics, 8\% (542 of 7176 SWRs) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
 This increase is not surprising, since the shuffle used for the radon method penalizes shorter events.
@@
 Conclusions about the representational content of events were similarly variable across methods.
-In addition to the radon and linear regression methods, we also computed  the MAP estimate (the maximum probability position of each 20 ms time bin) \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+In addition to the radon and linear regression methods, we also computed  the MAP estimate, the maximum probability position of each 20 ms time bin \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
@@
 In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
-This is to be expected because the MAP method imposes minimal smoothing (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
+This is to be expected because the MAP method imposes less smoothing (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
 Our decoder imposes more smoothing than the MAP estimate but much less than the linear methods and therefore is closer to the MAP estimates.
@@
 \section*{Discussion}
-We developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
+We developed a state space model that identifies and quantifies replay in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
 This model is robust: it is relatively insensitive to a range of plausible transition parameters between dynamics and instead strongly reflects the hippocampal place representation structure.
@@
 
-Previous work reported that less than half of SWRs or high-population-activity events contained replay \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work reported that less than half of SWRs or high-population-activity events contained replay events \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats experience the environment.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats move through their environment.
 The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 Furthermore, these studies found trajectories on average move at much higher speeds (20x normal rat movement speed) than we report here.
-These previous findings were based largely on a single linear fits to the data for each event, while our model allows the trajectory to change speed during the SWR.
+These previous findings were based largely on a single linear fits to the data for each event, while our model allows the trajectory to change speed during the SWR, which can lead to very different estimates of the speed of the event.
 Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics for some portion of the event(see for example \FIG{1}).
@@
 
-Similarly, it is possible that slow and fast dynamics are supported by different subcircuits.
+Similarly, it is possible that slow and fast dynamics are supported by different subcircuits within the hippocampal formation.
 \cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
@@
 
-Additionally, as the start and end time of each SWR event are also detection method- and parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
+One key aspect to understanding how the hippocampus engages with other brain structures is identifying the times when the neural dynamics have changed.
+The times of SWRs or increases in multiunit activity  has provided an important, but coarse, landmark to this change in neural dynamics.
+Because our method allows for moment-by-moment characterization of the neural dynamic, we suggest that evaluating the start and end of each dynamic may prove a powerful way identify periods of interest.
 Examination of such periods may advance our understanding of when and why the hippocampus may be representing current position, past experience, or possible future locations, how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
@@
 Code and algorithms for decoding hippocampal replay have not typically been made accessible or easy to use.
-We believe this is suboptimal because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
+This is problematic because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
 Accordingly, we have made code for this model publicly available as a python package at the following URL \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
@@
 
-State-space models like the one we use here can enable a richer set of analyses of generative events that take advantage of all of the available data.
+State-space models like the one we use here can enable a richer set of analyses of events that take advantage of all of the available data.
 These approaches can be applied not just to SWRs but to all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.

### 2021-06-02 Article.tex

@@
 For next 190 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
-Finally, for the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
+For the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
 \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
@@
 \figsupp[The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms)]{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
-\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
+\textbf{(A)} Each panel shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
 The corresponding expected duration of staying in the dynamic is listed as duration.
@@
 the case where there is no probability of switching to another dynamic.
-Shaded regions correspond to the classification as in \FIG{1}.
+Shaded regions correspond to the classification as in \FIG{1}F.
 \textbf{(B)} The probability of position over time for each diagonal value.
@@
 We use acausal decoding, meaning that we use all information from both past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position; \FIG{1}E, left plot) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, right plot).
+With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position; \FIG{1}E, left panel) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, right panel).
 
@@
 \figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.
-\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
-The middle plot shows the probability of each dynamic over time from the 2D decode.
+\textbf{(A)} The left panel shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
+The middle panel shows the probability of each dynamic over time from the 2D decode.
 Shaded regions indicate classification category as in \FIG{1}F and \FIG{2}.
-The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
+The rightmost panel shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
 The latent position posterior summed over time is shown in the purple shading.
@@
 Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019} and often requires inclusion decisions that can vary widely across experimenters.
-We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
+We therefore adopted a "clusterless" approach which directly uses multiunit spikes and their spike waveform features to decode position without spike sorting (see Methods).
 Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
@@
 \subsection*{Most replays contain coherent spatial content}
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}; position linearized to 1D).
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
@@
 \figsupp[More examples of SWRs with non-constant speed trajectories]{A-I.
-More examples of SWRs that would not be well-characterized by using the standard decoder.
+More examples of SWRs with non-constant speed trajectories.
 Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}}\label{figsupp:fig3sup1}
@@
 Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
-In contrast, smaller values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD region provides a complementary measure of spatial information, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
+In contrast, smaller values of the HPD region size indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
+Thus, the HPD region size provides a complementary measure of spatial information, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
 
@@
 % Figure 4-Supplemental 1
-\figsupp[Average 95\% HPD size increases on the 95\% highest posterior density on position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-position shuffled data for two recording sessions from different animals.
-Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
+\figsupp[Shuffling the position with replacement increases the average 95\% HPD region size]{Average 95\% HPD region size increases on position shuffled data for two recording sessions from different animals.
+Red line represents the average 95\% HPD region size for all ripples.
 The histogram represents the distribution after 50 shuffles of the position data.
@@
 % Figure 4-Supplemental 2
-\figsupp[Average 95\% HPD size increases on run-swapped, circularly shuffled position data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
-run-swapped, circularly shuffled position data for two recording sessions from different animals.
+\figsupp[Shuffling the position data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Average 95\% HPD region size increases for run-swapped, circularly shuffled position data compared to real data.
+Data from two recording sessions from different animals.
 Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
@@
 The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
-This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
+This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
 
@@
 Not surprisingly, the fraction of significant events was strongly related to which dynamic they contained.
-Of the SWRs that contained a period of continuous dynamics, 31\% (1386 of 4511 SWRs) were significant.
-Of all SWRs that contained a period of stationary dynamics, 8\% (542 of 7176 SWRs) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
+Of the SWRs that contained a period of continuous dynamics, 31\% (1,386 of 4,511 SWRs) were significant.
+Of all SWRs that contained a period of stationary dynamics, 8\% (542 of 7,176 SWRs) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
 This increase is not surprising, since the shuffle used for the radon method penalizes shorter events.
@@
 These events were mostly decoded representations where the most probable positions were highly spatially concentrated (i.e.
-had small HPD size), indicating that they were driven by spatially specific firing.
+had small HPD region size), indicating that they were driven by spatially specific firing.
 This was in contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
@@
 These previous findings were based largely on a single linear fits to the data for each event, while our model allows the trajectory to change speed during the SWR, which can lead to very different estimates of the speed of the event.
-Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics for some portion of the event(see for example \FIG{1}).
+Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics for some portion of the event(see for example \FIG{2}).
 Further, because the linear fit includes the whole SWR event, it is highly sensitive to the definition of the start and end of the SWR event, an often arbitrary definition which varies across research groups.
@@
 The model can be extended to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
-It can also be extended to account for task and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+It can also be extended to account for task conditions (such as the inbound and outbound conditions in our spatial alternation task) and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
 Finally, models can be built to estimate any covariate, including movement direction \citep{KayConstantSubsecondCycling2020}.
@@
 By using only the top values, this measurement of spread is not influenced by multimodal distributions (whereas an alternative measure like the quantiles of the distribution would be).
-In this manuscript, we use the HPD size---the total area of the track covered by the 95\% HPD region---to evaluate the uncertainty of the posterior probability of position.
-
-In order to calculate the HPD size, we first calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
+In this manuscript, we use the HPD region size---the total area of the track covered by the 95\% HPD region---to evaluate the uncertainty of the posterior probability of position.
+
+In order to calculate the HPD region size, we first calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
 $$
@@
 The 95\% HPD region is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
-The HPD size is calculated by taking the integral of the members of this set:
+The HPD region size is calculated by taking the integral of the members of this set:
 $$
@@
 
-\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics and HPD Size}
+\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics and HPD region size}
 In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
@@
 Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
-Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
+Code for the decoder is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
 Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn \citep{Waskomseabornstatisticaldata2021}.
 All code is open-source and licensed under the MIT Software License.
-Classifier code can be easily installed as a python package with all requisite dependencies using pip or conda.
+Decoder code can be easily installed as a python package with all requisite dependencies using pip or conda.
 See software repositories for specific details.

### 2021-06-02 src/visualization.py
Add plot for plotting run decode

@@ -1429,3 +1429,139 @@ def plot_upset_classification(replay_info, intersection_frac_threshold=0.01):
         ax_dict["shading"].add_patch(rect)
 
     return ax_dict, upset
+
+
+def plot_run_slice(results, data, classifier, start_time, end_time,
+                   cmap="bone_r",
+                   figsize=(7, 9.72)):
+    time_slice = slice(start_time, end_time)
+
+    fig, axes = plt.subplots(
+        5,
+        1,
+        sharex=False,
+        constrained_layout=True,
+        figsize=figsize,
+        gridspec_kw={"height_ratios": [0.5, 0.5, 0.5, 3, 0.5]},
+    )
+
+    # axis 0
+    lfp_start = start_time - pd.Timedelta(100, unit="ms")
+    lfp_end = end_time + pd.Timedelta(100, unit="ms")
+    lfps = data["lfps"].loc[lfp_start:lfp_end]
+    tetrode_id = data['tetrode_info'].loc[
+        (data['tetrode_info'].area == 'CA1') &
+        (data['tetrode_info'].validripple == 1.0)].iloc[0].tetrode_id
+
+    axes[0].plot(
+        (lfps.index) / np.timedelta64(1, "s"),
+        np.asarray(lfps.loc[:, tetrode_id]),
+        color="black",
+    )
+    axes[0].set_xlim((start_time / np.timedelta64(1, 's'),
+                      end_time / np.timedelta64(1, 's')))
+    axes[0].axis("off")
+
+    # axis 1
+    multiunit_rate = data['multiunit_firing_rate'].loc[time_slice]
+    axes[1].fill_between(multiunit_rate.index / np.timedelta64(1, 's'),
+                         np.asarray(multiunit_rate).squeeze(),
+                         color='black')
+    axes[1].set_ylabel("Mulitunit Rate\n[spikes/s]")
+    axes[1].set_xlim((start_time / np.timedelta64(1, 's'),
+                      end_time / np.timedelta64(1, 's')))
+    axes[1].set_xticks([])
+    axes[1].set_ylim([0, int(np.max(np.asarray(multiunit_rate)))])
+    axes[1].set_yticks([0, int(np.max(np.asarray(multiunit_rate)))])
+    sns.despine(ax=axes[1], offset=5)
+    axes[1].spines["bottom"].set_visible(False)
+
+    # axis 2
+    probability = results.sel(
+        time=time_slice).acausal_posterior.sum(["position"])
+
+    for state, prob in zip(results.state.values, probability.values.T):
[…]

### 2021-06-02 src/figure2_supplement3.py
Update scripts

@@ -1,0 +1,67 @@
+import logging
+
+import xarray as xr
+from loren_frank_data_processing.position import (EDGE_ORDER, EDGE_SPACING,
+                                                  make_track_graph)
+from replay_trajectory_classification import ClusterlessClassifier
+from sklearn.model_selection import KFold
+from src.load_data import load_data
+from src.parameters import (ANIMALS, TRANSITION_TO_CATEGORY,
+                            continuous_transition_types, discrete_diag, model,
+                            model_kwargs, movement_var, place_bin_size,
+                            replay_speed)
+from src.visualization import plot_run_slice
+
+logging.basicConfig(
+    level='INFO', format='%(asctime)s %(message)s',
+    datefmt='%d-%b-%y %H:%M:%S')
+
+
+def main():
+    epoch_key = 'bon', 5, 2
+
+    data = load_data(epoch_key)
+    track_graph, center_well_id = make_track_graph(epoch_key, ANIMALS)
+
+    cv = KFold()
+    results = []
+
+    for fold_ind, (train, test) in enumerate(
+            cv.split(data["position_info"].index)):
+        logging.info(f'Fitting Fold #{fold_ind + 1}...')
+        classifier = ClusterlessClassifier(
+            place_bin_size=place_bin_size,
+            movement_var=movement_var,
+            replay_speed=replay_speed,
+            discrete_transition_diag=discrete_diag,
+            continuous_transition_types=continuous_transition_types,
+            model=model,
+            model_kwargs=model_kwargs)
+        classifier.fit(
+            position=data["position_info"].iloc[train].linear_position,
+            multiunits=data["multiunit"].isel(time=train),
+            track_graph=track_graph,
+            center_well_id=center_well_id,
+            edge_order=EDGE_ORDER,
+            edge_spacing=EDGE_SPACING,
+        )
+
+        logging.info('Predicting posterior...')
+        results.append(
+            classifier.predict(
+                data["multiunit"].isel(time=test),
+                time=data["position_info"].iloc[test].index,
+            )
+        )
+
+    results = (xr.concat(results, dim="time")
+               .assign_coords(state=lambda ds:
+                              ds.state.to_index().map(TRANSITION_TO_CATEGORY)))
+
+    start_time, end_time = data['position_info'].index[[340_000, 346_000]]
+
+    plot_run_slice(results, data, classifier, start_time, end_time)
+
+
+if __name__ == "__main__":
+    main()

### 2021-06-02 src/figure5.py

@@ -1,4 +1,6 @@
+import ast
 import os
+import re
 from glob import glob
 
 import matplotlib.pyplot as plt
@@ -7,8 +9,9 @@ import pandas as pd
 import seaborn as sns
 from loren_frank_data_processing import make_tetrode_dataframe
 from src.figure_utilities import PAGE_HEIGHT, TWO_COLUMN, save_figure
-from src.parameters import (_BRAIN_AREAS, ANIMALS, PROCESSED_DATA_DIR,
-                            SHORT_STATE_ORDER, STATE_COLORS, STATE_ORDER)
+from src.parameters import (_BRAIN_AREAS, ANIMALS, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR, SHORT_STATE_ORDER,
+                            STATE_COLORS, STATE_ORDER)
 from src.visualization import (SHORT_STATE_NAMES, _plot_category,
                                plot_category_duration,
                                plot_linear_position_markers,
@@ -16,21 +19,28 @@ from src.visualization import (SHORT_STATE_NAMES, _plot_category,
 from upsetplot import UpSet
 
 
-def load_replay_info():
+def load_replay_info(
+    n_unique_spiking=2,
+    data_type="clusterless",
+    dim="1D",
+    probability_threshold=PROBABILITY_THRESHOLD,
+    speed_threshold=4,
+    exclude_interneuron_spikes=False
+):
     tetrode_info = make_tetrode_dataframe(ANIMALS)
-
-    data_type, dim = "clusterless", "1D"
-    n_unique_spiking = 2
-
-    file_paths = glob(
-        os.path.join(PROCESSED_DATA_DIR,
-                     f"*_{data_type}_{dim}_replay_info.csv")
-    )
+    prob = int(probability_threshold * 100)
+    if exclude_interneuron_spikes:
+        interneuron = 'no_interneuron_'
+    else:
+        interneuron = ''
+    file_regex = f"*_{data_type}_{dim}_{interneuron}replay_info_{prob:02d}.csv"
+    file_paths = glob(os.path.join(PROCESSED_DATA_DIR, file_regex))
     replay_info = pd.concat(
         [pd.read_csv(file_path) for file_path in file_paths], axis=0,
     ).set_index(["animal", "day", "epoch", "ripple_number"])
     replay_info = replay_info.loc[
-        replay_info.n_unique_spiking >= n_unique_spiking
[…]

### 2021-06-02 Article.tex
Update on Overleaf.

@@
 Each dot corresponds to the average probability of that dynamic for a given constant speed sequence.
-We use a 0.8 threshold (dotted line) to classify each sequence based on the dynamic or dynamics which contribute maximally to the posterior (shaded colors).
+We use a 0.80 threshold (dotted line) to classify each sequence based on the dynamic or dynamics which contribute maximally to the posterior (shaded colors).
 }
@@
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \cite{DengClusterlessDecodingPosition2015}.
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \citep{DengClusterlessDecodingPosition2015}.
 Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.
@@
 To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 cm/s.
-From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
+From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.80 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.80.
 In this manuscript, we will refer to these intermediate speeds as mixture dynamics.
@@
 We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
+Our goal in choosing 0.80 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
 Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.
@@
 
-Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
+Using a 0.80 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
 To ensure that this reflects the spatially tuned firing of the neurons, we trained the encoding model with positions resampled with replacement, a shuffling procedure which disrupts the relationship between spiking and position, for two recording sessions.
@@
 \caption{A-F.
-Six examples of SWRs with non-constant speed trajectories.
+Examples of SWRs with non-constant speed trajectories.
 Figure conventions are the same as in \FIG{2}.
@@
 % Figure 3-Supplemental 2
-\figsupp[Percentage of SWRs classified decreases on position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
+\figsupp[Shuffling the position data with replacement decreases the percent of SWRs classified]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
 position shuffled data for two recording sessions from different animals.
@@
 % Figure 3-Supplemental 3
-\figsupp[Percentage of spatially incoherent SWRs increases for run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
-run-swapped, circularly shuffled position.
+\figsupp[Shuffling the data by swapping the runs and circularly permuting the position increases the percentage of spatially incoherent SWRs and decreases the spatially coherent SWRs]{Shuffling the data by swapping the runs and circularly permuting the position increases the percentage of spatially incoherent SWRs and decreases the spatially coherent SWRs.
 The red line represents the percent of spatially coherent or incoherent SWRs in that recording session for actual data.
@@
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
-We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
+We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% (22,170 of 25,844) of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
@@
 
-We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content.
-Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
+We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.80.
+We find that, while this change slightly reduced the total fraction of classified SWRs (19,478 of 26,159 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,317 of 19,478 or 99\% classified SWRs) included spatially coherent content.
+Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (490 of 19,478 or 3\% classified SWRs).
 
@@
 % Figure 4-Supplemental 2
-\figsupp[Shuffling the position data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Average 95\% HPD region size increases for run-swapped, circularly shuffled position data compared to real data.
+\figsupp[Shuffling the position data by swapping the runs and circularly permuting the data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Shuffling the position data by swapping the runs and circularly permuting the data increases the average 95\% HPD region size for spatially coherent and incoherent classified times.
 Data from two recording sessions from different animals.
@@
 % Figure 5-Supplemental 2
-\figsupp[Distribution of dynamics is insensitive to parameters]{Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}
+\figsupp[Control analyses for distribution of dynamics]{Control analyses for distribution of dynamics.
+Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}
 
@@
 
-The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary).
+The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary for the entire duration of the SWR).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C).
@@
 \subsection*{Classification of Dynamics}
-We used a threshold of 0.8 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
-Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.8.
-Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.8, respectively.
+We used a threshold of 0.80 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
+Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.80.
+Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.80, respectively.
 Time periods where none of these criterion are met are considered unclassified.

### 2021-06-02 Article.tex
Update on Overleaf.

@@
 
-\title{Hippocampal replay of experience at real world speeds}
+\title{Hippocampal replay of experience at real-world speeds}
 
@@
 Representations related to past experiences play a critical role in memory and decision-making processes.
-The rat hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
+The rat hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x the movement speed of the animal.
 Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.

### 2021-06-02 elife_orignal_draft.tex
Add original draft tex

--- 2021-06-02 Article.tex
@@
 
-\title{Hippocampal replay of experience at real world speeds}
+\title{Hippocampal replay of experience at real-world speeds}
 
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[1, 2, 3]{Eric L. Denovellis}
+\author[1]{Eric L. Denovellis}
 \author[2, 3]{Anna K. Gillespie}
@@
 \begin{abstract}
-Representations related to past experiences play a critical role in memory and decision-making processes.
-The rat hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x movement speed of the animal.
+Representations of past and possible future experiences play a critical role in memory and decision-making processes.
+The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at \texttildelow20x real-world speeds.
 Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay.
@@
 Using this model, we find that the large majority of replay events contain spatially coherent, interpretable content.
-Furthermore, many events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
+Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.
 \end {abstract}
@@
 
-The brain has the remarkable ability to store and retrieve representations of past events, enabling memories of the past to influence future behavior.
-These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumConditioningConsciousRecollection2004}.
-Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking is thought to contribute to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018, CarrHippocampalreplayawake2011}.
+The brain has the remarkable ability to retrieve representations of past events and generate representations of possible future events.
+These "generative" \citep{KayConstantSubsecondCycling2020} mental events may not occur in the same time span as actual events: memories can be retrieved and used in less time than was required for the original experience, and mental simulations can span a series of actions more quickly than would be required to perform those actions.
+Thus, a potential neural substrate for retrieval and mental simulation would be expected to support rapid reinstatement of patterns of brain activity associated with past or potential future experience.
 
-The canonical example of "replay" is seen in the rodent hippocampus, where cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+Hippocampal "replay" events are well suited to that role \citep{CarrHippocampalreplayawake2011}.
+As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
-A subset of SWRs contain sequential firing similar to that seen during a previous experience, and thus are thought to "replay" these previous experiences.
+A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences.
 Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009, Tingleymethodsreactivationreplay2020}.
-One might therefore conclude that all memories that are stored or updated by SWRs correspond to sequences of locations.
-Our subjective experience, however, suggests that we can retrieve memories of individual locations without having to mentally traverse a long distance.
-Such memories would seem to be useful, in that they could encode the stimuli and values associated with a given place irrespective of the path used to get there.
-If the rodent hippocampus is capable of storing those sorts of memories, then one might expect to see SWRs where the neural activity corresponds not to a rapid trajectory through space, but instead to either a stable pattern associated with a single location or perhaps a brief pattern more similar to one that occurs during a real experience.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009}.
+One might therefore conclude that the retrieval of stored representation during SWRs is limited to sequences of locations, and that all such retrieval events are sequential in nature.
+Our subjective experience is that we can retrieve memories of locations without having to mentally traverse a long distance.
+If the rodent hippocampus is capable of that sort of retrieval, then one might expect to see SWRs where the neural activity corresponds not to a very rapid trajectory through space but instead to either a stable pattern associated with a single location or perhaps a pattern more similar to one seen during a real experience.
+Importantly, this sort of activity could still be useful, as it would take place during the short duration of an SWR (\texttildelow100 ms) and could thus support rapid retrieval and use of representations.
 
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
-Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
+Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.
 
 Our uncertainty stems in part from the dominant approaches used to identify the content of replay events.
-These approaches typically involve multiple steps and assumptions about the nature of replay, which are most commonly characterized using the standard "Bayesian" decoder.
+These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are most commonly characterized using the standard "Bayesian" decoder.
 First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons).
-Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
+Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}.
 A decoding algorithm is then applied to the spikes within these events, yielding a set of position probability distributions for each time bin.
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009, TingleyRoutingHippocampalRipples2020, BhattaraiDistincteffectsreward2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DrieuNestedsequenceshippocampal2018, OlafsdottirTaskDemandsPredict2017, TangHippocampalPrefrontalReactivationLearning2017, CarrTransientSlowGamma2012, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
-A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent an implicit definition of a "random" sequence.
+Finally, the most commonly used approach for detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}.
+A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
 
-While the standard approach identifies constant speed events, it does not consider events that are rejected by the statistical test.
-This is problematic because it has the potential to mischaracterize real events that do not move at constant speeds, such as those that change direction or are discontinuous \citep{LiuMethodsAssessmentMemory2018}.
-Furthermore, the use of large, fixed-size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
-For example, with 20 ms time bins and 3 cm position bins, an event can only move in 1.5 m/s increments (1 or more bins in 20 ms) between time steps.
-The linear fit is also highly dependent on the estimation of event boundaries, such as the start and end times of the SWR, because the fit depends on all the data over the course of the event.
+While the standard decoder approach identifies constant speed events, it does not consider events that are rejected by the statistical test, and the use of a fixed size temporal bins acts as a boxcar smoother that limits the potential movement speeds of the representation.
+The linear fit is also problematic: it has the potential to reject or mischaracterize real events that do not move at constant speeds.
 Approaches that focus on the order of cell activity within each event \citep{LeeMemorySequentialExperience2002, GuptaHippocampalReplayNot2010} allow for a relaxation of that linear assumption, but replace it with a loss of statistical power due to either ignoring all but the first spike from each cell or using an arbitrarily smoothed version of the spike train.
 These approaches also do not provide information about the dynamics of the underlying spatial representation.
-Moreover, these approaches often exclude events that have stationary representations of a single location \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
+Moreover, these approaches exclude events that  have stationary representations of a single location \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}.
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{KaeferReplayBehavioralSequences2020, CareyRewardrevaluationbiases2019, GrosmarkDiversityneuralfiring2016, WuHippocampalReplayCaptures2014, PfeifferHippocampalplacecellsequences2013}.
-For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found awake replay events containing alternation between representations of a single location and sequential spatial trajectories.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using an estimated represented location at each time bin and connecting each with a line.
+For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found that awake replays can alternate between representing a single location and sequential spatial trajectories.
 On the other hand, \cite{StellaHippocampalReactivationRandom2019} reported that replays during sleep are spatially continuous and follow Brownian diffusion dynamics.
-However, both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, making it hard to identify the source of the different conclusions.
+Both methods still used large time bins and neither took into account the uncertainty of the decoded estimates, however, making it hard to identify the source of the different conclusions.
 
-An ideal approach to identifying and quantifying the dynamics of replay would circumvent these problems.
+An ideal approach to identifying and quantifying the dynamics of generative representations would circumvent these problems.
 It would use all of the available spiking data to yield the most accurate decoded positions.
-It would provide a moment-by-moment estimate of position and dynamics that would not be dependent on the estimation of SWR start and end times and could rapidly adjust to changes in dynamics.
-It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and would provide information about the certainty of the decoded estimates.
+It would be applicable to either thresholded events or to all of the data to permit an unbiased assessment of the prevalence and nature of generative activity.
+It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and provide information about the certainty of the decoded estimates.
 It would be able to capture a range of movement dynamics from stationary or unchanging representations to trajectories that progress through space at constant or variable speeds to disorganized, spatially incoherent events.
 It would provide a robust statistical assessment of confidence for each dynamic.
-Finally, where assumptions are made, it would provide well-defined parameters whose values could be explored systematically to understand their influence on the results.
+Finally, where assumptions are made, it would provide well defined parameters whose values can be explored systematically to understand their influence on the results.
 
-We therefore developed a state space model that achieves all of those goals.
+We therefore developed a state space model of generative activity that achieves all of those goals.
 State space models are a well-understood, well-known statistical solution to the problems described above.
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing the spatial representations during SWRs as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing represented trajectories as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 We find that the large majority of SWRs contain spatially coherent content; that is, trajectories that are spatially concentrated at each moment in time and have no large discontinuities in position.
-Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at slower speeds, more consistent with real-world experiences.
+Surprisingly, while the expected high-speed, sequential replay events were identified, the most common category of events expressed representations that moved at much slower speeds, more consistent with real-world experiences, suggesting that high speed replay is not the most common manifestation of generative activity.
 These findings illustrate the power of state space models and provide a new understanding of the nature of hippocampal replay.
@@
 \begin{figure}
-\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v7}
+\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v6}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 For next 190 ms, the cells fire in sequence, representing a rapid continuous trajectory along the virtual track.
-For the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
+Finally, for the last 30 ms, cells fire randomly, out of spatial order, representing an fragmented spatial sequence.
 \textbf{(B)} Like the standard decoder, the state space model uses estimates of cells' place fields from when the animal is moving and combines them with the observed spikes in (A) to compute the likelihood of position for each time step.
@@
 This can be summarized by marginalizing over latent position (left panel) to get the probability of each dynamic over time.
-The shaded colors indicate the category of the speed of the trajectory at that time (Stat.
-= Stationary, S-C-M = Stationary-Continuous-Mixture, Cont.
-= Continuous, F-C-M = Fragmented-Continuous-Mixture, Frag.
-= Fragmented), which is determined from the probability.
+The shaded colors indicate the category of the speed of the trajectory at that time, which is determined from the probability.
 Marginalizing the posterior across dynamics also provides an estimate of latent position over time (right panel).
-Red dotted line in the right panel is the best fit line from the standard decoder using the Radon transform.
 \textbf{(F)} The probability of each dynamic depends heavily on the speed of the trajectory, as we show using a range of simulated spiking sequences each moving at a constant speed.
@@
 \figsupp[The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms)]{The model is robust to change of the probability of persisting in the same dynamic for a wide range of plausible expected durations (25 ms to 150 ms).
-\textbf{(A)} Each panel shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
+\textbf{(A)} Each plot shows the probability of each dynamic on simulated data example from \FIG{1} with a different diagonal value---which governs the probability of remaining in that dynamic.
 The corresponding expected duration of staying in the dynamic is listed as duration.
@@
 the case where there is no probability of switching to another dynamic.
-Shaded regions correspond to the classification as in \FIG{1}F.
+Shaded regions correspond to the classification as in \FIG{1}.
 \textbf{(B)} The probability of position over time for each diagonal value.
@@
 We simulate 19 Poisson spiking cells with Gaussian place fields on a 180 cm virtual linear track.
-Each place field has a 36 cm variance and a 15 Hz peak firing rate, and fields are spaced every 10 cm along the virtual track.
+Each place field has a 36 cm variance and a 15 Hz peak firing rate, which is spaced every 10 cm along the virtual track.
 We then construct a 280 ms spiking sequence (\FIG{1}A) and apply our model to the sequence.
@@
 20 ms or more) because we are able to take advantage of the prior placed on the dynamics by the state space model.
-This allows us to detect changes on smaller time scales than would be possible with the standard decoder and incorporate information about times when there is no spiking \cite{DengClusterlessDecodingPosition2015}.
-Further, because place estimates from spikes within a single bin are combined, our small time bins allow us to measure the spatial information conveyed by single spikes, rather than assuming that a downstream neuron would integrate spikes from multiple neurons on the timescale of 20 ms or more.
+This allows us to detect changes on smaller time scales than would be possible with the standard decoder.
 
-Next, we specify movement dynamic models that describe a variety of ways that the latent position---the "mental" position of the animal represented by the cells--- could evolve over time.
-(\FIG{1}C).
+Next, we specify movement dynamic models for how latent position---the "mental" position of the animal represented by the cells---evolves over time for a given dynamic (\FIG{1}C).
 We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
-Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
-We therefore define movement dynamic models to capture each of these possibilities.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
+We therefore define state transition models to capture each of these movement dynamics.
 
 In the stationary movement dynamic, the latent position does not change between time steps.
-The state transition matrix can thus be defined as an identity matrix, which predicts that the next position will be the same as the last position (at the resolution of the position bin).
+The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the resolution of the position bin).
 In the continuous movement dynamic, the latent position is most likely to be "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix.
@@
 This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).
-Last, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
+Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously.
 We model this using a uniform state transition matrix, which makes transitions to each position equally likely.
-Importantly, the fragmented dynamic provides a well-specified definition of what we mean by random, or unstructured activity.
 
-Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic via another state transition matrix between the movement dynamics (\FIG{1}D).
-In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms on average, which is approximately the duration of a SWR event.
+Finally, we specify how likely each movement dynamic is to persist in time versus change to another dynamic (\FIG{1}D) via a state transition matrix between dynamics.
+In order to be conservative with respect to switching between dynamics, we assume that each movement dynamic is likely to dominate for \texttildelow100 ms, which is approximately the duration of a SWR event.
 There is, however, a small probability of switching to one of the other movement dynamics.
 Accordingly, we set the probability of staying in a dynamic to 0.98 for each 2 ms time step, which corresponds to an expected duration of 100 ms for staying in a particular dynamic (the Markov assumption of the model means that the probability of staying in a dynamic follows a geometric distribution).
-Importantly, the data drives the estimated dynamic, so even if the probability of staying in a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within a time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
-In line with this, we show below that our results are relatively insensitive to the value of this parameter.
+Importantly, the data drives the estimated dynamic, so even if the probability of a particular movement dynamic is high, data providing clear evidence for a change in dynamic can, within the time frame of \texttildelow10 ms, drive a change to a different estimated dynamic, as in our simulated example.
+In line with this, note that below we show that our results are relatively insensitive to the value of this parameter.
 
@@
 We use acausal decoding, meaning that we use all information from both past and future spikes, to estimate the joint posterior probability of position and dynamic (see Methods).
-With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position; \FIG{1}E, left panel) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, right panel).
+With this, we can summarize the resulting posterior probability with two quantities: the probability of each movement dynamic over time (by integrating out position) and the probability of latent position over time, irrespective of movement dynamic (by summing over the movement dynamics; \FIG{1}E, left and right plot respectively).
 
@@
 
-This illustrates two key features of the model.
-First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle.
-Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the radon transform).
-The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation.
-To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 cm/s.
+Importantly, this approach allows the model to to capture a wide range of movement speeds for the latent position.
+The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the movement of the underlying position representation.
+To demonstrate this, we applied the model to 10,000 simulated trajectories, each trajectory proceeding at a constant speed (\FIG{1}F) from 1 cm/s to 10,000 m/s.
 From this, we can see that not only are there regions of speed that correspond to each of our three movement dynamics being highly probable (where we define highly probable to be greater than or equal to 0.8 probability), but there are also intermediate speeds where two of the dynamics exhibit relatively high probability; and where the sum of two of the dynamics' probabilities exceeds 0.8.
@@
 Correspondingly, if the continuous dynamic has a probability of 0.5 and the fragmented dynamic has a probability of 0.4, then we would call this a fragmented-continuous-mixture and this indicates the trajectory is moving very quickly, but not as quickly as the fragmented dynamic dictates.
-In summary, we can characterize the speed or set of speeds that occur within a SWR based on the probability of each of the three movement dynamics over time.
+In summary, we can characterize the speed or set of speeds in a SWR based on the probability of each of the three movement dynamics over time.
 We further classify the probability of each movement dynamic as being part of one of five speed categories: stationary, stationary-continuous-mixtures, continuous, fragmented-continuous mixtures, and fragmented.
 
-We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
+We note here that the choice of any particular threshold for classifying a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
 Our goal in choosing 0.8 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
@@
 To investigate this we decoded the spiking sequence in \FIG{1}A with different probabilities of staying in the same dynamic versus switching to another dynamic (\FIGSUPP[1]{fig1sup1}).
-We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993, corresponding to an expected duration between 25 and 150 ms), the model identified the movement dynamics consistently, with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
-Furthermore, the most probable latent positions remained relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
+We found that for a large range of plausible probabilities of staying in one of the dynamics (between 0.96 and 0.993 which corresponds to an expected duration between 25 and 150 ms), the model still correctly identified the sequence dynamics with high probability (\FIGSUPP[1]{fig1sup1}A), demonstrating that data itself is the most influential element in the model.
+Furthermore, the probability of latent position remains relatively consistent across the range of these probabilities as well (\FIGSUPP[1]{fig1sup1}B).
 
@@
 The top panel shows 31 cells on a W-track ordered according to linearized position by their place field peaks.
-The middle panel shows the probability of each dynamic over time as in \FIG{1}E, left panel.
-Shaded regions correspond to the speed classifications as in \FIG{1}F.
+The middle panel shows the probability of each dynamic over time as in \FIG{1}F.
+Shaded regions correspond to the speed classifications as in \FIG{1}G.
 The bottom panel shows the estimated probability of latent position over the course of the SWR as it travels down the center arm toward the right arm.
@@
 % Figure 2-Supplemental 1
-\figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding.
-\textbf{(A)} The left panel shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
-The middle panel shows the probability of each dynamic over time from the 2D decode.
-Shaded regions indicate classification category as in \FIG{1}F and \FIG{2}.
-The rightmost panel shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
+\figsupp[Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.]{Decoding the same SWR in \FIG{2} with 2D position using sorted spikes and clusterless decoding, respectively.
+\textbf{(A)} The left plot shows the spikes from cells arranged by the linear position of the peak of place field as in \FIG{2}.
+The middle plot shows the probability of each dynamic over time from the 2D decode.
+Shaded regions indicate classification category as in \FIG{1}G and \FIG{2}.
+The rightmost plot shows the most probable estimate of the latent position (MAP estimate) with color indicating time.
 The latent position posterior summed over time is shown in the purple shading.
-The light grey lines represent the position of the animal over the entire recording session and the magenta dot represents the animal's position.
+The light grey lines represent the position of the animal over the entire epoch and the magenta dot represents the animal's position.
 \textbf{(B)} Same as in (A), but with clusterless decoding.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}}\label{figsupp:fig2sup1}
@@
 Color indicates time.
-Light grey lines indicate the animal's position over the entire recording session.
+Light grey lines indicate the animal's position over the entire epoch.
 Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}}\label{figsupp:fig2sup2}
-
-% Figure 2-Supplemental 3
-\figsupp[Population firing rate on the track is spatially uniform and consistent for each animal.]{Multiunit rate of encoding spikes over track positions for each animal.
-Each grey line represents a recording session for that animal.
-}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
-
-% Figure 2-video1
-\videosupp{Example of a SWR with continuous content.
-Magenta dot represents the animal's position.
-Green dot represents the most likely decoded position projected from 1D back to the linearized 2D position.
-Green line represents the decoded positions in the last 10 ms.
-Grey lines represent the position of the animal over the course of the recording session.}\label{videosupp:fig2supv1}
 
@@
 \subsection*{Identification of continuous dynamics in a SWR with sorted or clusterless spikes}
-
-We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
-To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel).
-Here the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
-As expected, we observed that the probability of the continuous dynamic is high throughout this event, but the probability of the stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
-Using our speed classification scheme, this indicates that the speed of the replay is initially slower--- a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
-This is also evident in the posterior probability of latent linear position over time, which shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
-We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
-Importantly, when we apply the same model using the 2D position of the animal we get a similar result (\FIGSUPP[2]{fig2sup1}A).
-
-One of our criteria for a more optimal method is for it to use all of the available spiking data.
-Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
-Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019} and often requires inclusion decisions that can vary widely across experimenters.
-We therefore adopted a "clusterless" approach which directly uses multiunit spikes and their spike waveform features to decode position without spike sorting (see Methods).
-Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
-Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
-As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
-This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (\FIG{2}D vs C).
-
-\subsection*{Most replays contain coherent spatial content}
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}; position linearized to 1D).
-We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
-Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
-This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}.
-Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during movement, demonstrating the power of our algorithm.
-For each animal, we also verified that the population spiking rate was relatively consistent over all positions on the track, suggesting that we have comparable sampling of all locations in the environment (\FIGSUPP[2]{fig2sup3}).
-
-Having established the validity of our decoder in tracking the animal's position, we next assessed the prevalence of spatial content across SWRs.
-We detected SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large-amplitude events as well as the much more common small-amplitude events.
-As expected, and providing further support for the validity of the model, we observed replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using standard approaches (\FIGSUPP[2]{fig2sup2}A-F).
-However, we also observed many events with spatially coherent content that do not have this structure.
-For example, there are trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), representations that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
-We also observed SWRs where the content is spatially incoherent throughout (\FIGSUPP[3]{fig3sup1}A, D).
-
-Using a 0.8 threshold, 89\% (23,071 of 25,844) of SWRs contain at least one of the three dynamics or dynamic mixtures.
-To ensure that this reflects the spatially tuned firing of the neurons, we trained the encoding model with positions resampled with replacement, a shuffling procedure which disrupts the relationship between spiking and position, for two recording sessions.
-We then decoded the same SWR events, containing the original spikes.
-Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
-This shows that our model does not impose movement dynamics in the absence of spatial information in the data.
 
@@
 % Figure 3-Supplemental 1
-\figsupp[More examples of SWRs with non-constant speed trajectories]{A-I.
-More examples of SWRs with non-constant speed trajectories.
+\figsupp[More examples of SWRs that would not be well-characterized by using the standard decoder]{A-I.
+More examples of SWRs that would not be well-characterized by using the standard decoder.
 Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}}\label{figsupp:fig3sup1}
@@
 % Figure 3-Supplemental 2
-\figsupp[Percentage of SWRs classified decreases on position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
-position shuffled data for two recording sessions from different animals.
-Red line represents the percent of SWR events classified in that recording session for real data.
+\figsupp[Comparison of percentage of SWRs classified on real vs.
+position shuffled data]{Comparison of percentage of SWRs classified (that is, a SWR containing at least one of the five classifications) on real vs.
+position shuffled data for two epochs from different animals.
+Red line represents the percent of SWR events classified in that epoch for real data.
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
-
-% Figure 3-Supplemental 3
-\figsupp[Percentage of spatially incoherent SWRs increases for run-swapped, circularly shuffled position data]{Comparison of percentage of spatially coherent and incoherent SWRs on real vs.
-run-swapped, circularly shuffled position.
-The red line represents the percent of spatially coherent or incoherent SWRs in that recording session for actual data.
-The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
-This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1_supplemental3}}\label{figsupp:fig3sup3}
-
-% Figure 3-video1
-\videosupp{Example of a SWR with that is not purely continuous. Conventions the same as Figure 2-video1}\label{videosupp:fig3supv1}
-
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
 \end{figure}
 
+We next sought to validate the model on real hippocampal data recorded from awake, behaving rats.
+To ensure that we could capture rapid trajectories many times the speed of the animal---as described by most studies of hippocampal replay---we first decoded a single SWR with sequential population activity (\FIG{2}A, top panel) from 31 cells recorded in hippocampus while the rat was performing a spatial alternation task on a W-shaped track (see Methods for details).
+As expected, we observed that the probability of being in the continuous dynamic is high throughout this event, but the probability of being in a stationary dynamic was also noticeable at the beginning and end of the SWR (\FIG{2}A, middle panel).
+Using our speed classification scheme, this means that the speed of the replay starts slower---as a mixture of continuous and stationary dynamics---and then speeds up and slows down again.
+This is also evident in the posterior probability of latent linear position over time.
+This shows that the replay most likely travels down the center arm and up the right arm (\FIG{2}A, bottom panel).
+We can also see this when we project the maximum of the posterior of this trajectory (the most probable "mental" position) to 2D to better see the spatial trajectory on the maze (\FIG{2}C).
+Importantly, when we apply the same model using the 2D position of the animal, we get a similar result as when we use the 1D linearized position (\FIGSUPP[2]{fig2sup1}A).
+
+One of our criteria for a more optimal method is for it to use all of the available spiking data.
+Using only clustered spikes discards any spike events that cannot be uniquely assigned to a putative single neuron, substantially reducing the amount of information that the resultant decoding can use.
+Additionally, spike sorting is not necessary to recover the underlying neural population dynamics \citep{TrautmannAccurateEstimationNeural2019}, is time-consuming, and often requires inclusion decisions that can vary widely across experimenters.
+We therefore adopted a "clusterless" approach which directly relates spikes and their multiunit spike waveform features to decode position without spike sorting (see Methods).
+Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
+Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
+As predicted, the spatial extent of the event is longer and the estimate of the posterior probability of latent position is narrower for the clusterless model.
+This reflects the clusterless model's access to a larger pool of data that provides more information about the extent of the event and more certainty in the latent position and the dynamic (\FIG{2}D vs C).
+
+\subsection*{Most replays contain coherent spatial content}
+
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}, position linearized to 1D).
+One major goal of our work is to assess the overall prevalence of spatial content across SWRs, so we detect SWRs using a permissive threshold (see Methods) to ensure that we include both the rare large amplitude events as well as the much more common small amplitude events.
+
+As expected, and providing further support for the validity of the model, we observe many replays that are classified as continuous throughout the entirety of the SWR, similar to those seen using the standard decoder (\FIGSUPP[2]{fig2sup2}A-F).
+However, we also observe many events with spatially coherent content that do not have this structure.
+For example, we observe trajectories that start in one direction and reverse back to the original position (\FIG{3}A, \FIGSUPP[3]{fig3sup1}B), trajectories that remain fixed in one position (\FIG{3}B, \FIGSUPP[3]{fig3sup1}G), and trajectories that jump between arms and between dynamics (\FIG{3}C, D, F, \FIGSUPP[3]{fig3sup1}F, H, I).
+We also observe SWRs that are spatially incoherent throughout the SWR (\FIGSUPP[3]{fig3sup1}A, D).
+
+Using a 0.8 threshold, we are able to classify 89\% (23,382 of 26,160) of SWRs as containing at least one of the three dynamics or common dynamic mixtures.
+To ensure that this depends on the sequential firing of spatially tuned cells in the hippocampus, we resample position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+We then decode the same SWR events containing the same spikes.
+Only 9\% of the SWRs are classified in the shuffled datasets, a value that is significantly less than that seen for the real data (p=0.02 for recording session 1, p=0.02 for recording session 2, \FIGSUPP[3]{fig3sup2}).
+
 Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
-We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
-We find that 96\% (22,170 of 23,071) of classified SWRs and 86\% of all SWRs include spatially coherent structure, and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
+We therefore asked what fraction of classified events contain spatially coherent content, which we define as SWRs containing any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
+We find that 96\% (22,478 of 23,382) of classified SWRs and 85\% of all SWRs include spatially coherent structure and that this prevalence of spatially coherent structure is consistent across animals (\FIG{3}G).
 We then asked what fraction of events contained spatially incoherent content, defined as a SWR containing any times with fragmented or fragmented-continuous mixture dynamics.
-We find that only 14\% (3,295 of 23,071) of classified SWRs include any times with spatially incoherent structure (\FIG{3}G).
-To further validate this result, we performed a shuffling procedure that, in contrast to our previous shuffle, preserves the local correlations between spikes but reduces global spatial structure.
-This controls for dynamics that may have been induced by bursts or other local features of the spike train while randomizing the position to spike relationship.
-To do this, we randomized the order of runs (from one reward well to another) and then circularly permuted the resulting segments of data across all tetrodes uniformly.
-Using this shuffle, we found that there was still less spatially coherent and more spatially incoherent content compared to the real data (recording session \#1, spatially coherent: 99\% vs.
-71\%, p=0.02, real vs.
-shuffled; spatially incoherent: 22\% vs.
-61\%, p=0.02, real vs.
-shuffled; recording session \#2, spatially coherent: 95\% vs.
-63\%, p=0.02, real vs.
-shuffled; spatially incoherent: 16\% vs.
-79\%, p=0.02, real vs.
-shuffled; \FIGSUPP[3]{fig3sup3}).
+We find that only 14\% (3,353 of 23,382) of classified SWRs have time points with spatially incoherent structure (\FIG{3}G).
 
-To more directly compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
-Here our findings were consistent with previous reports: in our datasets, 4,449 of 23,071 or 19\% of classified SWRs included time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
-Thus, focusing on only high-speed, linear-fit trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
-We emphasize here that we did not limit our analysis to only those SWRs that had continuous dynamics for the entire SWR, as is assumed by line-fitting approach of the standard decoder.
-The consideration of this broader class of SWRs allows us to take advantage of the ability of our decoder to capture a range different speeds within each SWR event.
+To better compare our findings to previous work, we quantified the percentage of classified SWRs that contained continuous content, as would typically be analyzed when using the standard decoder.
+Here our findings were consistent with previous reports: in our datasets, 4,511 of 23,382 or 19\% of classified SWRs had time periods where the decoded position was classified as continuous (\FIG{3}G, 17\% of all SWRs).
+Thus, focusing on only high speed trajectories excludes a large fraction of events where there is evidence for spatially coherent, but not continuously changing, content.
 
 We repeated our classification analysis with a higher classification threshold of 0.95 to ensure that our result was not dependent on the threshold of 0.8.
-We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of all SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) included spatially coherent content.
-Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
+We find that, while this change slightly reduced the total fraction of classified SWRs (19,180 of 25,845 or 74\% of SWRs), an even higher fraction of the classified SWRs (19,020 of 19,180 or 99\% classified SWRs) had spatially coherent content.
+Similarly, SWRs containing spatially incoherent content were a small fraction of the classified SWRs (484 of 19,180 or 3\% classified SWRs).
 
@@
 To do so, we can compute the cumulative spatial size of the 95\% highest posterior density (HPD) region of the latent position estimate.
-The 95\% HPD region corresponds to the set of most probable latent positions as estimated by the model.
 Larger values of the HPD region size indicate the model is less certain about position, because the most probable decoded positions are distributed over more of the track at a given time point.
-In contrast, smaller values of the HPD region size indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
-Thus, the HPD region size provides a complementary measure of spatial information, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
+In contrast, smaller values indicate that the model is more certain about the estimate of position because the extent of the HPD is more concentrated and covers less of the track.
+Thus, the HPD region provides a complementary measure of spatial coherence, and evaluating it allows us to verify that the events we defined as spatially coherent also correspond to events where there is high certainty around the position estimates.
 
 We find that spatially coherent events indeed have smaller HPD regions than events with fragmented dynamics.
-\FIG{4}A and \FIG{4}B show two example SWRs that are classified as having continuous and stationary dynamics, respectively.
+\FIG{4}A and \FIG{4}D show two example SWRs that are classified as having stationary and continuous dynamics, respectively.
 The most probable positions (as estimated by the model) at each time step in these SWRs is concentrated in a small portion of the track and correspondingly, the HPD region is small throughout the SWRs.
@@
 The HPD region size also provides insights into the moment-by-moment structure of each event, which can change over the time course of a SWR.
-An example of this change is shown in \FIG{4}D, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
+An example of this change is shown in \FIG{4}B, where the HPD region is small for most of the SWR until the end, at which point the uncertainty of position becomes much higher, reflecting a transition from a spatially coherent to a spatially incoherent representation.
 Overall, when we examine the average HPD region size for each SWR, grouped by dynamic, we find a clear bimodal split between spatially coherent dynamics and spatially incoherent dynamics (\FIG{4}E).
@@
 median 238 cm, spatially incoherent, p=2.2e-16, one-sided Mann-Whitney-U).
-The HPD region for the unclassified SWRs was similarly large.
-
+Additionally, the HPD region for the unclassified SWRs was similarly large.
 We note here that while the size of the HPD region will be influenced by the dynamic estimated from the data, it remains possible for continuous or stationary dynamics to correspond to a large HPD region (see outliers in \FIG{4}E), indicating less spatial certainty for those specific events.
 In general, if the data does not exhibit strong place specific firing, the HPD region will be large, regardless of dynamic classification.
-To show this, we used the same shuffle as above, resampling position with replacement for two recording sessions and shuffling the relationship between spiking and spatial information when fitting the encoding model.
-The shuffled decodes have much larger HPD regions than the real data (recording session \#1: 55 cm vs.
+To show this, we resampled position with replacement for two recording sessions, shuffling the relationship between spiking and spatial information and then fitting the encoding model.
+The shuffled decodes have much larger HPD regions than the real data (epoch \#1: 55 cm vs.
 229 cm, p=0.02, real vs.
-shuffled; recording session \#2: 72 cm vs.
+shuffled; epoch \#2: 72 cm vs.
 231 cm, p=0.02, real vs.
-shuffled, one-sided; \FIGSUPP[4]{fig4sup1}).
-We also compared the HPD regions between the real data and a shuffle that swapped the runs and circularly shuffled position, as above.
-This also resulted in larger HPD regions than the real data for spatially coherent and incoherent SWRs (recording session \#1, spatially coherent: 29 cm vs.
-52 cm, p=0.02, real vs.
-shuffled; spatially incoherent: 181 cm vs.
-243 cm, p=0.02, real vs.
-shuffled; recording session \#2, spatially coherent: 26 cm vs.
-53 cm, p=0.02, real vs.
-shuffled; spatially incoherent: 210 cm vs.
-263 cm, p=0.02, real vs.
-shuffled; \FIGSUPP[4]{fig4sup2}).
+shuffled; one-sided Mann-Whitney-U, \FIGSUPP[4]{fig4sup1}.
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v7}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v6}
 \caption{Validation of classification using the 95\% Highest Posterior Density.
@@
 % Figure 4-Supplemental 1
-\figsupp[Shuffling the position with replacement increases the average 95\% HPD region size]{Average 95\% HPD region size increases on position shuffled data for two recording sessions from different animals.
-Red line represents the average 95\% HPD region size for all ripples.
-The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
-
-% Figure 4-Supplemental 2
-\figsupp[Shuffling the position data increases the average 95\% HPD region size for spatially coherent and incoherent classified times]{Average 95\% HPD region size increases for run-swapped, circularly shuffled position data compared to real data.
-Data from two recording sessions from different animals.
+\figsupp[Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data]{Comparison of average spatial coverage of the 95\% highest posterior density on real vs.
+position shuffled data for two epochs from different animals.
 Red line represents the average spatial position spanned by the highest posterior density on real data for all ripples.
 The histogram represents the distribution after 50 shuffles of the position data.
-The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
-This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
-
+Position data was shuffled by resampling with replacement from the set of all observed positions in that epoch, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v6}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v4}
 \caption{
@@
 \textbf{(A)} UpSet plot \citep{LexUpSetVisualizationIntersecting2014}---which is similar to a Venn diagram with more than 3 sets---of the most common sets of classifications within each SWR.
-Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at different times).
+Each row represents a classification and each column represents a set of classifications, where filled-in black dots with an edge between the dots indicates that multiple classifications are present in the SWR (at temporally distinct times).
 The sets are ordered by how often they occur as indicated by the bar plot above each category.
@@
 \textbf{(F)} Average tetrode multiunit spike rates for each dynamic category within each SWR (excluding classifications < 20 ms).
-\textbf{(G)} Probability density of animal movement speeds, illustrating prevalence of slower speed real-world movement consistent with stationary and stationary-continuous mixture replay events.
 }
@@
 % Figure 5-Supplemental 1
-\figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.
-Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
-
-% Figure 5-Supplemental 2
-\figsupp[Distribution of dynamics is insensitive to parameters]{Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}
-
-% Figure 5-Supplemental 3
-\figsupp[Further quantification of spiking and ripple properties by dynamic]{\textbf{(A)} Further quantification of the dynamics with respect to the spiking and ripple properties.
-\textbf{(B)} Speed and number of tetrodes with spikes for entire SWR }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5_v1_supplemental3}}\label{figsupp:fig5sup3}
-
-% Figure 5-video1
-\videosupp{Example of a stationary-continuous-mixture. Conventions the same as Figure 2-video1}\label{videosupp:fig5supv1}
+\figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
 
@@
 We therefore examined these events in more detail.
-We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) were classified as having only a single dynamic or dynamic mixture (\FIG{5}A, for specific example SWR see \FIG{3}B), whereas SWRs with multiple dynamics or dynamic mixtures (such as those in \FIG{3}A, C, D, E, and F) were less common.
-Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
-These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
-continuous 94 ms, \FIG{5}B).
-Nonetheless, both the slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 52 cm vs.
-continuous 43 cm, \FIG{5}C).
+We first note that most of the SWRs (16454 of 23382 or 70\% of classified SWRs) are well described by a single dynamic or mixtures of dynamics (\FIG{5}A).
+Surprisingly, the most common category is stationary-continuous mixtures (9860 of 23382 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
+These events contain representations that move at slower speeds (\FIG{5}D), and are slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs.
+continuous 95 ms, \FIG{5}B).
+Nonetheless, both these slow-moving events and continuous events frequently represent locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs.
+continuous 42 cm, \FIG{5}C).
 This indicates that the content of these SWRs, like those that are classified as continuous, do not represent the animal's actual position.
-We note that roughly a third of these events persisted for the entire duration of the SWR, but the other two thirds included some time where the model was uncertain about the dynamic.
-This shows the power of our state space decoder because it allows us identify periods of time when the model has high confidence with each SWR, and use these periods to characterize the SWR overall or to focus on specifically for further analysis.
 
-The second most prevalent classification was exclusively stationary events (4,858 of 21,433 or 23\% classified SWRs; 47\% of SWRs with stationary dynamics were entirely stationary).
+The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs).
 Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (\FIG{5}C).
 There were, however, a number of stationary events that represented positions far from the animal.
-We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in \texttildelow7\% of classified SWRs (1,437 of 21,433 classified SWRs).
-Of these, 46\% (664 of 1,437 stationary SWRs) were stationary throughout the duration of the SWR.
-These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{YuDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
+We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (\texttildelow7\%) of classified SWRs.
+About half of these (731 classified SWRs) were stationary throughout the duration of the SWR.
+These non-local stationary events most commonly represented reward wells or choice points (\FIG{5}E), consistent with \citep{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track.
 This suggests that these representations can be flexibly deployed to represent discrete locations.
 
-\subsection*{Control analyses}
-
-We then carried out a series of control analyses to determine whether our results could be due to differences in spiking statistics across dynamics or interneuron-specific activity patterns.
-We first calculated the multiunit spiking rates and found that these were similar across the different dynamics.
-(\FIG{5}F).
-This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
-
-We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007}).
-First, we increased our tetrode participation threshold from two up to five tetrodes active during the SWR, to eliminate the possibility that events with a single tetrode with a high rate interneuron might drive the classification of stationary or stationary-continuous mixture dynamics during SWRs.
-We found that our results were robust to this increase in threshold (\FIGSUPP[5]{fig5sup2}A).
-Indeed, most of our events and most periods of dynamics within each event included spikes from upwards of 10 tetrodes (\FIGSUPP[5]{fig5sup3}).
-Next, we removed spikes that had spike widths of less than 0.3 ms to reduce the potential contribution of narrow-waveform inhibitory interneurons \citep{FoxLocalizationanatomicalidentification1975}.
-Similarly, this had little effect: we still observed a high proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}B).
-Using clustered data, we can more robustly categorize units as putative interneurons; thus, we then repeated this analysis on clustered data for 9 of the 10 animals, excluding putative interneurons based on spike width and firing rate and requiring that at least three putative pyramidal cells be active during the SWR.
-Although there were many fewer events to examine, we still observed many stationary and stationary-continuous mixtures in SWRs (\FIGSUPP[5]{fig5sup2}D).
-Finally, if interneurons with less spatial specificity were driving the classification of slower dynamics, we would expect to see the posterior be less spatially concentrated during these events.
-Instead, we found that stationary and stationary-continuous mixtures tended to have narrow posteriors (as characterized by 95\% HPD region size in \FIG{4}E).
-This provides further evidence that our large fraction of slow-speed events were not the solely the result of the firing of interneurons.
-
-We then asked whether the slow spatially coherent events (stationary-continuous mixtures) potentially reflect a slow theta sweep rather than a replay event.
-We noted that this seems very unlikely given that the stationary-continuous mixtures tended to represent locations far from the animal that would require rapid, long-distance theta sequences (median 50 cm \FIG{5}C).
-In addition, the majority of our events occur when the animal is moving at speeds < 2 cm/s, as expected for SWRs (\FIGSUPP[5]{fig5sup3}B).
-Further, when we restrict the analyses to SWRs where the animal was moving at speeds < 1 cm/s, we find similar proportion of SWRs with stationary and stationary-continuous mixture dynamics (\FIGSUPP[5]{fig5sup2}C).
-
-A third possible concern is that burst firing from a single cell is driving the stationary dynamics.
-There are several lines of evidence against this.
-First, as mentioned above, most of our SWRs contained spikes from multiple tetrodes, including those with stationary dynamics (\FIGSUPP[5]{fig5sup3}A), so it is unlikely that a single cell could drive these dynamics.
-Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
-bursts which have a typical duration of 6-24 ms, \citep{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
-Third, only \texttildelow50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode, rather than single units.
-
-Finally, we wished to assess whether the spiking information from different hippocampal subfields (CA1, CA2, and CA3), could influence the types of dynamics we observed.
-This seemed unlikely, given that CA1 and CA3 spiking is tightly coordinated within and across hemispheres during SWR replay \citep{CarrTransientSlowGamma2012} and multiple previous papers have combined recordings across subfields and obtained results similar to those that restricted their analyses to CA1 \citep{DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009}.
-However, we verified that when we used only CA1 tetrodes to decode (\FIGSUPP[5]{fig5sup2}E), we get similar proportions of dynamics as when using all the tetrodes.
-
-\subsection*{Comparison to standard approaches}
-
-% Figure 6
-\begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_v2}
-\caption{
-Event speeds calculated using three common "Bayesian" decoder approaches (A-C) compared to using the state space model (D).
-For each panel, the top five rows show the probability density of the estimated speed for SWRs that contain that dynamic or combination of dynamics.
-Note that these categories are not mutually exclusive since a SWR can include more than one dynamic or combination of dynamics.
-The sixth row shows all SWRs containing unclassified dynamics.
-The final row is the estimated speed for all SWRs.
-\textbf{(A)} Radon transform  \textbf{(B)} MAP estimate  \textbf{(C)} Linear regression  \textbf{(D)} State space decoder presented in this manuscript.
-}
-\label{fig:6}
-\end{figure}
-
-Finally, we asked how our method compares to the standard approaches to identifying significant replay events via a non-parametric shuffle test and characterizing their content.
-This comparison is complicated due to the variety of parameters and inference methods used previously, so here we chose three commonly used approaches and a reasonable parameter set for the comparison.
-We began with an examination of events based on statistical significance of a line fit to a circularly permuted spatial distribution, which has been the standard in the field.
-We applied the clusterless algorithm in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
-We then used the radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
-We found that \texttildelow12\% of the classified events (as defined by our state-space decoder) were identified as significant.
-Not surprisingly, the fraction of significant events was strongly related to which dynamic they contained.
-Of the SWRs that contained a period of continuous dynamics, 31\% (1,386 of 4,511 SWRs) were significant.
-Of all SWRs that contained a period of stationary dynamics, 8\% (542 of 7,176 SWRs) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
-This increase is not surprising, since the shuffle used for the radon method penalizes shorter events.
-An alternative approach, the linear regression method, which samples from the posterior probability and uses linear regression to fit a line to the samples \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, detects only non-zero slopes as significant, and thus by definition cannot identify stationary events as significant.
-
-Conclusions about the representational content of events were similarly variable across methods.
-In addition to the radon and linear regression methods, we also computed  the MAP estimate, the maximum probability position of each 20 ms time bin \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
-We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
-We found that the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
-In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
-This is to be expected because the MAP method imposes less smoothing (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
-Our decoder imposes more smoothing than the MAP estimate but much less than the linear methods and therefore is closer to the MAP estimates.
-These results indicate that the methods that made more assumptions about the structure of replay (radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
-The substantial variability in both the proportions of "significant" events and the speed of the events demonstrates that key conclusions about the events could be very different depending on the specific method and parameters used.
+Finally, we examined the overall levels of neuronal spiking across dynamics.
+We found that all the dynamics had similar average multiunit spiking rates (\FIG{5}F).
+This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by sustained spiking information and could not be explained by the absence of spiking.
 
 \section*{Discussion}
-We developed a state space model that identifies and quantifies replay in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
+We developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented.
 This model is robust: it is relatively insensitive to a range of plausible transition parameters between dynamics and instead strongly reflects the hippocampal place representation structure.
-We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far more than have been previously associated with identifiable activity patterns.
-This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
-In fact, these slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical approaches that assume that events progress at constant speed.
+We show that this model is interpretable: it has well-defined parameters that capture intuitive movement models and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than has been previously captured.
+This model is also flexible: it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allows us to confidently discover not only SWRs with spatial representations that move more or less linearly at a constant speed, but also representations that vary in speed, have slower speeds, or that are stationary.
+These slower-moving representations constitute the majority of representations seen during SWRs, but would not be found with the typical standard decoder.
 The prevalence of these events indicates that hippocampal replay can recapitulate experience at both accelerated and closer to real-world speeds.
 
-Previous work reported that less than half of SWRs or high-population-activity events contained replay events \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work reported that less than half of SWRs or high-population-activity events contained replay \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
-Our results indicate that this is not the case: in our dataset, the large majority of SWRs (86\%) contained spatially coherent content, as defined by some combination of stationary and continuous dynamics.
-These events were mostly decoded representations where the most probable positions were highly spatially concentrated (i.e.
-had small HPD region size), indicating that they were driven by spatially specific firing.
-This was in contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
+Our results indicate that this is not the case: in our dataset, the large majority of SWRs (85\%) contained spatially coherent content, as defined by some combination of stationary or continuous dynamics.
+These events included decoded representations where the most probable positions were highly spatially concentrated (i.e.
+had small HPD size).
+This was in stark contrast to unclassified events or events that included fragmented dynamics, which were associated with highly distributed decoded spatial representations.
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats move through their environment.
-The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
-Events with slower dynamics tended to persist for approximately the same length of time as events  with faster continuous dynamics (\texttildelow50-100 ms), consistent with a brief activation of a representation of a location or a short trajectory through space.
-Events with slower dynamics most often represented the animal's current location, similar to continuous events identified in previous work \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}; however, slower events were also capable of representing locations far from the animal.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s.
+The next most common category were stationary events \citep{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
+Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space.
+Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
 
 These results challenge the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \citep{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009}.
-Instead, replay events encompass a much richer variety of event speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
+Instead, there is a much richer variety of speeds the animal can use to retrieve information about prior experience or generate possible future events, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
 Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
-Our results strongly suggest that they do not disappear, but are a very common occurrence with adult rats.
-We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events and using a restrictive decoding model which requires the decoded position to be stationary for the entire event.
-We also note that we chose to include a stationary dynamic to explicitly evaluate the prevalence of stationary events in adult animals, but in principle, a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
+Here, we show that they do not disappear, but are a very common occurrence with adult rats.
+We believe that the difference in these findings is a result of defining "chance" levels of occurrence of these events, using large time bins, and using a restrictive decoding model.
+We also note that we chose to include a stationary distribution to explicitly evaluate the prevalence of stationary events in adult animals, but in principle a model with just continuous and fragmented dynamics could capture these events as well, as the continuous dynamic includes movement speeds close to 0 m/s.
 
-The differences between our findings and previous approaches also highlight the importance of identifying the specific question that an analysis seeks to answer.
-Initial studies of replay \citep{LeeMemorySequentialExperience2002, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009} focused on the question of whether the observed patterns of spiking seen during replay reflected structured content or "random" spiking as defined by a comparison to one or more shuffles.
-These and subsequent studies answered the question "are replay events the result of random spiking?" with a clear "no".
-These studies did not, however, establish that the significance criteria used to show non-random structure has anything to do with the presence of spatial structure or events that are potentially useful for memory processes.
+We hypothesize that this large set of events supports the ability to rapidly retrieve representations of individual places and snippets of real-world experience.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{JaiDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to evaluate the value of specific locations or short movement trajectories without having to generate full trajectories to or through them \citep{YuHippocampalcorticalinteraction2015}.
 
-Our method aims to identify structure within events and to capture their complex dynamics to determine what sort of information replay events could provide in service of memory functions.
-We precisely define patterns of spatial representation that are, and are not, interpretable in terms of the structure of a given environment and the spatial activity seen during movement.
-In so doing, we find that most, but not all, events contain spiking that corresponds to a location or a trajectory through the animal's environment.
-
-The different questions addressed by our method compared to previous methods can also explain the apparent differences in findings.
-Our results showed that 19\% of SWRs contained continuous trajectories---meaning that portions of the SWR had spatially coherent trajectories that moved at high velocities---but only 4\% were continuous throughout the entire SWR.
-This, \textit{prima facie}, may seem at odds with previous results, which found between 5\% and 45\% of SWRs were "significant", meaning that they were well fit by a line compared to various shuffled versions of the dataset.
-Furthermore, these studies found trajectories on average move at much higher speeds (20x normal rat movement speed) than we report here.
-These previous findings were based largely on a single linear fits to the data for each event, while our model allows the trajectory to change speed during the SWR, which can lead to very different estimates of the speed of the event.
-Many of the events that had continuous dynamics, and would be classified as having high speed trajectories, also exhibited stationary-continuous mixture dynamics for some portion of the event(see for example \FIG{2}).
-Further, because the linear fit includes the whole SWR event, it is highly sensitive to the definition of the start and end of the SWR event, an often arbitrary definition which varies across research groups.
-In contrast, our model makes an estimate for each time point and only depends on the previous and future time step, which minimizes the dependence of the event on those boundaries and any possible "noise" activity within the SWR.
-This, along with our more permissive threshold for SWRs, allows us to characterize more SWRs than previous studies, rather than exclude them from analysis.
-
-Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that a subset of replay trajectories contained systematic jumps in position.
-We found that while replay speeds are, on average, distributed log normally, a small subset of the replays include definitive "jumps" from one position to another.
+Our results may also explain the seeming conflict between \cite{StellaHippocampalReactivationRandom2019}, who reported that hippocampal replay trajectories followed Brownian-like diffusion dynamics, and \cite{PfeifferAutoassociativedynamicsgeneration2015}, who reported that replay trajectories contained jumps in position.
+We found that while replay speeds are on average distributed log normally, there are still definitive "jumps" from one position to another in a small subset of the replays.
 Because our model makes it possible to identify different dynamics within the same SWR, we are able to identify both types of sequences.
 
-We hypothesize that this large set of events supports the ability to store and update representations of individual places and snippets of experience.
-Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
-Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to store or update the value of specific locations or short movement trajectories without having to generate full trajectories representing one of multiple ways to arrive at a location \citep{YuHippocampalcorticalinteraction2015}.
-Events that contain multiple dynamics (e.g.
-two continuous representations separated by a brief fragmented representation) might also help bind together experiences that occur in the same environment but at different places, whereas sustained spatially incoherent representations might represent coherent representations in other spatial environments \citep{KarlssonAwakereplayremote2009} or non-spatial experiences.
-Alternatively, downstream structures may be are attuned to the shifts in dynamics, and that each period of coherent hippocampal dynamics drives its own set of responses elsewhere in the brain.
-We further note that recent results from a study of activity patterns during movement indicates that individual theta cycles can have different dynamics as well \citep{WangAlternatingsequencesfuture2020a}, suggesting that the underlying sequences seen during replay and theta may overlap in their dynamics as well as their content.
-
-Similarly, it is possible that slow and fast dynamics are supported by different subcircuits within the hippocampal formation.
-\cite{YamamotoDirectMedialEntorhinal2017} found that blocking medial entorhinal cortex (MEC) output to CA1 resulted in shorter (and therefore slower) replay trajectories and \cite{OlivaOriginGammaFrequency2018} showed that layer 2/3 MEC neurons increased their firing 50 ms before the start of longer duration SWRs (>100 ms).
-This raises the possibility that these slower replays are driven by CA3-CA1 interaction while the faster trajectories are more externally driven by MEC-CA1, and that these different events could also engage different circuits outside of the hippocampus and entorhinal cortex.
-
-One key aspect to understanding how the hippocampus engages with other brain structures is identifying the times when the neural dynamics have changed.
-The times of SWRs or increases in multiunit activity  has provided an important, but coarse, landmark to this change in neural dynamics.
-Because our method allows for moment-by-moment characterization of the neural dynamic, we suggest that evaluating the start and end of each dynamic may prove a powerful way identify periods of interest.
-Examination of such periods may advance our understanding of when and why the hippocampus may be representing current position, past experience, or possible future locations, how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
+Additionally, as the start and end time of each SWR event are also detection method- and parameter-dependent, we suggest that instead, evaluating the start and end of each dynamic as an alternative landmark system may prove a powerful way identify periods of interest.
+Examination of such periods may shed light on when and why the hippocampus may be representing current position, past experience, or possible future locations, to understand how inputs to the hippocampus influence these events, and how these events may in turn influence downstream structures.
 
@@
 First, we sought to describe as many SWRs as possible, instead of setting a high threshold for event detection.
-This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \citep{YuDistincthippocampalcorticalmemory2017}.
+This more permissive approach was also critical for the identification of events that activate specific, immobility-associated locations in previous work \citep{JaiDistincthippocampalcorticalmemory2017}.
 More broadly, there is no evidence that SWR and high population activity events constitute clearly distinct classes of events that can be identified with a single fixed threshold.
-Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \citep{YuDistincthippocampalcorticalmemory2017}.
+Instead, the sizes of these events are well described by a unimodal, long-tailed distribution \citep{ChengNewExperiencesEnhance2008}.
 High thresholds can therefore exclude large numbers of real events and potentially lead to erroneous conclusions.
 
-Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenTransductiveneuraldecoding2012, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
-We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology could reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
+Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenBayesiannonparametricmethods2016, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology may reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
 
-Third, we built an explicit and flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
+Third, we built an explicit but flexible model that allowed us to identify not just one dynamic, but multiple dynamics consistent with different speeds of motion.
 We used these dynamics to describe the SWR events, rather than declaring the events as significant or not.
-
-Fourth, our model avoided using large temporal bins and does not make assumptions about the linear progression of events.
-Instead, the model allows for movement in any direction allowed by geometry of the environment, and further, still accounts for the uncertainty of the data.
-Because the model gives a moment-by-moment estimate of the position and dynamic, this also minimizes the effect of misidentifying the SWR start and end times.
-Combined with an acausal estimation of latent position, this approach allows us to identify the specific timing of transitions in dynamics at a fine timescale.
-In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
+Fourth, our model avoids large temporal bins and does not make assumptions about the linear progression of events.
+Instead, the model can use any size of bin and allows for movement in any direction.
+Finally, we used an acausal estimator of latent position, which allowed us to identify the specific timing of transitions in dynamics at a fine timescale.
 
@@
 Our model also represents a middle ground between Hidden Markov-style models \citep{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
-In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms).
+In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms)
 
 Code and algorithms for decoding hippocampal replay have not typically been made accessible or easy to use.
-This is problematic because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
+We believe this is not ideal because it can lead to severe variation or errors in code and parameters, limits reproducibility of results, and slows down scientific progress.
 Accordingly, we have made code for this model publicly available as a python package at the following URL \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}.
@@
 
-State-space models like the one we use here can enable a richer set of analyses of events that take advantage of all of the available data.
+State-space models like the one we use here can enable a richer set of analyses of generative events that take advantage of all of the available data.
 These approaches can be applied not just to SWRs but to all times, providing a moment-by-moment estimate of the nature of the spatial representation in the hippocampus, including important information about the spatial coherence of that representation.
 The model can be extended to incorporate other previously experienced environments by training place field models on those environments and including the appropriate movement transition matrices for those environments.
-It can also be extended to account for task conditions (such as the inbound and outbound conditions in our spatial alternation task) and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
+It can also be extended to account for trajectory task context and forward/reverse sequences as in \cite{DengRapidclassificationhippocampal2016}.
 Finally, models can be built to estimate any covariate, including movement direction \citep{KayConstantSubsecondCycling2020}.
@@
 Animals were free to choose any arm at any time.
-Only run recording sessions with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
+Only run epochs with at least 9 putative hippocampal pyramidal cells that fired at least 100 spikes were included in the analysis.
 
@@
 The code used for ripple detection can be found at \url{https://github.com/Eden-Kramer-Lab/ripple_detection}.
-We only analyzed SWRs with spikes from at least 2 tetrodes.
 
@@
 
+\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics}
+In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
+We fit the clusterless encoding model on the resampled data and then decoded the immobility periods.
+Like with the non-resampled data, we then extracted the SWR times and determined their classification based on our classification scheme.
+We repeated this shuffle analysis 50 times and compared this distribution to the real data for two epochs on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
+
 \subsection*{Highest Posterior Density}
 The 95\% highest posterior density (HPD) is a measurement of the spread of the posterior probability and is defined as the region of the posterior that contains the top 95\% of values of the posterior probability \citep{CasellaStatisticalinference2001}.
-By using only the top values, this measurement of spread is not influenced by multimodal distributions (whereas an alternative measure like the quantiles of the distribution would be).
-In this manuscript, we use the HPD region size---the total area of the track covered by the 95\% HPD region---to evaluate the uncertainty of the posterior probability of position.
-
-In order to calculate the HPD region size, we first calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
+In this manuscript, we use the HPD size to evaluate the uncertainty of the posterior probability of position.
+We calculate the 95\% HPD region by determining the maximum threshold value $h$ that fulfills the follow equality:
 $$
@@
 The 95\% HPD region is the set of position bins with posterior values greater than the threshold $\{x : p(x_{k} \mid O_{1:T}) > h\}$.
-The HPD region size is calculated by taking the integral of the members of this set:
+The HPD size is calculated by taking the integral of the members of this set:
 $$
@@
 
-\subsection*{Shuffle Analysis of the Effect of Place Encoding on Classification of Dynamics and HPD region size}
-In order to confirm that the model classification of dynamics depended on hippocampal place specific encoding, we resampled the position during movement with replacement, but preserved the spike times and spike waveform amplitudes.
-We fit the clusterless encoding model on the resampled data and then decoded the immobility periods.
-Like with the non-resampled data, we then extracted the SWR times and determined their classification based on our classification scheme.
-We repeated this shuffle analysis 50 times and compared this distribution to the real data for two recording sessions on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
-
-We also performed another shuffle that preserved more of the local correlation structure of spiking by shuffling the order of the well-to-well runs and then circularly shuffling the position within the runs.
-A well-to-well run starts when the animal moves more than 5 cm from a well and ends when the animal moves back within 5 cm of a well.
-This largely reflects runs from one well to another (e.g.
-from the center well to the left well), but for a small number of well-to-well runs, the animal did turn back to the same well (e.g.
-went back to the center well after leaving the center well).
-As above, after permuting the order of the well-to-well runs and circularly shuffling the position, we fit the clusterless encoding model and decoded SWRs.
-We repeated this shuffle analysis 50 times and compared this distribution to the real data for two recording sessions on two different animals (animal Bond, day 3, epoch 2 and animal Remy, day 35, epoch 2).
-
 \subsection*{Distance from animal}
-The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is the center of the position bin with the greatest posterior value.
+The distance of the decoded position from the animal's is defined as the shortest path distance along the track graph between the animal's 2D position projected on to the track graph (see Linearization) and the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
 The shortest path is found using Dijkstra's algorithm \citep{Dijkstranotetwoproblems1959} as implemented in NetworkX \citep{HagbergExploringNetworkStructure2008}.
@@
 \subsection*{Estimation of Speed}
-In order to estimate the speed over time, we first found the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is the center of the position bin with the greatest posterior value.
+In order to estimate the speed over time, we first found the MAP estimate of the posterior probability of position, $\argmax_{x_k} p(x_{k} \mid O_{1:T})$, which is defined as the center of the position bin with the greatest posterior value.
 We then computed the first derivative using the Python library Numpy's gradient function, which computes the difference in the forward and backward time direction and then averages them for all points except the boundaries.
@@
 
-\subsection*{Standard "Bayesian" Decoding and Significance Estimation}
-For the standard decoder analysis, we first fit a clusterless encoding model in the same manner as described in the subsection Encoding - Clusterless.
-We then decoded the posterior probability of each position, using 20 ms time bins, using the clusterless likelihood equation as described in the subsection The Model and assuming a uniform prior probability.
-Finally, to get an estimate of speed of the replay, we used either the radon transform method, the linear regression method, or the MAP estimate method.
-
-For the radon transform method, we densely sampled potential line trajectories and picked the line that maximized the probability along that line.
-The sum of the probability along the line is called the radon score.
-We picked the maximum radon score on lines fit on the combination of center arm and left arm posterior versus center arm and right arm posteriors.
-P-values were computed by circularly shuffling the posterior probabilities 1000 times and comparing the real data vs.
-the shuffled distribution.
-
-For the linear regression method, we drew 1000 position samples from each time bin according to the probability of position in that time bin.
-A linear regression was then fit to the sampled positions using time as a covariate.
-Fits were made separately for the posterior probabilities corresponding to the center arm with the left arm and the center arm with the right arm.
-Quality of the fit was assessed using the R-squared and the best fit between the arms was selected based on this metric.
-
-For the MAP estimate, we pick the position bin with the maximum value for each time bin, which corresponds to the most probable position.
-
 \subsection*{Software and Code Availability}
 Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}.
-Code for the decoder is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
-Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn \citep{Waskomseabornstatisticaldata2021}.
+Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification} (doi: 10.5281/zenodo.3713412).
+Both code bases rely on the following python packages: Numpy \citep{vanderWaltNumPyArrayStructure2011}, Numba \citep{LamNumbaLLVMbasedPython2015}, Matplotlib \citep{HunterMatplotlib2DGraphics2007}, xarray \citep{HoyerxarrayNDlabeled2017}, NetworkX \citep{HagbergExploringNetworkStructure2008}, Pandas \citep{McKinneyDataStructuresStatistical2010} and Seaborn (doi: 10.5281/zenodo.883859).
 All code is open-source and licensed under the MIT Software License.
-Decoder code can be easily installed as a python package with all requisite dependencies using pip or conda.
+Classifier code can be easily installed as a python package with all requisite dependencies using pip or conda.
 See software repositories for specific details.
@@
 \end{document}
-\label{fig:5}

### 2021-06-02 Article.tex
Update on Overleaf.

@@
 First, as mentioned above, the fragmented dynamic gives us a way to directly identify times when the position representation is spatially incoherent, a higher resolution and less assumption-dependent alternative to the more commonly used non-parametric shuffle.
-Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the radon transform).
+Second, this approach allows the model to to capture a wide range of movement speeds for the latent position in contrast to the standard decoder line fit (red dashed line in \FIG{1}E, using the Radon transform).
 The model is defined in terms of a mixture of movement dynamics, as summarized by the probability of each movement dynamic, and which dynamic or dynamics are dominant at a given moment is related to the temporal evolution of the underlying position representation.
@@
 We applied the clusterless algorithm in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
-We then used the radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
+We then used the Radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
 We found that \texttildelow12\% of the classified events (as defined by our state-space decoder) were identified as significant.
@@
 Of all SWRs that contained a period of stationary dynamics, 8\% (542 of 7,176 SWRs) were significant, and when we restricted the analysis to long stationary events (>100 ms duration), we found that 28\% were significant.
-This increase is not surprising, since the shuffle used for the radon method penalizes shorter events.
+This increase is not surprising, since the shuffle used for the Radon method penalizes shorter events.
 An alternative approach, the linear regression method, which samples from the posterior probability and uses linear regression to fit a line to the samples \citep{KarlssonAwakereplayremote2009, CarrHippocampalreplayawake2011, ShinDynamicsAwakeHippocampalPrefrontal2019}, detects only non-zero slopes as significant, and thus by definition cannot identify stationary events as significant.
@@
 Conclusions about the representational content of events were similarly variable across methods.
-In addition to the radon and linear regression methods, we also computed  the MAP estimate, the maximum probability position of each 20 ms time bin \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+In addition to the Radon and linear regression methods, we also computed  the MAP estimate, the maximum probability position of each 20 ms time bin \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
-We found that the methods that rely on fitting a line for the entire SWR (the radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
+We found that the methods that rely on fitting a line for the entire SWR (the Radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
 In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
@@
 Our decoder imposes more smoothing than the MAP estimate but much less than the linear methods and therefore is closer to the MAP estimates.
-These results indicate that the methods that made more assumptions about the structure of replay (radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
+These results indicate that the methods that made more assumptions about the structure of replay (Radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
 The substantial variability in both the proportions of "significant" events and the speed of the events demonstrates that key conclusions about the events could be very different depending on the specific method and parameters used.
@@
 We then decoded the posterior probability of each position, using 20 ms time bins, using the clusterless likelihood equation as described in the subsection The Model and assuming a uniform prior probability.
-Finally, to get an estimate of speed of the replay, we used either the radon transform method, the linear regression method, or the MAP estimate method.
-
-For the radon transform method, we densely sampled potential line trajectories and picked the line that maximized the probability along that line.
-The sum of the probability along the line is called the radon score.
-We picked the maximum radon score on lines fit on the combination of center arm and left arm posterior versus center arm and right arm posteriors.
+Finally, to get an estimate of speed of the replay, we used either the Radon transform method, the linear regression method, or the MAP estimate method.
+
+For the Radon transform method, we densely sampled potential line trajectories and picked the line that maximized the probability along that line.
+The sum of the probability along the line is called the Radon score.
+We picked the maximum Radon score on lines fit on the combination of center arm and left arm posterior versus center arm and right arm posteriors.
 P-values were computed by circularly shuffling the posterior probabilities 1000 times and comparing the real data vs.

### 2021-06-02 scripts/make_diff.sh
Script to make diff

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+latexdiff ../elife_orignal_draft.tex ../Article.tex > ../diff.tex

### 2021-06-03 Article.tex
Update on Overleaf.

@@
 \begin{figure}
-\includegraphics[width=0.75\linewidth]{figures/Figure1/Figure1_v7}
+\includegraphics[width=0.70\linewidth]{figures/Figure1/Figure1_final}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 \textbf{(B)} The probability of position over time for each diagonal value.
-Conventions the same as in (A).}{\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1_v3_supplemental1}}\label{figsupp:fig1sup1}
+Conventions the same as in (A).}{\includegraphics[width=1.0\linewidth]{figures/Figure1-supplemental1/Figure1-supplemental1_final}}\label{figsupp:fig1sup1}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1.0\linewidth]{figures/Figure2/Figure2_v5}
+\includegraphics[width=0.8\linewidth]{figures/Figure2/Figure2_final}
 \caption{The model can decode hippocampal replay trajectories using either sorted and clusterless spikes from the same SWR event.
@@
 The light grey lines represent the position of the animal over the entire recording session and the magenta dot represents the animal's position.
-\textbf{(B)} Same as in (A), but with clusterless decoding.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental1/Figure2_v3-supplemental1}}\label{figsupp:fig2sup1}
+\textbf{(B)} Same as in (A), but with clusterless decoding.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental1/Figure2-supplemental1_final}}\label{figsupp:fig2sup1}
 
@@
 Light grey lines indicate the animal's position over the entire recording session.
-Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2_v4-supplemental2}}\label{figsupp:fig2sup2}
+Magenta dashed line represents the animal's position during the SWR.}{\includegraphics[width=1.0\linewidth]{figures/Figure2-supplemental2/Figure2-supplemental2_final}}\label{figsupp:fig2sup2}
 
@@
 Each grey line represents a recording session for that animal.
-}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2_v1-supplemental3}}\label{figsupp:fig2sup3}
+}{\includegraphics[width=0.5\linewidth]{figures/Figure2-supplemental3/Figure2-supplemental3_final}}\label{figsupp:fig2sup3}
 
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_v5}
+\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_final}
 \caption{A-F.
@@
 More examples of SWRs with non-constant speed trajectories.
-Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure3-supplemental1/Figure3_v4_supplemental1}}\label{figsupp:fig3sup1}
+Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure3-supplemental1/Figure3-supplemental1_final}}\label{figsupp:fig3sup1}
 
@@
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3_v2-supplemental2}}\label{figsupp:fig3sup2}
+Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental2/Figure3-supplemental2_final}}\label{figsupp:fig3sup2}
 
@@
 The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
-This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3_v1_supplemental3}}\label{figsupp:fig3sup3}
+This preserves local spatial correlations in the data but breaks the global spatial relationships between the spikes and the data.}{\includegraphics[width=0.8\linewidth]{figures/Figure3-supplemental3/Figure3-supplemental3_final}}\label{figsupp:fig3sup3}
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_v7}
+\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_final}
 \caption{Validation of classification using the 95\% Highest Posterior Density.
@@
 The histogram represents the distribution after 50 shuffles of the position data.
-Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4_v1-supplemental1}}\label{figsupp:fig4sup1}
+Position data was shuffled by resampling with replacement from the set of all observed positions in that recording session, destroying position information but preserving spiking timing and position occupancy.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental1/Figure4-supplemental1_final}}\label{figsupp:fig4sup1}
 
@@
 The histogram represents the distribution after 50 shuffles of the run from well to well as well as the circularly shuffled position (for each tetrode).
-This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4_v1-supplemental2}}\label{figsupp:fig4sup2}
+This preserves local correlations between spikes but reduces global spatial structure.}{\includegraphics[width=0.8\linewidth]{figures/Figure4-supplemental2/Figure4-supplemental2_final}}\label{figsupp:fig4sup2}
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_v6}
+\includegraphics[width=0.80\linewidth]{figures/Figure5/Figure5_final}
 \caption{
@@
 \figsupp[More examples of stationary-continuous-mixtures.]{More examples of stationary-continuous-mixtures.
-Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5_v1_supplemental1}}\label{figsupp:fig5sup1}
+Conventions are the same as in \FIG{3}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental1/Figure5-supplemental1_final}}\label{figsupp:fig5sup1}
 
@@
 \figsupp[Control analyses for distribution of dynamics]{Control analyses for distribution of dynamics.
-Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5_v1_supplemental2}}\label{figsupp:fig5sup2}
+Conventions are the same as in \FIG{5}.}{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental2/Figure5-supplemental2_final}}\label{figsupp:fig5sup2}
 
@@
 \figsupp[Further quantification of spiking and ripple properties by dynamic]{\textbf{(A)} Further quantification of the dynamics with respect to the spiking and ripple properties.
-\textbf{(B)} Speed and number of tetrodes with spikes for entire SWR }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5_v1_supplemental3}}\label{figsupp:fig5sup3}
+\textbf{(B)} Speed and number of tetrodes with spikes for entire SWR }{\includegraphics[width=1.0\linewidth]{figures/Figure5-supplemental3/Figure5-supplemental3_final}}\label{figsupp:fig5sup3}
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_v2}
+\includegraphics[width=0.80\linewidth]{figures/Figure6/Figure6_final}
 \caption{

### 2021-07-12 src/analysis.py
Allow loading of multiunit HSE data

@@ -582,6 +582,7 @@ def load_all_replay_info(
     probability_threshold=PROBABILITY_THRESHOLD,
     speed_threshold=4,
     exclude_interneuron_spikes=False,
+    use_multiunit_HSE=False,
     brain_areas=None,
 ):
     tetrode_info = make_tetrode_dataframe(ANIMALS)
@@ -596,6 +597,9 @@ def load_all_replay_info(
         epoch_identifier += f'_{area_str}'
     else:
         brain_areas = _BRAIN_AREAS
+
+    if use_multiunit_HSE:
+        epoch_identifier += '_multiunit_HSE'
 
     file_regex = f"{epoch_identifier}_replay_info_{prob:02d}.csv"
     file_paths = glob(os.path.join(PROCESSED_DATA_DIR, file_regex))

### 2021-07-13 Article.tex

@@
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019}.
 Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
@@
 \begin{figure}
-\includegraphics[width=0.70\linewidth]{figures/Figure1/Figure1_final}
+\includegraphics[width=0.65\linewidth]{figures/Figure1/Figure1_final}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
-Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
+Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (7 cm median difference, 5-9 cm 95\% CI, for all times where the animal was moving greater than 4 cm/s, 5 fold cross validation).
 This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}.
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{figures/Figure3/Figure3_final}
+\includegraphics[width=0.70\linewidth]{figures/Figure3/Figure3_final}
 \caption{A-F.
@@
 \label{fig:3}
+
+\figdata{Table of replay statistics for each SWR for Figure 3G.}
 
@@
 \label{fig:4}
+
+\figdata{Table of replay statistics for each SWR for Figure 4E.}
 
@@
 
+\figdata{Table of replay statistics for each SWR for Figure 5.}
+
 % Figure 5-Supplemental 1
@@
 \label{fig:6}
+
+\figdata{Table of replay statistics for each SWR for Figure 6.}
 \end{figure}
@@
 
-\section{Contributions}
-E.L.D.
-conceived of the model and did the analysis with help from U.T.E, A.K.G., and M.E.C.
-E.L.D and L.M.F wrote the manuscript with help from A.K.G.
-and M.E.C.
-A.K.G, M.S., J.E.C.
-contributed data to this manuscript.
-
 \bibliography{refs}

### 2021-07-19 src/figure6_supplemental1.py
Handle one ripple number

@@ -380,6 +380,8 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
 
 
 def plot_figure(epoch_key, ripple_numbers):
+    if isinstance(ripple_numbers, int):
+        ripple_numbers = [ripple_numbers]
 
     set_figure_defaults()
     (clusterless_state_space_results, sorted_spikes_state_space_results,

### 2021-07-20 Article.tex

@@
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
+Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics \FIGSUPP[5]{fig5sup3}F).
 
@@
 In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
+Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \cite{}.
 

### 2021-07-23 Article.tex
Update on Overleaf.

@@
 Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
-These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
+These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G, median 17 cm/s for run periods and 4 cm/s for all times)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
 continuous 94 ms, \FIG{5}B).
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speed that rats move through their environment.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speeds that rats experience their environment.
 The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 Events with slower dynamics most often represented the animal's current location, similar to continuous events identified in previous work \citep{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}; however, slower events were also capable of representing locations far from the animal.
+Interestingly, Louie and Wilson \citeyearpar{LouieTemporallyStructuredReplay2001} found that during REM sleep, the longer time scale population activity (on the order of seconds) is highly correlated with firing during times when the animal is running, resembling the animal's experience.
+Our results are similar but on the timescale of SWRs, which last 10s to 100s of milliseconds, and during awake behavior.
 
@@
 In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
-Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \cite{}.
+Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \cite{DragoiDistinctpreplaymultiple2013}.
 
@@
 We used a threshold of 0.80 to classify the probability of each state $Pr(I_{k} \mid O_{1:T})$ into 5 categories.
-Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.80.
+Time periods during sharp wave ripples are labeled as stationary, continuous, or fragmented when the probability of each individual state is above 0.80, regardless of duration.
 Time periods are labeled as stationary-continuous-mixture or fragmented-continuous-mixture when the sum of stationary and continuous or fragmented and continuous are above 0.80, respectively.

### 2021-07-24 src/figure6_supplemental1.py
Code for figure6-supplement1

@@ -9,7 +9,7 @@ import xarray as xr
 from replay_trajectory_classification import SortedSpikesClassifier
 from src.analysis import (get_is_classified, get_place_field_max,
                           get_probability)
-from src.figure_utilities import ONE_COLUMN, PAGE_HEIGHT, set_figure_defaults
+from src.figure_utilities import PAGE_HEIGHT, TWO_COLUMN, set_figure_defaults
 from src.parameters import (PROBABILITY_THRESHOLD, PROCESSED_DATA_DIR,
                             STATE_COLORS)
 from src.standard_decoder import (fit_mark_likelihood, load_data,
@@ -231,7 +231,7 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
     MILLISECONDS_TO_SECONDS = 1000
 
     fig, axes = plt.subplots(
-        6, 1, figsize=(ONE_COLUMN, PAGE_HEIGHT / 2),
+        6, 1, figsize=(TWO_COLUMN / 3, PAGE_HEIGHT / 2),
         constrained_layout=True,
         sharex=False, sharey=False,
         gridspec_kw={"height_ratios": [2, 3, 2, 3, 3, 1]},)
@@ -258,7 +258,7 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
            robust=True))
     axes[1].set_xlabel('')
     axes[1].set_title(
-        'Sorted Spikes Standard Decoder, 20 ms bins', fontsize=8)
+        'Sorted Spikes Standard Decoder, 20 ms bins', fontsize=7)
     axes[1].set_ylabel('Pos. [cm]')
     axes[1].plot(sorted_spikes_posterior.time *
                  MILLISECONDS_TO_SECONDS, sorted_spikes_radon_prediction,
@@ -297,7 +297,7 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
      .plot(x='time', y='position', ax=axes[3], add_colorbar=False, cmap=cmap,
            robust=True))
     axes[3].set_xlabel('')
-    axes[3].set_title('Clusterless Standard Decoder, 20 ms bins', fontsize=8)
+    axes[3].set_title('Clusterless Standard Decoder, 20 ms bins', fontsize=7)
     axes[3].set_ylabel('Pos. [cm]')
     axes[3].plot(clusterless_posterior.time *
                  MILLISECONDS_TO_SECONDS, clusterless_radon_prediction,
[…]

### 2021-07-24 src/figure6_supplemental1.py
Add save figure code

@@ -9,9 +9,10 @@ import xarray as xr
 from replay_trajectory_classification import SortedSpikesClassifier
 from src.analysis import (get_is_classified, get_place_field_max,
                           get_probability)
-from src.figure_utilities import PAGE_HEIGHT, TWO_COLUMN, set_figure_defaults
-from src.parameters import (PROBABILITY_THRESHOLD, PROCESSED_DATA_DIR,
-                            STATE_COLORS)
+from src.figure_utilities import (PAGE_HEIGHT, TWO_COLUMN, save_figure,
+                                  set_figure_defaults)
+from src.parameters import (FIGURE_DIR, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR, STATE_COLORS)
 from src.standard_decoder import (fit_mark_likelihood, load_data,
                                   normalize_to_posterior,
                                   predict_clusterless_wtrack,
@@ -218,7 +219,8 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
                     clusterless_posterior, sorted_spikes_radon_prediction,
                     sorted_spikes_map_prediction, clusterless_radon_prediction,
                     clusterless_map_prediction,
-                    clusterless_state_space_results, multiunit_spike_times):
+                    clusterless_state_space_results, multiunit_spike_times,
+                    is_save_figure=False):
     cmap = copy.copy(plt.cm.get_cmap('bone_r'))
     cmap.set_bad(color="lightgrey", alpha=1.0)
 
@@ -374,8 +376,20 @@ def plot_posteriors(ripple_number, start_time, end_time, position_info,
 
     axes[-1].set_xlabel('Time [ms]')
 
+    # Save Plot
+    if is_save_figure:
+        figure_dir = os.path.join(FIGURE_DIR, "Figure6-supplemental1")
+        os.makedirs(figure_dir, exist_ok=True)
+        animal, day, epoch = epoch_key
+        fig_name = (
+            "figure6_supplemental1_"
+            f"{animal}_{day:02d}_{epoch:02d}_{ripple_number:04d}_"
+            f"decoder_comparison"
+        )
+        save_figure(os.path.join(figure_dir, fig_name))
 
-def plot_figure(epoch_key, ripple_numbers):
+
+def plot_figure(epoch_key, ripple_numbers, is_save_figure=False):
     if isinstance(ripple_numbers, int):
         ripple_numbers = [ripple_numbers]
 
[…]

### 2021-07-26 Article.tex
Update on Overleaf.

@@
 Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.
+We do not know if downstream neurons can explicitly responds to these dynamics based on either threshold (there is probably no hard boundary between these dynamics), but the dynamics correspond to fundamentally different sets of speeds and can be used as a convenient tool in order to understand and summarize the SWR.
 
@@
 Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (490 of 19,478 or 3\% classified SWRs).
+This shows that
 
@@
 
+% Figure Data
 \figdata{Table of replay statistics for each SWR for Figure 6.}
+
+% Figure 6-Supplemental 1
+\figsupp[Examples of fits from the standard "Bayesian" decoder and state space model.]{Examples of fits from the standard "Bayesian" decoder and state space model.
+(A-F) For each panel, the top row is a raster plot of the sorted cells, the second row is the decoded posterior probability of position in 20 ms time bins for the sorted spikes and the corresponding Radon line fit (blue line) and the MAP fit (green line) on this posterior probability, the third row is a raster plot of the multiunit spike times used for clusterless decoding, the fourth row is the posterior probability in 20 ms time bins using clusterless data as well as the accompanying Radon fit and MAP fit, the fifth row is the posterior probability from the clusterless state space model, and the sixth row is the probability of each dynamic (lines) and the color shading indicates the category of dynamic as before.
+}{\includegraphics[width=1.0\linewidth]{figures/Figure6-supplemental1/Figure6-supplemental1_final}}\label{figsupp:fig6sup1}
+
 \end{figure}
@@
 We applied the clusterless algorithm in 20 ms bins to produce probability densities of represented locations, as this binning is comparable to previous methods.
-We then used the Radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event.
-We found that \texttildelow12\% of the classified events (as defined by our state-space decoder) were identified as significant.
+We then used the Radon transform method, which finds the line with the highest summed posterior probability of position along the line \citep{DavidsonHippocampalReplayExtended2009, FarooqEmergencepreconfiguredplastic2019}, to calculate, for each SWR, a measure of significance based on the assumption that the probability density moves at a constant speed throughout the event (see examples in \FIGSUPP[6]{fig6sup1}A-F, fourth row, blue line).
+We found that \texttildelow12\% of the classified events (as defined by our state-space decoder) were identified as significant by this Radon method.
 Not surprisingly, the fraction of significant events was strongly related to which dynamic they contained.
@@
 Conclusions about the representational content of events were similarly variable across methods.
-In addition to the Radon and linear regression methods, we also computed  the MAP estimate, the maximum probability position of each 20 ms time bin \citep{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}.
+In addition to the Radon and linear regression methods, we also computed the MAP estimate, the maximum probability position of each 20 ms time bin (see examples in \FIGSUPP[6]{fig6sup1}A-F, fourth row, green line, \cite{PfeifferAutoassociativedynamicsgeneration2015, StellaHippocampalReactivationRandom2019, KaeferReplayBehavioralSequences2020}).
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
 We found that the methods that rely on fitting a line for the entire SWR (the Radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
-In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by our decoder (\FIG{6}D).
-This is to be expected because the MAP method imposes less smoothing (only the 20 ms time bins) whereas the linear methods impose a larger amount of smoothing.
-Our decoder imposes more smoothing than the MAP estimate but much less than the linear methods and therefore is closer to the MAP estimates.
+In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by the state space decoder (\FIG{6}D).
+This is to be expected because the MAP method imposes less smoothing (although some is added by using 20 ms time bins) whereas the linear methods impose a larger amount of smoothing and cannot change directions (see \FIGSUPP[6]{fig6sup1}A for an example where this is beneficial because the trajectory is linear and the smoothing ignores a lack of spiking and \FIGSUPP[6]{fig6sup1}B for an example where this is not beneficial because the linear fit cannot change direction).
+Our state space decoder imposes much less smoothing than the linear methods and therefore is closer to the MAP estimates (\FIGSUPP[6]{fig6sup1}B), but also takes into account the uncertainty of each time bin (\FIGSUPP[6]{fig6sup1}C).
 These results indicate that the methods that made more assumptions about the structure of replay (Radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).

### 2021-07-28 elife_revision1.tex
Add diff

--- 2021-07-20 Article.tex
@@
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{StellaHippocampalReactivationRandom2019}.
 Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
@@
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
-Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (7 cm median difference, 5-9 cm 95\% CI, for all times where the animal was moving greater than 4 cm/s, 5 fold cross validation).
+Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (4 cm median difference, 3-6 cm 95\% CI, for all time in sessions, 5 fold cross validation).
 This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}.
@@
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
-Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics \FIGSUPP[5]{fig5sup3}F).
 
@@
 In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
-Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \cite{}.
 

### 2021-07-28 scripts/make_diff.sh

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-latexdiff ../elife_orignal_draft.tex ../Article.tex > ../diff.tex
+latexdiff ../elife_revision1.tex ../Article.tex > ../diff.tex

### 2021-07-30 Article.tex
Revert "Revert "Update on Overleaf.""
This reverts commit 7bfb406708996248245fb81365e047950d24c3e1.

@@
 These memory processes depend critically on the hippocampus, where rapid plasticity during an experience is thought to drive the initial encoding of events \citep{EichenbaumConditioningConsciousRecollection2004}.
-Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking is thought to contribute to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, Joohippocampalsharpwave2018, CarrHippocampalreplayawake2011}.
-
-The canonical example of "replay" is seen in the rodent hippocampus, where cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
+Subsequently, hippocampal "replay" of stored representations during both slow-wave sleep and periods of immobility during waking is thought to contribute to the longer term storage and updating of event memories in distributed hippocampal-neocortical circuits \citep{Franklandorganizationrecentremote2005, CarrHippocampalreplayawake2011, Joohippocampalsharpwave2018}.
+
+The canonical example of "replay" is seen in the rodent, where hippocampal cells preferentially fire at specific locations in an environment, and thus ensembles of cells fire in sequence as the animal moves through a series of locations.
 When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event.
@@
 
-While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{MichonPostlearningHippocampalReplay2019, KaeferReplayBehavioralSequences2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DavidsonHippocampalReplayExtended2009, Tingleymethodsreactivationreplay2020}.
+While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow5-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay \citep{DavidsonHippocampalReplayExtended2009, MichonPostlearningHippocampalReplay2019,
+ShinDynamicsAwakeHippocampalPrefrontal2019, KaeferReplayBehavioralSequences2020, Tingleymethodsreactivationreplay2020}.
 One might therefore conclude that all memories that are stored or updated by SWRs correspond to sequences of locations.
@@
 Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019}.
+Other studies have reported activity that, when combined across an entire event, seems to correspond roughly to a single location, although the dynamics of these events is typically not investigated (REF Dupret, Csicvari papers with cheeseboard task).
 Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
@@
 Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter.
-Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, KarlssonAwakereplayremote2009, TingleyRoutingHippocampalRipples2020, BhattaraiDistincteffectsreward2020, ShinDynamicsAwakeHippocampalPrefrontal2019, DrieuNestedsequenceshippocampal2018, OlafsdottirTaskDemandsPredict2017, TangHippocampalPrefrontalReactivationLearning2017, CarrTransientSlowGamma2012, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
+Finally, the most commonly used approaches for detecting sequential replay involve fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant speed \citep{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007a, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009, CarrTransientSlowGamma2012, OlafsdottirTaskDemandsPredict2017, TangHippocampalPrefrontalReactivationLearning2017, DrieuNestedsequenceshippocampal2018,  ShinDynamicsAwakeHippocampalPrefrontal2019, TingleyRoutingHippocampalRipples2020, BhattaraiDistincteffectsreward2020}.
 A statistical test is then used to determine whether the line fit to the data is better than a line fit to shuffled versions of the data, where the shuffled version of the data represent an implicit definition of a "random" sequence.
@@
 
-Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{KaeferReplayBehavioralSequences2020, CareyRewardrevaluationbiases2019, GrosmarkDiversityneuralfiring2016, WuHippocampalReplayCaptures2014, PfeifferHippocampalplacecellsequences2013}.
+Recognizing the problems with the linear fit, several studies have moved away from the constant velocity assumption, using the most probable location at each time bin \citep{PfeifferHippocampalplacecellsequences2013, WuHippocampalReplayCaptures2014, GrosmarkDiversityneuralfiring2016, CareyRewardrevaluationbiases2019,  KaeferReplayBehavioralSequences2020}.
 For example, using this approach, \cite{PfeifferAutoassociativedynamicsgeneration2015} found awake replay events containing alternation between representations of a single location and sequential spatial trajectories.
@@
 It would use very small temporal bins (1 or 2 ms) to allow for very rapid representational movement and would provide information about the certainty of the decoded estimates.
-It would be able to capture a range of movement dynamics from stationary or unchanging representations to trajectories that progress through space at constant or variable speeds to disorganized, spatially incoherent events.
+It would be able to capture a range of movement dynamics including stationary or unchanging representations, trajectories that progress through space at constant or variable speeds, and disorganized, spatially incoherent events.
 It would provide a robust statistical assessment of confidence for each dynamic.
@@
 
-We therefore developed a state space model that achieves all of those goals.
+We developed a state space model that achieves all of those goals.
 State space models are a well-understood, well-known statistical solution to the problems described above.
 By mathematically modeling the relationship between the data and latent dynamics, state space models make the assumptions of the model explicit and interpretable.
-Our model goes beyond previous approaches \citep{MaboudiUncoveringtemporalstructure2018, DengRapidclassificationhippocampal2016} by characterizing the spatial representations during SWRs as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
+Our model goes beyond previous approaches \citep{DengRapidclassificationhippocampal2016, MaboudiUncoveringtemporalstructure2018} by characterizing the spatial representations during SWRs as a mixture of three underlying patterns of movement dynamics: stationary trajectories, continuous trajectories that can progress at many times the typical speed of the animal, and spatially fragmented trajectories.
 We show how this model can take advantage of clusterless decoding---which relates multiunit spike waveform features to position without spike sorting---giving us more information about the population spiking activity.
@@
 \begin{figure}
-\includegraphics[width=0.65\linewidth]{figures/Figure1/Figure1_final}
+\includegraphics[width=0.95\linewidth]{figures/Figure1/Figure1_final}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.
-We do not know if downstream neurons can explicitly responds to these dynamics based on either threshold (there is probably no hard boundary between these dynamics), but the dynamics correspond to fundamentally different sets of speeds and can be used as a convenient tool in order to understand and summarize the SWR.
+We do not know if downstream neurons can explicitly responds to these dynamics based on either threshold, and there is probably no hard boundary between these dynamics, but our approach makes it possible to ask that question in a systematic way.
 
@@
 We therefore adopted a "clusterless" approach which directly uses multiunit spikes and their spike waveform features to decode position without spike sorting (see Methods).
-Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{KloostermanBayesiandecodingusing2014, ChenTransductiveneuraldecoding2012,DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
+Clusterless decoding has previously been used to successfully identify theta sequences and replay sequences in the hippocampus \citep{ChenTransductiveneuraldecoding2012, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016, KayConstantSubsecondCycling2020}.
 Applying a clusterless decoder to the same SWR event, we get similar classification of the sequence (\FIG{2}B, D), both with 1D linearized position and 2D position (\FIGSUPP[2]{fig2sup1}B).
@@
 \subsection*{Most replays contain coherent spatial content}
-After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, CarrTransientSlowGamma2012}; position linearized to 1D).
+After testing our model on a single SWR event, we applied our clusterless decoding algorithm to hippocampal recordings from 10 rats performing the W-track spatial alternation task (\# tetrodes range: [10, 24], brain areas = [CA1, CA2, CA3]; some data previously used in \cite{KarlssonAwakereplayremote2009, CarrTransientSlowGamma2012, KayConstantSubsecondCycling2020}; position linearized to 1D).
 We first confirmed that our cell population for each animal could closely track the position of the animal during behavior by comparing the most probable decoded position to the position of the animal.
-Even including theta sequences and replay sequences, which can represent positions other than where the animal is, we found that the average median difference between actual and decoded location is small (7 cm median difference, 5-9 cm 95\% CI, for all times where the animal was moving greater than 4 cm/s, 5 fold cross validation).
-This level of accuracy is just above the size of our 3 cm position bins, and is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}.
+We found that the average median difference between actual and decoded location is small (7 cm median difference, 5-9 cm 95\% CI, for all times where the animal was moving greater than 4 cm/s, 5 fold cross validation).
+This level of accuracy is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}, and reflects in part the presence of "theta sequences" [REF] where position deviates behind and ahead of the animal across individual theta cycles.
 Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during movement, demonstrating the power of our algorithm.
@@
 \begin{figure}
-\includegraphics[width=0.70\linewidth]{figures/Figure3/Figure3_final}
+\includegraphics[width=0.95\linewidth]{figures/Figure3/Figure3_final}
 \caption{A-F.
@@
 
-Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work focusing on spatially sequential replay reported that only a minority of events contain sequential spatial content \citep{FosterReversereplaybehavioural2006, KarlssonAwakereplayremote2009, DavidsonHippocampalReplayExtended2009}.
 We therefore asked what fraction of classified events contain spatially coherent content, which we define as any times with stationary, stationary-continuous mixture, or continuous dynamics (see Methods and \FIG{1}F).
@@
 Similarly, SWRs containing spatially incoherent content comprised a small fraction of the classified SWRs (490 of 19,478 or 3\% classified SWRs).
-This shows that
 
@@
 \begin{figure}
-\includegraphics[width=0.80\linewidth]{figures/Figure4/Figure4_final}
+\includegraphics[width=0.95\linewidth]{figures/Figure4/Figure4_final}
 \caption{Validation of classification using the 95\% Highest Posterior Density.
@@
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
-Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics \FIGSUPP[5]{fig5sup3}F).
-
-We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{HangyaComplementaryspatialfiring2010, WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007}).
+Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs to identify events, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics \FIGSUPP[5]{fig5sup3}F).
+
+We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007, HangyaComplementaryspatialfiring2010}).
 First, we increased our tetrode participation threshold from two up to five tetrodes active during the SWR, to eliminate the possibility that events with a single tetrode with a high rate interneuron might drive the classification of stationary or stationary-continuous mixture dynamics during SWRs.
@@
 These results indicate that the methods that made more assumptions about the structure of replay (Radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
-The substantial variability in both the proportions of "significant" events and the speed of the events demonstrates that key conclusions about the events could be very different depending on the specific method and parameters used.
+The substantial variability in both the proportions of "significant" events and the speed of the events demonstrates that key conclusions about the events could be very different depending on the which of the previous methods and parameter were used.
 
@@
 
-Previous work reported that less than half of SWRs or high-population-activity events contained replay events \citep{KarlssonAwakereplayremote2009, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009}.
+Previous work reported that less than half of SWRs or high-population-activity events contained replay events \citep{FosterReversereplaybehavioural2006, KarlssonAwakereplayremote2009, DavidsonHippocampalReplayExtended2009}.
 This could lead one to the assumption that most events are not spatially coherent or meaningful.
@@
 
-Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speeds that rats experience their environment.
+Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speeds at which rats traverse their environment.
 The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
@@
 Interestingly, Louie and Wilson \citeyearpar{LouieTemporallyStructuredReplay2001} found that during REM sleep, the longer time scale population activity (on the order of seconds) is highly correlated with firing during times when the animal is running, resembling the animal's experience.
-Our results are similar but on the timescale of SWRs, which last 10s to 100s of milliseconds, and during awake behavior.
+Our results are similar but were observed on the timescale of SWRs, which last 10s to 100s of milliseconds, and during pauses in active behavior.
 
@@
 Instead, replay events encompass a much richer variety of event speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
-Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
+This is consistent with observations of events that, when decoded in a single time bin, represent a small neighborhood in the environment (CITE Dupret, etc.) Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
 Our results strongly suggest that they do not disappear, but are a very common occurrence with adult rats.
@@
 
-The differences between our findings and previous approaches also highlight the importance of identifying the specific question that an analysis seeks to answer.
+The differences between our findings and those reported in some previous papers also highlight the importance of identifying the specific question that an analysis seeks to answer.
 Initial studies of replay \citep{LeeMemorySequentialExperience2002, FosterReversereplaybehavioural2006, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009} focused on the question of whether the observed patterns of spiking seen during replay reflected structured content or "random" spiking as defined by a comparison to one or more shuffles.
@@
 We hypothesize that this large set of events supports the ability to store and update representations of individual places and snippets of experience.
-Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{YuDistincthippocampalcorticalmemory2017, JadhavCoordinatedExcitationInhibition2016}.
+Previous findings demonstrated that SWRs with movement- or immobility-related activity engage specific patterns of spiking activity outside the hippocampus \citep{JadhavCoordinatedExcitationInhibition2016, YuDistincthippocampalcorticalmemory2017}.
 Thus, we would expect that the large set of events with slow dynamics would also engage specific patterns of activity outside the hippocampus, perhaps allowing animals to store or update the value of specific locations or short movement trajectories without having to generate full trajectories representing one of multiple ways to arrive at a location \citep{YuHippocampalcorticalinteraction2015}.
@@
 Our model goes beyond the Johnson model in that it explicitly permits a mixture between the movement speeds, can work for arbitrary track geometries, and uses clusterless decoding.
-Our model also represents a middle ground between Hidden Markov-style models \citep{MaboudiUncoveringtemporalstructure2018, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, ChenUncoveringspatialtopology2012}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
+Our model also represents a middle ground between Hidden Markov-style models \citep{ChenUncoveringspatialtopology2012, ChenBayesiannonparametricmethods2016, LindermanBayesiannonparametricapproach2016, MaboudiUncoveringtemporalstructure2018}, which seek to be environment-agnostic detectors of sequential patterns, and the typical standard decoder, which typically use arbitrarily chosen bin sizes and restrictive assumptions about the nature of the trajectories.
 In particular, our approach allows for a variety of dynamics while still yielding spatially interpretable results and makes it possible to use bin sizes of any size (here 2 ms).
@@
 Ten male Long Evans rats (500-700 g, 4-9 months old) were trained on a W-track spatial alternation task.
-9 rats contributed to previous studies \citep{KarlssonAwakereplayremote2009, KayConstantSubsecondCycling2020, Kayhippocampalnetworkspatial2016, CarrTransientSlowGamma2012}.
+9 rats contributed to previous studies \citep{KarlssonAwakereplayremote2009, CarrTransientSlowGamma2012, Kayhippocampalnetworkspatial2016, KayConstantSubsecondCycling2020}.
 Neural activity was recorded in CA1, CA2, CA3, MEC, Subiculum, and DG depending on the rat.
@@
 
-Likewise, it has been shown \citep{ChenTransductiveneuraldecoding2012, DengRapidclassificationhippocampal2016, KloostermanBayesiandecodingusing2014} that the clusterless likelihood can be computed as:
+Likewise, it has been shown \citep{ChenTransductiveneuraldecoding2012, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016} that the clusterless likelihood can be computed as:
 $$

### 2021-07-30 Article.tex
Update on Overleaf.

@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{figures/Figure1/Figure1_final}
+\includegraphics[width=0.65\linewidth]{figures/Figure1/Figure1_final}
 \caption{The model can capture different sequence dynamics on simulated data.
@@
 We note here that the choice of any particular threshold for classifying the movement dynamic of a SWR is arbitrary, and that the power of our approach lies in part on the ability to assign a probability for each dynamic or combinations of dynamics to each moment in time.
-Our goal in choosing 0.80 was to use a threshold that corresponds to highly probable events and to explicitly avoid the arbitrary choice of 0.95, as the 0.95 threshold often leads to the conclusion that only events that are statistically significant at the 0.05 level are meaningful in the brain.
+Our goal in choosing 0.80 was to use a threshold that corresponds to highly probable events that roughly partition the latent position trajectory into interpretable categories of speed.
 Nonetheless, we also verify that our central results hold with a higher threshold of 0.95.
@@
 We found that the average median difference between actual and decoded location is small (7 cm median difference, 5-9 cm 95\% CI, for all times where the animal was moving greater than 4 cm/s, 5 fold cross validation).
-This level of accuracy is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}, and reflects in part the presence of "theta sequences" [REF] where position deviates behind and ahead of the animal across individual theta cycles.
+This level of accuracy is comparable to that of other studies \citep{DavidsonHippocampalReplayExtended2009, ShinDynamicsAwakeHippocampalPrefrontal2019, FarooqEmergencepreconfiguredplastic2019}, and reflects in part the presence of "theta sequences" \citep{DragoiTemporalEncodingPlace2006, FosterHippocampalthetasequences2007} where position deviates behind and ahead of the animal across individual theta cycles.
 Importantly, we achieved this using small time bins (2 ms) compared to the commonly used 250 ms time bin that most studies use for decoding during movement, demonstrating the power of our algorithm.
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{figures/Figure3/Figure3_final}
+\includegraphics[width=0.70\linewidth]{figures/Figure3/Figure3_final}
 \caption{A-F.
@@
 
-We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007, HangyaComplementaryspatialfiring2010}).
+We also verified that our classifications were very unlikely to be driven by the firing of interneurons, which are less likely to project to other brain areas and exhibit less spatial specificity than principal neurons in the hippocampus \citep{WilentDiscretePlaceFields2007, JinnoNeuronalDiversityGABAergic2007, HangyaComplementaryspatialfiring2010}.
 First, we increased our tetrode participation threshold from two up to five tetrodes active during the SWR, to eliminate the possibility that events with a single tetrode with a high rate interneuron might drive the classification of stationary or stationary-continuous mixture dynamics during SWRs.
@@
 Second, the stationary dynamics have a longer duration than one would expect from a burst alone (\FIG{5}B, 54 ms median duration, 38-74 ms quartiles vs.
-bursts which have a typical duration of 6-24 ms, \citep{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
+bursts which have a typical duration of 6-24 ms; \cite{RanckJrStudiessingleneurons1973, HarrisTemporalInteractionSingle2001, TroppSneiderDifferentialbehavioralstatedependence2006}).
 Third, only \texttildelow50\% of the spikes had interspike intervals (ISIs) of less than 6 ms during the stationary, stationary-continuous, and continuous dynamics (\FIGSUPP[5]{fig5sup3}A), indicating that rapid bursts made up at most half of the observed spikes, and likely fewer given that these ISIs are computed for each tetrode, rather than single units.
@@
 Second, we used clusterless decoding, which enabled us to take advantage of information from neurons farther away from the tetrodes and not just those that could be cleanly and somewhat subjectively separated \citep{ChenTransductiveneuraldecoding2012, KloostermanBayesiandecodingusing2014, DengRapidclassificationhippocampal2016}.
-We note that more fully automated modern spike sorting algorithms (such as MountainSort \citep{ChungFullyAutomatedApproach2017}) in combination with better recording technology could reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
+We note that more fully automated modern spike sorting algorithms (such as MountainSort; \cite{ChungFullyAutomatedApproach2017}) in combination with better recording technology could reduce the differences in results when using sorted spikes or clusterless decoding, as these sorting methods help reduce  experimenter subjectivity involved in spike sorting and identify larger numbers of putative single neurons.
 

### 2021-07-30 Article.tex
Update on Overleaf.

@@
 \subsection*{Encoding - Clusterless}
-In order to relate each tetrode's unsorted spiking activity and waveform features to position, we used kernel density estimation (KDE) to estimate the following distributions: $\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$.
+In order to relate each tetrode's unsorted spiking activity and waveform features to position during the encoding period, which we define as all movement times (time periods when the running speed is greater than 4 cm/s), we used kernel density estimation (KDE) to estimate the following distributions: $\pi(x)$, $p_{i}(x_k)$, and $p_{i}(x_k, \vec{m}_k)$.
 We used KDEs of the form:

### 2021-07-30 Article.tex
Update on Overleaf.

@@
 Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019}.
-Other studies have reported activity that, when combined across an entire event, seems to correspond roughly to a single location, although the dynamics of these events is typically not investigated (REF Dupret, Csicvari papers with cheeseboard task).
+Other studies have reported activity that, when combined across an entire event, seems to correspond roughly to a single location, although the dynamics of these events is typically not investigated \citep{Dupretreorganizationreactivationhippocampal2010}.
 Thus, while it would seem useful to replay memories associated with single locations or a small region of the environment, the prevalence of that type of replay in awake animals remains unclear.
@@
 Instead, replay events encompass a much richer variety of event speeds that could promote storage and updating of memories on different spatial and temporal scales, including slower moving representations that activate a spatially compact set of nearby locations and stationary representations of single locations.
-This is consistent with observations of events that, when decoded in a single time bin, represent a small neighborhood in the environment (CITE Dupret, etc.) Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
+This is consistent with observations of events that, when decoded in a single time bin, represent a small neighborhood in the environment \citep{Dupretreorganizationreactivationhippocampal2010}.
+Interestingly, a previous paper \citep{FarooqEmergencepreconfiguredplastic2019} reported that these stationary representations were most prevalent in juvenile rats and decreased to "chance" levels when the rats reached adulthood.
 Our results strongly suggest that they do not disappear, but are a very common occurrence with adult rats.

### 2021-08-24 src/standard_decoder.py
Add dt as an argument to the fitting

@@ -99,6 +99,7 @@ def fit_mark_likelihood(
     model=model,
     model_kwargs=model_kwargs,
     place_bin_size=place_bin_size,
+    dt=0.020
 ):
 
     (
@@ -126,7 +127,6 @@ def fit_mark_likelihood(
     ground_process_intensities = []
     joint_pdf_models = []
 
-    dt = 0.020  # seconds
     total_time = np.diff(np.asarray(
         linear_position.index.total_seconds())[[0, -1]])
     n_total_time_bins = total_time // dt

### 2021-08-24 src/figure6_supplemental1.py
Add which ripples were used and the time step

@@ -1,6 +1,5 @@
 import copy
 import os
-import sys
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -74,7 +73,7 @@ def fit_and_load_models(epoch_key):
         place_bin_center_ind_to_edge_id,
         nodes_df,
     ) = fit_mark_likelihood(
-        linear_position, multiunit_dfs, track_graph, center_well_id)
+        linear_position, multiunit_dfs, track_graph, center_well_id, dt=0.020)
 
     place_fields = np.asarray(sorted_spikes_classifier.place_fields_)
 
@@ -95,7 +94,7 @@ def get_clusterless_posteriors_and_fits(
         start_time, end_time, place_bin_centers, occupancy,
         joint_pdf_models, multiunit_dfs, ground_process_intensities,
         mean_rates, is_track_interior, place_bin_edges,
-        track_graph1, place_bin_center_ind_to_node):
+        track_graph1, place_bin_center_ind_to_node, dt):
     clusterless_likelihood, clusterless_time = predict_mark_likelihood(
         start_time,
         end_time,
@@ -106,6 +105,7 @@ def get_clusterless_posteriors_and_fits(
         ground_process_intensities,
         mean_rates,
         is_track_interior,
+        dt
     )
 
     (
@@ -131,6 +131,7 @@ def get_clusterless_posteriors_and_fits(
         place_bin_edges,
         track_graph1,
         place_bin_center_ind_to_node,
+        dt
     )
 
     clusterless_posterior = xr.DataArray(
@@ -160,9 +161,10 @@ def get_clusterless_posteriors_and_fits(
 def get_sorted_spikes_posteriors_and_fits(
         start_time, end_time, place_fields, spike_times, is_track_interior,
         place_bin_centers, place_bin_edges, track_graph1,
-        place_bin_center_ind_to_node):
+        place_bin_center_ind_to_node, dt):
     sorted_spikes_likelihood, sorted_spikes_time = predict_poisson_likelihood(
-        start_time, end_time, spike_times, place_fields, is_track_interior)
+        start_time, end_time, spike_times, place_fields, is_track_interior,
+        dt)
 
     (
         sorted_spikes_time,
@@ -187,6 +189,7 @@ def get_sorted_spikes_posteriors_and_fits(
         place_bin_edges,
         track_graph1,
[…]

### 2021-08-24 src/figure6_supplemental2.py
Add examples with 2 ms bins

@@ -1,0 +1,484 @@
+import copy
+import os
+
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+import xarray as xr
+from replay_trajectory_classification import SortedSpikesClassifier
+from src.analysis import (get_is_classified, get_place_field_max,
+                          get_probability)
+from src.figure_utilities import (PAGE_HEIGHT, TWO_COLUMN, save_figure,
+                                  set_figure_defaults)
+from src.parameters import (FIGURE_DIR, PROBABILITY_THRESHOLD,
+                            PROCESSED_DATA_DIR, STATE_COLORS)
+from src.standard_decoder import (fit_mark_likelihood, load_data,
+                                  normalize_to_posterior,
+                                  predict_clusterless_wtrack,
+                                  predict_mark_likelihood,
+                                  predict_poisson_likelihood)
+
+
+def fit_and_load_models(epoch_key):
+    # clusterless model
+    data_type, dim = "clusterless", "1D"
+    animal, day, epoch = epoch_key
+
+    clusterless_state_space_results = xr.open_dataset(
+        os.path.join(PROCESSED_DATA_DIR, f"{animal}_{day:02}_{epoch:02}.nc"),
+        group=f"/{data_type}/{dim}/classifier/ripples/",
+    )
+
+    # sorted spikes model
+    data_type, dim = "sorted_spikes", "1D"
+    animal, day, epoch = epoch_key
+
+    sorted_spikes_state_space_results = xr.open_dataset(
+        os.path.join(PROCESSED_DATA_DIR, f"{animal}_{day:02}_{epoch:02}.nc"),
+        group=f"/{data_type}/{dim}/classifier/ripples/",
+    )
+
+    model_name = os.path.join(
+        PROCESSED_DATA_DIR,
+        f"{animal}_{day:02}_{epoch:02}_{data_type}_{dim}_model.pkl"
+    )
+    sorted_spikes_classifier = SortedSpikesClassifier.load_model(model_name)
+
+    (
+        linear_position,
+        multiunit_dfs,
+        position_info,
+        ripple_times,
+        track_graph,
+        center_well_id,
+        spike_times,
+    ) = load_data(epoch_key)
+
+    (
+        occupancy,
+        joint_pdf_models,
+        multiunit_dfs,
+        ground_process_intensities,
+        mean_rates,
+        place_bin_centers,
+        place_bin_edges,
+        is_track_interior,
[…]

### 2021-08-25 src/figure6_supplemental1.py
More exactly specify which ripples were used as examples

@@ -462,22 +462,18 @@ def plot_figure(epoch_key, ripple_numbers, is_save_figure=False):
 
 
 if __name__ == '__main__':
-    epoch_key = 'bon', 3, 4
-    ripple_numbers = [12, 200]
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'bon', 3, 6
-    ripple_numbers = [5, 27, 90, 93]
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'bon', 4, 2
-    ripple_numbers = 158
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'Cor', 1, 4
-    ripple_numbers = 84
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'fra', 6, 6
-    ripple_numbers = 151
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
+    plot_figure(epoch_key=('bon', 4, 2),
+                ripple_numbers=158,
+                is_save_figure=True)
+    plot_figure(epoch_key=('fra', 6, 6),
+                ripple_numbers=151,
+                is_save_figure=True)
+    plot_figure(epoch_key=('Cor', 1, 4),
+                ripple_numbers=84,
+                is_save_figure=True)
+    plot_figure(epoch_key=('bon', 3, 4),
+                ripple_numbers=12,
+                is_save_figure=True)
+    plot_figure(epoch_key=('bon', 3, 6),
+                ripple_numbers=[5, 93],
+                is_save_figure=True)

### 2021-08-25 src/figure6_supplemental2.py

@@ -463,22 +463,18 @@ def plot_figure(epoch_key, ripple_numbers, is_save_figure=False):
 
 
 if __name__ == '__main__':
-    epoch_key = 'bon', 3, 4
-    ripple_numbers = [12, 200]
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'bon', 3, 6
-    ripple_numbers = [5, 27, 90, 93]
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'bon', 4, 2
-    ripple_numbers = 158
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'Cor', 1, 4
-    ripple_numbers = 84
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
-
-    epoch_key = 'fra', 6, 6
-    ripple_numbers = 151
-    plot_figure(epoch_key, ripple_numbers, is_save_figure=True)
+    plot_figure(epoch_key=('bon', 4, 2),
+                ripple_numbers=158,
+                is_save_figure=True)
+    plot_figure(epoch_key=('fra', 6, 6),
+                ripple_numbers=151,
+                is_save_figure=True)
+    plot_figure(epoch_key=('Cor', 1, 4),
+                ripple_numbers=84,
+                is_save_figure=True)
+    plot_figure(epoch_key=('bon', 3, 4),
+                ripple_numbers=12,
+                is_save_figure=True)
+    plot_figure(epoch_key=('bon', 3, 6),
+                ripple_numbers=[5, 93],
+                is_save_figure=True)

### 2021-08-25 elife_revision2.tex
Add revision tex

--- 2021-07-30 Article.tex
@@
 Interestingly, there is evidence for SWRs where the spiking corresponds to a single location.
-Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although another report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019}.
+Specifically, some SWRs contain spiking of neurons associated with single locations where animals are immobile \citep{YuDistincthippocampalcorticalmemory2017}, although two report suggested that events that represent a single location are only seen in young animals \citep{FarooqEmergencepreconfiguredplastic2019, MuessigCoordinatedEmergenceHippocampal2019}.
 Other studies have reported activity that, when combined across an entire event, seems to correspond roughly to a single location, although the dynamics of these events is typically not investigated \citep{Dupretreorganizationreactivationhippocampal2010}.
@@
 We do this by defining a state transition matrix that defines how the latent position can move from the previous time step (in our case, 2 ms).
-Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
+Previous findings suggest that replay may exhibit at least three distinct types of movement dynamics: stationary \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019, MuessigCoordinatedEmergenceHippocampal2019}, continuous \citep{DavidsonHippocampalReplayExtended2009}, and fragmented, which could correspond to both extended spatially incoherent representations and representations that jump from one position to another in a short time period \citep{PfeifferAutoassociativedynamicsgeneration2015}.
 We therefore define movement dynamic models to capture each of these possibilities.
@@
 We then carried out a series of control analyses to determine whether our results could be due to differences in spiking statistics across dynamics or interneuron-specific activity patterns.
-We first calculated the multiunit spiking rates and found that these were similar across the different dynamics.
-(\FIG{5}F).
+We first calculated the multiunit spiking rates and found that these were similar across the different dynamics (\FIG{5}F).
 This indicates that all dynamics, including stationary and stationary-continuous mixtures, were driven by comparable levels of sustained spiking information and could not be explained by the absence of spiking.
-Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs to identify events, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics \FIGSUPP[5]{fig5sup3}F).
+Further corroborating this, we found that when we re-ran our analysis using periods of high multiunit activity instead of SWRs to identify events, we still found a larger proportion of stationary and stationary-continuous dynamics relative to the continuous dynamics (\FIGSUPP[5]{fig5sup3}F).
 
@@
 Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of less than 1 m/s (<100 cm/s), which is consistent with the speeds at which rats traverse their environment.
-The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
+The next most common category were stationary events \citep{YuDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019, MuessigCoordinatedEmergenceHippocampal2019} that activated a persistent representation of a single location, while the "classical" replay events---continuous, extended trajectories through space---made up only about 19\% of classified events.
 Events with slower dynamics tended to persist for approximately the same length of time as events  with faster continuous dynamics (\texttildelow50-100 ms), consistent with a brief activation of a representation of a location or a short trajectory through space.
@@
 In addition to our SWR-based analyses, this approach could also be used to decode activity during movement, where we might expect that place-specific activity would express combinations of stationary and continuous dynamics reflecting different speeds of representational movement during each cycle of the theta rhythm.
-Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \cite{DragoiDistinctpreplaymultiple2013}.
+Similarly, because our approach uses the same encoder-decoder approach as the standard decoder, it can be used to investigate sequence dynamics before experience \citep{DragoiDistinctpreplaymultiple2013}.
 
@@
 
+\subsection*{Identifying Events of High Multiunit Activity}
+We identified times of high multiunit activity when the animal was immobile as a control analysis.
+Our approach was similar to \cite{DavidsonHippocampalReplayExtended2009}.
+High multiunit periods were identified as times when the z-scored multiunit population spiking activity was greater than 2 standard deviations for at least 15 ms and the animal was moving at speeds less than 4 cm/s.
+The code used for detection of high multiunit time periods can be found at \url{https://github.com/Eden-Kramer-Lab/ripple_detection}.
+
 \subsection*{Standard "Bayesian" Decoding and Significance Estimation}

### 2021-08-25 scripts/make_diff.sh
Diff with revision 2

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-latexdiff ../elife_revision1.tex ../Article.tex > ../diff.tex
+latexdiff ../elife_revision2.tex ../Article.tex > ../diff.tex

### 2021-08-25 elife_revision2.tex
Fix to correct revision

@@
 We first note that most of the SWRs (14,989 of 21,433 or 70\% of classified SWRs) were classified as having only a single dynamic or dynamic mixture (\FIG{5}A, for specific example SWR see \FIG{3}B), whereas SWRs with multiple dynamics or dynamic mixtures (such as those in \FIG{3}A, C, D, E, and F) were less common.
-Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}).
+Interestingly, the most common category of SWR were those with only stationary-continuous mixtures (8,944 of 21,433 or 42\% of classified SWRs, \FIG{5}A, \FIGSUPP[5]{fig5sup1}, note that this percentage ignores the unclassified category).
 These events contain representations that move at slower speeds (\FIG{5}D)---similar to those speeds experienced by the animal in the environment (\FIG{5}G, median 17 cm/s for run periods and 4 cm/s for all times)---and are sustained, but slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 73 ms vs.
@@
 % Figure 6-Supplemental 1
-\figsupp[Examples of fits from the standard "Bayesian" decoder and state space model.]{Examples of fits from the standard "Bayesian" decoder and state space model.
+\figsupp[Examples of fits from the standard "Bayesian" decoder with 20 ms bins and state space model.]{Examples of fits from the standard "Bayesian" decoder  with 20 ms bins and state space model.
 (A-F) For each panel, the top row is a raster plot of the sorted cells, the second row is the decoded posterior probability of position in 20 ms time bins for the sorted spikes and the corresponding Radon line fit (blue line) and the MAP fit (green line) on this posterior probability, the third row is a raster plot of the multiunit spike times used for clusterless decoding, the fourth row is the posterior probability in 20 ms time bins using clusterless data as well as the accompanying Radon fit and MAP fit, the fifth row is the posterior probability from the clusterless state space model, and the sixth row is the probability of each dynamic (lines) and the color shading indicates the category of dynamic as before.
 }{\includegraphics[width=1.0\linewidth]{figures/Figure6-supplemental1/Figure6-supplemental1_final}}\label{figsupp:fig6sup1}
+
+% Figure 6-Supplemental 2
+\figsupp[Examples of fits from the standard "Bayesian" decoder with 2 ms bins and state space model.]{Examples of fits from the standard "Bayesian" decoder  with 2 ms bins and state space model.
+(A-F) For each panel, the top row is a raster plot of the sorted cells, the second row is the decoded posterior probability of position in 2 ms time bins for the sorted spikes and the corresponding Radon line fit (blue line) and the MAP fit (green line) on this posterior probability, the third row is a raster plot of the multiunit spike times used for clusterless decoding, the fourth row is the posterior probability in 2 ms time bins using clusterless data as well as the accompanying Radon fit and MAP fit, the fifth row is the posterior probability from the clusterless state space model, and the sixth row is the probability of each dynamic (lines) and the color shading indicates the category of dynamic as before.
+}{\includegraphics[width=1.0\linewidth]{figures/Figure6-supplemental2/Figure6-supplemental2_final}}\label{figsupp:fig6sup2}
 
@@
 We then compared the speed of the replay for each dynamic and for the entire SWR (\FIG{6}).
-We found that the methods that rely on fitting a line for the entire SWR (the Radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, can capture both fast and slow speeds.
+We found that the methods that rely on fitting a line for the entire SWR (the Radon transform and the linear regression), tend to estimate that the replays proceed at faster speeds (\FIG{6}A,C), whereas the MAP method (\FIG{6}B), which only relies on picking the most probable position for a given time bin and allows for more variable speeds, meaning it can capture both fast and slow speeds.
 In addition, the speeds from the MAP method (\FIG{6}B) are quite similar to the speeds estimated by the state space decoder (\FIG{6}D).
-This is to be expected because the MAP method imposes less smoothing (although some is added by using 20 ms time bins) whereas the linear methods impose a larger amount of smoothing and cannot change directions (see \FIGSUPP[6]{fig6sup1}A for an example where this is beneficial because the trajectory is linear and the smoothing ignores a lack of spiking and \FIGSUPP[6]{fig6sup1}B for an example where this is not beneficial because the linear fit cannot change direction).
-Our state space decoder imposes much less smoothing than the linear methods and therefore is closer to the MAP estimates (\FIGSUPP[6]{fig6sup1}B), but also takes into account the uncertainty of each time bin (\FIGSUPP[6]{fig6sup1}C).
+This is to be expected because the MAP method imposes less smoothing (although some is added by using 20 ms time bins, see \FIGSUPP[6]{fig6sup2}A-F for the same examples with 2 ms bins) whereas the linear methods impose a larger amount of smoothing and cannot change directions (see \FIGSUPP[6]{fig6sup1}A for an example where this is beneficial because the trajectory is linear and the smoothing ignores a lack of spiking and \FIGSUPP[6]{fig6sup1}B for an example where this is not beneficial because the linear fit cannot change direction).
+Our state space decoder with 2 ms bins imposes much less smoothing than the linear methods and therefore is closer to the MAP estimates with 20 ms bins (\FIGSUPP[6]{fig6sup1}B), but also takes into account the uncertainty of each time bin (\FIGSUPP[6]{fig6sup1}C) and is able to change direction more quickly than the 20 ms time bins (\FIGSUPP[6]{fig6sup1}E).
+Using 2 ms bins with the MAP method would allow for faster change in the direction of the latent position but also results in much more variable estimates (\FIGSUPP[6]{fig6sup2}).
 These results indicate that the methods that made more assumptions about the structure of replay (Radon and linear regression approaches) yield different conclusions about which events are potentially meaningful and how those events progress over time from methods that make fewer assumptions (the MAP and state-space approaches).
